### 2023-01-04 Lab1_R_Review_Answers.Rmd
Add files via upload

---
title: '784 Lab 1: R Refresher: Answer Key'
author: "Eric Parajon"
output:
  pdf_document:
    keep_tex: TRUE
  html_document:
    df_print: paged
---

Code from today's lab was partially adapted from labs by Simon Hoellerbauer and Isabel Laterzo.

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, include = TRUE)
```

Today's lab will help you refresh some of what you learned in R in 783. It will also introduce you to some topics in R that you may not be familiar with, but are very useful for this semester.

For a refresher on how to use R Markdown, please see this cheat sheet: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf.

Within your code chunks, comment our your code describing what you are doing in each step. This is an important habit to establish.

#R Review
\medskip
1. First, it is important to save the document you are using in a central folder. Then, set your working directory to that folder. It's also important to clear your environment before working in a new document.
```{r, include = TRUE, eval = FALSE}
getwd() #what is your current working directory?

setwd() #in the parentheses, set the working directory to the folder where
#you save this file. Realistically, this isn't super relevant now because
#you're not reading in external data, but it's a good habit to get into!

rm(list=ls()) #clear global environment
```

\bigskip
2. First, create a vector of ten 5s using the function `rep()`. Second, create a vector that counts from 1 to 10 - do this in two ways, one using `c()` and one using `seq()`. Third, create a vector of length 10 filled with NAs. Finally, uses `cbind()` to bind theses four vectors together to create a 10 x 14 (10 rows, 4 columns) matrix. Remember, if you don't know how to use a function, type ? function name into your console to see the help file (e.g., `?seq()`).

Try binding the vectors together with `rbind()` instead. What happens?

```{r}

vec1 <- rep(5, 10) #vector of ten 5s

vec2 <- c(1:10) #vector counting from 1 to 10

vec3 <- seq(from = 1, to = 10, by = 1) #vector counting from 1 to 10

vec4 <- rep(NA, 10) #vector of 10 NAss

all_vecs <- cbind(vec1, vec2, vec3, vec4) #bind vectorss together using cbind
#we get four columns and ten rows

all_vecs2 <- rbind(vec1, vec2, vec3, vec4) #bind vectors together using rbind
#we get ten columns and four rows

```

\bigskip
3. Set your seed to 425. Simulate 100 draws from a Normal distribution with mean 2 and standard deviation 4, assigning the result to an object.

What proportion of the draws is above 3.1 in our sample? Think about using the functions `sum()` and `length()` to solve this, although there are other ways as well.

What is the expected proportion of draws above 3.1? Hint, look into the function `pnorm()`

Repeat this process, but this time take 10,000 draws from the same Normal distribution. How does our result compare with the expected proportion of draws above 3.1 now?

Thinking back to last semester, what did we just calculate using simulation?

```{r}
set.seed(425)

#simulate 100 draws from N(2, 4)
draws <- rnorm(100, 2, 4)

#calculate proportion of draws above 3.1
sum(draws > 3.1)/length(draws)
#0.39

#using pnorm to retrieve the expected proportion of draws above 1.5
1 - pnorm(3.1, 2, 4)
#just about matches above!

#repeating the above with 100,000 draws
draws2 <- rnorm(10000, 2, 4)
sum(draws2 > 3.1)/length(draws2)
#0.3864, very similar

#We simulated 1 - CDF of N(2, 4) at x = 3.1

```

\bigskip
4. Another important concept in R is a `for` loop. If you need a refresher on these, visit: https://www.r-bloggers.com/2015/12/how-to-write-the-first-for-loop-in-r/.

Using the 100 draws from the Normal distribution from Q3, create a new vector that is 1 when the vector of draws is greater than 3.1 and 0 when it is less than or equal to 3.1. First, do this using a `for` loop with an embedded `if` `else` sequence. What is the sum of this vector?

In addition to loops, `ifelse()` is also an important function. Look at the documentation of this function and repeat the above activity using `ifelse()`.

```{r}
#creating a container (here, a vector) into which we can dump the 1s and 0s. 
#We could do this with c() but it's more efficient to create
#a container of the proper size first 
draws_new <- vector("integer", length(draws))

#for loop that fills the container with 1 when draws > 3.1 and 0 elsewise
for (i in 1:length(draws)){
  if(draws[i] > 3.1){
    draws_new[i] <- 1
  } else draws_new[i] <- 0
}

sum(draws_new) #the number here looks familiar, no?

#repeat above using ifelse
draws_new2 <- ifelse(draws > 3.1, 1, 0)

#checking to make sure we have the same number
sum(draws_new2)
#yay! it is the same.

```

\bigskip
5. Load in the `swiss` dataset. A description for this data can be found here: https://stat.ethz.ch/R-manual/R-patched/library/datasets/html/swiss.html How many variables does it have? How many observations?

The variable `Catholic` is the \% of Catholic individuals, as opposed to Protestant in each province. First look at the summary of this variable. Then, create a new variable `Protestant` that is the percent of Protestant individuals in each province, just going with the assumption that all non-Catholics are Protestant here.

What is the mean percentage of this new variable?

```{r}
#loading in swiss dataset
data(swiss)

#find number of variables and number of observations
ncol(swiss); nrow(swiss)
#six variables, 47 observations

#summary
summary(swiss$Catholic)

#creating protestant variable
swiss$Protestant <- 100 - swiss$Catholic

#summary
summary(swiss$Protestant)

#finding mean of Protestant
mean(swiss$Protestant)
#58.86

```
\bigskip
6. Now, find out what type of object `swiss` is. Then, coerce `swiss` to a list by using `as.list()` and assign it to an objected named `swiss_list`. Look at the object with `View()` What happened? Use the `length()` function on `swiss`. What is the output? Why?

Create a matrix filled with random values that has 4 rows and 2 columns. Use `length()` on this matrix. Why might this be? Use `?` to read the documentation for `length()`.
```{r}
#finding the type of swisss
class(swiss) 
typeof(swiss)

#corece swiss to a list
swiss_list <- as.list(swiss)

#View
View(swiss_list)

#get length of swiss dataset 
length(swiss)

#7 - this is because swiss is a dataframe and a dataframe is really a list of
#same length vectors. This fact helps dataframes mimic matrices.

#this is an example matrix
test_mat <- matrix(c(1, 0, 3, 2, 4, 5, 6, 8), nrow = 4, ncol = 2)
#this is the length
length(test_mat)
#8

#?length()
```
\bigskip
7. Continuing to use the original `swiss` data, use a for loop to calculate the median of each of the variables. Store the result in a *named* vector (i.e., provide a name to each of the values so they are not just a bunch of random numbers with no context!).

Some useful functions here are `vector()`, `seq_along()`, and `names()`. Remember to use the help function `?` for further information about these functions.
```{r}
#creating container for for loop
medians <- vector("double", ncol(swiss))

#for loop that finds the median for each variable (column)
for (i in seq_along(swiss)){
  medians[i] <- median(swiss[, i])
}

#assigning the names of swiss (variable names/column headers) to the median
#vector
names(medians) <- names(swiss)

#print
print(medians)
```
\bigskip
8. Look into the three functions `apply()`, `sapply()`, and `lapply()`. What are the differences between these functions?

Repeat 7, but this time use the three above functions. Using each of these functions, examine what type of object they produce. How are they different from using a `for` loop? How are they different from each other in terms of what they produce? A useful function for examining different object types is `class()`.
```{r}
#examine functions
?apply()
?sapply()
?lapply()

#using apply
medians2 <- apply(swiss, 2, median)
print(medians2)
typeof(medians2)
#it created a named vector

#using lapply
medians3 <- lapply(swiss, median)
print(medians3)
typeof(medians3)
#outputs a named list

#using sapply
medians4 <- sapply(swiss, median)
print(medians4)
typeof(medians4)
#outputs a named vector
```
\bigskip
9. Now we will explore `ggplot2`, which is a great (but sometimess tedious) package for plotting data and results. You will probably grow to have a love/hate relationship with `ggplot2`. For help with this package, reference this cheat sheet: https://rstudio.com/wp-content/uploads/2016/11/ggplot2-cheatsheet-2.1.pdf

Using `ggplot`, create a scatterplot with the `swiss` dataset that plots `Education` against `Examination`. Add axes labels and a title. Does it look like there might be a relationship between these two variables?
```{r}
library(ggplot2)

ggplot(swiss, aes(x = Education, y = Examination)) +
  geom_point() + 
  labs(x = "Education", y = "Examination", title = "Scatterplot") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

#looks there there could be a positive relationship between Examination and Education 
# as Examination score increases, Education also increases

```

\newpage

\bigskip
10. It is also useful to be able to write your own functions in `R`. Sometimes, `R` doesn't have a function for exactly what you are looking for. Doing so is pretty straightforward. For a quick review, see:  https://swcarpentry.github.io/r-novice-inflammation/02-func-R/

For example, if I wanted to write a function that evaluated the equation \[f(x, y) = xy \] I could write one of the following options:

```{r, include = TRUE}
#example 1
fun_ex1 <- function(params) {
  x <- params[1]
  y <- params[2]

  result <- x*y

  return(result)
}

#evaluate
fun_ex1(c(3,4))

#example 2
fun_ex2 <- function(x, y) {

  result <- x*y

  return(result)
}

#evaluate
fun_ex2(3,4)

```

Now your turn. Write a function that evaluates the following equation:
\[f(x, y, z) = x^2 + 3xy + \sin(z)
\]
Have this function takes a vector as its single argument (see example 1), made up of the three parameters ($x, y, z$). Evaluate the function at $x = 3/4, y = 13, z = 3.4$.

```{r}
#function for the above equation
my_fun <- function(params){
  x <- params[1]
  y <- params[2]
  z <- params[3]

  result <- x^2 + 3*x*y + sin(z)

  return(result)
}

#evaluating the function at x = 3/4, y = 13, z = 3.4
my_fun(c(3/4, 13, 3.4))
```

\newpage
# Math Review:
\bigskip
11. Let's try out some linear algebra. Below are 2 vectors, a and b, and two matrices A and B. Use them for this section.
Here is a link to a linear algebra cheat sheet in R (and other languages): 
https://github.com/scalanlp/breeze/wiki/Linear-Algebra-Cheat-Sheet
```{r, include = TRUE}
#a
a <- c(4, -5, 4, 1)
#b
b <- c(3, 2, -7, 17)

#A
A <- matrix(c(1, -3, 4, 5, 6, 8, 9, -10), ncol = 4)

#B
B <- matrix(c(4:1, 8, 14, -2, 1), ncol = 2)
```

What is the dot product of $\mathbf{a}$ and $\mathbf{b}$? 
Does the result change if we take the dot product of $\mathbf{b}$ and $\mathbf{a}$?

What are the dimensions for the matrix $\mathbf{A}\mathbf{B}$? Is this new matrix invertible?

What are the eigenvectors and eigenvalues of this new matrix?

Verify that the values produced are actually eigenvalues.

```{r}
#dot product
a %*% b
b %*% a

#AB
AB <- A %*% B
solve(AB)
dim(AB)
#yes it is invertible

#eigen vectors and values
eigen_AB <- eigen(AB)

#verifying that they are eigenvalues
sapply(eigen_AB$values, function(x) det(AB - diag(2)*x))

```

### 2023-01-04 Lab1_R_Review.Rmd
Add files via upload

---
title: '784 Lab 1: R Refresher'
author: "Eric Parajon"
output:
  pdf_document:
    keep_tex: TRUE
  html_document:
    df_print: paged
---

Code from today's lab was partially adapted from labs by Simon Hoellerbauer and Isabel Laterzo.

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, include = FALSE)
```

Today's lab will help you refresh some of what you learned in R in 783. It will also introduce you to some topics in R that you may not be familiar with, but are very useful for this semester.

For a refresher on how to use R Markdown, please see this cheat sheet: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf.

Further see: https://posit.co/resources/cheatsheets/
For a list of all R studio cheatsheets (super helpful!)

Within your code chunks, comment our your code describing what you are doing in each step. This is an important habit to establish.

# R Review
\medskip
1. First, it is important to save the document you are using in a central folder. Then, set your working directory to that folder. It's also important to clear your environment before working in a new document.
```{r, include = TRUE, eval = FALSE}
getwd() #what is your current working directory?

setwd(____) #in the parentheses, set the working directory to the folder where
#you save this file. Realistically, this isn't super relevant now because
#you're not reading in external data, but it's a good habit to get into!

rm(list=ls()) #clear global environ
```

\bigskip
2. First, create a vector of ten 5s using the function `rep()`. 
Second, create a vector that counts from 1 to 10 - do this in two ways, one using `c()` and one using `seq()`.

Third, create a vector of length 10 filled with NAs. Finally, uses `cbind()` to bind theses four vectors together to create a 10 x 14 (10 rows, 4 columns) matrix.

Remember, if you don't know how to use a function, type ? function name into your console to see the help file (e.g., `?seq()`).

Try binding the vectors together with `rbind()` instead. What happens?

```{r}
vec1 <- rep(5, 10) #vector of ten 5s

vec2 <- c(1:10) #vector counting from 1 to 10

vec3 <- seq(from = 1, to = 10, by = 1) #vector counting from 1 to 10

vec4 <- rep(NA, 10) #vector of 10 NAs

all_vecs <- cbind(vec1, vec2, vec3, vec4) #bind vectors together using cbind

all_vecs2 <- rbind(vec1, vec2, vec3, vec4) #bind vectors together using rbind

```
\bigskip
3. Set your seed to 425. Simulate 100 draws from a Normal distribution with mean 2 and standard deviation 4, assigning the result to an object.

What proportion of the draws is above 3.1 in our sample? Think about using the functions `sum()` and `length()` to solve this, although there are other ways as well.

What is the expected proportion of draws above 3.1? Hint, look into the function `pnorm()`

Repeat this process, but this time take 10,000 draws from the same Normal distribution. How does our result compare with the expected proportion of draws above 3.1 now?

Thinking back to last semester, what did we just calculate using simulation?

```{r}
set.seed(425)

#simulate 100 draws from N(2, 4)
draws <- rnorm(100, 2, 4)

#calculate proportion of draws above 3.1
sum(draws > 3.1)/length(draws)

#using pnorm() to retrieve the expected proportion of draws above 1.5
1 - pnorm(3.1, 2, 4)
#Does this match the above?

#repeating the above with 100,000 draws
draws2 <- rnorm(10000, 2, 4)
sum(draws2 > 3.1)/length(draws2)
#Again is the value similar?

```

\bigskip
4. Another important concept in R is a `for` loop. If you need a refresher on these, visit: https://www.r-bloggers.com/2015/12/how-to-write-the-first-for-loop-in-r/.

Using the 100 draws from the Normal distribution from Q3, create a new vector that is 1 when the vector of draws is greater than 3.1 and 0 when it is less than or equal to 3.1. First, do this using a `for` loop with an embedded `if` `else` sequence. What is the sum of this vector?

In addition to loops, `ifelse()` is also an important function. Look at the documentation of this function and repeat the above activity using `ifelse()`.

```{r}

#creating a container (here, a vector) into which we can dump the 1s and 0s. 
#We could do this with c() but it's more efficient to create
#a container of the proper size first 
draws_new <- vector("integer", length(draws))

#for loop that fills the container with 1 when draws > 3.1 and 0 elsewise
for (i in 1:length(draws)){
  if(draws[i] > 3.1){
    draws_new[i] <- 1
  } else draws_new[i] <- 0
}

sum(draws_new) #the number here looks familiar, no?

#repeat above using ifelse
draws_new2 <- ifelse(____ > 3.1, 1, 0)

#checking to make sure we have the same number
sum(draws_new2)
#Is the value the same?

```

\bigskip
5. Practice: Load in the `swiss` dataset. A description for this data can be found here: https://stat.ethz.ch/R-manual/R-patched/library/datasets/html/swiss.html

How many variables does it have? How many observations?

The variable `Catholic` is the \% of Catholic individuals, as opposed to Protestant in each province. First look at the summary of this variable. Then, create a new variable `Protestant` that is the percent of Protestant individuals in each province, just going with the assumption that all non-Catholics are Protestant here.

What is the mean percentage of this new variable?

```{r}
#loading in swiss dataset
data(swiss)

#find number of variables and number of observations
ncol(____); nrow(____)
#How many variables and how many total observations?

#summary of Catholic
summary(____)

#creating protestant variable
swiss$Protestant <- 100 - swiss$____

#summary of Protestant
____(swiss$Protestant)

#finding mean of Protestant
____(swiss$Protestant)
#What is the mean value?

```
\bigskip
6. Now, find out what type of object `swiss` is. Then, coerce `swiss` to a list by using `as.list()` and assign it to an objected named `swiss_list`. Look at the object with `View()` What happened? Use the `length()` function on `swiss`. What is the output? Why?

Create a matrix filled with random values that has 4 rows and 2 columns. Use `length()` on this matrix. Why might this be? Use `?` to read the documentation for `length()`.

```{r}

#finding the type of swiss
____(swiss)

```

\bigskip
7. Continuing to use the original `swiss` data, use a for loop to calculate the median of each of the variables. Store the result in a *named* vector (i.e., provide a name to each of the values so they are not just a bunch of random numbers with no context!).

Some useful functions here are `vector()`, `seq_along()`, and `names()`. Remember to use the help function `?` for further information about these functions.
```{r}

#creating container for the loop
medians <- vector("double", ncol(____))

#for loop that finds the median for each variable (column)
for (i in seq_along(swiss)){
  medians[i] <- ____(swiss[, i])
}

#assigning the names of swiss (variable names/column headers) to the median
#vector
names(medians) <- names(swiss)

#Output your medians
print(____)

```

\bigskip
8. Look into the three functions `apply()`, `sapply()`, and `lapply()`. What are the differences between these functions?

Repeat 7, but this time use the three above functions. Using each of these functions, examine what type of object they produce. How are they different from using a `for` loop? How are they different from each other in terms of what they produce? A useful function for examining different object types is `class()`.
```{r}
#examine functions
?apply()
?sapply()
?lapply()

#using apply (2 is required because apply needs an arguement to margin (2 indicates what?))
medians2 <- ____(swiss, 2, median)
print(medians2)
typeof(medians2)

#using lapply

#using sapply

```

\bigskip
9. Now we will explore `ggplot2`, which is a great package for plotting data and results. You will probably grow to have a love/hate relationship with `ggplot2`. For help with this package, reference this cheat sheet: https://res.cloudinary.com/dyd911kmh/image/upload/v1666806657/Marketing/Blog/ggplot2_cheat_sheet.pdf.
Here is another useful link to a variety of graphs created using ggplot2: 
http://r-graph-gallery.com/

Reminder, load a package using library().

Using `ggplot`, create a scatterplot with the `swiss` dataset that plots `Education` against `Examination`. Add axes labels and a title.

Does it look like there might be a relationship between these two variables?
```{r}
library(____)

data(swiss)

P1<-ggplot(swiss, aes(x = ____, y = Examination)) +
  geom_point() +
  labs(x = "Edu", ____) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

P1

```

\newpage

\bigskip
10. It is also useful to be able to write your own functions in `R`. Sometimes, `R` doesn't have a function for exactly what you are looking for. Doing so is pretty straightforward. For a quick review, see:  https://swcarpentry.github.io/r-novice-inflammation/02-func-R/

For example, if I wanted to write a function that evaluated the equation \[f(x, y) = xy \] I could write one of the following options:

```{r, include = TRUE}
#example 1
fun_ex1 <- function(params) {
  x <- params[1]
  y <- params[2]

  result <- x*y

  return(result)
}

#evaluate
fun_ex1(c(3,4))

#example 2
fun_ex2 <- function(x, y) {

  result <- x*y

  return(result)
}

#evaluate
fun_ex2(3,4)

```

Now your turn. Write a function that evaluates the following equation:
\[f(x, y, z) = x^2 + 3xy + \sin(z)
\]

Have this function takes a vector as its single argument (see example 1), made up of the three parameters ($x, y, z$). Evaluate the function at $x = 3/4, y = 13, z = 3.4$.

```{r, echo = FALSE}

my_fun <- function(______){

  x <- ______
  y <- ______
  z <- ______

  result <- x^2 + 3*x*y + sin(z)

  return(result)
}

#evaluating the function at x = 3/4, y = 13, z = 3.4

my_fun(c(3/4, 13, ____))

```

\newpage
# Math Review:
\bigskip
11. Let's try out some linear algebra. Below are 2 vectors, a and b, and two matrices A and B. Use them for this section.

Here is a link to a linear algebra cheat sheet in R (and other languages): 
https://github.com/scalanlp/breeze/wiki/Linear-Algebra-Cheat-Sheet

```{r, include = TRUE}
#a
a <- c(4, -5, 4, 1)
#b
b <- c(3, 2, -7, 17)

#A
A <- matrix(c(1, -3, 4, 5, 6, 8, 9, -10), ncol = 4)

#B
B <- matrix(c(4:1, 8, 14, -2, 1), ncol = 2)
```

What is the dot product of $\mathbf{a}$ and $\mathbf{b}$? 
Does the result change if we take the dot product of $\mathbf{b}$ and $\mathbf{a}$?

What are the dimensions for the matrix $\mathbf{A}\mathbf{B}$? Is this new matrix invertible?

What are the eigenvectors(vector whose direction remains unchanged when a linear transformation is applied to it) and eigenvalues (scalars associated with a linear system of equations) of this new matrix?

Verify that the values produced are actually eigenvalues.

```{r, echo = FALSE}
#dot product
a %*% b
___ %*% a

#AB
AB <- A %*% ___
___(AB)

#eigen vectors and values
eigen_AB <- ___(AB)

#verifying that they are eigenvalues
sapply(eigen_AB$___, function(x) det(AB - diag(2)*x))

```

### 2023-01-04 Lab 1  R Review/Lab1_R_Review.tex
Add files via upload

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={784 Lab 1: R Refresher},
  pdfauthor={Eric Parajon},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{784 Lab 1: R Refresher}
\author{Eric Parajon}
\date{}

\begin{document}
\maketitle

Code from today's lab was partially adapted from labs by Simon
Hoellerbauer and Isabel Laterzo.

Today's lab will help you refresh some of what you learned in R in 783.
It will also introduce you to some topics in R that you may not be
familiar with, but are very useful for this semester.

For a refresher on how to use R Markdown, please see this cheat sheet:
\url{https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf}.

Further see: \url{https://posit.co/resources/cheatsheets/} For a list of
all R studio cheatsheets (super helpful!)

Within your code chunks, comment our your code describing what you are
doing in each step. This is an important habit to establish.

\#R Review \medskip
1. First, it is important to save the document you are using in a
central folder. Then, set your working directory to that folder. It's
also important to clear your environment before working in a new
document.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{getwd}\NormalTok{() }\CommentTok{\#what is your current working directory?}

\FunctionTok{setwd}\NormalTok{(\_\_\_\_) }\CommentTok{\#in the parentheses, set the working directory to the folder where}
\CommentTok{\#you save this file. Realistically, this isn\textquotesingle{}t super relevant now because}
\CommentTok{\#you\textquotesingle{}re not reading in external data, but it\textquotesingle{}s a good habit to get into!}

\FunctionTok{rm}\NormalTok{(}\AttributeTok{list=}\FunctionTok{ls}\NormalTok{()) }\CommentTok{\#clear global environ}
\end{Highlighting}
\end{Shaded}

\bigskip
2. First, create a vector of ten 5s using the function \texttt{rep()}.
Second, create a vector that counts from 1 to 10 - do this in two ways,
one using \texttt{c()} and one using \texttt{seq()}.

Third, create a vector of length 10 filled with NAs. Finally, uses
\texttt{cbind()} to bind theses four vectors together to create a 10 x
14 (10 rows, 4 columns) matrix.

Remember, if you don't know how to use a function, type ? function name
into your console to see the help file (e.g., \texttt{?seq()}).

Try binding the vectors together with \texttt{rbind()} instead. What
happens?

\bigskip
3. Set your seed to 425. Simulate 100 draws from a Normal distribution
with mean 2 and standard deviation 4, assigning the result to an object.

What proportion of the draws is above 3.1 in our sample? Think about
using the functions \texttt{sum()} and \texttt{length()} to solve this,
although there are other ways as well.

What is the expected proportion of draws above 3.1? Hint, look into the
function \texttt{pnorm()}

Repeat this process, but this time take 10,000 draws from the same
Normal distribution. How does our result compare with the expected
proportion of draws above 3.1 now?

Thinking back to last semester, what did we just calculate using
simulation?

\bigskip
4. Another important concept in R is a \texttt{for} loop. If you need a
refresher on these, visit:
\url{https://www.r-bloggers.com/2015/12/how-to-write-the-first-for-loop-in-r/}.

Using the 100 draws from the Normal distribution from Q3, create a new
vector that is 1 when the vector of draws is greater than 3.1 and 0 when
it is less than or equal to 3.1. First, do this using a \texttt{for}
loop with an embedded \texttt{if} \texttt{else} sequence. What is the
sum of this vector?

In addition to loops, \texttt{ifelse()} is also an important function.
Look at the documentation of this function and repeat the above activity
using \texttt{ifelse()}.

\bigskip
5. Practice: Load in the \texttt{swiss} dataset. A description for this
data can be found here:
\url{https://stat.ethz.ch/R-manual/R-patched/library/datasets/html/swiss.html}

How many variables does it have? How many observations?

The variable \texttt{Catholic} is the \% of Catholic individuals, as
opposed to Protestant in each province. First look at the summary of
this variable. Then, create a new variable \texttt{Protestant} that is
the percent of Protestant individuals in each province, just going with
the assumption that all non-Catholics are Protestant here.

What is the mean percentage of this new variable?

\bigskip
6. Now, find out what type of object \texttt{swiss} is. Then, coerce
\texttt{swiss} to a list by using \texttt{as.list()} and assign it to an
objected named \texttt{swiss\_list}. Look at the object with
\texttt{View()} What happened? Use the \texttt{length()} function on
\texttt{swiss}. What is the output? Why?

Create a matrix filled with random values that has 4 rows and 2 columns.
Use \texttt{length()} on this matrix. Why might this be? Use \texttt{?}
to read the documentation for \texttt{length()}.

\bigskip
7. Continuing to use the original \texttt{swiss} data, use a for loop to
calculate the median of each of the variables. Store the result in a
\emph{named} vector (i.e., provide a name to each of the values so they
are not just a bunch of random numbers with no context!).

Some useful functions here are \texttt{vector()}, \texttt{seq\_along()},
and \texttt{names()}. Remember to use the help function \texttt{?} for
further information about these functions.

\bigskip
8. Look into the three functions \texttt{apply()}, \texttt{sapply()},
and \texttt{lapply()}. What are the differences between these functions?

Repeat 7, but this time use the three above functions. Using each of
these functions, examine what type of object they produce. How are they
different from using a \texttt{for} loop? How are they different from
each other in terms of what they produce? A useful function for
examining different object types is \texttt{class()}.

\bigskip
9. Now we will explore \texttt{ggplot2}, which is a great package for
plotting data and results. You will probably grow to have a love/hate
relationship with \texttt{ggplot2}. For help with this package,
reference this cheat sheet:
\url{https://res.cloudinary.com/dyd911kmh/image/upload/v1666806657/Marketing/Blog/ggplot2_cheat_sheet.pdf}.
Here is another useful link to a variety of graphs created using
ggplot2: \url{http://r-graph-gallery.com/}

Reminder, load a package using library().

Using \texttt{ggplot}, create a scatterplot with the \texttt{swiss}
dataset that plots \texttt{Education} against \texttt{Examination}. Add
axes labels and a title.

Does it look like there might be a relationship between these two
variables?

\newpage

\bigskip
10. It is also useful to be able to write your own functions in
\texttt{R}. Sometimes, \texttt{R} doesn't have a function for exactly
what you are looking for. Doing so is pretty straightforward. For a
quick review, see:
\url{https://swcarpentry.github.io/r-novice-inflammation/02-func-R/}

For example, if I wanted to write a function that evaluated the equation
\[f(x, y) = xy \] I could write one of the following options:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#example 1}
\NormalTok{fun\_ex1 }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(params) \{}
\NormalTok{  x }\OtherTok{\textless{}{-}}\NormalTok{ params[}\DecValTok{1}\NormalTok{]}
\NormalTok{  y }\OtherTok{\textless{}{-}}\NormalTok{ params[}\DecValTok{2}\NormalTok{]}

\NormalTok{  result }\OtherTok{\textless{}{-}}\NormalTok{ x}\SpecialCharTok{*}\NormalTok{y}

  \FunctionTok{return}\NormalTok{(result)}
\NormalTok{\}}

\CommentTok{\#evaluate}
\FunctionTok{fun\_ex1}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{3}\NormalTok{,}\DecValTok{4}\NormalTok{))}

\CommentTok{\#example 2}
\NormalTok{fun\_ex2 }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x, y) \{}

\NormalTok{  result }\OtherTok{\textless{}{-}}\NormalTok{ x}\SpecialCharTok{*}\NormalTok{y}

  \FunctionTok{return}\NormalTok{(result)}
\NormalTok{\}}

\CommentTok{\#evaluate}
\FunctionTok{fun\_ex2}\NormalTok{(}\DecValTok{3}\NormalTok{,}\DecValTok{4}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Now your turn. Write a function that evaluates the following equation:
\[f(x, y, z) = x^2 + 3xy + \sin(z)
\]

Have this function takes a vector as its single argument (see example
1), made up of the three parameters (\(x, y, z\)). Evaluate the function
at \(x = 3/4, y = 13, z = 3.4\).

\newpage

\hypertarget{math-review}{%
\section{Math Review:}\label{math-review}}

\bigskip
11. Let's try out some linear algebra. Below are 2 vectors, a and b, and
two matrices A and B. Use them for this section.

Here is a link to a linear algebra cheat sheet in R (and other
languages):
\url{https://github.com/scalanlp/breeze/wiki/Linear-Algebra-Cheat-Sheet}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#a}
\NormalTok{a }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{4}\NormalTok{, }\SpecialCharTok{{-}}\DecValTok{5}\NormalTok{, }\DecValTok{4}\NormalTok{, }\DecValTok{1}\NormalTok{)}
\CommentTok{\#b}
\NormalTok{b }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DecValTok{2}\NormalTok{, }\SpecialCharTok{{-}}\DecValTok{7}\NormalTok{, }\DecValTok{17}\NormalTok{)}

\CommentTok{\#A}
\NormalTok{A }\OtherTok{\textless{}{-}} \FunctionTok{matrix}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\SpecialCharTok{{-}}\DecValTok{3}\NormalTok{, }\DecValTok{4}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{6}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{9}\NormalTok{, }\SpecialCharTok{{-}}\DecValTok{10}\NormalTok{), }\AttributeTok{ncol =} \DecValTok{4}\NormalTok{)}

\CommentTok{\#B}
\NormalTok{B }\OtherTok{\textless{}{-}} \FunctionTok{matrix}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{4}\SpecialCharTok{:}\DecValTok{1}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{14}\NormalTok{, }\SpecialCharTok{{-}}\DecValTok{2}\NormalTok{, }\DecValTok{1}\NormalTok{), }\AttributeTok{ncol =} \DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

What is the dot product of \(\mathbf{a}\) and \(\mathbf{b}\)? Does the
result change if we take the dot product of \(\mathbf{b}\) and
\(\mathbf{a}\)?

What are the dimensions for the matrix \(\mathbf{A}\mathbf{B}\)? Is this
new matrix invertible?

What are the eigenvectors(vector whose direction remains unchanged when
a linear transformation is applied to it) and eigenvalues (scalars
associated with a linear system of equations) of this new matrix?

Verify that the values produced are actually eigenvalues.

\end{document}

### 2023-01-04 Lab 2  Tidyverse Review/Lab2_Tidyverse_Review_Answers.Rmd
Add files via upload

---
title: '784 Lab 2: Review of Tidyverse and ggplot Answer key'
author: "Eric Parajon"
output:
  pdf_document:
    keep_tex: TRUE
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, include = FALSE)
```

# Tidyverse Review
\medskip
1. Again, it is important to save the document you are using in a central folder. Then, set your working directory to that folder. It's also important to clear your environment before working in a new document.
```{r, include = TRUE, eval = FALSE}
getwd() #what is your current working directory?

setwd(____) #in the parentheses, set the working directory to the folder where
#you save this file. Realistically, this isn't super relevant now because
#you're not reading in external data, but it's a good habit to get into!

rm(list=ls()) #clear global environment

#The pacman package is required to load/install the additional packages. Uncomment the line below to install the package.
#install.packages("pacman")

#Installing and loading packages for use
pacman::p_load(tidyverse,ggthemes,nycflights13)

```

2.The tidyverse allows us to use the piping operator `%>%`, which is a useful piece of code (base R can do this too with |>) .

It pipes an object (either by itself or created by another function) into the next function in the sequence, as the first argument. This makes the code much easier to read.

For example, we could calculate the standard deviation of the `mpg` variable in the `mtcars` dataset in the following ways:

```{r, include = TRUE, echo = TRUE}
library(tidyverse)

#normal
sd(mtcars$mpg)

#piping
mtcars$mpg %>% 
  sd()

#Does it return the same value?
```
That is a very simple example, of course, and using a pipe in this case does not really gain you much. However, pipes and the myriad of tools in the `tidyverse` can make our work much, much easier. The [`dplyr`](https://dplyr.tidyverse.org/) package in the `tidyverse` allows us to manipulate data much more easily than in base R.

`dplyr` allows us to manipulate our data in a structured way. The key functions are `mutate()`, which creates new variables, `select()`, which allows us to pick variable by characteristics of their names, `filter()`, which allows us to choose rows according to criteria we define, `summarise()`, which allows us to "reduce multiple values down to a single summary", and `arrange()`, which allows us to order the observations according values of 1 or more variables.

RStudio makes useful cheat sheets for their packages, including [`dplyr`](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf).

### Star Wars example
```{r}
#1. Star Wars example
data(starwars)
  #Let's make our data into a tibble

#Tibbles are evolution of data.frame
#Inherits nominally, but not strictly!
  #Does not match partial variable names
  #Does not convert character vectors to factors
  #Always returns a tibble when [ is used
  #Allows readable variable names

dplyr::as_tibble(starwars)

  #Now that we have our data imported, we should start exploring it to see exactly what we have

  #the glimpse() function is quite similar to base R's str function
glimpse(starwars)
 #What if we want to pull a specific piece of data?
snip<-slice(starwars, 10:11)

  #What if we want some summary stats on specific variables?
starwars %>%
  group_by(homeworld) %>%
    count() %>% 
arrange(desc(n))

starwars %>%
  group_by(hair_color) %>%
  count() %>% 
  arrange(desc(n))

```

```{r}
#We should start to do some column manipulation now that we have some variables we could be interested in 
d1<-select(starwars, name, species, mass,height,gender,films)

#So now that we have a smaller tibble let's try and do some analysis.
  #What species are the heaviest and which are the lightest?

d1 %>%
  group_by(species) %>%
  summarise(avg = mean(mass, na.rm=TRUE)) %>%
  arrange(desc(avg))
#Hutts are the heaviest!

#Okay great, but what if we want some information that's not explicitly contained in the data? 
  #What if for example, you wanted to know the BMI of a bunch of characters?
  #You can use mutate to create new variables.

Star_BMI<-d1 %>%
  mutate(bmi = mass/((height/100)^2)) %>%
  arrange(desc(bmi)) %>%
  select(name, species, height, mass, bmi)

#Now what if we only want to BMI of Humans characters?

Human_BMI<-Star_BMI %>% 
  filter(species=="Human")

#What if we wanted to create a new variable that combined a characters name with where they were born?
#This is easy using tidyr (another package in tidyverse suite)
starwars %>%
  unite(name_homeworld, name, homeworld) %>% 
  select(name_homeworld)

```
## Joins

```{r}

# Load data 1
(dates <- read_csv("https://bit.ly/DM22_dates"))

# Load data 2
(works <- read_csv("https://bit.ly/DM22_works"))

#Inner join
#With an inner_join(x, y), only observations that appear in both x and y datasets are returned

# Do inner join
dates_works <- inner_join(dates, works) 
(dates_works)

#Left & Right Join
#A left_join(x, y) call returns all observations in x, and drops observations that only appear in y
#Variables in y that do not exist in x are appended to x in the returned tibble.

#The same is true for a right_join(x, y), but with the roles of x and y reversed.

dates_someworks <- left_join(dates, works) 
(dates_someworks)

dates_allworks <- right_join(dates, works) 
print(dates_allworks, width=70)

#Full join
#Finally, a full_join(x, y) returns all observations and all variables in either in x or in y (filled with NAs)

alldates_allworks <- full_join(dates, works) 
print(alldates_allworks, width=70)

```

## Pivoting
Pivoting: Wide -> Long
With pivot_longer(), you can take values of a variable that are stored as variable names pivot them onto the cells of a single (implicit) variable column

```{r}
(relig_income)
#This data set has counts based on a survey which (among other things) asked people about their religion and annual income:

relig_income_long<-relig_income %>% 
  pivot_longer(!religion, names_to = "income", values_to = "count")

#The first argument is the dataset to reshape, relig_income.

#The second argument describes which columns need to be reshaped. In this case, it’s every column apart from religion.

#The names_to gives the name of the variable that will be created from the data stored in the column names, i.e. income.

#The values_to gives the name of the variable that will be created from the data stored in the cell value, i.e. count.

```
Pivoting: Long -> Wide

With pivot_wider(), we move in the opposite direction: your units are distributed over multiple rows, so pivot until you have a single case per row of your tibble

```{r}
(fish_encounters) 
#This dataset, contributed by Myfanwy Johnston, describes when fish swimming down a river are detected by automatic monitoring stations:

fish_encounters_wide<- fish_encounters %>% 
          pivot_wider(names_from = station, values_from = seen)

#In this case the NAs mean no fish was recorded so we can convert that to zero like this

fish_encounters %>% pivot_wider(
  names_from = station, 
  values_from = seen,
  values_fill = 0
)

```
#Part 2: ggplot

`ggplot` is very important, and requires some time and patience to get to know well. We won't have time to dive fully into it, but we will begin reviewing the following code so you become more familiar. R has its own built-in plot function, but ggplot is far more flexible and creates far more descriptive (and thus useful!) plots.

For a comprehensive 'cheat sheet', see: https://www.maths.usyd.edu.au/u/UG/SM/STAT3022/r/current/Misc/data-visualization-2.1.pdf

```{r}
#we have to load the package first
library(ggplot2)
```

To begin, we can use the 'qplot' or 'ggplot' commands. I use ggplot. It's more customization.

```{r}
data(mtcars)

qplot(x = disp, y = mpg, data = mtcars, color = cyl, geom = "point") 
#x and y are independent + dependent variables, 
#data can be specified (or use mtcars$var),
#color distinguishes a characteristic over which to map color
#and geom is a plot type

ggplot(data = mtcars, aes(x = disp, y = mpg)) + geom_point(aes(color = cyl))
#Note the similarities above. 
#ggplot is more flexible in that you map to different `aesthetics'- 
#hence why we use 'aes.' Plenty of examples below to dig into this more! 
```

What are some ways we can make this plot clearer?

```{r}
ggplot(data = mtcars, aes(x = disp, y = mpg)) + #we add 'layers' with +
  geom_point(aes(colour = cyl), size = 2) + #point=scatterplot
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon")+ 
  #label x and y axes, give the plot a main title. Note the quotes!
  scale_colour_continuous(name = "Cylinders")+ #label the scale
  theme_bw() #make the plot background blank w/grid lines
```

You can also assign this plot to an object and save that object as a .pdf file.

```{r, message=FALSE, warning=FALSE}
carsplot <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()

#One method for saving plots:
#pdf("Plot1.pdf")
#carsplot
#dev.off()

#Where do these plots go?!

#Unique to ggplot:
#ggsave("Plot1a.pdf") #can edit image sizes as well
```
## Themes
As you've hopefully seen ggplot is highly customization. One way to standardize your graphs and avoid repeating your code to control ancillary parameters (e.g. color palettes, axes labels, etc.) is to use a ggplot() theme that includes bundles of code. This is what we were doing using theme_bw() above.

You can also create custom themes yourself!

```{r}
#Here is a theme I use often

plot_theme = theme(text = element_text(size=11),  #set text size to 11
                   panel.grid.minor = element_blank(), #these next two lines set the grid lines to null
                   panel.grid.major = element_blank(),
                   panel.background = element_blank(),#these lines make the background of the plot blank
                   plot.background = element_rect(fill = "transparent",colour = NA),
                   axis.text = element_text(color="black"), #setting axis text to black 
                   axis.line = element_line(colour = "black"),#setting axis lines to black 
                   strip.text.y = element_text(angle = 0), axis.title.y = element_text(angle=90, hjust=.5, vjust=.5),#adjusting the angle of text
                   axis.title = element_text(lineheight=1.1), #increasing height of axis title
                   legend.position = "bottom",  #Placing the legend at the bottom of the figure
                   legend.key = element_rect(colour = "transparent",  fill = "transparent") #setting legend to transparent
)

#Using the theme
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  plot_theme

#As you can see there are tons of options that you can set yourself. Alternatively you can use existing gplot_themes. Here are some of the most popular
#bw
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()

#minimal
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_minimal()

#classic
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_classic()

#If you want even more themes you can use the ggthemes library!
#install.packages("ggthemes")
library(ggthemes)

#WSJ
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_wsj()

#538
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_fivethirtyeight()

#The Economist
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_economist()

```
## Facets
Using facets lets you create multi-panel plots based off a given variable.
There are two different ggplot functions to use to facet. 
```{r}
#1. facet_wrap() "wraps" a 1d ribbon of panels into 2d. Essentially, creating a square grid of panels, with rows and columns defined by combinations of two categorical variables (even if some combinations of faceting values do not exist in data). This is most useful when you want to facet by a single variable (or two)

carsplot_facet <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_wrap(~ am)

#2. facet_grid() forms a matrix of panels defined by row and column faceting variables, creating panels based on observed combinations of values in faceting variables

#facet_grid(.~variable) will return facets equal to the levels of variable distributed horizontally
#facet_grid(variable~.) will return facets equal to the levels of variable distributed vertically.

carsplot_facet_2 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_grid(vs ~ am)

carsplot_facet_3 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_grid(am ~ .)

carsplot_facet_4 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_grid(. ~ am)
```

What are other types of common geoms?

geom_histogram(): plots a histogram of single value; usually needs modification to number of bins.

geom_density(): density plot of single variable.

geom_bar(): vertical barplot by categorical variable

geom_point(): typical two-way scatterplot

geom_smooth(): simple line interpolator

#Part 3: Practice
## Tidyverse practice

Use the data below to first split the complete dataset into individual planes and then summarise each plane 
by counting the number of flights (count = n()) and computing the mean distance flown.

Next, determine what city is flown to the most often from NYC.

Finally, create a plot to create a barplot visualizing if flights are on time or delayed (defined as later than 5 minutes) by month. Additionally, ensure your plot has meaningful labels.

```{r}
#First, we need to install our data. This dataset includes airline on-time data for all flights departing NYC in 2013. 
#install.packages("nycflights13")

library(nycflights13)

data("flights")
dplyr::as_tibble(flights)

```

```{r}
by_tailnum <- group_by(flights, tailnum) %>% 
  summarise(count = n(), dist = mean(distance, na.rm = TRUE))

```

```{r}
#What city is flown to the most often from NYC?

destinations <- group_by(flights, dest) %>%
  summarise(planes = n_distinct(tailnum),
            flights = n())%>%
  arrange(desc(flights))

#calling to view top 10 
d1<-head(destinations,10)

```

```{r}
#Barplot
flight_delays<-flights %>% 
  mutate(dep_type = ifelse(dep_delay < 5, "on time", "delayed")) %>% 
   dplyr::group_by(dep_type,month) %>%
    dplyr::summarise(n = n()) %>% 
    filter(dep_type!="")

plot1<-ggplot(data=flight_delays, aes(x = month,y=n, fill = dep_type)) +
  geom_bar(stat="identity")+
  labs(main="Frequency of On Time or Delayed Arrivals by Month", y="Count", fill="Type")

```

### 2023-01-04 Lab 10 PoissonNBmodels/Lab10_PoissonNBmodels.Rmd
Add files via upload

---
title: "Lab 10: Poisson and Negative Binomial Models"
author: "Eric Parajon"
output: pdf_document
---

Code has been adapted from Simon Hoellerbauer and Isabel Laterzo.

Load necessary packages:
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, messsage = FALSE)
rm(list=ls())

#Load in packages
pacman::p_load(tidyverse,
               haven,#for importing data
               labelled,
               MASS,
               AER,
               foreign)
```

### Poisson and Negative Binomial Models

We can apply the types of exercises we've used for logits and multinomial logits to poisson models.

First, get our data. These data are counts from Maoz (2006) of interstate war.

(These data are available on Sakai.)

```{r, message=F, warning=F, include=F}
count <- read_dta("Maoz.dta")

```

We'll be using the variable "nowars20" as our DV - this counts the number of wars in the international system. Let's plot a histogram:

```{r, message=F, warning=F}
ggplot(data = count , aes(x= nowars20 )) +
  geom_histogram(bins=25, fill="turquoise") +
  labs(x="Number of MIDS") +
  theme_bw()
```

This variable is an interesting one in that it's possibly a bit larger (higher range of counts) than many we'd encounter.

Nonetheless, let's build our model. We're going to look at the number of militarized interstate wars based on 'trade network polarization' (lgtrdnpi2) and the number of democratic cliques (clqdemreg) in the system:

```{r}
m1 <- glm(nowars20 ~ lgtrdnpi2 + clqdemreg,
          data = count,
          family = "poisson")
summary(m1)
```
Note that this tells us "Dispersion parameter for poisson family taken to be 1."

So, our constant gives the prediction if all X variables at zero (not always substantively meaningful, since some variables are not/cannot be 0).

The coefficient of each x is the expected difference in y (on the logarithmic scale) for each additional unit of the X. To interpret coefficients, we can exponentiate them and treat them as multiplicative effects.

Before we delve further into this model, we might want to check for overdispersion (what's this?).

One approach is to plot the mean of the data against the estimated variance. Try doing this by using the `fitted()` function below (look up what it is!), and then following the code.

```{r, message=F, warning=F}
fits <- fitted(m1) #fitted()

over <- data.frame(cbind(fits, (model.frame(m1)$nowars20-fits)^2)) #Use squared residuals to approx. value of the variance

ggplot(data=over, aes(x=fits, y=V2))+
  geom_point()+
  geom_smooth(aes(x=fits, y=V2), method=lm, se=FALSE, color="blue")+ #line for mean of estimates against the variance of estimates
  geom_smooth(aes(x=fits, y=fits), method=lm, se=FALSE, color='red')+ #assumes mean = variance
  theme_bw()
  #Add legend
```

We can also check with the dispersiontest() in the AER package:

```{r, message=F, warning=F}
dispersiontest(m1)
```
So, our ALTERNATIVE hypothesis is that the true dispersion is greater than 1. What is our p-value? Do we reject the null? (What is the null?)

--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----
--- don't look ----

YES! Looks like we might be dealing with some overdispersion. What can we do to correct this? Let's try a new model - a negative binomial model. While Poisson assumes that the mean and variance are the same, the Negative Binomial model better fits data that shows variation which is greater than the mean. The NB model has an additional parameter compared to the Poisson, and this parameter allows us to adjust for more dispersion in the data. In fact, the Poisson model is a special case of the NB.

The `glm` function does not handle NB models. We now need to use a new function - `glm.nb` - which is especially for NB models.

```{r, message=F, warning=F}

m2<- glm.nb(nowars20 ~ lgtrdnpi2 + clqdemreg, 
            data = count)
summary(m2)

coefs<-cbind(coef(m1), coef(m2))
coefs

```
What happened to the coefficients?

We also can look at AIC - let's compare the two models:

```{r}
#compare models using AIC
#which is better?

AIC(m1)
AIC(m2)

#AIC in m2 is lower
```

Let's also compare these models using a rootogram. We're going to install a package here from a repo which makes it easy to do this. Rootograms help us understand the predictive accuracy of a model based on actual data. It will show us the expected counts given the model vs. the observed counts. It does so by square rooting the counts, which allows us to more easily examine departures from our expectations even at smaller frequencies.

```{r}
install.packages("countreg", repos="http://R-Forge.R-project.org")
library(countreg)
library(cowplot)

root.pois <- rootogram(m1, style = "hanging", plot = FALSE)
root.nb <- rootogram(m2, style = "hanging", plot = FALSE)

ylims <- ylim(-5, 11)  # common scale for comparison
plot_grid(autoplot(root.pois) + ylims, autoplot(root.nb) + ylims, 
          ncol = 2, labels = c("Poisson", "Neg. Binom"))

```
What does this tell us? The rootogram reflects fitted counts. If a bar doesn't reach the zero line, the model is overpredicting a certain count bin. If it exceeds the zero line, it underpredicts. So, neither of these models are perfect. BUT - the poisson model seems to both over and underpredict a lot. The negative binomial model on the other hand definitely over predicts, but not as much.

## Predicted Counts

For the sake of illustration (and because we see it is a better fitting model), let's use the NB model to plot predicted counts, using simulation.

```{r, message=F, warning=F}
set.seed(8970)
Betas <- m2$coefficients
vcv <- vcov(m2)

#Set a number of simulations. 
m2_sims <-1000

#Create our matrix of simulated data- we'll limit our dataset to do so.
sims_mat <- mvrnorm(m2_sims, Betas, vcv)
dim(sims_mat)

temp_data <- model.matrix(m2)

pc_sims <- matrix(NA, nrow = m2_sims, ncol = nrow(temp_data))

#Fill in our predicted probabilities with a for() loop! 
for(i in 1:m2_sims){
  pc_sims[i, ] <- exp(temp_data%*%sims_mat[i, ]) 
}

#What if we wanted a single point estimate of our predicted count of wars given our IVs?
mean(pc_sims)

#And the upper and lower boundaries of a confidence interval based on this simulation?
quantile(pc_sims, prob= 0.025) 
quantile(pc_sims, prob= 0.975) 
```

How do we predict counts over the range of a variable? Like we always have! Let's say we want to know what happens to the predicted number of counts as the average number of democratic cliques varies from its min to its max.
```{r}
temp_data <- count

dem_cl <- seq(from = min(count$clqdemreg, na.rm = T), 
              to = max(count$clqdemreg, na.rm = T),
              length.out = 50)

pc_dem_cl <- purrr::map_dfr(dem_cl, function(x){

  temp_data$clqdemreg <- x

  pc_df <- exp(model.matrix(m2, data = temp_data) %*% t(sims_mat))

  pc_val <- as.data.frame(t(colMeans(pc_df)))

})

#we could do this in one step, but I like keeping all of the simulated counts so
#we can do whatever we want with the distribution of predicted counts
plot_data <- t(apply(pc_dem_cl, 1, quantile, probs = c(.1, .5, .9))) %>% as.data.frame()
plot_data$dem_cl <- dem_cl
```

To do on your own:

Below, try using ggplot to plot this information we just generated. Follow these guidelines:
- X axis: Average number of Democratic Cliques
- Y axis: Predicted count of Intl. Wars
- Include a geom_line() for Dem_Cliques
- Include a geom_ribbon() with CIs (80% confidence interval)
- Include a legend which specifies what the line is and what the CIs are, try putting that legend on the bottom

```{r}
#Now to plot: 
colnames(plot_data)<-c("low", "med", "high", "Dem_Cliques")

#Collapse information into 3 vectors: for each value of trade in our data, 
#give us the average point estimate (expected count), the low bound of a confidence interval, and the high bound.

pcplot <- ggplot(_____)

pcplot

```

What do you find?

### 2023-01-04 Lab 10 PoissonNBmodels/Lab10_PoissonNBmodels_AnswerKey.Rmd
Add files via upload

--- 2023-01-04 Lab 10 PoissonNBmodels/Lab10_PoissonNBmodels.Rmd
@@
 ---
-title: "Lab 10: Poisson and Negative Binomial Models"
-author: "Eric Parajon"
+title: "Lab 10: Poisson and Negative Binomial Models: Answers"
+author: "TA: Eric Parajon"
 output: pdf_document
@@
 rm(list=ls()) 
-
 #Load in packages
@@
 
-### Poisson and Negative Binomial Models
+### Poisson
 
@@
 
-(These data are available on Sakai.)
+*You'll need to change the directory to bring in the file.*
 
@@
 count <- read_dta("Maoz.dta")
-
+#count2 <-na.omit(count) # <- don't do this with your data unless you are absolutely
+#sure that you want to drop all nas 
 ```
 
-We'll be using the variable "nowars20" as our DV - this counts the number of wars in the international system. Let's plot a histogram:
+We'll be using the variable "nowars20" as our DV - this counts the number of wars in the international system. Here's our histogram:
 
 ```{r, message=F, warning=F}
-ggplot(data = count , aes(x= nowars20 )) +
+ggplot(data=count, aes(x=nowars20)) +
   geom_histogram(bins=25, fill="turquoise") +
@@
 ```{r}
-m1 <- glm(nowars20 ~ lgtrdnpi2 + clqdemreg,
-          data = count,
-          family = "poisson")
+m1 <- glm(data = count, nowars20 ~ lgtrdnpi2 + clqdemreg, family="poisson")
 summary(m1)
@@
 
-One approach is to plot the mean of the data against the estimated variance. Try doing this by using the `fitted()` function below (look up what it is!), and then following the code.
+One approach is to plot the mean of the data against the estimated variance.
 
 ```{r, message=F, warning=F}
-fits <- fitted(m1) #fitted()
+fits <- fitted(m1) #what does fitted do, do you think?
 
@@
 ```
-So, our ALTERNATIVE hypothesis is that the true dispersion is greater than 1. What is our p-value? Do we reject the null? (What is the null?)
+So, our ALTERNATIVE hypothessis is that the true dispersion is greater than 1. What is our p-value? So what is supported - null or alternative hyptohesis? Do we have overdispersion?
 
@@
 Let's try a new model - a negative binomial model.
-While Poisson assumes that the mean and variance are the same, the Negative Binomial model better fits data that shows variation which is greater than the mean.
+While Poisson assumes that the mean and variance are the same, the Negative Binomial model better fits data shows variation which is greater than the mean.
 The NB model has an additional parameter compared to the Poisson, and this parameter allows us to adjust for more dispersion in the data.
@@
 
-m2<- glm.nb(nowars20 ~ lgtrdnpi2 + clqdemreg, 
-            data = count)
+m2<- glm.nb(data = count, nowars20~lgtrdnpi2+clqdemreg)
 summary(m2)
@@
 ```
-What happened to the coefficients?
+What happened to the coefficients? What happened to the standard errors?
 
@@
 ```{r}
-#compare models using AIC
-#which is better?
-
 AIC(m1)
 AIC(m2)
-
-#AIC in m2 is lower
+#which is better?
 ```
 
-Let's also compare these models using a rootogram.
-We're going to install a package here from a repo which makes it easy to do this.
-Rootograms help us understand the predictive accuracy of a model based on actual data.
-It will show us the expected counts given the model vs.
-the observed counts.
-It does so by square rooting the counts, which allows us to more easily examine departures from our expectations even at smaller frequencies.
+Let's also compare these models using a rootogram:
 
 ```{r}
-install.packages("countreg", repos="http://R-Forge.R-project.org")
+#install.packages("countreg", repos="http://R-Forge.R-project.org")
 library(countreg)
@@
 
-## Predicted Counts
+#Predicted Counts
 
@@
 #Set a number of simulations. 
-m2_sims <-1000
+m2_sims<-1000
 
 #Create our matrix of simulated data- we'll limit our dataset to do so.
-sims_mat <- mvrnorm(m2_sims, Betas, vcv)
+sims_mat <- mvrnorm(m2_sims, Betas, vcv) #asymptotic sampling distribution of the coefs is the same as in any other GLM
 dim(sims_mat)
@@
 
-#Fill in our predicted probabilities with a for() loop! 
+#Fill in our expected counts with a for() loop
 for(i in 1:m2_sims){
@@
 
-To do on your own:
-
-Below, try using ggplot to plot this information we just generated. Follow these guidelines:
-- X axis: Average number of Democratic Cliques
-- Y axis: Predicted count of Intl. Wars
-- Include a geom_line() for Dem_Cliques
-- Include a geom_ribbon() with CIs (80% confidence interval)
-- Include a legend which specifies what the line is and what the CIs are, try putting that legend on the bottom
-
 ```{r}
@@
 
-pcplot <- ggplot(_____)
-
-pcplot
-
+pc_plot <- ggplot(data=plot_data) + 
+  labs(x = "Average Number of Democratic Cliques", 
+       y = "Predicted Count of Intl. Wars") +
+  geom_line(aes(x = Dem_Cliques, y = med, 
+                color = "Number of Interstate Wars"), size = 1) +
+  geom_ribbon(aes(x = Dem_Cliques, ymin = low, ymax = high, 
+                  fill = "80% Confidence Interval (Simulation)"), alpha = .5) +
+  theme_bw() +
+  theme(legend.title = element_text(), 
+        legend.position = "bottom", panel.grid.minor.x=element_blank(),
+        panel.grid.minor.y = element_blank()) +
+  scale_fill_manual(values = "steelblue3", name = "")+
+  scale_color_manual(values = "black", name = "")
+pc_plot
 ```
-
-What do you find?

### 2023-01-04 Lab 11 Missing data/Lab11_MissingData.Rmd
Add files via upload

---
title: 'Lab 11 - Missing Data and Imputation'
author: "Eric Parajon"
header-includes:
  - \usepackage{enumerate}
  - \usepackage{graphicx}
output: pdf_document
---

This week we will be learning ways to handle missing data. Often times you will come across, or collect, data with a significant amount of missingness across variables of interest. If the amount of data is quite small, typically you can ignore it. However, you will likely come across cases where your data has a pretty significant amount of missingness. A common rule of thumb for "too much" missingness - indicating you need to deal with it in some way - is if > 5% of your data is missing.

Today, we will be examining how we would deal with data that has a significant amount of missingness. We will walk through listwise deletion, mean/median/mode substitution, and something called "imputation."

**Data!**
First, let's get our data set up. 
```{r}
#clear global environ
rm(list=ls())

#load in airquality data
data_clean <- airquality
summary(data_clean)

#here's our "true model" for future reference
model_true <- lm(Temp~ Ozone + Solar.R + Wind, data = data_clean)

#lets copy this df and then add some missingness
data_miss <- data_clean
data_miss[4:9,3] <- rep(NA,6)
data_miss[1:8,4] <- NA
summary(data_miss) #check out those *~* NAs ~*~

```

Alright, let's take a look at those NA's more deeply
```{r}
#what does this do?
is.na(data_miss$Ozone)

#and this?
which(is.na(data_miss$Ozone))

#and this?
sum(is.na(data_miss$Ozone))

#apply this to other variables!

#An easy way to look at this across the whole DF
colSums(is.na(data_miss))
```

**Listwise Deletion**
Some people just delete the observations with missing values. Although this is definitely a viable approach, depending on how much missing data you have, you might be getting rid of a lot of information!

Below, use the function `na.omit()` to get rid of observations with missing values, then repeat the above `lm()` model with this new data.
```{r, eval = FALSE}
# subset with complete.cases to get complete cases
data_listwise <- na.omit(data_miss)
summary(data_listwise)

#let's create an lm with this model to compare to later iterations
#model Temp as your y, Ozone, Solar.R, and Wind as your x values
model_listwise <- lm(Temp ~ Ozone + Solar.R + Wind, data = data_listwise)

summary(model_listwise)
```

**Mean/Median Recoding**

Alternatively, we can replace missing values with other values - such as the mean, median, or mode. Mode is appropriate if you're dealing with categorical variables, but since we're dealing with continuous here, we'll use one of the others.

Although this is a pretty easy approach, mean/median recoding decreases and changes the variance of your data. Especially in the case of time series data, this might not be great - we might be replacing a value with a mean of *all* the existing values, instead of one appropriate for the time period it belongs to! But, let's try it anyways.

```{r, eval = FALSE}
#make a copy of the data
data_recode <- data_miss

#uses if else statements to recode NA values as the mean of the rest of that column
#is.an()
data_recode$Ozone <- ifelse(is.na(data_recode$Ozone), 
                            mean(data_recode$Ozone, na.rm = TRUE),
                            data_recode$Ozone)

data_recode$Solar.R<- ifelse(is.na(data_recode$Solar.R),
                            mean(data_recode$Solar.R, na.rm = TRUE),
                            data_recode$Solar.R)

data_recode$Wind<- ifelse(is.na(data_recode$Wind),
                            mean(data_recode$Wind, na.rm = TRUE),
                            data_recode$Wind)

data_recode$Temp <- ifelse(is.na(data_recode$Temp),
                            mean(data_recode$Temp, na.rm = TRUE),
                            data_recode$Temp)
#summarize
summary(data_recode)

#linear model (same as above, but rename "model_recode")
model_recode <- lm(Temp~ Ozone + Solar.R + Wind, data = data_recode)

summary(model_recode)
```

**Multiple Imputation**

Now multiple imputation, a super useful technique. First, we'll discuss `mice`. This package uses multivariate imputations to estimate missing values. What does this mean? Well, if a value is identified as missing, the package will then regress over the other variables and predict the various missing values. This is pretty cool. Further, using *multiple* imputations, or iterations of this process, instead of just one, helps us reduce uncertainty.

Let's take a look!

```{r, eval = FALSE}
#Load in packages
pacman::p_load(mice,
               VIM) #Visualization

#mice's df visualization
md.pattern(data_miss)

#vis using aggr() from VIM
aggr_plot <- aggr(data_miss, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE,
                  labels=names(data_miss), cex.axis=.7, gap=3,
                  ylab=c("Histogram of missing data","Pattern"))
#what is this showing us?

#MICE TIME
mice_data <- mice(data_miss, #call our data
                  m=5, #number of multiple imputations
                  maxit=50, #number of iterations
                  # method = ? mice will decide the most appropriate method based on your data
                  seed=500)

summary(mice_data)

#try this out with 25 iterations and 10 imputations
#select animputation method (try pmm) - what does it do differently?
mice_data_2 <- mice(data_miss, #call our data
                  m=_____, #number of multiple imputations
                  maxit=_____, #number of iterations
                  method = _____,
                  seed=500)

summary(mice_data_2)
```

So this is great, but then what do we do if we want to run a model? We have five data sets. One option is to pool the data this way:
```{r, eval = FALSE}
#pooling
model_mice <- with(mice_data,lm(Temp~ Ozone + Solar.R + Wind))
summary(pool(model_mice))
```
So mice is great. Let's also learn about Amelia II (my personal favorite, named after Amelia Earhart, because she is missing). Amelia is a little different because it assumes that your data is jointly distributed as multivariate normal. It uses the "expectation-maximization with bootstrapping" as an algorithm for imputation. The EM algorithm alternates between something called the expectation (E-step) and maximization (M-step) steps until it converges on values that were missing - it converges when the current and previous values are quite close (sounds a bit like Newton Raphson, no?). It brings in bootstraping by doing this process on multiple bootstrapped samples drawn from your original (and incomplete) data.

Great, let's try it out. Examine the function by calling `?amelia()` - look at the different arguments. Notice that amelia can handle time series data uniquely. Let's factor that in by creating a running time variable.

```{r, message = FALSE}

#Load in packages
pacman::p_load(Amelia,
               lubridate,
               xtable)#For tables

#create running time variable, lets just assume we're in 2021
data_miss$time <- make_datetime(month = data_miss$Month,
                                day = data_miss$Day,
                                year = 2023)

#impute
amelia_data <- amelia(x = data_miss,
                 m = 5,
                 time = "time")

#let's run the model again, but this time using a for loop
models_amelia <- vector("list", 5)  #vector for our 5 newly imputed data sets

for(i in 1:5){
  models_amelia[[i]] <- lm(Temp ~ Ozone + Solar.R + Wind,
                        data=amelia_data$imputations[[i]]) #why do we index like this?
}

#Pooling models (using mice)
summary(pool(models_amelia))

#Pulling into DF
pooled_df <- data.frame(summary(pool(models_amelia)))

#Displaying in table
xtable(pooled_df)

```

Compare how models from mice, Amelia, and the above methods perform to our true data:
```{r, eval = FALSE}
#true model
summary(model_true)

#listwise
summary(model_listwise)

#recoding
summary(model_recode)

#mice
summary(pool(model_mice))

#Amelia
summary(pool(models_amelia))
```

**On your own:**

Try Amelia again, but this time let's use more of its functionality. Read in data from the `africa` package and examine its missingness. Then, using Amelia, impute missing values. Note that the dataset is cross-sectional by country, time series by year, and should have the `gdp_pc` variable specified as logged.

After imputing the data, write a linear model that evaluates the effect of `infl`, `trade`, `civlib` and `population` on `gdp_pc`. Report the pooled results!

```{r, eval = FALSE}
library(Amelia)
data("africa") #read in

#explore missingness
apply(africa, 2, function(x) sum(is.na(x))) #missingness by column

#use Amelia to generate 5 imputed datasets
#use ts and cs so imputation algerithem knows which variables represent countries and years

africa_imp <- amelia(x = __________, 
                     cs = __________ 
                     ts = __________
                     logs = "gdp_pc")

models_africa <- vector(__)
for(i in 1:length(africa_imp$imputations)){
  __________________________________) 
}

#report results, including standard errors

```

### 2023-01-04 Lab 11 Missing data/Lab11_MissingData_AnswerKey.Rmd
Add files via upload

--- 2023-01-04 Lab 11 Missing data/Lab11_MissingData.Rmd
@@
 ---
-title: 'Lab 11 - Missing Data and Imputation'
+title: 'Lab 11 - Missing Data and Imputation:Answers'
 author: "Eric Parajon"
@@
 
-Below, use the function `na.omit()` to get rid of observations with missing values, then repeat the above `lm()` model with this new data.
-```{r, eval = FALSE}
+Below, use the function `na.omit()` to get rid of observations with missing values
+```{r}
 # subset with complete.cases to get complete cases
@@
 #model Temp as your y, Ozone, Solar.R, and Wind as your x values
-model_listwise <- lm(Temp ~ Ozone + Solar.R + Wind, data = data_listwise)
+model_listwise <- lm(Temp~ Ozone + Solar.R + Wind, data = data_listwise)
 
@@
 
-```{r, eval = FALSE}
+```{r}
 #make a copy of the data
@@
 #uses if else statements to recode NA values as the mean of the rest of that column
-#is.an()
 data_recode$Ozone <- ifelse(is.na(data_recode$Ozone),
@@
                             data_recode$Temp)
+
 #summarize
@@
 **Multiple Imputation**
-
 Now multiple imputation, a super useful technique.
@@
 #vis using aggr() from VIM
-aggr_plot <- aggr(data_miss, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE,
-                  labels=names(data_miss), cex.axis=.7, gap=3,
-                  ylab=c("Histogram of missing data","Pattern"))
+aggr_plot <- aggr(data_miss, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(data_miss), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))
 #what is this showing us?
@@
 mice_data_2 <- mice(data_miss, #call our data
-                  m=_____, #number of multiple imputations
-                  maxit=_____, #number of iterations
-                  method = _____,
+                  m=10, #number of multiple imputations
+                  maxit=25, #number of iterations
+                  method = "pmm",
                   seed=500)
@@
 summary(mice_data_2)
+
 ```
@@
 So this is great, but then what do we do if we want to run a model? We have five data sets. One option is to pool the data this way:
-```{r, eval = FALSE}
+```{r}
 #pooling
@@
 summary(pool(model_mice))
+
 ```
@@
 
-```{r, message = FALSE}
-
+```{r}
 #Load in packages
 pacman::p_load(Amelia,
-               lubridate,
-               xtable)#For tables
+               lubridate)
 
@@
 #let's run the model again, but this time using a for loop
-models_amelia <- vector("list", 5)  #vector for our 5 newly imputed data sets
+models_amelia <- list() #create with as many elements that you know exist
+#create a vector of type list of size five
 
@@
 
-#Pooling models (using mice)
 summary(pool(models_amelia))
-
-#Pulling into DF
-pooled_df <- data.frame(summary(pool(models_amelia)))
-
-#Displaying in table
-xtable(pooled_df)
 
@@
 Compare how models from mice, Amelia, and the above methods perform to our true data:
-```{r, eval = FALSE}
+```{r}
 #true model
@@
 summary(pool(models_amelia))
+
 ```
@@
 Then, using Amelia, impute missing values.
-Note that the dataset is cross-sectional by country, time series by year, and should have the `gdp_pc` variable specified as logged.
+Not that the dataset is cross-sectional by country, time series by year, and should have the `gdp_pc` variable specified as logged.
 
@@
 
-```{r, eval = FALSE}
+```{r}
 library(Amelia)
@@
 
-#explore missingness
 apply(africa, 2, function(x) sum(is.na(x))) #missingness by column
@@
 #use Amelia to generate 5 imputed datasets
-#use ts and cs so imputation algerithem knows which variables represent countries and years
+#use ts and cs so imputation algs knows countries and years
 
-africa_imp <- amelia(x = __________, 
-                     cs = __________ 
-                     ts = __________
-                     logs = "gdp_pc")
+africa_imp <- amelia(x = africa, cs = "country", ts = "year", logs = "gdp_pc")
 
-models_africa <- vector(__)
+models_africa <- list() 
 for(i in 1:length(africa_imp$imputations)){
-  __________________________________) 
+  models_africa[[i]] <- lm(gdp_pc ~ infl + trade + civlib + population,
+                        data=africa_imp$imputations[[i]]) #why do we index like this?
 }
 
-#report results, including standard errors
+summary(pool(models_africa))
 
+#alternatively!
+all_coefs <- sapply(models_africa, coef) #get coefs
+all_ses <- sapply(models_africa, function(x) sqrt(diag(vcov(x)))) #get ses
+
+mi.meld(q = all_coefs, se = all_ses, byrow = F) 
 ```

### 2023-01-04 Lab 12  Declare Design/Lab12-declaredesign.Rmd
Add files via upload

---
title: "Lab 12: Declare Design"
author: "Eric Parajon with code adapted from Colin Case"
output: pdf_document
header-include:
  - \usepackage{amsbsy}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, message = FALSE, warning = FALSE)
```

## Pre-Registration, and Registred Reports
In recent years, there has been a shift within political science, particularly in political behavior, towards the use of pre-registration. Much of this motivation can be stemmed from the replication crisis and publication bias that has existed within the social sciences more broadly. That is to say, statistically significant results and hypothesis testing has driven hypothesis and theory generation. By moving towards pre-registration, where researchers develop hypotheses and an analysis plan before analyzing or collecting data, there can be an increased [distinctiveness between hypothesis generation and hypothesis testing](doi/10.1073/pnas.1708274114) (as well as an increased on the former *a priori*) that helps to improve research credibility.

A number of political science journals have embraced pre-registration and [registered reports](https://www.cambridge.org/core/journals/journal-of-experimental-political-science/information/faqs-for-registered-reports). The Journal of Experimental Political Science accepts the registered report format where you receive conditional acceptance before collecting data. The [Journal of Politics](https://www.journals.uchicago.edu/journals/jop/pre-registration) recently made the requirement for all experimental work to be pre-registered. JOP also [recently announced](https://mobile.twitter.com/The_JOP/status/1544712044594724864) it would be accepting registered reports starting in January 2023. Given this, it is likely some other journals will follow suit and this will become a more common practice across the discipline.

As a result of this shift, it is important to be familiar with tools that can help you develop strong pre-analysis plans, especially for registered reports (it is also just a good idea regardless of this shift)! In particular, simulating results, although not required, is a good practice for helping to diagnose potential issues with the research design. It also makes it easy to see what an eventual paper will look like with the results and eliminates some the arbitrariness that can be associated with unclear pre-analysis plans. Plus, you can then have all your code written *before* collecting the data and just click run once you have the data! Today, we are going to look at one of the more common tools, `DeclareDesign`, for describing research designs and simulating them without coding simulations from scratch. For our lab today, we are going to be producing simulated results for two separate analyses: one just simulating a model and the other simulating random assignment that is often associated with pre-analysis plans.

## Declare Design Software

Let's start first by loading the `DeclareDesign` package. If you have not used this before, make sure to install the package. Within `DeclareDesign`, there are three subsequent packages that are related to different steps of the research process: `fabricatr` can be used to simulate data, `randomizr` can be used for random sampling and random assignment, and `estimatr` for design-based estimators. For our lab, we are going to work on two concurrent model simulations. In one, we will be simulating a model with observational data and in the second we will use random assignment to estimate population average treatment effects.

```{r}
# Clear Environment
rm(list = ls())
# Load Packages
pacman::p_load(DeclareDesign,tidyverse,MASS)
# Set Seed for Reproducibility 
set.seed(100)
```

## Simulating Data

We are first going to work on simulating data to use in our observational analysis. To do this, we will use `fabricatr`. Note, `fabricatr` also lets you work with pre-existing data. Depending on the data structure might be more conducive (i.e. in Qualtrics, given a survey design, you can simulate survey results and use this pre-existing data versus hand coding everything).

`fabricatr` allows you to create common variable types, including assignment to treatment, count data, ordinal data (including “Likert scale” data, popular in surveys and survey experiments), categorical data (popular for modeling demographic characteristics). It also supports the creation of data with fixed intra-cluster correlations, so individual observations can be modeled as being part of groups or regions.

Let's first start by creating a *sample*, df1, with 500 observations. We are going to focus on the following variable characteristics to try and generate data that might look like our actual data from the real world:

$X_1$: Continuous value with mean of 0 and standard deviation of 1.\\

$X_2$: Continuous value with mean of 4 and standard deviation of 2. $X_1$ and $X_2$ should have a covariance of 0.6.\\

For our outcome variable, $Y$, we are going to assume a relationship between our independent variables of  $.5*X_1 + 3*X_2$. Because we don't want a perfect relationship for our simulation, allow $\beta_1$ (the effect of $X_1$ on $Y$) to follow a normal distribution with a mean of  and a standard deviation of 10 and $\beta_2$ (the effect of $X_2$ on $Y$) to follow a normal distribution with a mean of 3 and a standard deviation of 3. You can think about these as the true effect and standard deviation for the population.

Check the characteristics of your data after you simulate it. Do they roughly match the parameters we set above here?

```{r}

df1 <- ________(
  N = ___, # Specify sample size. All subsequent variables need to have this 
           # number of observations for fabricate to work.
  draw_multivariate(c(___, ____) ~ _______( # Create variable names and simulate
    _______, # Specify sample size
    _________, # Specify means
    ______ = matrix(c(_,__,__,__), 2, 2))), # Specify covariance matrix.
                                              # Think carefully about the 
                                              # structure of the matrix 
                                              # and the values.
  Y = _______________
)

# Summary statistics of X_1
mean(df1$X_1)
sd(df1$X_1)
# Summary statistics of X_2
mean(df1$X_2)
sd(df1$X_2)
# Correlation between X_1 and X_2
cor(df1$X_1, df1$X_2)
```

`df1` will be our dataframe for the observational data. As mentioned above, we will look at a second design to estimate average treatment effects. However, we will create that data in the next section as we go through the steps to see a different way of using `DeclareDesign` following the potential outcomes framework.

## Building Steps

Now that we have our data for the first model, we can start to build our research design. Research designs within `DeclareDesign` are constructed by building "steps." Almost all steps take a dataset as input and return a dataset as output. Each step of the research design process can be written as `declare_*` functions for the following steps: model, inquiry, sampling, assignment, measurement, and estimator. We are going to work through these steps and then simulate a design. We are going to work through two concurrent processes: one where we have pre-existing data that we simulated above and are specifying an observational model, and another where we are interested in the PATE of a treatment effect. It should be noted, you do not have to use all steps of the process. The use of certain steps and not others will depend on your research design.

### Model

The model defines the structure of the world, both its size and background characteristics as well as how interventions in the world determine outcomes. The model defines the number of units in the population, any multilevel structure to the data, and its background characteristics. We can define the population in several ways. In some cases, you may start a design with data on the population. When that happens, we do not need to simulate it at this step. We can simply declare the data as our population.

Because we already built simulated data above, we are going to use df1 for the observational model, m1. Our model can be specified as:

$$ Y = \beta_1X_1 + \beta_2X_2 + U$$

To specify the model, all you need to do is use the data argument of `declare_model`.

```{r}
# Declare Model 1, specify the data argument
m1 <- ________________
```

For our second design, we will simulate data where we are interested in the PATE for a treatment assignment, $Z$, on an outcome, $Y$. In this design, we will define a potential outcomes framework where the PATE is 0.1. Allow this PATE to have a normal distribution and a standard deviation of 0.2. It is important to note, as we are declaring this model to create the simulated data, we are creating a *population* of 1,000,000 people, *not* a sample. This is important for later on in the process.

```{r}
# Declare Model 2
m2 <- 
  declare_model(
     N = 1000000, # Specify population size
    # Define potential_outcomes and specify formula with uncertainly around PATE
    potential_outcomes(Y ~ Z*________________) 
  )

```

You will see this at the end of the lab, but this type of declare model setup will create two rows in the dataframe, Y_Z_1 (treatment outcome) and Y_Z_0 (control outcome), for each row in your dataframe. You will need call these variables later in the lab.

### Inquiry

Now that we have declared the model we plan on using, we are going to specify our inquiry that we are interested in. Our use of this step depends on the structure of our data for the previous step and our research design. For our first design, because we are using observational data without a random treatment, we do not have to specify the quantity of inquiry. For our second design, we are interested in randomly assigning a treatment, $Z$, and we specified the potential outcomes framework. To do this, specify the PATE as the average difference between the treatment and control groups.

```{r}
# Declare Inquiry for PATE (Think -- what do we want to calculate)
inquiry2 <- declare_inquiry(__________________)
```

### Data Strategy

Now it is time to set up our data strategy. This part is made up of a few steps. Again, we will be working primarily with our second design because our first design (1) is already a sample (2) lacks random assignment and (3) already has a specified outcome for the model whereas the second design has both Y_Z_1 and Y_Z_0 for each observation.

As noted above, m2 is generating a dataframe at the *population* level. In many cases, we don't have the capability to study the whole population, so we need to specify a sampling procedure. To provide substantive understanding for why this might be the case, recall the example used in class discussing random assignment of GOTV mailers. We may be interested in specifying our sampling procedure if we have data at the population level. For the purposes of this lab, we will create a simple sample of 100 people from our population chosen completely at random. There are a number of different sampling techniques you can use here if your design calls for it (i.e. stratified sampling).

```{r}
# Use declare_sample and specify S as completely random (complete_rs)
sample2 <- _______________________
```

Next, we need to randomly assign our treatment. To do this, we will use declare_assignment and specify Z to be completely random. As with our sampling procedure above, you can also specify specific arguments for assignment depending on your research design. This is helpful if you are using blocked randomization. Here, we will just randomly assign treatment and control groups with probability 0.5

```{r}
# Define random assignment and specify Z as completely random with prob = 0.5
assignment2 <- __________________________
```

Finally, for our second design, we need to specify which outcome, Y_Z_1 or Y_Z_0 we want included in our SATE. While in our simulated data we are able to observe both outcomes for each individual, no design has this capability (recall the fundamental problem of causal inference). As a result, we need to use the treatment assignment, $Z$, to specify which of the outcomes we actually observe in the estimate for $Y$, our observed outcome.

```{r}
# Declare measurement, specify Y to reveal outcomes of assigned treatment
measurement2 <- ________________________
```

### Answer Strategy

Now that we have our model, quantity of inquiry, and design strategy, we need to specify the estimator we are interested in. For our observational model, we are going to declare an estimator that looks at the effect of $X_1$ and $X_2$ on $Y$ using an OLS regression.

```{r}
# Declare estimator
estimator1 <- __________________(_______, # Formula
                                term = ________, # Use term to identify which
                                                       # Variables you want to see
                                ___________) # Specify .method (what model?)
```

For our experimental study, we are interested in comparing the difference in mean values from the control group ($Z = 0$) and the treatment group ($Z = 1$). Same as above, declare the estimator and specify the variables you are interested in. Because we are not using a model this time, we need to specify our method as difference in means. We also need to specify our quantity of inquiry from above (PATE) so the model knows which variables in the dataframe to use.

```{r}
# Declare estimator
estimator2 <- declare_estimator(_______, # Specify formula
                                .method = ______________, # Specify measure
                                inquiry = _______) # Identify QOI
```

## Build Design From Steps

From all the steps above, declare design creates values that need to be combined into a single element. To do this, you simply take the steps from above and add them to create the full design with + between each step. Be advised, the order of the steps matters! 
```{r}
# Create Design1 object by adding m1 and estimator1
design1 <- ___________________

# Create Design2 object by adding all steps for design2 above
design2 <- ___________________

```

## Simulate Research Design

Now that we have our design elements with all steps included, it is time to see the simulated results. There are a few different components you can pull out of this. The first is the simulated dataframe. Call draw_data on the design objects for the first ten rows. Does the data look like you expect? Take note, if we had not specified measurement2 in the prior step, we would not have a single observable $Y$.

```{r}
# Use draw_data to view first ten rows of data for design1
___________________

# Use draw_data to view first ten rows of data for design2
___________________
```

We can also see the results of our analysis given the simulated data. To see this, call `draw_estimates` on each design object. How closely do the quantities of interest match the relationships we specified in the above parts?

```{r}
# View Estimates for design1. Do they match expectations?
___________________

# View PATE for design2. Do they match expectations?
___________________
```

One of the great advantages of declare design is your ability to play with different parameters. For example, how does sample size change the outcomes you observe? The same goes with anticipated effect size. Especially in the experimental where you are randomly assigning a treatment, this can help with research design decisions.

### 2023-01-04 Lab 12  Declare Design/Lab12_declaredesign_Answers.Rmd
Add files via upload

--- 2023-01-04 Lab 12  Declare Design/Lab12-declaredesign.Rmd
@@
 ---
-title: "Lab 12: Declare Design"
+title: "Lab 12: Declare Design: Answers"
 author: "Eric Parajon with code adapted from Colin Case"
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, eval = FALSE, message = FALSE, warning = FALSE)
+knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
 ```
@@
 
-df1 <- ________(
-  N = ___, # Specify sample size. All subsequent variables need to have this 
+df1 <- fabricate(
+  N = 500, # Specify sample size. All subsequent variables need to have this 
            # number of observations for fabricate to work.
-  draw_multivariate(c(___, ____) ~ _______( # Create variable names and simulate
-    _______, # Specify sample size
-    _________, # Specify means
-    ______ = matrix(c(_,__,__,__), 2, 2))), # Specify covariance matrix.
+  draw_multivariate(c(X_1, X_2) ~ mvrnorm( # Create variable names and simulate
+    n = N, # Specify sample size
+    mu = c(0, 4), # Specify means
+    Sigma = matrix(c(1, 1.2, 1.2, 4), 2, 2))), # Specify covariance matrix.
                                               # Think carefully about the 
@@
                                               # and the values.
-  Y = _______________
+  Y = rnorm(500, mean = .5, sd = 10)*X_1 + rnorm(500, mean = 3, sd = 3)*X_2
 )
@@
 # Declare Model 1, specify the data argument
-m1 <- ________________
+m1 <- declare_model(data = df1)
 ```
@@
     # Define potential_outcomes and specify formula with uncertainly around PATE
-    potential_outcomes(Y ~ Z*________________) 
+    potential_outcomes(Y ~ Z*rnorm(N, mean  = (0.1), sd = .2)) 
   )
@@
 # Declare Inquiry for PATE (Think -- what do we want to calculate)
-inquiry2 <- declare_inquiry(__________________)
+inquiry2 <- declare_inquiry(PATE = mean(Y_Z_1 - Y_Z_0))
 ```
@@
 # Use declare_sample and specify S as completely random (complete_rs)
-sample2 <- _______________________
+sample2 <- declare_sampling(S = complete_rs(N, n = 100))
 ```
@@
 # Define random assignment and specify Z as completely random with prob = 0.5
-assignment2 <- __________________________
+assignment2 <- declare_assignment(Z = complete_ra(100, prob = 0.5))
 ```
@@
 # Declare measurement, specify Y to reveal outcomes of assigned treatment
-measurement2 <- ________________________
+measurement2 <- declare_measurement(Y = reveal_outcomes(Y ~ Z))
 ```
@@
 # Declare estimator
-estimator1 <- __________________(_______, # Formula
-                                term = ________, # Use term to identify which
+estimator1 <- declare_estimator(Y ~ X_1+X_2, # Formula
+                                term = c('X_1','X_2'), # Use term to identify which
                                                        # Variables you want to see
-                                ___________) # Specify .method (what model?)
+                                .method = lm) # Specify .method (what model?)
 ```
@@
 # Declare estimator
-estimator2 <- declare_estimator(_______, # Specify formula
-                                .method = ______________, # Specify measure
-                                inquiry = _______) # Identify QOI
+estimator2 <- declare_estimator(Y ~ Z, # Specify formula
+                                .method = difference_in_means, # Specify measure
+                                inquiry = "PATE") # Identify QOI
 ```
@@
 # Create Design1 object by adding m1 and estimator1
-design1 <- ___________________
+design1 <- m1 + estimator1
 
 # Create Design2 object by adding all steps for design2 above
-design2 <- ___________________
+design2 <- m2 + inquiry2 + sample2 + assignment2 + measurement2 + estimator2
 
@@
 # Use draw_data to view first ten rows of data for design1
-___________________
+draw_data(design1)[1:10,]
 
 # Use draw_data to view first ten rows of data for design2
-___________________
+draw_data(design2)[1:10,]
 ```
@@
 # View Estimates for design1. Do they match expectations?
-___________________
+draw_estimates(design1)
 
 # View PATE for design2. Do they match expectations?
-___________________
+draw_estimates(design2)
+
 ```

### 2023-01-04 Lab 13  CBPS/Lab13_covariatebalance.Rmd
Add files via upload

---
title: "Lab 13: Covariate Balancing and CBPS"
author: "Eric Parajon with code adapted from Colin Case"
output: html_document
header-include:
  - \usepackage{amsbsy}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \DeclareMathOperator*{\argmin}{argmin} % thin space, limits underneath in displays
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, message = FALSE, warning = FALSE)
```

# Weighting

Randomized controlled trials (RCTs) are often seen as the ``gold standard'' for estimating the effects of treatments, interventions, and exposures on outcomes. However, in the social sciences, it is not always possible to achieve randomization (and even in experimental studies, this doesn't happen either!), and we may have unobserved confounding variables that affect someone's likelihood of receiving treatment. Weighting helps adjust for the covariate imbalance between the treated and the control groups in a study. The weights on observations can help us "balance-out" the influence of each observation on the causal relationship. In this manner, we are making it so the treatment and control groups are *balanced* on average when considering observable characteristics (we obviously cannot balance on unobservable characteristics).

In this exercise we will use the `lalonde` dataset from the `cobalt` package, which is data on the [National Supported Work Demonstration](https://www.mdrc.org/publication/summary-and-findings-national-supported-work-demonstration), an experimental public employment assistance program in the US in the 1970s. The question we are trying to answer here is: does the participation in the national supported work demonstration (`treat`) help workers gain a higher wage? In thinking about this question, consider who is likely to opt-into this program if it is optional and not randomly assigned?

The outcome variable of interest is workers' earnings in 1978 `re78`. We are going to control for a couple of things such as pre-treatment wages `re74`, `re75`, if the person earned a degree `nodegree`, marital status `married`, as well as other demographic variables.

## Examining Covariate Balance

To start, we are going to look at covariate balance between the treated and the control groups. We are going to use the `bal.tab` function from the `cobalt` package, and we specify that we are interested in the average treatment effect on the treated (ATT):

```{r}
# Clear Environment
rm(list = ls())
# Load Packages

pacman::p_load(WeightIt,
               cobalt,
               tidyverse,
               stargazer,
               survey,
               jtools,
               CBPS,
               lmtest,
               sandwich)
# Load dataset Lalonde from cobalt package
data('lalonde', package = 'cobalt')

# Evaluate covariate balance on age, educ, married, nodegree, re74, and re75
________(treat ~ age + educ + married + nodegree + re74 + re75,
        data = lalonde, estimand = "ATT", m.threshold = .05)

```

As you can see from the output, none of the variables are balanced. Those who are older, married, and had higher income in 1974 and 1975 are much less likely to attend training (i.e. receive the treatment). Those without a high school degree and those who are more educated are more likely to attend job training. This raises questions regarding the ATT of the treatment in 1978 : Is it a treatment effect, or a function of the pre-existing differences that have not been taken into account in the treatment assignment process.

## Propensity Scores

We will now try to address this by computing propensity score weights. To recall, propensity scores are probabilities of a unit (e.g. a subject) being assigned to a particular treatment given a set of factors/ covariates:

$$e(x) = Pr(T=1|X=x)$$

for $T$ is the treatment assignment dummy and $X$ is a matrix of covariates.

Propensity Score weighting uses this quantity to adjust for the influence of a particular observation in the treated and control groups. It turns out the weights for the ATT are defined as $w = 1$ for the treated group and $w = \frac{\hat{e(x)}}{1- \hat{e(x)}}$ for the control group.

Let's try to compute this by hand. We will first fit a simple GLM regressing treatment on the covariates:

```{r}
# Estimate glm predicting treatment status
ps.mod <- glm(__________________)
# Create predictions from above model
ps <- ______(______, type = "response")
# View propensity scores
head(ps)
```

These probabilities are our propensity scores $\hat{e(x)}$. Now, let's compute the weights:

```{r}
# Create variable name weights in lalonde df
lalonde$weights <- ifelse(lalonde$treat ==1, 1, # w=1 for treatment
                  ps/(1-ps)) # propensity score/(1-propensity score)
```

We should always check the distribution of our weights. If there are extreme values in the weights computed, we might need to use another weighting method, or trim some of the cases:

```{r}
# Subset to treatment and weights
weights_dist <- lalonde %>% dplyr::select(treat, weights)
# Add labls
weights_dist$treat <- ifelse(weights_dist$treat == 0, "Control", "Treatment")

# Plot weight density by condition looking for outliers  
ggplot(weights_dist, aes(x=weights))+
  geom_density()+
  facet_wrap(~treat)+
  theme_bw()

```

As we can see, the control weights are OK - we don't have extreme values. The computed weights range between 0-2. For the treatment group, all weights are equal to 1, so we don't have to worry about it in this case. If the causal estimand of interest is not ATT, then weights are computed differently for the treatment group and you may see some variations in the calculated weights for the treated.

And these are then the weights for the ATE for each observation. We can now refit the model with these weights using the `weights=` parameter in `lm()`. Since we didn't weight by race, we also want to make sure we control for that factor in this model. Let's compare the results between the model without weighting and with weighting:

```{r}
# Estimate and View model without weights
lalonde_model_noW <- lm(____ ~ _________, data=lalonde)
summary(lalonde_model_noW)
# Estimate and view model with weights
lalonde_model_w <- lm(____ ~ _________, data = lalonde, weights = weights)
summary(lalonde_model_w)
```

As you can see, after weighting our observations, the treatment is now significant whereas without weights it would be not significant!

## The `WeightIt` package

Now that you understand how weighting works under the hood, we will turn to the `WeightIt` package that simplifies this process. `WeightIt` is a very powerful package for computing different kinds of weights and examining the weighting result. Let's look at how you can compute weights with `WeightIt`.

You compute the weights with the `weightit` function, which uses the same set of parameters from the `bal.tab` function:

```{r}
# Use weightit to calculate weights
W.out <- ________(______ ~ ____________________________,
        data = lalonde, estimand = "ATT", method = "ps")
W.out
```

This creates a `Weightit` object, which gives you a summary of the weight computation outcome. You can get more information out of it by calling `summary()` on it:

```{r}
# View w.out
summary(W.out)
```

Now let's reexamine covariate balance on the weightit object:

```{r}
# Re-examine covariate balance after weighting
_________(W.out, m.threshold = .05, disp.v.ratio = TRUE)
```

Now you can see that every covariate is balanced between the treated and the control group. Now, let's try to compare the hand-computed weights and the weights generated by `WeightIt` to validate our calculations:

```{r}
# Merge hand and weightit weights
wdf <- data.frame(hand = lalonde$weights,
                  weightit = get.w(W.out))
# Calculate correlation
cor(____)

# Plot Weights
ggplot(wdf, aes(x=___, y=_____))+
  geom_point()+
  geom_abline(slope=1, lty=2)+
  xlab("Weights computed by hand")+
  ylab("Weights computed by WeightIt")+
  theme_bw()
```

And the correlation matrix confirms our calculations - they are exactly the same as those generated by `WeightIt` and they fall perfectly along the 45 degrees diagonal line (perfect correlation).

There is an alternative method, called Covariate Balancing Propensity Score Weighting (`CBPS`), that performs a similar procedure than the `ps` method we used above. The difference comes in that `CPBS` estimates propensity scores such that both covariate balance and prediction of treatment assignment are maximized. The method, therefore, avoids an iterative process between model fitting and balance checking and implements both simultaneously. Let's see how these weights compare to the `ps` process we used in `WeightIt`.

```{r}
# Create CPBS Weights
CPBS.out <-  ______(_____ ~ ____________________, 
			data = lalonde, ATT = TRUE)

# Merge hand and weightit weights
wdf <- data.frame(cbps = CPBS.out$weights,
                  weightit = get.w(W.out))
# Calculate correlation
cor(wdf)

# Plot Weights
ggplot(____, aes(x=_____, y=______))+
  geom_point()+
  geom_abline(slope=1, lty=2)+
  xlab("Weights computed by CBPS")+
  ylab("Weights computed by WeightIt")+
  theme_bw()

```
As you can see, the two are still highly correlated, but with some slight differences this time. Let's now see what this looks like when we include the weights in the regression output. We first need add the weights from `W.out` to the data frame, and then use `lm` to estimate the model.

```{r}
# Model using WeightIt Weights and controlling for race
fit1 <- lm(____ ~ ______, data = lalonde, weights = get.w(W.out))
summary(fit1)

# Model using CBPS Weights and controlling for race
fit2 <- lm(____ ~ ______, data = lalonde, weights = CPBS.out$weights)
summary(fit2)

```

## Survey Weighted Regression

How can we make inference out of these weights? We have explored `lm()` and `glm()` with the weights option, which finds:

$$\argmin_\beta \ w_i \sum^N_{i=1}(y_i- X_i\beta)^2$$
where $w_i$ refers to the propensity score weights for individual $i$.

It turns out using `lm()` is not giving us an accurate estimation of the ATT because it only applies the weights on the errors in the least squares estimation (aka Weighted Least Squares - WLS). However, we should also consider the loss of precision due to the presence of sampling weights:

$$\argmin_\beta \ s_iw_i \sum^N_{i=1}(y_i- X_i\beta)^2$$

where $s_i$ is the sampling weight in a survey setting. This type of survey-weighted linear regression model can be fitted with the `survey` package:

```{r}
# Specify svydesign with no clusters
d.w <- svydesign(ids = ~1, weights = get.w(W.out),
                     data = lalonde)
# Fit model taking into account survey weights
fit <- svyglm(re78 ~ treat, design = d.w)
# View Fit 
summ(fit, confint = TRUE) 
```

And it turns out after adjusting for survey-weights, we found the effect of the treatment is actually not statistically significant after considering for survey sampling weights on top of propensity score weights.

If you have survey weights already, one way you could incorporate them in `WeightIt` is to specify the `s.weights` argument in `weightit()`. Then, the ATT weights are the product of the propensity score weights and the survey weights supplied. Using these ATT weights, you can then fit the WLS as usual but the standard errors will have to be adjusted by using the sandwich estimator (`sandwich::vcovHC`).

# Exercise: Priming Norwegian Voters about the Municipal Reform, 2014

In Norway, there is a lot of small municipalities (similar to U.S. Counties) scattered across the country. There has been a debate on merging these small municipalities for roughly 2 decades. Proponents for the reform argue that this helps achieve the economies of scale of providing services to the localities, whereas opponents argue that each locality is unique. Given that the center-periphery cleavage is quite salient in Norway and overlaps with the linguistic conflict between Nynorsk and Bokmål speakers, municipal reform was a highly salient political issue in Norway at the time.

In 2014, the Norwegian Citizens Panel (NCP) (Wave 3) conducted a randomized experiment on the respondents by coincidence. There was a question about whether the participant would pay 1000 Norwegian Krone (~$115) to maintain the current municipal structure. This primes the respondent that the current municipal structure can be costly and merging them could save the taxpayers' money. Note that the assignment of this group of questions about municipal reforms is random, so even if the priming is not intentional, it remains a valid experimental treatment nonetheless.

In this exercise, you are going to explore **if priming respondents the cost of the existing municipal structure would lead to support for the Conservative Party**. The Conservative Party of Norway supports municipal reforms of cutting down the number of municipalities, along with the Center Party and Progress Party. The Conservative Party would go on to form a center-right cabinet after emerging second in the 2017 parliamentary election. They will go on to implement the reform after forming a center-right government. The reforms were eventually completed on 1/1/2020.

Since the treatment assignment is unintentional, it is unclear if the covariates are balanced for the treated and control group. Therefore, it is necessary to check covariate balance and generate weights to balance out the influence of particular observations given their covariate values.

In the dataset `ncp_final.rds`, I have attached the survey weights from the NCP for you. The variables are as follows:

- `conservative` : Self-reported affinity to the Conservative Party of Norway. Continuous. 1 = Intensely dislike; 7= Intensely like.
- `treat` : If the respondent has been asked the question : "Would you rather pay NOK 1000 per year to maintain the current municipal structure?". Dummy.
- `interest` : Self-reported Political Interest. Continuous. 1= Very Interested; 5 = Not Interested.
- `gender` : Biological Sex. Dummy. 1 = Male, 2= Female.
- `region` : Region of the respondent. 1 = Oslo. 2= Østlandet. 3 = Sørlandet. 4= Vestlandet. 5=Trøndelag. 6 = Nord-Norge.
- `edu` : Highest level of education attained. 1 = No Education/ Elementary School. 2= Upper Secondary School. 3= Univesity/ Community College.
- `yob` : Year of Birth. 1 = 1939 or earlier. 2= 1940-1949. 3= 1950-1959. ... 7= 1990 or later.
- `weight` : Survey weights from NCP. Continuous.

Take note that this dataset already has survey weights which are different from the weights we will calculate.

Using bal.tab, look at the balance of demographic variables (`gender`, `region`, `edu`, `yob`) on the treatment before and after weighting. Focus on the ATT. How do they look?

```{r}
# Load data
ncp_final <- readRDS("ncp_final.rds")

# Look at balance of demographic variables
bal.tab(______ ~ _____ + _____ + ____ + _____, 
        data = ncp_final, estimand = ______, m.threshold = .05)
```

The balance table suggests that education and year of birth are unbalanced between the treated and the control group. Use weightit to weight them (you should do this even on variables that are listed as balanced).

```{r}
# Create weights
W.out <- weightit(_____ ~ ______ + _____ + ____ + ______,
                  data = _______, estimand = ______, method = "ps")
# Re-check balance with weights
bal.tab(W.out, m.threshold = .05, disp.v.ratio = TRUE)
```

Regress the outcome variable on the treatment, `interest`, and the demographic variables. Fit two models: 1) One with no adjustment; 2) a Survey-Weighted GLM with propensity score weights. Then, use `summ` to take a look at the 95% Confidence Interval on the effect of the treatment from the survey-weighted GLM.

```{r}
# without any adjustment
# without any adjustment
tmod1 <- lm(conservative ~ treat + interest + as.factor(gender) + 
              as.factor(region)+ as.factor(edu) + as.factor(yob) , ncp_final)

# With Survey-weights & PS weights adjustment

# Create survey weights using svydesign
design <- ________________________

# Estimate model
tmod2 <- svyglm(____________________)
# 95% CI
summ(tmod2, confint = TRUE) 
```

Compute propensity score weights, INCLUDING survey weights provided (`weights`). Then, fit a WLS with weights computed and correct their standard errors with the sandwich estimator.

To include our survey weights, we specify them with the `s.weights` parameter in `weightit()`. Then, we multiply the resulting weights with the survey weights. When fitting the WLS, we use this product of weights, then we will correct the standard errors with the sandwich estimator.

```{r}
library(lmtest)
library(sandwich)
# include survey weight variants:

ncp_clean <- ncp_final %>% dplyr::select(treat, interest, 
                                         gender, region, edu, yob, 
                                         conservative, weights)

# remove NAs, otherwise there will be NAs in the weights
ncp_clean <- na.omit(ncp_clean)

# Calculate weights with survey weights
W.out <- weightit(_______________________, s.weights="weights")
# Check balance
bal.tab(W.out, m.threshold = .05, disp.v.ratio = TRUE, un=TRUE)
```

Again, the covariates are all balanced now.

```{r}

# Multiply the survey weights with the balance weights to create att.weights
att.weights <- W.out$weights * ncp_clean$weights #product of weights

# Estimate the model using lm and specifying weights
tmod3 <- lm(____________________, weights = att.weights)

#correct SE with the sandwich estimator:
corr_se <- sandwich::vcovHC(tmod3) %>% diag %>% sqrt
```

Report all models in the same table using stargazer (make sure to specify html document). What is your conclusion?

```{r results="asis"}
#Putting it all together:

stargazer(_____, _____, _____, type="html",
          covariate.labels = c("Primed",
                               "Political Interest",
                               "Female",
                               "Østlandet",
                               "Sørlandet",
                               "Vestlandet",
                               "Trøndelag",
                               "Nord-Norge",
                               "Completed Upper Secondary Education",
                               "Completed University/ College Education",
                               "Born between 1940-1949",
                               "Born between 1950-1959",
                               "Born between 1960-1969",
                               "Born between 1970-1979",
                               "Born between 1980-1989",
                               "Born in or after 1990",
                               "Intercept"),
          se=list(NULL, NULL, corr_se),
          column.labels = c("No Weights OLS", "PS Weights SWLS", "PS/NCP Weights SWLS"),
          star.cutoffs = c(.10,.05,.01,.001),
          star.char = c("+","*","**","***"),
          column.sep.width = "10pt",
          notes = c("p<.10+, p<.05*, p<.01**, p<.001***"),
          single.row = TRUE,
          notes.append=FALSE)
```

From the three models reported, we see that there is a small positive effect of being primed the cost of the existing municipal structure. Across three models, the effect is significant, p<.10+. These primed respondents are likely to increase their rating of the conservative party .1 point more on a 1-7 scale. The effect size is modest but it is statistically significant in all three models, regardless of the weighting method used.

### 2023-01-04 Lab 13  CBPS/Lab13_covariatebalance_Answers.Rmd
Add files via upload

--- 2023-01-04 Lab 13  CBPS/Lab13_covariatebalance.Rmd
@@
 ---
-title: "Lab 13: Covariate Balancing and CBPS"
+title: 'Lab 13: Covariate Balancing and CBPS: Answers'
 author: "Eric Parajon with code adapted from Colin Case"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 header-include:
@@
                sandwich)
+
 # Load dataset Lalonde from cobalt package
@@
 # Evaluate covariate balance on age, educ, married, nodegree, re74, and re75
-________(treat ~ age + educ + married + nodegree + re74 + re75,
+bal.tab(treat ~ age + educ + married + nodegree + re74 + re75,
         data = lalonde, estimand = "ATT", m.threshold = .05)
@@
 # Estimate glm predicting treatment status
-ps.mod <- glm(__________________)
-# Create predictions from above model
-ps <- ______(______, type = "response")
+ps.mod <- glm(treat~ age + educ + married + nodegree + re74 + re75, 
+              data = lalonde, family = "binomial")
+# Create predictions
+ps <- predict(ps.mod, type = "response")
 # View propensity scores
@@
 # Estimate and View model without weights
-lalonde_model_noW <- lm(____ ~ _________, data=lalonde)
+lalonde_model_noW <- lm(re78 ~ treat + race, data=lalonde)
 summary(lalonde_model_noW)
 # Estimate and view model with weights
-lalonde_model_w <- lm(____ ~ _________, data = lalonde, weights = weights)
+lalonde_model_w <- lm(re78 ~ treat + race, data = lalonde, weights = weights)
 summary(lalonde_model_w)
@@
 # Use weightit to calculate weights
-W.out <- ________(______ ~ ____________________________,
+W.out <- weightit(treat ~ age + educ + married + nodegree + re74 + re75,
         data = lalonde, estimand = "ATT", method = "ps")
@@
 # Re-examine covariate balance after weighting
-_________(W.out, m.threshold = .05, disp.v.ratio = TRUE)
+bal.tab(W.out, m.threshold = .05, disp.v.ratio = TRUE)
 ```
@@
 # Calculate correlation
-cor(____)
+cor(wdf)
 
 # Plot Weights
-ggplot(wdf, aes(x=___, y=_____))+
+ggplot(wdf, aes(x=hand, y=weightit))+
   geom_point()+
@@
 ```{r}
-# Create CPBS Weights
-CPBS.out <-  ______(_____ ~ ____________________, 
+CPBS.out <-  CBPS(treat ~ age + educ + married + nodegree + re75 + re74, 
 			data = lalonde, ATT = TRUE)
@@
 # Plot Weights
-ggplot(____, aes(x=_____, y=______))+
+ggplot(wdf, aes(x=cbps, y=weightit))+
   geom_point()+
@@
 ```{r}
-# Model using WeightIt Weights and controlling for race
-fit1 <- lm(____ ~ ______, data = lalonde, weights = get.w(W.out))
+# Model using WeightIt Weights
+fit1 <- lm(re78 ~ treat + race, data = lalonde, weights = get.w(W.out))
 summary(fit1)
 
-# Model using CBPS Weights and controlling for race
-fit2 <- lm(____ ~ ______, data = lalonde, weights = CPBS.out$weights)
+# Model using CBPS Weights
+fit2 <- lm(re78 ~ treat + race, data = lalonde, weights = CPBS.out$weights)
 summary(fit2)
@@
 # Look at balance of demographic variables
-bal.tab(______ ~ _____ + _____ + ____ + _____, 
-        data = ncp_final, estimand = ______, m.threshold = .05)
+bal.tab(treat~ gender + region + edu + yob, 
+        data = ncp_final, estimand = "ATT", m.threshold = .05)
 ```
@@
 # Create weights
-W.out <- weightit(_____ ~ ______ + _____ + ____ + ______,
-                  data = _______, estimand = ______, method = "ps")
+W.out <- weightit(treat ~ gender + region + edu + yob,
+                  data = ncp_final, estimand = "ATT", method = "ps")
 # Re-check balance with weights
@@
 # without any adjustment
-# without any adjustment
 tmod1 <- lm(conservative ~ treat + interest + as.factor(gender) + 
@@
 # Create survey weights using svydesign
-design <- ________________________
+design <- svydesign(ids = ~1, weights = get.w(W.out),
+                 data = ncp_final)
 
 # Estimate model
-tmod2 <- svyglm(____________________)
+tmod2 <- svyglm(conservative ~ treat + interest + as.factor(gender) + 
+                as.factor(region)+ as.factor(edu) + as.factor(yob), 
+                design = design)
 # 95% CI
@@
 # Calculate weights with survey weights
-W.out <- weightit(_______________________, s.weights="weights")
+W.out <- weightit(treat ~ gender + region + edu + yob,
+                  data = ncp_clean, estimand = "ATT", method = "ps",
+                  s.weights="weights")
 # Check balance
@@
 # Estimate the model using lm and specifying weights
-tmod3 <- lm(____________________, weights = att.weights)
+tmod3 <- lm(conservative ~ treat + interest + as.factor(gender) + 
+              as.factor(region)+ as.factor(edu) + as.factor(yob) , 
+            data = ncp_clean, weights = att.weights)
 
@@
 
-Report all models in the same table using stargazer (make sure to specify html document). What is your conclusion?
+5. Report all models in the same table using stargazer (make sure to specify html document). What is your conclusion?
 
@@
 
-stargazer(_____, _____, _____, type="html",
+stargazer(tmod1, tmod2, tmod3, type="html",
           covariate.labels = c("Primed",

### 2023-01-04 Lab 2  Tidyverse Review/Lab2_Tidyverse_Review.Rmd
Add files via upload

---
title: '784 Lab 2: Review of Tidyverse and ggplot'
author: "Eric Parajon"
output:
  pdf_document:
    keep_tex: TRUE
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, include = FALSE)
```

# Tidyverse Review
\medskip
1. Again, it is important to save the document you are using in a central folder. Then, set your working directory to that folder. It's also important to clear your environment before working in a new document.
```{r, include = TRUE, eval = FALSE}
getwd() #what is your current working directory?

setwd(____) #in the parentheses, set the working directory to the folder where
#you save this file. Realistically, this isn't super relevant now because
#you're not reading in external data, but it's a good habit to get into!

rm(list=ls()) #clear global environment

#The pacman package is required to load/install the additional packages. Uncomment the line below to install the package.
#install.packages("pacman")

#Installing and loading packages for use
pacman::p_load(tidyverse,ggthemes,nycflights13)

```

2.The tidyverse allows us to use the piping operator `%>%`, which is a useful piece of code (base R can do this too with |>) .

It pipes an object (either by itself or created by another function) into the next function in the sequence, as the first argument. This makes the code much easier to read.

For example, we could calculate the standard deviation of the `mpg` variable in the `mtcars` dataset in the following ways:

```{r, include = TRUE, echo = TRUE}
library(tidyverse)

#normal
sd(mtcars$mpg)

#piping
mtcars$mpg %>% 
  sd()

#Does it return the same value?
```
That is a very simple example, of course, and using a pipe in this case does not really gain you much. However, pipes and the myriad of tools in the `tidyverse` can make our work much, much easier. The [`dplyr`](https://dplyr.tidyverse.org/) package in the `tidyverse` allows us to manipulate data much more easily than in base R.

`dplyr` allows us to manipulate our data in a structured way. The key functions are `mutate()`, which creates new variables, `select()`, which allows us to pick variable by characteristics of their names, `filter()`, which allows us to choose rows according to criteria we define, `summarise()`, which allows us to "reduce multiple values down to a single summary", and `arrange()`, which allows us to order the observations according values of 1 or more variables.

RStudio makes useful cheat sheets for their packages, including [`dplyr`](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf).

### Star Wars example

```{r}
#1. Star Wars example
data(starwars)
  #Let's make our data into a tibble

#Tibbles are evolution of data.frame
#Inherits nominally, but not strictly!
  #Does not match partial variable names
  #Does not convert character vectors to factors
  #Always returns a tibble when [ is used
  #Allows readable variable names

dplyr::as_tibble(starwars)

  #Now that we have our data imported, we should start exploring it to see exactly what we have

  #the glimpse() function is quite similar to base R's str function
glimpse(starwars)
 #What if we want to pull a specific piece of data?
snip<-slice(starwars, 10:11)

  #What if we want some summary stats on specific variables?
starwars %>%
  group_by(homeworld) %>%
    count() %>% 
arrange(desc(n))

starwars %>%
  group_by(hair_color) %>%
  count() %>% 
  arrange(desc(n))

```

```{r}
#We should start to do some column manipulation now that we have some variables we could be interested in 
d1<-select(starwars, name, species, mass,height,gender,films)

#So now that we have a smaller tibble let's try and do some analysis.
  #What species are the heaviest and which are the lightest?

d1 %>%
  group_by(species) %>%
  summarise(avg = mean(mass, na.rm=TRUE)) %>%
  arrange(desc(avg))
#Hutts are the heaviest!

#Okay great, but what if we want some information that's not explicitly contained in the data? 
  #What if for example, you wanted to know the BMI of a bunch of characters?
  #You can use mutate to create new variables.

Star_BMI<-d1 %>%
  mutate(bmi = mass/((height/100)^2)) %>%
  arrange(desc(bmi)) %>%
  select(name, species, height, mass, bmi)

#Now what if we only want to BMI of Humans characters?

Human_BMI<-Star_BMI %>% 
  filter(species=="Human")

#What if we wanted to create a new variable that combined a characters name with where they were born?
#This is easy using tidyr (another package in tidyverse suite)
starwars %>%
  unite(name_homeworld, name, homeworld) %>% 
  select(name_homeworld)

```
## Joins

```{r}

# Load data 1
(dates <- read_csv("https://bit.ly/DM22_dates"))

# Load data 2
(works <- read_csv("https://bit.ly/DM22_works"))

#Inner join
#With an inner_join(x, y), only observations that appear in both x and y datasets are returned

# Do inner join
dates_works <- inner_join(dates, works) 
(dates_works)

#Left & Right Join
#A left_join(x, y) call returns all observations in x, and drops observations that only appear in y
#Variables in y that do not exist in x are appended to x in the returned tibble.

#The same is true for a right_join(x, y), but with the roles of x and y reversed.

dates_someworks <- left_join(dates, works) 
(dates_someworks)

dates_allworks <- right_join(dates, works) 
print(dates_allworks, width=70)

#Full join
#Finally, a full_join(x, y) returns all observations and all variables in either in x or in y (filled with NAs)

alldates_allworks <- full_join(dates, works) 
print(alldates_allworks, width=70)

```

## Pivoting
Pivoting: Wide -> Long
With pivot_longer(), you can take values of a variable that are stored as variable names pivot them onto the cells of a single (implicit) variable column

```{r}
(relig_income)
#This data set has counts based on a survey which (among other things) asked people about their religion and annual income:

relig_income_long<-relig_income %>% 
  pivot_longer(!religion, names_to = "income", values_to = "count")

#The first argument is the dataset to reshape, relig_income.

#The second argument describes which columns need to be reshaped. In this case, it’s every column apart from religion.

#The names_to gives the name of the variable that will be created from the data stored in the column names, i.e. income.

#The values_to gives the name of the variable that will be created from the data stored in the cell value, i.e. count.

```
Pivoting: Long -> Wide

With pivot_wider(), we move in the opposite direction: your units are distributed over multiple rows, so pivot until you have a single case per row of your tibble

```{r}
(fish_encounters) 
#This dataset, contributed by Myfanwy Johnston, describes when fish swimming down a river are detected by automatic monitoring stations:

fish_encounters_wide<- fish_encounters %>% 
          pivot_wider(names_from = station, values_from = seen)

#In this case the NAs mean no fish was recorded so we can convert that to zero like this

fish_encounters %>% pivot_wider(
  names_from = station, 
  values_from = seen,
  values_fill = 0
)

```
# Part 2: ggplot

`ggplot` is very important, and requires some time and patience to get to know well. We won't have time to dive fully into it, but we will begin reviewing the following code so you become more familiar. R has its own built-in plot function, but ggplot is far more flexible and creates far more descriptive (and thus useful!) plots.

For a comprehensive 'cheat sheet', see: https://www.maths.usyd.edu.au/u/UG/SM/STAT3022/r/current/Misc/data-visualization-2.1.pdf

```{r}
#we have to load the package first
library(ggplot2)
```

To begin, we can use the 'qplot' or 'ggplot' commands. I use ggplot. It's more customization.

```{r}
data(mtcars)

qplot(x = disp, y = mpg, data = mtcars, color = cyl, geom = "point") 
#x and y are independent + dependent variables, 
#data can be specified (or use mtcars$var),
#color distinguishes a characteristic over which to map color
#and geom is a plot type

ggplot(data = mtcars, aes(x = disp, y = mpg)) + geom_point(aes(color = cyl))
#Note the similarities above. 
#ggplot is more flexible in that you map to different `aesthetics'- 
#hence why we use 'aes.' Plenty of examples below to dig into this more! 
```

What are some ways we can make this plot clearer?

```{r}
ggplot(data = mtcars, aes(x = disp, y = mpg)) + #we add 'layers' with +
  geom_point(aes(colour = cyl), size = 2) + #point=scatterplot
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon")+ 
  #label x and y axes, give the plot a main title. Note the quotes!
  scale_colour_continuous(name = "Cylinders")+ #label the scale
  theme_bw() #make the plot background blank w/grid lines
```

You can also assign this plot to an object and save that object as a .pdf file.

```{r, message=FALSE, warning=FALSE}
carsplot <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()

#One method for saving plots:
#pdf("Plot1.pdf")
#carsplot
#dev.off()

#Where do these plots go?!

#Unique to ggplot:
#ggsave("Plot1a.pdf") #can edit image sizes as well
```

## Themes
As you've hopefully seen ggplot is highly customization. One way to standardize your graphs and avoid repeating your code to control ancillary parameters (e.g. color palettes, axes labels, etc.) is to use a ggplot() theme that includes bundles of code. This is what we were doing using theme_bw() above.

You can also create custom themes yourself!

```{r}
#Here is a theme I use often

plot_theme = theme(text = element_text(size=11),  #set text size to 11
                   panel.grid.minor = element_blank(), #these next two lines set the grid lines to null
                   panel.grid.major = element_blank(),
                   panel.background = element_blank(),#these lines make the background of the plot blank
                   plot.background = element_rect(fill = "transparent",colour = NA),
                   axis.text = element_text(color="black"), #setting axis text to black 
                   axis.line = element_line(colour = "black"),#setting axis lines to black 
                   strip.text.y = element_text(angle = 0), axis.title.y = element_text(angle=90, hjust=.5, vjust=.5),#adjusting the angle of text
                   axis.title = element_text(lineheight=1.1), #increasing height of axis title
                   legend.position = "bottom",  #Placing the legend at the bottom of the figure
                   legend.key = element_rect(colour = "transparent",  fill = "transparent") #setting legend to transparent
)

#Using the theme
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  plot_theme #no () after the custom theme call
```
As you can see there are tons of options that you can set yourself. Alternatively you can use existing gplot_themes. Here are some of the most popular
```{r}
#bw
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()

#minimal
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_minimal()

#classic
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_classic()

#If you want even more themes you can use the ggthemes library!
#install.packages("ggthemes")
library(ggthemes)

#WSJ
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_wsj()

#538
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_fivethirtyeight()

#The Economist
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_economist()

```
## Facets
Using facets lets you create multi-panel plots based off a given variable.
There are two different ggplot functions to use to facet. 
```{r}
#1. facet_wrap() "wraps" a 1d ribbon of panels into 2d. Essentially, creating a square grid of panels, with rows and columns defined by combinations of two categorical variables (even if some combinations of faceting values do not exist in data). This is most useful when you want to facet by a single variable (or two)

carsplot_facet <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_wrap(~ am)

#2. facet_grid() forms a matrix of panels defined by row and column faceting variables, creating panels based on observed combinations of values in faceting variables

#facet_grid(.~variable) will return facets equal to the levels of variable distributed horizontally
#facet_grid(variable~.) will return facets equal to the levels of variable distributed vertically.

carsplot_facet_2 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_grid(vs ~ am)

carsplot_facet_3 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_grid(am ~ .)

carsplot_facet_4 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()+
  facet_grid(. ~ am)
```

What are other types of common geoms?

geom_histogram(): plots a histogram of single value; usually needs modification to number of bins.

geom_density(): density plot of single variable.

geom_bar(): vertical barplot by categorical variable

geom_point(): typical two-way scatterplot

geom_smooth(): simple line interpolator

# Part 3: Practice
## Tidyverse practice

Use the data below to first split the complete dataset into individual planes and then summarise each plane 
by counting the number of flights (count = n()) and computing the mean distance flown.

Next, determine what city is flown to the most often from NYC.

Finally, create a plot to create a barplot visualizing if flights are on time or delayed (defined as later than 5 minutes) by month. Additionally, ensure your plot has meaningful labels.

```{r}
#First, we need to install our data. This dataset includes airline on-time data for all flights departing NYC in 2013. 
#install.packages("nycflights13")

library(nycflights13)

data("flights")
dplyr::as_tibble(flights)

```

```{r}
by_tailnum <- ___(___, tailnum) %>% 
  summarise(count = n(), dist = ___)

```

```{r}
#What city is flown to the most often from NYC?

destinations <- ___(flights, dest) %>%
  ___(planes = n_distinct(tailnum),
            flights = n())%>%
  arrange(desc(flights))

#calling to view top 10 
d1<-___(destinations,___)

```

```{r}
#Barplot
flight_delays<-flights %>% 
  mutate(dep_type = ___(dep_delay < 5, "on time", "delayed")) %>% 
   ___(dep_type,month) %>%
    ___(n = n()) %>% 
    filter(dep_type!="")

plot1<-ggplot(data=___) +

```

### 2023-01-04 Lab 3  Fitting LMs/Lab3_LM.Rmd
Add files via upload

---
title: "Lab 3: The Linear Model in R"
author: "Eric Parajon"
output: pdf_document
---
Adapted  from code and text from Isabel Laterzo, Simon Hoellerbauer and Chelsea Estancona.

This lab will introduce you to `lm()`, the function we will be using to fit linear models (for now) in \texttt{R}. I will walk you through the different parts of this lab, but then we will largely be working independently. Chime in whenever you have questions!

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, include = TRUE, eval=FALSE)

#Clear workspace
rm(list=ls())

#The pacman package is required to load/install the additional packages. Uncomment the line below to install the package.
#install.packages("pacman")

#Installing and loading packages for use
pacman::p_load(xtable,stargazer,palmerpenguins,tidyverse)

```

Don't forget to set your working directory!
```{r, eval = FALSE}

setwd()

```

# Part 1: LM in R

First, we'll simulate some data to use. This will be a very useful skill in the future, so pay attention to what we do here.

We first create our variables and then stipulate what our effects should be.

Let's set our seed to 123.

For our $\mathbf{x}_1$ variable, let's take 20 draws from a $\mathcal{N}(1.3, 2.3)$ (using the `R` parameterization, where the second parameter is the standard deviation).

Let's set $\beta_0$ to be 17, and $\beta_1$ to be 1.3. We also can't forget to simulate $\boldsymbol{\epsilon}$! Let's $\boldsymbol{\epsilon} \sim \mathcal{N}(0, 3.4)$.

```{r, eval = FALSE}
set.seed(123)

##Explanatory variables
x1 <- rnorm(__, __, __)

##Coefficients
b0 <- 17 #Intercept
b1 <- 1.3 #Beta 1

##standard deviation
sigma <- 3.4

error <- rnorm(length(x1), 0, sigma) #Let's simulate some random error, too.

```

What would we do in order to simulate a linear relationship with the variables and effects above? Create the outcome variable `y`.
```{r, eval = FALSE}
y <- _____________________
```

What do $\mathbf{x}_1$ and $\mathbf{y}$ look like together? Use the `plot` function to plot a scatter plot of $\mathbf{x}_1$ against $\mathbf{y}$.

```{r, eval = FALSE}
plot(x1, y)
```

Now that we have simulated explanatory variables and set up a 'true' relationship between our explanatory variable (x1) and outcome variable, we can use R's built-in functions to evaluate our linear model. For a linear model, we will want to use the `lm()` function. 
The basic syntax for this function is `lm(Y ~ X1 + X2)`.

Use `summary()` on the model object (`m1` in this case) in order to see some of the most interesting information about the fitted model.

```{r, eval = FALSE}
m1 <- lm(___________) #the basic syntax here: lm(Y ~ X1 + X2)
#summarize the 'model object' we've created
summary(m1)
```

# Part 2: Combining LM and the tidyverse

We can also combine what we learned about using the tidyverse with linear modeling!

As we dive into learning linear models in `R`, we're going to use some data from the `mtcars`. But first, lets use the `tidyverse` to clean it up a bit. First, take a look at the mtcars data (its readily available so you don't really need to load it in!).

```{r}
library(tidyverse)
#examine data
summary(mtcars)
```

Great. Let's say we want to analyze the role that `disp` and `cyl` play in influencing `mpg`. However, we're only interested in observations where `cyl` is greater than 4. Let's use `dplyr` and the `tidyverse` to only keep observations where `cyl` is greater than 4. Then, only keep the three above variables in the new data set. Label this data set `new_data`.

Hint: the `dplyr` functions you are interested in here are `filter()` and `select()` - don't forget to call their help files if you need more information (e.g., `?filter()` in your console).
```{r, eval = FALSE}

new_data <- mtcars %>%
  filter(cyl > 4) %>%
  select(__, __, __)
```

Now that we have our data ready, we can use `lm()`.  Fill in the blank in the chunk below where Y is `mpg`, X1 is `cyl`, and X2 is `disp`. Use `summary()` on the model object (`m2` in this case) in order to see some of the most interesting information about the fitted model.

Remember, our data is `new_data`.

```{r, eval = FALSE}
#model
m2 <- lm(___________)

#summarize
summary(___)

```

The `summary` function provides lots of valuable information about the model we've created. We can also extract different information if we need, for example, just the coefficients, or just the residuals. Use `View()` on `m2`. What does it look like `m2` is? What do you think we could do if we wanted to access the different objects within `m2`?

```{r, eval=FALSE, echo=TRUE}
#Taking a look at m1
____________

#extract coefficients
m2$_____

#extract residuals
m2$______

#extract y hat 
m2$_______
```

We can also use indexing to get to parts of these vectors - the output from a model is a list object, which means we can use brackets with $ indexing.

```{r, eval = FALSE}
m2$coefficients[1] #What does this give us?
m2[2] #What about this?
```

What if we wanted to include an interaction term? Try interacting `disp` and `cyl`. 
```{r, eval = FALSE}
m3 <- lm(mp ~ disp + cyl + _______, data=new_data) #use * for interaction- include both variables

summary(m3)
```

We don't actually have to write all of the variables two times; the `lm` function will include the variables involved in the interaction even if we don't specify them individually in the formula. Try this out below, creating model object `m4` and compare it to `m3`.
```{r, eval = FALSE}
#m4:
m4<-lm(_________)

```

How might we present these results in a table? Both `xtable()` and `stargazer()` are functions that produce LaTeX code that could be copy-pasted into a .tex document or included in an Rmarkdown document. Try both of these on one of the model object `m4` we created above and see what happens. Note, these are VERY flexible and can be customized to make very nice tables, the ones we are generating here are just basic. 
```{r, eval=FALSE, echo=TRUE}
#These produce code for a .pdf that we can copy and paste
xtable(______)
stargazer(_______)

#you can also add lots of useful information onto these
#check out guides online!

```

When presenting our results, we might be largely concerned with prediction: if I have a model and provide it with 'new' data, what should I expect? How confident are we in that expectation? We can use the handy `predict()` function here, which works with almost any type of model object.

Check out `predict.lm` using ?, as that is the method that `predict` invokes when we feed it an lm object.

```{r}
#fitting our model
m5 <- lm(mpg ~ disp + cyl, data = mtcars)

#looking at our model
summary(m5)

#what are we doing here? Also, what does with allow us to do here?
x_vec_n <- with(mtcars, seq(min(disp), max(disp), length.out = 500))
#we've created a 500 length vector starting from the min value of disp
#through the max value of disp

#why am I choosing these values? What are other options?
new_dat_n_1 <- with(mtcars, data.frame(x_vec_n, 4)) 
new_dat_n_2 <- with(mtcars, data.frame(x_vec_n, 6))
new_dat_n_3 <- with(mtcars, data.frame(x_vec_n, 8))

#what are we doing here?
colnames(new_dat_n_1) <- c("disp", "cyl")
colnames(new_dat_n_2) <- c("disp", "cyl")
colnames(new_dat_n_3) <- c("disp", "cyl")
#we just created three data frames - all with our x_vec_n vaalues for "disp"
#and three values we have set for "cyl"

#this is very important for predict! See note below.

#getting our predicted values 
fit_n_1 <- predict(m5, newdata = new_dat_n_1, interval = "confidence")

fit_n_2 <- predict(m5, newdata=new_dat_n_2, interval="confidence")

fit_n_3 <- predict(m5, newdata=new_dat_n_3, interval="confidence")

#ggplot is rather particular about the format of the data you will be using
plot_dat_n <- data.frame(fit_n_1, fit_n_2, fit_n_3, new_dat_n_1)

#note we never call `cyl` on its own below so the iteration of `new_dat_n_x` we include
#here doesn't matter a ton, it just gives us all values of disp

#this looks like a monster plot, but there are just a lot of layers. A lot of 
#ggplot plot code will end up looking like this
ggplot(data=plot_dat_n) + 
  labs(x = "Displacement of Vehicle", y = "Miles Per Gallon")+
  geom_line(aes(x = disp, y = fit_n_1[, 1], color = "4 cylinders"),linetype = "solid") +
  geom_line(aes(x = disp, y = fit_n_2[, 1], color = "6 cylinders"),linetype = "solid") +
  geom_line(aes(x = disp, y = fit_n_3[, 1], color = "8 cylinders"), linetype = "solid") +

  #what are we doing with the ribbons? What does alpha do? Play around with it.
  geom_ribbon(aes(x = disp, ymin = fit_n_1[, 2], ymax = fit_n_1[, 3]), alpha = .3, fill = "red") +
  geom_ribbon(aes(x = disp, ymin = fit_n_2[, 2], ymax = fit_n_2[, 3]), alpha = .3, fill = "green") +
  geom_ribbon(aes(x = disp, ymin = fit_n_3[, 2], ymax = fit_n_3[, 3]), alpha = .3, fill = "blue") +
  theme(legend.title = element_text(size = 14),
        legend.position = "bottom",
        panel.grid.minor.x = element_blank(), #what does this do?
        panel.grid.minor.y = element_blank()) +
  scale_color_discrete(name="Number of Cylinders") + #what does this seem to do? what happens if you take it out?
  theme_bw()
```

Note that when using predict, it is very important that our new observations have variables that bear the same name as in the formula. Otherwise `predict` won't know what to do! Try out the following:
```{r, eval = FALSE}
#recreating new_dat_n_1
new_dat_n_1_a <- with(mtcars, data.frame(x_vec_n, 4))

#but this time we don't label the variables
predict(m5, newdata = new_dat_n_1_a, interval = "confidence")

#and this time we label them incorrectly
colnames(new_dat_n_1_a) <- c("cisp", "dyl")
predict(m5, newdata = new_dat_n_1_a, interval = "confidence")

#and this time we flip the variables
new_dat_n_1_a <- with(mtcars, data.frame(4, x_vec_n))
colnames(new_dat_n_1_a) <- c("disp", "cyl")
predict(m5, newdata = new_dat_n_1_a, interval = "confidence")
#this time it works, but our predictions are clearly incorrect (compare them to
#fit_n_1 from above)

#this works!
new_dat_n_1_a <- with(mtcars, data.frame(4, x_vec_n))
colnames(new_dat_n_1_a) <- c("cyl", "disp")
fit_n_1_a <- predict(m5, newdata = new_dat_n_1_a, interval = "confidence")

all(fit_n_1 == fit_n_1_a) #what does all() do here?

```

# Part 3: LM practice
Using the palmerpenguines dataset, create an OLS model using multiple predictors and then display the results in a publication quality table (using stargazer).
```{r}
library(palmerpenguins)
library(stargazer)

#Check out the data file
(penguins)

#Create OLS file

model1<-lm(______,data=penguins)

#Create table

stargazer(____)
```

### 2023-01-04 Lab 3  Fitting LMs/Lab3_LM_Answers.Rmd
Add files via upload

--- 2023-01-04 Lab 3  Fitting LMs/Lab3_LM.Rmd
@@
 ---
+
 Adapted  from code and text from Isabel Laterzo, Simon Hoellerbauer and Chelsea Estancona.
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(cache = TRUE, include = TRUE, eval=FALSE)
+knitr::opts_chunk$set(cache = TRUE, include = FALSE, echo=TRUE)
 
@@
 
-```{r, eval = FALSE}
+```{r}
 set.seed(123)
@@
 ##Explanatory variables
-x1 <- rnorm(__, __, __)
+x1 <- rnorm(20, 1.3, 2.3)
 
@@
 error <- rnorm(length(x1), 0, sigma) #Let's simulate some random error, too. 
-
+#Random error is always included in a regression model because unless there is a linear relationship between the variables in the model, there will always be some amount of error (or variation) which is not totally unaccounted for by the model.
 ```
@@
 What would we do in order to simulate a linear relationship with the variables and effects above? Create the outcome variable `y`.
-```{r, eval = FALSE}
-y <- _____________________
+```{r}
+y <- b0 + b1 * x1 + error
 ```
@@
 
-```{r, eval = FALSE}
+```{r}
 plot(x1, y)
@@
 The basic syntax for this function is `lm(Y ~ X1 + X2)`.
-
 Use `summary()` on the model object (`m1` in this case) in order to see some of the most interesting information about the fitted model.
 
-```{r, eval = FALSE}
-m1 <- lm(___________) #the basic syntax here: lm(Y ~ X1 + X2)
+```{r}
+m1 <- lm(y ~ x1) #the basic syntax here: lm(Y ~ X1 + X2)
 #summarize the 'model object' we've created
@@
 
-We can also combine what we learned about using the tidyverse with linear modeling!
-
-As we dive into learning linear models in `R`, we're going to use some data from the `mtcars`.
-But first, lets use the `tidyverse` to clean it up a bit.
-First, take a look at the mtcars data (its readily available so you don't really need to load it in!).
-
 ```{r}
-library(tidyverse)
 #examine data
@@
 Hint: the `dplyr` functions you are interested in here are `filter()` and `select()` - don't forget to call their help files if you need more information (e.g., `?filter()` in your console).
-```{r, eval = FALSE}
+
+```{r}
 
@@
   filter(cyl > 4) %>%
-  select(__, __, __)
+  select(cyl, disp, mpg)
 ```
 
-Now that we have our data ready, we can use `lm()`.
+Now that we have our data ready, we can use R's built-in functions to evaluate our linear model.
+For a linear model, we will want to use the `lm()` function.
+The basic syntax for this function is `lm(Y ~ X1 + X2)`.
 Fill in the blank in the chunk below where Y is `mpg`, X1 is `cyl`, and X2 is `disp`.
-Use `summary()` on the model object (`m2` in this case) in order to see some of the most interesting information about the fitted model.
+Use `summary()` on the model object (`m1` in this case) in order to see some of the most interesting information about the fitted model.
 
@@
 
-```{r, eval = FALSE}
+```{r}
 #model
-m2 <- lm(___________)
+m2 <- lm(mpg~disp+cyl, data=new_data)
 
 #summarize
-summary(___)
+summary(m2)
 
@@
 ```{r, eval=FALSE, echo=TRUE}
-#Taking a look at m1
-____________
+#Taking a look at m2
+View(m2)
 
 #extract coefficients
-m2$_____
+m2$coefficients
 
 #extract residuals
-m2$______
+m2$residuals
 
 #extract y hat 
-m2$_______
+m2$fitted.values
 ```
@@
 
-```{r, eval = FALSE}
+```{r}
 m2$coefficients[1] #What does this give us?
@@
 
-What if we wanted to include an interaction term? Try interacting `disp` and `cyl`. 
-```{r, eval = FALSE}
-m3 <- lm(mp ~ disp + cyl + _______, data=new_data) #use * for interaction- include both variables
+What if we wanted to include an interaction term?
+```{r}
+m3<-lm(mpg~disp+cyl+disp*cyl, data=new_data) #use * for interaction- include both variables
 
@@
 Try this out below, creating model object `m4` and compare it to `m3`.
-```{r, eval = FALSE}
-#m4:
-m4<-lm(_________)
+```{r}
+#m3:
+m4 <- lm(mpg ~ disp*cyl, data=new_data)
 
+summary(m4)
 ```
@@
 ```{r, eval=FALSE, echo=TRUE}
-#These produce code for a .pdf that we can copy and paste
-xtable(______)
-stargazer(_______)
+
+xtable(m4)
+stargazer(m4)
 
@@
 #check out guides online!
-
 ```
@@
 We can use the handy `predict()` function here, which works with almost any type of model object.
-
 Check out `predict.lm` using ?, as that is the method that `predict` invokes when we feed it an lm object.
@@
 m5 <- lm(mpg ~ disp + cyl, data = mtcars)
-
 #looking at our model
@@
 colnames(new_dat_n_3) <- c("disp", "cyl")
-#we just created three data frames - all with our x_vec_n vaalues for "disp"
+#we just created three data frames - all with our x_vec_n values for "disp"
 #and three values we have set for "cyl"
@@
 plot_dat_n <- data.frame(fit_n_1, fit_n_2, fit_n_3, new_dat_n_1)
-
 #note we never call `cyl` on its own below so the iteration of `new_dat_n_x` we include
@@
   geom_line(aes(x = disp, y = fit_n_3[, 1], color = "8 cylinders"), linetype = "solid") +
-
   #what are we doing with the ribbons? What does alpha do? Play around with it.
@@
 
-all(fit_n_1 == fit_n_1_a) #what does all() do here?
+all(fit_n_1 == fit_n_1_a) #what does all() do here? #check in a vector whether all the values are true or not.
 
@@
 # Part 3: LM practice
-Using the palmerpenguines dataset, create an OLS model using multiple predictors and then display the results in a publication quality table (using stargazer).
+Using the palmerpenguines dataset, create an OLS using multiple predictors and then display the results in a publication quality table (using stargazer).
 ```{r}
@@
 
-#Check out the data file
-(penguins)
+#Running model
+bill_model <- lm(bill_length_mm ~ species+sex+body_mass_g, penguins)
 
-#Create OLS file
+summary(bill_model)
 
-model1<-lm(______,data=penguins)
+#Putting results in table
+stargazer(bill_model,title="Effect of Predictors on Bill Length",digits=3,star.cutoffs = c(0.1,0.05,0.01),
+          label="bill_model",
+          covariate.labels = c("\\textbf{Species} (reference= Adélie) &   \\\\
+                      \\hspace{1cm}Chinstrap",
+                      "\\hspace{1cm}Gentoo",
+                    "\\textbf{Sex} (reference= Female) &   \\\\
+                      \\hspace{1cm}Male",
+                      "\\textbf{Body mass (grams)}"),
+          notes="\\parbox[t]{12cm}{Coefficients reported from OLS regression models. Significance codes:*$<$ 0.10,**p$<$ 0.05, ***p$<$ 0.01, two-tailed tests.}",
+          notes.append=FALSE,
+          style="apsr",
+          dep.var.caption = c(""),
+          dep.var.labels.include = FALSE,
+          table.layout ="-c-!t-s-n",
+          header=FALSE,
+          notes.align= "l",
+          column.sep.width = "1pt",
+          font.size = "scriptsize",
+          no.space = TRUE)
 
-#Create table
-
-stargazer(____)
 ```

### 2023-01-04 Lab 4  Graphical Interpretation of LMs/Lab4_Graphical_interpretation_LM.Rmd
Add files via upload

---
title: "Lab 4: Graphical interpretation of LMs"
author: "Eric Parajon"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, include = TRUE)

#Clear workspace
rm(list=ls())

#The pacman package is required to load/install the additional packages. Uncomment the line below to install the package.
#install.packages("pacman")

#Installing and loading packages for use
pacman::p_load(tidyverse,palmerpenguins,ggeffects)

#Creating a ggplot theme for use in several figures throughout the code.
plot_theme = theme(text = element_text(size=11),  
                   panel.grid.minor = element_blank(),
                   panel.grid.major = element_blank(),
                   panel.background = element_blank(),
                   plot.background = element_rect(fill = "transparent",colour = NA),
                   axis.text = element_text(color="black"),
                   axis.line = element_line(colour = "black"),
                   strip.text.y = element_text(angle = 0), axis.title.y = element_text(angle=90, hjust=.5, vjust=.5),
                   axis.title = element_text(lineheight=1.1), 
                   legend.position = "bottom", 
                   legend.key = element_rect(colour = "transparent",  fill = "transparent") 
)

```
Today's lab will focus on graphical interpretation of LMs using a variety of R tools.

Don't forget to set your working directory!
```{r, eval = FALSE}

setwd()

```

#Part 1: Ggplot visualizations
Lets begin by returning to the mtcars data set and creating a basic scatterplot of displacement (disp) vs mpg.

```{r, message=FALSE, warning=FALSE}
carsplot <- ggplot(data = mtcars, aes(x = ____, y = ____)) + 
  ____(aes(colour = cyl), size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_colour_continuous(name = "Cylinders")+
  theme_bw()

```

Okay great, but what if we wanted to plot the linear relationship between our x and y?

```{r}
carsplot + geom_smooth(method = "lm", se = FALSE, 
                       color = "darkgrey", size = 0.75)
#geom_smooth is for fitted lines.
#se=TRUE plots confidence intervals
#We can adjust the color and size of the line plotted.
#Also, note that we just added a layer over carsplot - did we change the carsplot object itself?
```

Right now, we're using color as the aesthetic mapped to a specific explanatory variable (color represents the number of cylinders). We can play around with this:

```{r}
ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(size = cyl), colour = "red") +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") +
  theme_bw()

ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
  scale_size_continuous(name = "Cylinders") +
  theme_bw()
```

What are the benefits/drawbacks of each of these strategies? Think about how well the plot communicates information about the types of variables plotted as well as the relationship between them.

The above are great places to start out visualizing any potential relationships between your variables.

Now let's jump back into the palmerpenguins dataset to perform some more in-depth visualization of regression models.

Here we will evaluate the association between penguin body mass (measured in grams) and flipper length (measured in mm).

Let's suppose our main research question is to determine whether body mass was associated with flipper length We will set the body mass as our dependent variable and flipper length as our independent variable (or predictor of interest).

<div align="left">$\large E[Body_mass_{i} | Flipper_length_{i}] = \beta_0 + \beta_1  Flipper_length_{i} + \epsilon,$</div>

where $E[Body_mass_{i} |  Flipper_length_{i}]$ denotes the expected body mass for penguin $i$ given the flipper length of penguin $i$.

```{r}
#We could start by plotting the association between the penguin's body mass and flipper length

mass_flipper <- ggplot(data = penguins, 
                       aes(x = flipper_length_mm,
                           y = body_mass_g))+
  geom_point()

#Does there seem to be an association?

#Yes! as flipper length increases, the body mass also increases. There appears to be a positive relationship between penguin body mass and flipper length.

```

Now, we can construct a linear regression model with body mass as the dependent variable and flipper length. as the independent variable (or predictor of interest).

```{r}
linear.model1 <- lm(body_mass_g ~ flipper_length_mm, data = penguins)
summary(linear.model1)

#Using the broom package to tidy the model data and format into a more easily usable form.
temp_lm = broom::tidy(linear.model1, se = 'standard', conf.int = TRUE,conf.level = 0.95) %>% 
  filter(term=="flipper_length_mm")

#Visualize the results in a coefficient plot
Penguin_results <- ggplot(data=temp_lm, aes(x=term, y=estimate)) + 
    geom_hline(yintercept=0, color="red", size=.5) +
    geom_errorbar(aes(ymin=conf.low, ymax=conf.high, width=0), size=.5, position=position_dodge(width=0.5)) +
    geom_point(aes(y=estimate), size=1.75, position = position_dodge(width=0.5)) +
    ylab("Effect of Flipper Length on Penguin Body Mass") +
    xlab("")+
    coord_flip()+
  geom_text(aes(y=estimate,label=round(estimate, digits=2)), size=2.5, vjust=-1.5, position = position_dodge(width=.2))

#Okay great so there's an effect! But what about confounders? Perhaps the penguins species is also important.

```
Lets again plot it but this time break out the results by species.

```{r}
mass_flipper_species <- ggplot(data = penguins, 
                       aes(x = flipper_length_mm,
                           y = body_mass_g)) +
  geom_point(aes(color = species, 
                 shape = species),
             size = 3,
             alpha = 0.8) +
  scale_color_manual(values = c("darkorange","purple","cyan4")) +
  labs(title = "Flipper length and body mass",
       x = "Flipper length (mm)",
       y = "Body mass (g)",
       color = "Penguin species",
       shape = "Penguin species")

#Definitely seems like species could be important to capture in our model. So let's add to our current model by including a confounder.

```
Now our model looks like this.
<div align="left">$\large E[Body\_mass_{i} | Flipper\_length_{i}], Species_{i}] = \beta_0 + \beta_1  Flipper\_length_{i}+ \beta_2  Species_{i} + \epsilon$</div>

where $E[Body\_mass_{i} |  Flipper\_length_{i}], Species{i}]$ denotes the expected body mass for penguin $i$ given the flipper length of penguin $i$ and controlling for species of the penguin $i$.

```{r}
linear.model2 <- lm(body_mass_g ~ flipper_length_mm+species, data = penguins)
summary(linear.model2)

#Using the broom package to tidy the model data and format into a more easily usable form.
temp_lm_2 = broom::tidy(linear.model2, se = 'standard', conf.int = TRUE,conf.level = 0.95) %>% 
  filter(term!="(Intercept)")

#Visualize the results in a coefficient plot
Penguin_results_2 <- ggplot(data=temp_lm_2, aes(x=term, y=estimate)) + 
    geom_hline(yintercept=0, color="red", size=.5) +
    geom_errorbar(aes(ymin=conf.low, ymax=conf.high, width=0), size=.5, position=position_dodge(width=0.5)) +
    geom_point(aes(y=estimate), size=1.75, position = position_dodge(width=0.5)) +
    ylab("Effect of Flipper Length and Species on Penguin Body Mass") +
    xlab("")+
    coord_flip()+
  geom_text(aes(y=estimate,label=round(estimate, digits=2)), size=2.5, vjust=-1.5, position = position_dodge(width=.2))+
  plot_theme

```

Okay great, but what if we want to visualize substantive effects? For this we can use our old friend predict()!

```{r}

newData <- expand.grid(flipper_length_mm = seq(190, 200, by = 5),
                       species=c("Adelie","Chinstrap","Gentoo"))

plotData <- as.data.frame(predict(linear.model2, newdata=newData, interval='confidence'))

#Pulling it all together
int_dat <- data.frame(newData, plotData)

subeffects_plot <- ggplot(int_dat, aes(x=flipper_length_mm, y=fit, ymin=lwr, ymax=upr,color=species)) +
        geom_pointrange() +
 scale_color_manual(values = c("darkorange","purple","cyan4")) +
  labs(title = "Flipper length and body mass",
       x = "Flipper length (mm)",
       y = "Body mass (g)",
       color = "Penguin species",
       shape = "Penguin species")
subeffects_plot+
  plot_theme

```

## Interaction analysis
Additionally, in other models we may be interested in exploring the interaction between variables. Often these interaction effects are difficult to interpret without graphic aids.

In this example, suppose we are interested in knowing if the effect of body mass on bill length is attenuated by the sex of the penguin. We begin by plotting the data and then perform interaction analysis.

Our regression equation looks like this:

$body\_mass = B0 + B1*bill\_length + B2*sex + B3*bill\_length*sex+ \epsilon$

```{r}
#First let's plot the raw data to see if there's anything there.
penguins %>% 
  # drop rows with missing data
  drop_na(sex) %>% 
  ggplot(aes(x = body_mass_g, y = bill_length_mm, color = sex)) +
  geom_point() +
  facet_wrap(~ sex)

penguin_interact<-lm(bill_length_mm~body_mass_g*sex,data=penguins)

tidy_penguin_interact = broom::tidy(penguin_interact, se = 'standard', conf.int = TRUE,conf.level = 0.95)

#Next we'll use ggpredict to get the marginal effects for the main effects and the interaction term.
#There are a ton of different packages that do similar things make_predictions from jtools is also great.

y_hat <- ggpredict(penguin_interact, terms = c("body_mass_g", "sex"))

#This returns predicted values of the dv
y_hat

#Plotting interaction effects

  #First extracting data from the model
model_data <- penguin_interact$model
  #Next, rename group to sex 
y_hat <- 
  y_hat %>% 
  rename(sex = group)

#Plotting
ggplot(data = y_hat, aes(x = x, y = predicted, fill = sex)) +
  #plot the fitted line
  geom_line() +
  #plot the confidence intervals
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.3) +
  #plot the model data
  geom_point(data = model_data, aes(x = body_mass_g, y = bill_length_mm),
             size = 2, shape = 21, alpha = 0.5) +
  facet_wrap(~ sex, scales = "free") +
  labs(title = "The Effect of Body Mass on Bill length by Sex Among Penguins",
       x = "Body mass (grams)",
       y = "Predicted Bill length",
       fill = "Sex",
       caption = "Fitted line estimate with OLS. \n The shaded region shows the 95% confidence interval.")

```

#Part 3: LM practice
Use the msleep dataset to model how body weight (bodywt) and brain weight (brainwt) influences total REM sleep (sleep_rem). Then create a model including an additional predictor (of your choice). Create a coefficient plot of the results of the two models and compare them.

```{r}
glimpse(msleep)

#Creating model 1
model1 <- lm(______)

#Tidying model 1
model1_tidy<-______
  mutate(model = "Model 1")

#Creating model 2
model2 <- lm(______)

#Tidying model 2
model2_tidy<-______ %>% 
  mutate(model = "Model 2")

#Combine results into a single dataframe
sleep_results <- bind_rows(model1_tidy, model2_tidy)

#Plotting (make sure to assign color and shape to the model variable in the ggplot call)
sleep_results_plot <- ggplot(_____)

#Interpret 
```

### 2023-01-04 Lab 4  Graphical Interpretation of LMs/Lab4_Graphical_interpretation_LM_Answers.Rmd
Add files via upload

--- 2023-01-04 Lab 4  Graphical Interpretation of LMs/Lab4_Graphical_interpretation_LM.Rmd
@@
 #Part 1: Ggplot visualizations
-Lets begin by returning to the mtcars data set and creating a basic scatterplot of displacement (disp) vs mpg.
+Lets begin by returning to the mtcars data set and creating a basic plot of Displacement vs mpg.
 
 ```{r, message=FALSE, warning=FALSE}
-carsplot <- ggplot(data = mtcars, aes(x = ____, y = ____)) + 
-  ____(aes(colour = cyl), size = 2) +
+carsplot <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
+  geom_point(aes(colour = cyl), size = 2) +
   labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
@@
 
+#One method for saving plots:
+#pdf("Plot1.pdf")
+#carsplot
+#dev.off()
+
+#Where do these plots go?!
+
+#Unique to ggplot:
+#ggsave("Plot1a.pdf") #can edit image sizes as well
 ```
@@
 
-#Okay great so there's an effect! But what about confounders? Perhaps the penguins species is also important.
+#Okay great so huge effect! But what about confounders? Perhaps the penguins species is also important.
 
@@
 ```
-
 #Part 3: LM practice
@@
 #Creating model 1
-model1 <- lm(______)
+model1 <- lm(sleep_rem ~ bodywt + brainwt, data = msleep)
 
 #Tidying model 1
-model1_tidy<-______
+model1_tidy<-broom::tidy(model1, se = 'standard', conf.int = TRUE,conf.level = 0.95) %>% 
   mutate(model = "Model 1")
@@
 #Creating model 2
-model2 <- lm(______)
+model2 <- lm(formula = sleep_rem ~ bodywt + brainwt + sleep_total, data = msleep)
 
 #Tidying model 2
-model2_tidy<-______ %>% 
+model2_tidy<-broom::tidy(model2, se = 'standard', conf.int = TRUE,conf.level = 0.95) %>% 
   mutate(model = "Model 2")
@@
 #Combine results into a single dataframe
-sleep_results <- bind_rows(model1_tidy, model2_tidy)
+sleep_results <- bind_rows(model1_tidy, model2_tidy) %>% 
+  filter(term!="(Intercept)")
 
 #Plotting (make sure to assign color and shape to the model variable in the ggplot call)
-sleep_results_plot <- ggplot(_____)
+sleep_results_plot <- ggplot(data=sleep_results, aes(x=term, y=estimate, 
+           color = model, shape = model)) + 
+    geom_hline(yintercept=0, color="red", size=.5) +
+    geom_errorbar(aes(ymin=conf.low, ymax=conf.high, width=0), size=.5, position=position_dodge(width=0.5)) +
+    geom_point(aes(y=estimate), size=1.75, position = position_dodge(width=0.5)) +
+  labs(title = "Model Estimates of Brain and Body Weight (and total sleep) on REM Sleep",
+       x = "Predictor",
+       y = "Coefficient Estimate",
+       caption = "Models fit with OLS. Error bars show the 95% confidence interval.",
+       color = "Model:",
+       shape = "Model:") +
+   scale_x_discrete(labels = c("Body Weight", "Brain Weight", "Total Sleep"))+ 
+  coord_flip()+
+  plot_theme
 
-#Interpret 
 ```

### 2023-01-04 Lab 5  Optim/Lab5_Optim.Rmd
Add files via upload

---
title: "Lab 5: Optimization"
author: "Eric Parajon"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, include = FALSE)

rm(list=ls())

pacman::p_load(tidyverse,broom)
```

Some of today's code has been adapted from Simon Hoellerbauer, Isabel Laterzo and Sean Norton.

At the end of the day, maximum likelihood estimation is an optimization problem. We're looking for the set of inputs that maximizes that likelihood function. (Alternatively, you can swap the sign and treat it as a minimization problem, because the actual numbers are meaningless - only the relationship between them matters.)

Let's say we have a function $f(x)$ - we want to know what value of $x$ will maximize or minimize that function. In other words, let us say that the local maximum (or minimum) of a functtion is called $x^*$ (because it's the star on top of the plot of the function). Then we could say, for maximization, that:

$$x^* = argmax f(x)$$

For minimization, we would say:
$$x^* = argmin f(x)$$

Today, we're going to learn how to do that with the function `optim()`, which we will apply to fitting normal linear models. You may be thinking that we already have a function, `lm()`, that does that without requiring any mental effort from you. This is true, but understanding optimization will both a) help you better understand the concepts behind OLS and MLE and b) make you appreciate the R devs who wrote convenient modeling functions so that you no longer have to do this every time you want to fit a model.

# Newton-Raphson for MLE

One method of optimization is to employ derivative-based methods. These can be solved analytically or computed numerically, i.e. iteratively. We'll predominantly take the second approach - computers are very bad at calculating derivatives analytically since this involves manipulation of symbols rather than numbers.

Recall that the root of a function is the values of the inputs when $f(.) = 0$. Also, remember that the maximum of a function is when its derivative is equal to zero. By approximating the root of the first derivative, we can find the maximum of the likelihood function. But why approximate it if we can solve it analytically? In addition to computers struggling to solve derivatives numerically, likelihood functions are not always particularly tractable - the derivative may be very complex.

Assumptions: our function of interest is smooth and has a single minimum.

We'll be using the Newton-Raphson algorithm to do so (yes, Isaac Newton). The basic format of the algorithm is: 
$$x_{n+1}=x_n-\frac{f(x_n)}{f'(x_n)}$$
Let's think about what that means. We're stating that given an unknown $x_{n+1}$ but a known $f(x_n)$, we can compute the unknown value using the output of the function for the known value and the first derivative. Why? Think about what that fraction is doing - you are dividing the output of the function at $x_n$ by the derivative of the function at $x_n$. What does the derivative represent?

But since we are dealing with maximum likelihood estimation, we define $f(x_n) = l'(\theta_n)$, where $l(\theta_n)$ is the appropriate log likelihood function. (We use log likelihoods because they are mathematically easier to work with.)

Since we are interested in optimizing, the derivatives in this case are very important. The first derivative tells us the slope of the function at any given point. If you're trying to maximize a function, you're looking for the point at which that derivative becomes zero. Think of climbing a hill - as you approach the top, the slope will remain positive but gradually decrease towards zero near the peak. The same thing applies for the derivative - you're trying to find the inflection point at which the first derivative is zero.

But since we are dealing with maximum likelihood estimation, we define $f(x_n) = l'(\theta_n)$, where $l(\theta_n)$ is the appropriate log likelihood function. Since we are interested in optimizing, the derivatives in this case are very important. The role of the first derivative should be simple to understand given the above explanation. The role of the second derivative is a bit different.

Let's suppose we're "hiking" up that mathematical hill trying to find the maximum, and using the first derivative to decide how big of a "step" we take from our current location. However, the first derivative only gives us information on the steepness of that hill. If all we're paying attention to is gaining elevation, we could take a very circular path up to the top. That's where the second derivative comes in.

The second derivative can be thought of as "the rate of change in the rate of change".Remember that as we approach the maximum, the first derivative should be decreasing. Likewise, the second derivative should be as well. (Remember - what's the second order condition for maximization?) We can use the first derivative to determine how big of a step to take, and the second derivative to tell us what direction is the most direct route to the top. (The second derivative encodes "curvature" information for the function.)

See \href{https://spia.uga.edu/faculty_pages/rbakker/pols8501/MLENotes1a.pdf}{these helpful notes} for more detail. So, we have:

$$\theta_{n+1}=\theta_n-\frac{l'(\theta_n)}{l''(\theta_n)}$$

Where $l''(\theta_n)$ is known as the Hessian of the log-likelihood function.

I will show this with a simple Poisson model. First, we observe data $x_1, x_2, ... , x_n \sim Poisson (\mu)$ and we want to estimate $\mu$ via maximum likliehood. The log-likelihood of the Poisson model is:

$$\mathcal{l}(\mu) \sum_{i = 1}^{n} x_{i}log(\mu) - \mu = n \bar{x}(log\mu) - n \mu$$

The first derivative of the Poisson log-likelihood (called the score function) is:

$$l^{'}(\mu) = \frac{n \bar{x}}{\mu} - n$$

The second derivative is:
$$l^{''}(\mu) = - \frac{n \bar{x}}{\mu^{2}}$$

Thus the Newton iteration becomes:

$$\mu_{n+1} = \mu_{n} - [- \frac{n \bar{x}}{\mu^{2}_n}]^{-1}(\frac{n \bar{x}}{\mu_{n}}-n)$$
$$= 2 \mu_n - \frac{\mu^{2}_n}{\bar{x}} $$

You would iterate Newton's method multiple times until it converges and finds the root. In theory, the algorithm would converge perfectly when we find that root - however, we'll never actually reach the root exactly. Why? (Think of Zeno's Paradox.) Thus, we need to either (or both) cap the number of interations or set some convergence criteria. For simplicity's sake, we'll just cap the maximum number of iterations here.

Here, we are going to set a starting point for the NR method of $\mu_0 = 10$. We will also set our iterations to 7, and then plot the score function with the values for each of these iterations. Don't worry so much about this code, but look at the plot. 
```{r, include = TRUE, message = FALSE, warning = FALSE}

#first generate our data (remember to set the seed)

set.seed(123)
x <- rpois(100, 5)
xbar <- mean(x)
n <- length(x)

#our score function (first derivative of the log likelihood)
score <- function(mu) {
        n * xbar / mu - n
}

Funcall <- function(f, ...) f(...)
Iterate <- function(f, n = 1) {
        function(x) {
                Reduce(Funcall, rep.int(list(f), n), x, right = TRUE)
        }
}

single_iteration <- function(x) {
        2 * x - x^2 / xbar #our second derivative here 
}
g <- function(x0, n) {
        giter <- Iterate(single_iteration, n)
        giter(x0)
}

g <- Vectorize(g, "n") #vectorize our iterator with respect to n
iterates <- g(10, 1:7)

ggplot(data.frame(x=c(seq(.35, 10))), aes(x)) + stat_function(fun = score) +
  ylab(expression(paste("Score(", mu, ")"))) + xlab(expression(paste(mu))) +
  geom_hline(yintercept = 2, color = "red", linetype = "dashed") +
  annotate(geom = "vline", x = c(10, iterates),
           xintercept = c(10, iterates), linetype = "dashed") +
  annotate(geom = "text", label = c(as.character(c(0:7))),
           x = c(10, iterates), y = c(rep(1510, 8)),
           angle = 0, vjust = 1, hjust = 2)

```

Here we can see that across the 7 iterations, we get closer and closer to the root (which is 5.1). The algorithm continues (for as many iterations as you specify) until the parameter values stabilize, or $\theta_{n+1}$ and $\theta_{n}$ become approximately equal, and the root is approximated. If this doesn't make sense to you, watch \href{https://www.youtube.com/watch?v=-5e2cULI3H8}{this video}.

We can see that by the 7th NR iteration, we are quite close to the root (5.1).This method is good for finding local optima as well, if you have a multimodal case for example. In those cases, you can specify intervals and find the local optimum (minimum or maximum).

The point here is that you understand the intuition, don't worry so much about performing this yourself. Because, as we will see, there's very easy ways to optimze using certain functions in R, namely `optim()`!

#Using `optim()` to maximize the log likelihood

Although it is useful to learn Newton Raphson, today we will mostly be focusing on learning how to use `R` to optimize for us. Specifically, we'll be using the function `optim()`. To see how `optim()` works, let's use it to optimize the mean of a normally distributed random variable. Note that since we are just optimizing a single parameter here, we'll use the cousin function `optimize()`, but the intuition is the same.

```{r, include = TRUE}

set.seed(123) #first set our seed 
n <-rnorm(100) #generate a normal random variable

fun_2 <- function(x) { #create a function
  y <- dnorm(x) #use the density of a normal distribution
}

plot(n, fun_2(n)) #plot now our "x" against our "y"

#now lets optimize to find the mean of this function!
max <- optimize(fun_2, #call our function
                lower = -4, upper = 4, #set our bounds
                maximum = TRUE) #say we want to maximize

max$maximum #look at that

#plot out
plot(n, fun_2(n)) + abline(v = max$maximum, col = "red")

```

Using this with a normal linear model is not that much more complicated. You can manually optimize your linear regressions in R using the `optim()` function. The `lm()` command engages in some optimization itself, but knowing how to use the `optim()` command itself is useful (and generates some slightly different estimates than `lm()`!). Let's walk through it:

First let's generate some data again:
```{r, include = TRUE}
set.seed(123)
X <- cbind(1, rnorm(100)) #what are we doing here? Binding together with a constant

#What are the parameters of the normal distribution?
theta_true <- c(1,1,1)
#first element is Beta0, the second element is Beta1, and the last element is sigma2

#just another way to simulate data
y <- rnorm(100, mean = X%*%theta_true[1:2], sd = theta_true[3])
```

Now, we will write out the function we want to optimize. In this case, we want to optimize our log likelihood function. That is, we want to find the values of our parameters that maximize this log likelihood function.

```{r, include = TRUE}
#write our log likelihood function
ols.lf <- function(theta, d, X){
  n <- nrow(X) #number of observations
  k <- ncol(X) #number of beta parameters to be estimated
  beta <- theta[1:k] #first k elements of theta are our betas
  sigma <- exp(theta[k+1]) #K+1 element is our sigma 2, exp makes sure
  #this is positive, we can't have a negative
  e <- d - X%*%beta #residuals

  logl <- sum(dnorm(e, mean = 0, sigma, log = TRUE)) #using dnorm again

  return(-logl) #we do negative because the optim fn automatically minimizes.
}

```

Now we will pass this function into `optim()`. If you look at the `optim()` help file, you'll see there a couple of main arguments:

First, `par` which are initial values for the parameters to be optimized over. You can specify these parameters using, for example, the means of the values, but here we'll just use arbitrary values of $`(1,1,1)`$.

Second, `fn` which is where our function is called.

Third, `hessian` which tells it whether or not we want the Hessian matrix to be returned. We do, this is needed for our standard errors.

Fourth, `method` which tells the function which optimization method to use - by default, it uses Nelder Mead. But, we specify "BFGS" which is a quasi-Newton method, similar to the Newton-Raphson algorithm we use above. Unlike the Newton-Raphson, it uses an approximation to the Hessian - the Hessian is often very difficult to compute and invert.

```{r, include = TRUE}
set.seed(123)

init_par <- rnorm(3) #values for our parameters, drawn from the normal
                      #just like our data

p <- optim(par = init_par, #our starting parameters

           ols.lf, #the function we want it to optimize

           hessian = TRUE, #return Hessian matrix (need for se)

           method = "BFGS", #calls a quasi-Newton method

           #control - read about this in the help file

           d = y, #passing our above data into the function
           X = X)

#we need hessian = T for the se
coef <- p$par #contains estimated arg mins
print(paste('Coefficients:', coef))

#the final element is the log of sigma2 
rse <- exp(coef[3]) #standard deviation of conditional distribution
rse

se2 <- sqrt(diag(solve(p$hessian))) #se
print(paste('Standard errors:', se2))

vals <- p$value #log likelihood min
print(paste('Value of ll at minimum', vals))

#did this converge in the min number of iterations?
converge <- p$convergence
converge #yes

```

Now that we have our estimates from optimization, let's see how they compare to the same model but called with `lm()`.
```{r, include = TRUE}
#checking our answer
m1 <- lm(y ~ X[,2])
m1$coef

library(broom) #sweeps things into dataframes
tidy(m1)

#how does this compare?
```

In addition, we could do the above process all in one function. Try filling in the function below and optimizing.
````{r, include = TRUE}
#Could do this all in one function
lm.optim <- function(par, y, X, fn = ols.lf){

  est <- optim(par = init_par, #our starting parameters
           ols.lf,
           hessian = TRUE,
           method = "BFGS",
           d = y, #passing our above data into the function
           X = X)

  pars <- est$par
  ses <- sqrt(diag(solve(est$hessian)))

  return(data.frame(params = pars, se = ses)) #put the optim vs lm estimates into a dataframe

}

lm.optim(init_par, y = y, X = X)

# notice how this doesn't clog up your workspace with objects?
# that's because the function creates an "enclosing environment" that discards 
# the temporary objects formed during function execution
# the function returns either the output of its last line OR whatever output 
# is specified using return() - using return() explicitly is good practice

```

\newpage
#Practice on your own:
Now, on your own, try doing this with `mtcars` data, instead of generated data. Let's say we want to set our $y$ value as `mpg`, $x1$ as `wt`, and $x2$ as `gear`.

```{r, eval = FALSE, include = TRUE}

y2 <- mtcars$______
X2 <-cbind(1,______,________)

#in practice, you can use other ways to select your starting values
#such as a grid search, but we will stick with selecting from a normal distribution
set.seed(2030)
params <- rnorm(4)

#now for the optim() part!
p2 <- optim(par = ____, #our starting parameters

           _____, #the function we want it to optimize

           ______, #return Hessian matrix (need for se)

           _______, #calls a quasi-Newton method

           d = y2, #passing our above data into the function
           X = X2)

#we need hessian = T for the se
coef2 <- ____
_____

#the final element is the log of sigma2 
rse2 <- ____ #standard deviation of conditional distribution
____

se2_2 <- _____ #se
____

vals <- _____ #log likelihood min
____

#did this converge in the min number of iterations?
converge <- ______
_________

#compare to lm() model
m2 <- lm(mpg ~________
__________

```

### 2023-01-04 Lab 5  Optim/Lab5_Optim_AnswerKey.Rmd
Add files via upload

--- 2023-01-04 Lab 5  Optim/Lab5_Optim.Rmd
@@
 ---
-title: "Lab 5: Optimization"
+title: "Lab 5: Optimization: Answers"
 author: "Eric Parajon"
@@
 
+#Installing and loading packages for use
 pacman::p_load(tidyverse,broom)
 ```
-
-Some of today's code has been adapted from Simon Hoellerbauer, Isabel Laterzo and Sean Norton.
 
@@
 
-Let's suppose we're "hiking" up that mathematical hill trying to find the maximum, and using the first derivative to decide how big of a "step" we take from our current location.
+Let's suppose we're "hiking" up that mathemetical hill trying to find the maximum, and using the first derivative to decide how big of a "step" we take from our current location.
 However, the first derivative only gives us information on the steepness of that hill.
@@
 Don't worry so much about this code, but look at the plot.
+
 ```{r, include = TRUE, message = FALSE, warning = FALSE}
@@
 #first generate our data (remember to set the seed)
-
 set.seed(123)
-x <- rpois(100, 5)
-xbar <- mean(x)
-n <- length(x)
+x <- rpois(100, 5) #sample from a poisson
+xbar <- mean(x) #mean of the poisson samples
+n <- length(x) #n obs
 
@@
 If this doesn't make sense to you, watch \href{https://www.youtube.com/watch?v=-5e2cULI3H8}{this video}.
-
 We can see that by the 7th NR iteration, we are quite close to the root (5.1).This method is good for finding local optima as well, if you have a multimodal case for example.
@@
   logl <- sum(dnorm(e, mean = 0, sigma, log = TRUE)) #using dnorm again 
+                                               #but note the differences
 
@@
 }
-
 ```
@@
 ```{r, include = TRUE}
-set.seed(123)
+set.seed(123118)
 
-init_par <- rnorm(3) #values for our parameters, drawn from the normal
-                      #just like our data
+init_par <- rnorm(3) #values for our paramters
 
@@
 converge #yes
-
 ```
@@
 In addition, we could do the above process all in one function. Try filling in the function below and optimizing.
-````{r, include = TRUE}
+
+```{r, include = TRUE}
+
 #Could do this all in one function
@@
 
-\newpage
-#Practice on your own:
+# Practice on your own:
 Now, on your own, try doing this with `mtcars` data, instead of generated data. Let's say we want to set our $y$ value as `mpg`, $x1$ as `wt`, and $x2$ as `gear`.
@@
 
-y2 <- mtcars$______
-X2 <-cbind(1,______,________)
+y2 <- mtcars$mpg
+X2 <-cbind(1, mtcars$wt, mtcars$gear)
 
 #in practice, you can use other ways to select your starting values
-#such as a grid search, but we will stick with selecting from a normal distribution
+#such as a grid search, but we will stick with selecting from a normal distrib
 set.seed(2030)
@@
 
-#now for the optim() part!
-p2 <- optim(par = ____, #our starting parameters
+p2 <- optim(par = params, #our starting parameters
 
-           _____, #the function we want it to optimize
+           ols.lf, #the function we want it to optimize
 
-           ______, #return Hessian matrix (need for se)
+           hessian = TRUE, #return Hessian matrix (need for se)
 
-           _______, #calls a quasi-Newton method
+           method = "BFGS", #calls a quasi-Newton method
 
@@
 #we need hessian = T for the se
-coef2 <- ____
-_____
+coef2 <- p2$par #contains estimated arg mins
+coef2
 
 #the final element is the log of sigma2 
-rse2 <- ____ #standard deviation of conditional distribution
-____
+rse2 <- exp(coef2[4]) #standard deviation of conditional distribution
+rse2
 
-se2_2 <- _____ #se
-____
+se2_2 <- sqrt(diag(solve(p2$hessian))) #se
+se2_2
 
-vals <- _____ #log likelihood min
-____
+vals2 <- p2$value #log likelihood min
+vals2
 
 #did this converge in the min number of iterations?
-converge <- ______
-_________
+converge2 <- p2$convergence
+converge2 #yes
 
 #compare to lm() model
-m2 <- lm(mpg ~________
-__________
-
+m2 <- lm(mpg ~ wt + gear, data = mtcars)
+summary(m2)
 ```

### 2023-01-04 Lab 6  Diagnostics and simulation based inferences/Lab6_Simulation_Uncertainty.Rmd
Add files via upload

---
title: "Lab 6 - Diagnostics and simulation based inferences "
author: "Eric Parajon"
output: pdf_document
---

This lab has been adapted from code by Simon Hoellerbauer and Isabel Laterzo.

So far in lab, we have often simulated data to allow us to test out various methods. Simulation can help us do much more, as well. Simulations are a useful tool that is being used more and more often within the discipline. They can help us:
1) Construct a hypothetical DGP and then see if the methods we use/develop can successfully retrieve the parameters of that DGP
2) Make our data go further with re-sampling 
3) Create plots to help us assess our results when our variance/covariance matrix gets complex (when distributional assumptions can be tricky)

Today, we will use simulation to help us assess the uncertainty around our parameter estimates and will then incorporate this uncertainty into the predicted values of our outcome variable.

We will be examining **Regression Diagnostics**, or ways to examine if our regression meets necessary assumptions, and if anything weird (e.g., outliers, etc.) is going on. Remember the four key assumptions of a linear regression:

1) Linearity of the data - the relationship between x and y is linear

2) Normality of residuals - the residual errors are normally distributed

3) Homogeneity of residual variance - the residuals have CONSTANT variance (homoscedasticity)

4) Independence of residual error terms

The following tests will help us see if these assumptions hold, and will examine a number of possible issues. These include (but are not limited to) non-linearity of the data, heteroscedasticity (non-constant variance in the errors), and presence of influential values (outliers and high-leverage points).

# Simulation
```{r, message = FALSE}
#clear global environ
rm(list=ls())

#the packages we will need

#Installing and loading packages for use
pacman::p_load(MASS,#required for multivariate normal
               tidyverse,
               AER) #For guns data

```

Download the "aid_data_95_ee.csv" in the Labs folder on Sakai and then read it into R:
```{r, warning = F, message = F}
#remember to set your working directory!
setwd("______")

#read_csv comes from readr from the tidyverse suite of packages
aid_data_95_ee <- read_csv("aid_data_95_ee.csv")
#also look into fread()
```

Let's take a look at this data very quickly:
```{r, warning = F, message = F, include = F}
#checking dimensions
dim(aid_data_95_ee)
#that's a lot of variables!

#which variables?
names(aid_data_95_ee)

#which countries?
unique(aid_data_95_ee$recipient)

#for which years?
unique(aid_data_95_ee$year)

#how many countries in years? dplyr can help us here
aid_data_95_ee %>% group_by(recipient) %>% tally
```

We are going to see how civil society strength in Eastern Europe varies according to lagged civil society assistance, EU accession leverage, and lagged GDP per capita.

Let's look at the distribution of the outcome variable (with some simple, barebones graphs):
```{r, warning = FALSE}
#distribution of civil society strength
ggplot(data = aid_data_95_ee, aes(x = civ_soc_strength)) + 
  geom_density()
#Is it necessarily a problem that our outcome variable is skewed?

#what does our main predictor of interest look like?
ggplot(data = aid_data_95_ee, aes(x = civ_soc_aid_l1_log)) + 
  geom_density()
#why do you think we log civil society aid?

#let's look at relationship between the two
ggplot(data = aid_data_95_ee, aes(x = civ_soc_aid_l1_log, y = civ_soc_strength)) +
  geom_point()

#what about EU accession?
ggplot(data = aid_data_95_ee, aes(x = EU_accession)) + geom_bar()
```

Let's fit our model:
```{r, message = F, warning = F, results = "hide"}
#first we have to make sure that EU_accession is a factor, and that the 
#reference category makes sense for us
aid_data_95_ee$EU_accession <- factor(aid_data_95_ee$EU_accession) %>%
  relevel(ref = "no leverage")
levels(aid_data_95_ee$EU_accession) #First one will be the reference catagory

#Estimating the model
cs_model <- lm(civ_soc_strength ~ GDPpc_ppp_WB_2011_l1_log +
                 EU_accession*civ_soc_aid_l1_log, data = aid_data_95_ee)
summary(cs_model)
```

Now simulation comes in! We can use simulation to get confidence intervals. Here, we will be using the function `mvrnorm` to simulate some data. Look at the help file for that function and what it does. 
```{r}
#set seed for replication
set.seed(123)

#We will take 500 samples from the sampling distribution
#of our model's coefficients
samp_beta <- mvrnorm(500,
                     coef(cs_model),
                     vcov(cs_model))
dim(vcov(cs_model))
dim(samp_beta)

# 87% Monte Carlo/simulation based confidence interval:
t(apply(samp_beta,
        2, #why 2? look at help file and the "MARGIN" argument
        quantile,
        probs = c(0.065, 0.935)))
```

Let's compare that to CIs generated from the function `confint()`. Both produce asymptotic confidence intervals, but the first uses resampling/Monte Carlo methods. I suggest using these methods, as it is easier for us to get CIs for slightly more complicated models via simulation.

```{r}
confint(cs_model, level = 0.87)
```

We are now going to use the "average value approach" to simulating predicted outcomes and uncertainty around them. We want to see how predicted civil society strength varies according to lagged civil society assistance. In order to do this, there are a few thing we have to do first:
```{r, results = "hide"}
# we are going to let lagged log civil society assistance vary along its 
# inner-quartile range. These will be the values over which we plot the predicted
# civil society strength
cs_aid_sim <- with(aid_data_95_ee,
                   #quantile() here produces sample quantities based on the given
                   #probability we provide it, here 0.25
                   seq(quantile(civ_soc_aid_l1_log, 0.25, na.rm = T),
                       quantile(civ_soc_aid_l1_log, 0.75, na.rm = T),
                       length.out = 100))

# we now need to create a matrix of hypothetical predictors. We will use the 
# mean of lagged log GDP per capita; in order to see the effect of EU accession
# leverage, we will also allow that to vary (otherwise, you usually use the mode
# of categorical variables)
x_mat_sim <- with(aid_data_95_ee,
                  model.matrix(~ GDPpc_ppp_WB_2011_l1_log + 
                                 EU_accession*civ_soc_aid_l1_log,
                               data = data.frame(GDPpc_ppp_WB_2011_l1_log = 
                                        mean(GDPpc_ppp_WB_2011_l1_log, 
                                             na.rm = T), #mean of lagged log GDP
                                      EU_accession = relevel(
                                        factor(c("no leverage",
                                                  "active", # vary EU accession
                                                  "member")), 
                                        ref = "no leverage"),
                                      #calling above data for civ soc assist
                                      civ_soc_aid_l1_log = rep(cs_aid_sim,
                                                               each = 3))))

#let's check dimensions
dim(x_mat_sim)

#how does it look? note the interaction effects included!
head(x_mat_sim)

```

Now we have to do XBeta to get our predicted values! But instead of beta being a k x 1 vector, it is now a k x 500 (number of samples from our multivariate normal) matrix. So, what will the dimensions of this new object be? 
```{r}
cs_str_hat <- x_mat_sim %*% t(samp_beta)

dim(cs_str_hat) #300 x 500

#we only need quantiles for our confidence intervals
cs_str_hat_quant <- apply(cs_str_hat, 1, #why 1? again look at MARGIN
                      quantile, c(0.065, 0.5, 0.935))

#creating data frame of predictions
pred_df <- data.frame(PCS = cs_str_hat_quant[2, ], #why calling 2, 3, and 1?
                      UB = cs_str_hat_quant[3, ],
                      LB = cs_str_hat_quant[1, ],
                      cs_aid = x_mat_sim[, "civ_soc_aid_l1_log"],
                      EU_Leverage = rep(c("no leverage", "active", "member"), 
                                         100)) #why can we do this?

#Creating data for rug plot
cs_aid_IQ <- dplyr::select(aid_data_95_ee, civ_soc_aid_l1_log, EU_accession) %>%
  filter(civ_soc_aid_l1_log > quantile(civ_soc_aid_l1_log, 0.25, na.rm = T) &
           civ_soc_aid_l1_log < quantile(civ_soc_aid_l1_log, 0.75, na.rm = T)) %>%
  rbind(data.frame(civ_soc_aid_l1_log = rep(NA, nrow(pred_df) - nrow(.)),
                  EU_accession = rep(NA, nrow(pred_df) - nrow(.))))
```
Now let's graph this!

```{r}
#a library with different colors for ggplot2
library("RColorBrewer")

ggplot(data = pred_df, aes(x = cs_aid, y = PCS, lty = EU_Leverage)) +

  geom_ribbon(aes(ymin = LB, ymax = UB, color = EU_Leverage),
              alpha = 0.5,
              fill = "gray70") +

  geom_line(aes(color = EU_Leverage)) +
  #what do you think this does?

  geom_rug(data = cs_aid_IQ, aes(x = civ_soc_aid_l1_log, color = EU_accession), 
           size = 1, inherit.aes = F) +
  #rug plot

  scale_color_brewer(palette = "Dark2") +
  #if you want to have some fun, change around the colors
  #scale_color_brewer(palette = "Set3") +
  #scale_color_brewer(palette = "Greys") +
  #scale_color_brewer(palette = "PuOr") +
  #scale_color_brewer(palette = "Set1") +

  xlab("Lagged Civil Society Assistance, Millions of Dollars, Log") +
  ylab("Predicted Civil Society Strength") +
  #labels

  theme_bw()
  #theme
```
# Regression Diagnostics

First we'll talk about (statistical) leverage! Points that are unusual (aka they are \textit{outliers}) with respect to the \textit{predictors}---that is, all of the $\mathbf{X}$---are said to have high \textit{leverage} because they can impact our coefficients. One way to assess this is called ``Cook's Distance," or a measure of an observation's influence.

We can write a function for Cook's Distance that computes a standard regression diagnostic and plots the results all in one. Useful if we want flexibility with different models and different observations!

$$D_i=\frac{\sum_{j=1}^{n}(\hat{y_{j(i)}}-\hat{y_j})^2}{k\hat{\sigma}^2}$$

Where $k$ is our number of coefficients, $\sigma^2$ is the variance of our model, and the numerator is an estimated mean of $\hat{y}$ when removing observation $i$.

When we have multiple regressors in our model (and with an eye to using R for computing Cook's D), it's useful to write this formula as such:

$$D_i=\frac{e_{i}^2}{k\hat{\sigma}^2}\frac{h_{i}}{(1-h_i)^2}$$

Where $h_{i}$ is the $i$'th diagonal element of the hat matrix - and each observation's leverage.

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=20)}
m1 <- lm(mpg ~ disp + wt, data = mtcars) #simple regression model

#Now, the function
cooks <- function(LM, x = "all") {
  #browser()
  #2 arguments
  if (is.character(x) && x == "all") x <- 1:nrow(model.matrix(LM))

  resid <- LM$residuals 
  k <- length(LM$coefficients) #number of coefficients estimated
  num_1 <- resid^2 #squared residuals
  denom_1 <- mean(resid^2) * k #setting up the numerator and denominator for Cook's Dist
  num_2 <- hatvalues(LM)
  denom_2 <- (1 - hatvalues(LM))^2
  cooks_d <- ((num_1 / denom_1) * (num_2 / denom_2))[x] #Calculate Cook's D
  plot_dat <- data.frame(x = x, 
                            cooks_d = cooks_d) #x allows for flex in the # of obsvs
  cooks_plot <- ggplot(plot_dat, aes(x = x, y = cooks_d)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = (4 / nrow(model.matrix(LM))), colour = "red") +
  labs(x = "Observation", y = "Cook's Distance") +
  theme_bw() #Plot with a standard cutoff for concern
  return(list(plot = cooks_plot, data = plot_dat))
}

#Apply function for our model and 1st 20 observations
test <- cooks(m1, 1:17)
test$plot
test$data
cooks(m1, "all")$plot
```

This is one rule of thumb with Cook's distance: looking at points above 4/$n$, where $n$ is our number of observations. Sometimes, you'll see this as 4/$n-k$.

Although it is useful to be able to write our own Cook's Distance function, do note there are pre-existing ``canned" functions. Your models won't always be compatible with these functions, but in simple cases they are useful.

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=20)}
#cooks.distance function from base R

test2 <- as.data.frame(cooks.distance(m1)) #make this into a df
test2$ident <- seq(1, nrow(test2)) #add observation index
colnames(test2) <- c("cooksd", "obs") #add meaningful column names

#plot
cooks_plot <- ggplot(test2, aes(y = cooksd, x = obs)) +
  geom_bar(stat= "identity") +
  geom_hline(yintercept = (4 / nrow(test2)), colour = "red")

cooks_plot

```

**More Assumptions**

Another assumption we may want to check is that our residuals (errors) are normally distributed (normality of errors assumption). We can do this by separating our residuals into quantiles and comparing their distribution to points produced randomly by a normal distribution (a q-q plot).

If both sets of quantiles came from the same distribution, we should see the points forming a line that's roughly straight.

```{r, eval=TRUE, echo=TRUE, results='asis', message=FALSE}
ggQQ <- function(LM) { # argument: a linear model
     y <- quantile(LM$resid[!is.na(LM$resid)], c(0.25, 0.75))
     x <- qnorm(c(0.25, 0.75))
     slope <- diff(y)/diff(x)
     int <- y[1] - slope * x[1]
     p <- ggplot(LM) +
     stat_qq(aes(sample=.resid), alpha = 0.5) +
     geom_abline(slope = slope, intercept = int, color="blue")
     return(p)
    }

ggQQ(m1)
```

As stated before, it's often useful to be able to write these functions ourselves. Moving forward (particularly with your own data!) you might be in a situation with missingness, multilevel data, or other data quirks, and 'canned' functions aren't always able to give us reliable regression diagnostics.

That being said, in addition to the Cook's Distance canned function shown before, there are some other useful functions to be aware of: like `tidy` and `augment` from the `broom` package.

1) `tidy` converts your model into a tidy tibble, and provides useful estimates about your model components. This includes the coefs and p-values for each term in the regression. It is similar to `summary`, but in a new format that is easier to work with, particularly if you want to do further analysis with your data.

2) `augment` adds columns to the original data that was modeled, providing information about predictions, residuals, and Cook's Distance! As you can see, it shows this information for each type of the original points in your regression, in this case the type of car. This function *augments* the original data with more information from the model.

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=20)}
library(broom)
tidy(m1)
head(augment(m1))
```

Neat, huh? So, this means that for our plot above (Cook's distance) we could also do the following: 
```{r, tidy=TRUE, tidy.opts=list(width.cutoff=20)}
vals <- augment(m1)

ggplot(data=vals, aes(seq_along(.cooksd),y=.cooksd))+
    geom_bar(stat="identity")+
    geom_hline(yintercept=4/length(vals$.cooksd), color="red")+
    labs(x="Observation", y="Cook's Distance") +
    theme_bw()
```

Other 'canned' functions that can be useful are found in the `stats` package and the `car` package (although most of these values can be gotten just by using `augment`):

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=20), message = FALSE, results = "hide"}
library(stats)

#Another cook's distance
cooks.distance(m1)

#dfbetas
dfbetas(m1)

#or, to get them all:
influence.measures(m1)
#note that this "marks" influential observations for you
#with an asterisk

library(car)
#variance inflation factors
#quantifies the severity of multicollinearity
vif(m1)
sqrt(vif(m1)) > 2 #look into this, is this a problem?

```

Other useful diagnostics include fitted value vs. residual plots and component + residual plots. The `plot` function allows you to get 4 standard diagnostic plots easily.

```{r}
plot(m1)
```

1) The Residuals vs Fitted values plot can help us with two things: 1) it can help us see if the constant variance assumption is appropriate (most often, if it isn't we would see a "funnel" in one direction or another), and 2) it can tell us whether we might need to transform the outcome variable in some way (if the line of best fit isn't reasonably straight.) In an ideal scenario, the red line should show no fitted pattern and be approximately horizontal at zero. Is this the case here?

2) We discussed the Q-Q plot above! It helps us visually check the normality assumption. How does this one look?

3) The Scale-Location plot is basically the first plot folded in half and compressed. It doesn't tell us much more, to be quite honest, but it does make it more clear if our residuals are spread equally along the range of our predictors. It is good to see a horizontal line with all points spread out equally. Is this the case here?

4) The Residuals vs Leverage plot, as the name suggests, plots the standardized residuals of each observation against its leverage (remember only if both are high do we worry about how it affects our model). You can't really see it here, but R plots .5 and 1 Cook's Distance "contours" on this plot as well, which indicates where Cook's Distances of .5 and 1 would be. Some people say a Cook's Distance of $>1$ is concerning, but some people say that is extremely conservative. In addition, you'll see in this plot it identifies the top 3 most extreme points - Toyota Corolla, Fiat, and Chrysler Imperial - which have standardized residuals > 2. As a rule of thumb, some say that observations with a standardized residual > 3 are concerning and possible outliers - we don't see any here. Good news!

Handily, all of these plots point out observations of which you may want to be wary.

And to wrap up, a couple other tests which will be useful for you to know!

1) The Durbin Watson test to see if we have autocorrelated errors (violating independence of errors). The DW statistic will always be between 0 and 4, with a value of 2 signifying that there is NO autocorrelation detected in your sample. Values < 2 means there is positive autocorrelation, values > 2 indicate there is negative autocorrelation.

2) Shapiro-Wilk's normality test. This test, generally speaking, examines normality. We can apply it to our residual terms to check for normality of errors. If the p value is > 0.05 the distribution of the data is NOT significantly different from the normal distribution (the normality assumption holds).

```{r, tidy=TRUE, tidy.opts=list(width.cutoff=30)}
#durbin watson
durbinWatsonTest(m1) #is this a problem?

#shapiro test
shapiro.test(m1$residuals) #how does it look?
```

# Practice
To do on your own:

Take the `Guns` dataset in the `AER` package. Take a look at the dataset and use `?Guns` to get the codebook. 
Regress the robbery rate on the percent of the state population that is male, the real per capita personal income in the state, and then an interaction between the population density and whether a shall carry law was in effect that year.

Use Monte Carlo simulation with 10,000 samples from the sampling distribution to get 80% confidence intervals around the coefficient estimates.

Then, use simulation to plot the predicted robbery rate as the population density varies from its 10th percentile to its 90th percentile dependent on if a shall carry law is in effect (don't worry about the rug plot).

```{r, eval = FALSE}
data(Guns)

#estimate model
guns_model <- lm(_________)

set.seed(123)

#sampling betas
samp_beta_guns <- mvrnorm(____, 
                          coef(____),
                          vcov(____))

#confidence intervals
t(apply(____, 2, quantile, prob = c(.1, .9)))
confint(guns_model, ____)

#getting range of density
density_sim <- with(Guns,
                    seq(quantile(density, .1), quantile(density, .9),
                        length.out = 100))

#hypothetical predictor matrix
 X_sim_guns <- with(Guns,
                   model.matrix(~ ____,
                      data = data.frame(male = mean(male),
                                        income = mean(income),
                                        law = c(____),
                                        density = rep(density_sim, each = 2))
                   ))

robbery_hat <- X_sim_guns %*% t(samp_beta_guns)

#we only need quantiles for our confidence intervals
robbery_hat_quant <- apply(robbery_hat, 1,
                      quantile, c(________))

pred_df_guns <- data.frame(PP = robbery_hat_quant[2, ],
                      UB = robbery_hat_quant[3, ],
                      LB = robbery_hat_quant[1, ],
                      Density = X_sim_guns[, "density"],
                      Law = c("No", "Yes")[X_sim_guns[, "law"] + 1])

ggplot(data =____) +
  geom_ribbon(aes(____fill = Law), alpha = 0.5) +
  geom_line()

```

### 2023-01-04 Lab 6  Diagnostics and simulation based inferences/Lab6_Simulation_Uncertainty_Answers.Rmd
Add files via upload

--- 2023-01-04 Lab 6  Diagnostics and simulation based inferences/Lab6_Simulation_Uncertainty.Rmd
@@
 ---
-title: "Lab 6 - Diagnostics and simulation based inferences "
-author: "Eric Parajon"
+title: "Lab 6 - Diagnostics and simulation based inferences: Answers"
+author: "TA: Eric Parajon"
 output: pdf_document
@@
 
-We will be examining **Regression Diagnostics**, or ways to examine if our regression meets necessary assumptions, and if anything weird (e.g., outliers, etc.) is going on.
-Remember the four key assumptions of a linear regression:
-
-1) Linearity of the data - the relationship between x and y is linear
-
-2) Normality of residuals - the residual errors are normally distributed
-
-3) Homogeneity of residual variance - the residuals have CONSTANT variance (homoscedasticity)
-
-4) Independence of residual error terms
-
-The following tests will help us see if these assumptions hold, and will examine a number of possible issues.
-These include (but are not limited to) non-linearity of the data, heteroscedasticity (non-constant variance in the errors), and presence of influential values (outliers and high-leverage points).
-
-# Simulation
-```{r, message = FALSE}
+```{r}
 #clear global environ
@@
                tidyverse,
-               AER) #For guns data
-
+               AER)
 ```
@@
 #remember to set your working directory!
-setwd("______")
 
-#read_csv comes from readr from the tidyverse suite of packages
+#read_csv comes from readr from the tidyverse
 aid_data_95_ee <- read_csv("aid_data_95_ee.csv")
-#also look into fread()
 ```
@@
 Let's take a look at this data very quickly:
-```{r, warning = F, message = F, include = F}
+```{r}
 #checking dimensions
@@
 Let's look at the distribution of the outcome variable (with some simple, barebones graphs):
-```{r, warning = FALSE}
+```{r}
 #distribution of civil society strength
@@
 Let's fit our model:
-```{r, message = F, warning = F, results = "hide"}
+```{r}
 #first we have to make sure that EU_accession is a factor, and that the 
@@
 In order to do this, there are a few thing we have to do first:
-```{r, results = "hide"}
+```{r}
 # we are going to let lagged log civil society assistance vary along its 
@@
                                  EU_accession*civ_soc_aid_l1_log,
+
                                data = data.frame(GDPpc_ppp_WB_2011_l1_log = 
@@
                                              na.rm = T), #mean of lagged log GDP
+
                                       EU_accession = relevel(
@@
                                         ref = "no leverage"),
+
                                       #calling above data for civ soc assist
@@
 head(x_mat_sim)
-
 ```
@@
 
-dim(cs_str_hat) #300 x 500
+dim(cs_str_hat)
 
@@
 
-#Creating data for rug plot
+#for rug plot
 cs_aid_IQ <- dplyr::select(aid_data_95_ee, civ_soc_aid_l1_log, EU_accession) %>%
@@
 ```
+
 Now let's graph this!
-
 ```{r}
@@
 ```
+
 # Regression Diagnostics
@@
 
-# Practice
-To do on your own:
+# To do on your own:
 
@@
 
-```{r, eval = FALSE}
+```{r}
+library(AER)
 data(Guns)
@@
 #estimate model
-guns_model <- lm(_________)
+guns_model <- lm(robbery ~ male + income + density*law,
+                 data = Guns)
 
-set.seed(123)
+set.seed(53)
 
 #sampling betas
-samp_beta_guns <- mvrnorm(____, 
-                          coef(____),
-                          vcov(____))
+samp_beta_guns <- mvrnorm(10000, 
+                          coef(guns_model),
+                          vcov(guns_model))
 
 #confidence intervals
-t(apply(____, 2, quantile, prob = c(.1, .9)))
-confint(guns_model, ____)
+t(apply(samp_beta_guns, 2, quantile, prob = c(.1, .9)))
+confint(guns_model, level = .8)
 
@@
  X_sim_guns <- with(Guns,
-                   model.matrix(~ ____,
+                   model.matrix(~ male + income + density*law,
                       data = data.frame(male = mean(male),
                                         income = mean(income),
-                                        law = c(____),
+                                        law = c(0, 1),
                                         density = rep(density_sim, each = 2))
@@
 #we only need quantiles for our confidence intervals
-robbery_hat_quant <- apply(robbery_hat, 1,
-                      quantile, c(________))
+robbery_hat_quant <- apply(robbery_hat, 1, #why 1?
+                      quantile, c(0.1, 0.5, 0.9))
 
@@
                       Density = X_sim_guns[, "density"],
-                      Law = c("No", "Yes")[X_sim_guns[, "law"] + 1])
+                      Law = c("No", "Yes")[X_sim_guns[, "law"] + 1]) #why can we do this?  `
 
-ggplot(data =____) +
-  geom_ribbon(aes(____fill = Law), alpha = 0.5) +
-  geom_line()
+ggplot(data = pred_df_guns, aes(x = Density, y = PP, lty = Law)) +
+  geom_ribbon(aes(ymin = LB, ymax = UB, fill = Law), alpha = 0.5) +
+  geom_line() + #what do you think this does?
+  xlab("Population Density") +
+  ylab("Predicted Number of Robberies") +
+  theme_bw()
 

### 2023-01-04 Lab 8  BinomialModels/Lab8_BinomialModels.Rmd
Add files via upload

---
title: 'Lab 7: Missing Data and Imputation:Answers'
author: "Eric Parajon"
header-includes:
  - \usepackage{enumerate}
  - \usepackage{graphicx}
output: pdf_document
---

Today's lab has been adapted from code by Chelsea Estancona, Isabel Laterzo, and Simon Hoellerbauer.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)

rm(list=ls()) # clear workspace
#setwd("____") # change to wherever you saved the data
set.seed(18765) # set seed for replication

#Load in packages
pacman::p_load(tidyverse,
               rio,#for importing data
               arm)#inverse link function

```

# Binomial Models

Today we'll work through a fairly standard problem: you have a dichotomous dependent variable and a set of independent variables, and you want to estimate a model. Let's walk through a full example.

Load the `Lab_8_data.dta` file. These are data from a previous project (Chelsea Estancona ('18)'s) about paramilitary groups in Colombia. Our DV is a dichotomous variable coding paramilitary presence in certain municipalities.

```{r}
colombia <- rio::import("Lab_8_data.dta")

```

Let's plot our DV. This is always a good practice for *all* variables (independent or dependent) in your work.

```{r}
ggplot(colombia, aes(x = as.factor(AUC))) +
  geom_bar() +
  labs(x = 'AUC Presence') +
  theme_bw()

```

Note that plotting in this way can also give us information about missingness (which there's a lot of on this variable). Remember - what does R do with our missing data if we leave it to handle it on its own?

Now we're ready to put together a simple model.

```{r}
#for a logit model, how should we specify family?

m1 <- glm(AUC ~ tribut_2_log + tribut_2_log_sq + pop_attacks + disbogota +
            drugs + pop_log, data=colombia, family = __________)

summary(m1)
```

If we wanted to use a probit link:

```{r}
#for a probit model, how should we specify family?

m2 <- glm(AUC ~ ., data = colombia, family = ___________)
#what is the . doing above?
summary(m2)
```

The sign (and generally, significance) of our coefficient estimates remain the same. The values differ due to these link functions.

This is all very well and good, but how useful are the coefficient estimates? How do we interpret them?

We might want to use our model to estimate predicted probabilities, as this is a readily-understood measure of how changes in our independent variable affect the probability of observing a certain outcome. In doing so, we'll want to think about a 'hypothetical case' for which we want to predict: for example, here, we'll calculate the predicted probability of the United Self-Defenses of Colombia or Autodefensas Unidas de Colombia (AUC) presence in a municipality with the mean tax income, mean attacks on population, mean distance from the capitol, drug crops, and the average population. *Note*: This is the so-called average case approach.

```{r}
## calculate value of the linear predictor for this variable profile
linpred <- as.numeric(with(colombia, coef(m1)[1] +
                             coef(m1)[2] * mean(tribut_2_log, na.rm = T) +
                             coef(m1)[3] * mean(tribut_2_log, na.rm = T)^2 +
                             coef(m1)[4] * mean(pop_attacks, na.rm = T) +
                             coef(m1)[5] * mean(disbogota, na.rm = T) +
                             coef(m1)[6] * (1) + #could set this to its mode
                             coef(m1)[7] * mean(pop_log, na.rm = T)))

#let's take a look!
invlogit #no parentheses from the arm package

## invlogit lets us transform our log-odds to the range (0,1) - now interpretable
## as a standard probability
invlogit(linpred)
```
We can also do this using the 'predict' function:

```{r}
## create a 'new data' frame: we want to predict the probability for a hypothetical situation.
new_data <- with(colombia, data.frame(tribut_2_log = mean(tribut_2_log, na.rm = T),
                                      tribut_2_log_sq = mean(tribut_2_log, na.rm = T)^2,
                                      pop_attacks = mean(pop_attacks, na.rm = T),
                                      disbogota = mean(disbogota, na.rm = T),
                                      drugs = c(0, 1),
                                      pop_log = mean(pop_log, na.rm = T)))
## note above that I include both possible values for drugs. the 'data.frame()'
## command is flexible and will intelligently duplicate the other values

#Here use predict()
predict(_____, newdata = ________, type = 'response') #The type="response" option tells R to output probabilities of the form P(Y = 1|X)
```

What if we were interested in a *range* of values' predicted probability? We could modify our newdata frame and our predict() function to allow for this.

```{r}
min_tax <- min(colombia$tribut_2_log, na.rm = T)
max_tax <- max(colombia$tribut_2_log, na.rm = T)
min_tax_2 <- min(colombia$tribut_2_log_sq, na.rm = T)
max_tax_2 <- max(colombia$tribut_2_log_sq, na.rm = T)

new_data_2 <- with(colombia,
                 data.frame(tribut_2_log = seq(from = min_tax, to = max_tax,
                                               length.out = 100),
                            tribut_2_log_sq = seq(from = min_tax_2, to = max_tax_2,
                                                  length.out = 100),
                            pop_attacks = mean(pop_attacks, na.rm = T),
                            disbogota = mean(disbogota, na.rm = T),
                            drugs = 1, pop_log = mean(pop_log, na.rm = T)))

new_data_3 <- with(colombia,
                 data.frame(tribut_2_log = seq(from = min_tax, to = max_tax,
                                               length.out = 100),
                            tribut_2_log_sq = seq(from = min_tax_2, to = max_tax_2,
                                                  length.out = 100),
                            pop_attacks = mean(pop_attacks, na.rm = T),
                            disbogota = mean(disbogota, na.rm = T),
                            drugs = 0, pop_log = mean(pop_log, na.rm = T)))

predict(m1, newdata = new_data_2, type = 'response') # what is the length of this output?
predict(m2, newdata = new_data_3, type = 'response')

#we can calculate the logit predicted probs by hand by doing 1/(1 + exp(-linpred))
#How could we calculate this for probit?

#We can use pnorm! (this is, after all, the difference between the logit and probit)
all(predict(m2, newdata = new_data_3, type = "response") ==
      pnorm(as.matrix(cbind(1, new_data_3)) %*% coef(m2)))
```

Note that in this case, it is perfectly ok to vary the tax variable and the the squared tax value from its min to its max, as the square roots of the minimum and maximum squared are equal to the minimum and maximum (unless for some reason you have missingness in one variable at the max and min, which should be unlikely).

We know from the numeric output that logit and probit links produce different coefficient estimates, but how do these differences translate into predicted probabilities? We can plot the predicted probabilities from each model to quickly inspect these differences.

```{r}
## generate predicted probabilities for municipality with drugs and all other
## variables at their means
pred_output <- data.frame(Logit = predict(_____, newdata = _______, type = _______),
                          Probit = predict(____, newdata = ________, type = ______),
                          tax = new_data_2$tribut_2_log)

plot_pred1 <- gather(pred_output, key = "link", value = "Pred_Probs", -tax)

## plot predicted probabilities
ggplot(data = plot_pred1, aes(x = tax, y = Pred_Probs, color = link)) +
  geom_line() +
  scale_color_discrete(name = 'Link Function') +
  labs(x = 'Tax Revenue', y = 'Predicted Probability') +
  theme_bw()
```

# Simulation

Simulation strikes again! This time around, we will do similar things, but with non-Normal outcomes, where simulations are even more helpful.

Today, we're going to simulate predicted probabilities for a model with an interaction term and plot our results. We will use both the average case and the observed-case approaches.

## Using Average-Case Approach

The following should all look familiar:
```{r}
## number of simulations for predicted probabilities
N_sim <- 1000

## extract beta hats and construct vcv from our first model
betas <- m1$coef
vcv <- vcov(m1)

## use the mvrnorm function with the betas, vcv, and # of simulations to simulate
## parameter estimates. will have 1,000 estimates of each beta
beta_sims <- mvrnorm(N_sim, betas, vcv)

## create a vector that is a sequence of 100 points, evenly spaced between the
## lowest and highest observed values of tax income in the data set. we have to
## do this for the (tax^2) variable, too!
tax_min <- with(colombia, min(tribut_2_log, na.rm = T))
tax_max <- with(colombia, max(tribut_2_log, na.rm = T))
tax_vec <- seq(from = tax_min, to = tax_max, length = 100)
trans_vec <- (tax_vec^2)
## these will be the x values over which we can plot the predicted probability.
## we're interested in how the predicted probability of AUC presence varies over
## the municipalities' values of tax revenue

## create a hypothetical independent variable as before: mean attacks on population,
## mean distance from the capitol, drug crops set to 1, and the average population.
hyp <- with(colombia, data.frame(intercept = 1, 
                                 tribut_2_log = tax_vec,
                                 tribut_2_log_sq = trans_vec,
                                 pop_attacks = mean(pop_attacks, na.rm = T),
                                 disbogota = mean(disbogota, na.rm = T),
                                 drugs = 1,
                                 pop_log = mean(pop_log, na.rm = T)))

#We could also have done this using model.matrix()

## create an empty matrix of values to fill with simulated predicted probabilities
pp_sims <- matrix(NA, nrow = N_sim, ncol = length(tax_vec))
#Using a for() loop to fill wth simulated pred. probs
for(i in 1:N_sim) {

  pp_sims[i, ] <- invlogit(as.matrix(hyp) %*% beta_sims[i, ])

}

#we could also do this with less code:
pp_sims2 <- invlogit(as.matrix(hyp) %*% t(beta_sims))

dim(pp_sims)
dim(pp_sims2)

all(t(pp_sims) == pp_sims2)

```

We're not going to worry a lot about confidence intervals today, but below is the code for them as well.

```{r}

## CIs
pe <- apply(pp_sims, 2, mean)
lo <- apply(pp_sims, 2, quantile, prob = .025)
hi <- apply(pp_sims, 2, quantile, prob = .975)

sim_output <- data.frame(tax = hyp[, 2], pe, lo, hi)

ggplot(sim_output, aes(x = tax, y = pe, ymin = lo, ymax = hi)) +
  geom_ribbon(aes(fill = '95% Confidence Interval (Simulation)'), alpha = .5) +
  geom_line(aes(color = 'Predicted Probability Over Values of Municipal Revenue'),
            size = 1) +
  labs(x = 'Revenue', y = 'Probability of Paramilitary Presence') +
  theme_bw() +
  theme(legend.title = element_text(), legend.position = 'bottom',
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()) +
  scale_fill_manual(values = 'steelblue3', name = '') +
  scale_color_manual(values = 'black', name = '')
```

## Using Observed-Case Approach - Average Predicted Effects

The code above created a hypothetical set of cases in which we might be interested, but what if we wanted more of an average effect? To this end, we can code and plot an 'average predictive effect' that covers the values of our key variable (we still generate predicted probabilities 'along' these values) and all possible values of the other variables in our dataset.

```{r, cache = TRUE}
## create a 'temporary' dataset - bind a 1 for the intercept and omit missingness
temporary <- model.matrix( ~ tribut_2_log + tribut_2_log_sq + pop_attacks + 
                             disbogota + drugs + pop_log, data = colombia)

## create an object and store predicted probabilities
## look into the ?map_dfc function from the purrr package (part of the tidyverse)

#note this will take a few minutes
pp <- purrr::map_dfc(tax_vec, function(j) {
    temporary[, "tribut_2_log"] <- j
    temporary[, "tribut_2_log_sq"] <- j^2

    pp <- invlogit(temporary %*% t(beta_sims))

    #getting mean for each set of betas
    pp <- apply(pp, 2, mean)

    return(pp)
    })

## find the mean (average predictive effect) for each observation, and bind with
## confidence measures (interval) 
plotdat <- t(apply(pp, 2, function(x) { # transposing w/ t() puts data into a column
    c(M = mean(x), quantile(x, c(0.025, 0.975)))
}))

# add x vector and convert to data frame
plotdat <- data.frame(tax_vec, plotdat)

# better names and show the first few rows (to check)
colnames(plotdat) <- c('tax', 'pe', 'lo', 'hi')
head(plotdat)

#now let's rowbind in the average-case predicted probabilities, so that we can 
#compare them
plotdat <- rbind(plotdat, sim_output) %>% 
  mutate(Approach = c(rep("Observed-Case",length(tax_vec)),
                      rep("Average-Case", length(tax_vec))))

## plot of predicted probabilities
ggplot(plotdat, aes(x = tax, y = pe, ymin = lo, ymax = hi, color = Approach,
                    fill = Approach)) +
  geom_ribbon(alpha = .25) +
  geom_line() +
  labs(x = 'Tax Revenue', y = 'Predicted Probability') +
  ylim(c(0, 1)) +
  theme_bw()
```

# Assignment

Read in the data 'turnout.csv.' Fit a logit model that predicts voting based on race, education and income.

Now, fit a second model that predicts voting based on race, education, income, and age. Assume that you've hypothesized that age has a curvilinear relationship with voting turnout, such that lower ages and higher ages are less likely to vote. Include this transformed variable.

Compare the model fit of these two options using lrtest, AIC, and BIC. Which do you choose as the better fitting model?

Simulate and plot predicted probabilities over the range of age, holding all other variables at their mean, for the second model. Plot this.

Generate an average predicted effect over the observed values of income for the first model. Plot this too.

```{r}

#Load in packages
pacman::p_load(lmtest)

turnout<-import("turnout.csv")
#Fit the two models
  #m1
m1 <- glm(vote ~ race + educate + income, turnout, family =______)

  #m2
m2 <- glm(vote ~ race + educate + income + poly(age, 2, raw = T), turnout, 
         family =______))

#Model fit
lrtest(______)
AIC()
BIC()
AIC()
BIC()

# Average-Case Approach
#When we combine the average-case approach with simulation, we create a hypothetical design matrix, where we hold all continuous variables at their mean or another justifiable value, and then let the predictor of interest vary along some range.

sim_dat <- with(______, data.frame(intercept = 1, race = 1,
                                    educate = mean(educate), income = mean(income),
                                    age = seq(min(age), max(age), length.out = 100),
                                    age2 = (seq(min(age), max(age), length.out = 100))^2))

sim_betas <- mvrnorm(______)

sim_pp <- apply(sim_betas, 1, function(x) invlogit(as.matrix(sim_dat) %*% x))

sim_gg <- data.frame(age = ______, pe = apply(sim_pp, 1, mean),
                     lo = ______,
                     hi = ______)

ggplot()

# Observed-Case Approach
#When taking the observed-case approach (see Hanmer and Kalkan (2013) for why this is the preferable approach) and using simulation to get confidence intervals, we have to do things slightly differently. Specifically, we will have to take an addition mean.

temporary <- model.matrix(m1)

#getting sim betas for m1
sim_betas_m1 <- mvrnorm(______)

#getting range of income
income_vec <- seq(from = ______,
                  length.out = 100)

#Here we use `map_dfc` function from `purrr` (part of the tidyverse) to we apply the function to each element in the vector, and then (because the function returns a vector) we get a dataframe in as output.

#This is doing a few things: 
#1. takes the temporary design matrix we created above, and then replaces *all* of the values of our predictor of interest with one value (taken from the hypothetical value vector).

#2. We then matrix multiply this new design matrix, with all actual values of our predictor of interest replaced by one hypothetical value of our predictor interest and with all the other values held at their observed values, by our simulated betas.

#3.  We then take the mean across the columns (`margins = 2`) because different columns represent different predicted probabilities, for each of the observations in our temporary design matrix, when we use different betas and when our predictor of interest has been replaced with one value.

#4. The call to `map_dfc` then returns a dataframe that has 1000 rows and 100 columns (1 column for each value of our hypothetical value vector).

pp <- map_dfc(income_vec, function(j) {
    temporary[, "income"] <- j

    pp <- invlogit(temporary %*% t(sim_betas_m1))

    pp <- apply(pp, 2, mean)

    return(pp)
    })

#Pulling it all into a df and getting the mean of predicted values for each value of our hypothetical value vector, and then use the `quantile` with a `probs` argument equal to the bounds of our desired confidence interval (for example, if we want a 95% confidence interval, we will do `quantile(x, probs = c(0.025, 0.975))`)

plotdat <- t(apply(pp, 2, function(x) { # transposing w/ t() puts data into a column
    c(M = mean(x), quantile(x, c(0.025, 0.975)))
}))

plotdat <- data.frame(plotdat, ______)

#We can then use these data to plot the mean predicted probability and our confidence bands.

#
colnames(______) <- c('pe', 'lo', 'hi', 'income')

ggplot()

```

### 2023-01-04 Lab 8  BinomialModels/Lab8_BinomialModels_AnswerKey.Rmd
Add files via upload

--- 2023-01-04 Lab 8  BinomialModels/Lab8_BinomialModels.Rmd
@@
 ---
-title: 'Lab 7: Missing Data and Imputation:Answers'
-author: "Eric Parajon"
-header-includes:
-  - \usepackage{enumerate}
-  - \usepackage{graphicx}
+title: 'Lab 8: Binomiaal Models in R'
+author: "TA: Eric Parajon"
 output: pdf_document
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
+knitr::opts_chunk$set(echo = TRUE)
 
@@
 pacman::p_load(tidyverse,
-               rio,#for importing data
-               arm)#inverse link function
-
+               rio,
+               arm)#for importing data
 ```
@@
 colombia <- rio::import("Lab_8_data.dta")
-
 ```
@@
   theme_bw()
-
 ```
@@
 ```{r}
-#for a logit model, how should we specify family?
-
 m1 <- glm(AUC ~ tribut_2_log + tribut_2_log_sq + pop_attacks + disbogota +
-            drugs + pop_log, data=colombia, family = __________)
+            drugs + pop_log, data=colombia, family = binomial(link = 'logit'))
 
@@
 ```{r}
-#for a probit model, how should we specify family?
-
-m2 <- glm(AUC ~ ., data = colombia, family = ___________)
+m2 <- glm(AUC ~ ., data = colombia, family = binomial(link = 'probit'))
 #what is the . doing above?
@@
 We might want to use our model to estimate predicted probabilities, as this is a readily-understood measure of how changes in our independent variable affect the probability of observing a certain outcome.
+
 In doing so, we'll want to think about a 'hypothetical case' for which we want to predict: for example, here, we'll calculate the predicted probability of the United Self-Defenses of Colombia or Autodefensas Unidas de Colombia (AUC) presence in a municipality with the mean tax income, mean attacks on population, mean distance from the capitol, drug crops, and the average population.
+
 *Note*: This is the so-called average case approach.
@@
 #let's take a look!
-invlogit #no parentheses from the arm package
+arm::invlogit #no parentheses from the arm package
 
@@
 
-#Here use predict()
-predict(_____, newdata = ________, type = 'response') #The type="response" option tells R to output probabilities of the form P(Y = 1|X)
+predict(m1, newdata = new_data, type = 'response')
 ```
@@
 ## variables at their means
-pred_output <- data.frame(Logit = predict(_____, newdata = _______, type = _______),
-                          Probit = predict(____, newdata = ________, type = ______),
+pred_output <- data.frame(Logit = predict(m1, newdata = new_data_2, type = 'response'),
+                          Probit = predict(m2, newdata = new_data_2, type = 'response'),
                           tax = new_data_2$tribut_2_log)
@@
 
-Simulation strikes again! This time around, we will do similar things, but with non-Normal outcomes, where simulations are even more helpful.
+We talked about simulations in Lab 4, when we talked about OLS. This time around, we will do similar things, but with non-Normal outcomes, where simulations are even more helpful!
 
-Today, we're going to simulate predicted probabilities for a model with an interaction term and plot our results. We will use both the average case and the observed-case approaches.
+Today, we're going to simulate predicted probabilities for a model with an interaction term, use the distribution of these simulated probabilities to define confidence intervals, and plot our results, using both the average case and the observed-case approaches.
 
@@
 
-We're not going to worry a lot about confidence intervals today, but below is the code for them as well.
+We're not going to worry about confidence intervals today, but below is the code for them as well.
 
 ```{r}
-
 ## CIs
@@
 
-Generate an average predicted effect over the observed values of income for the first model. Plot this too.
+Generate an average predicted effect over the observed values of income for the first model. Plot this.
 
-```{r}
-
+Required packages and data:
+```{r, message=FALSE}
 #Load in packages
@@
 turnout<-import("turnout.csv")
-#Fit the two models
-  #m1
-m1 <- glm(vote ~ race + educate + income, turnout, family =______)
 
-  #m2
+```
+
+First, we fit the models:
+```{r}
+m1 <- glm(vote ~ race + educate + income, turnout, family = binomial(link = 'logit'))
 m2 <- glm(vote ~ race + educate + income + poly(age, 2, raw = T), turnout, 
-         family =______))
+          family = binomial(link = 'logit'))
 
-#Model fit
-lrtest(______)
-AIC()
-BIC()
-AIC()
-BIC()
+lrtest(m1, m2)
+AIC(m1)
+AIC(m2)
+BIC(m1)
+BIC(m2)
+```
 
 # Average-Case Approach
-#When we combine the average-case approach with simulation, we create a hypothetical design matrix, where we hold all continuous variables at their mean or another justifiable value, and then let the predictor of interest vary along some range.
 
-sim_dat <- with(______, data.frame(intercept = 1, race = 1,
+When we combine the average-case approach with simulation, we create a hypothetical design matrix, where we hold all continuous variables at their mean or another justifiable value, and then let the predictor of interest vary along some range.
+
+```{r}
+sim_dat <- with(turnout, data.frame(intercept = 1, race = 1,
                                     educate = mean(educate), income = mean(income),
@@
                                     age2 = (seq(min(age), max(age), length.out = 100))^2))
+```
 
-sim_betas <- mvrnorm(______)
+In order to simulate different betas, we take a predetermined number of draws (1000 in this case here) from a multi-variate normal distribution centered around the coefficient estimates from our model, with a variance-covariance matrix also from our model.
+We then matrix multiply our hypothetical design matrix with our simulated betas.
+This means that we will have 1000 different predicted probabilities for each row in our hypothetical design matrix (in other words, we have 1000 different predicted probabilities for each element of the vector of values for our predictor of interest).
+We then have to make sure that we take the mean, and the quantiles we are interested in, along whichever dimension represents the elements of the hypothetical range vector.
+We can do this using `apply`.
+Note that in the example below, we are using `apply` along the rows of `sim_pp` because of how we constructed `sim_pp` (also using `apply`.) If we had not used `apply` and had instead simply done `as.matrix(sim_dat) %*% sim_betas`, the dimensions of `sim_pp` would have been reversed, and we would have to `apply` the mean and quantile functions along the columns, not the rows.
+```{r}
+sim_betas <- mvrnorm(1e3, coef(m2), vcov(m2))
 
@@
 
-sim_gg <- data.frame(age = ______, pe = apply(sim_pp, 1, mean),
-                     lo = ______,
-                     hi = ______)
+sim_gg <- data.frame(age = sim_dat$age, pe = apply(sim_pp, 1, mean),
+                     lo = apply(sim_pp, 1, quantile, probs = .025),
+                     hi = apply(sim_pp, 1, quantile, probs = .975))
+```
 
-ggplot()
+We can plot the predicted probabilities and our confidence bands.
+```{r}
+ggplot(sim_gg, aes(x = age, y = pe, ymin = lo, ymax = hi)) +
+  geom_ribbon(aes(fill = '95% Confidence Interval'), alpha = .5) +
+  geom_line(aes(color = 'Predicted Probability')) +
+  labs(x = 'Age', y = 'Probability of Turnout') +
+  theme_bw() +
+  theme(legend.title = element_text(), legend.position = 'bottom',
+        panel.grid.minor.x = element_blank(),
+        panel.grid.minor.y = element_blank()) +
+  scale_fill_manual('', values = 'dodgerblue4') +
+  scale_color_manual('', values = 'grey30')
+```
 
 # Observed-Case Approach
-#When taking the observed-case approach (see Hanmer and Kalkan (2013) for why this is the preferable approach) and using simulation to get confidence intervals, we have to do things slightly differently.
+
+When taking the observed-case approach (see Hanmer and Kalkan (2013) for why this is the preferable approach) and using simulation to get confidence intervals, we have to do things slightly differently.
 Specifically, we will have to take an addition mean.
+This is because whereas in the average-case approach, we had one row for each hypothetical value of our predictor of interest (however long we decided to make it), in the observed-case approach, we will have a whole matrix or dataframe for each hypothetical value of our predictor of interest.
 
+We start by creating a temporary design matrix. This while start out as being the same exact design matrix used to fit the model, which is why we can just use `model.matrix()` on our model object.
+
+```{r}
 temporary <- model.matrix(m1)
+```
 
+We then have to make sure to once again simulate our betas, using our model object.
+```{r}
 #getting sim betas for m1
-sim_betas_m1 <- mvrnorm(______)
+sim_betas_m1 <- mvrnorm(1e3, coef(m1), vcov(m1))
+```
 
+Just like with the average-case approach, we have to create the values we want our predictor of interest (*income* in this case) to take on.
+Just like in the average-case approach, these are really hypothetical values.
+Here, we will let income vary from its min to its max.
+```{r}
 #getting range of income
-income_vec <- seq(from = ______,
+income_vec <- seq(from = min(turnout$income), to = max(turnout$income),
                   length.out = 100)
+```
 
-#Here we use `map_dfc` function from `purrr` (part of the tidyverse) to we apply the function to each element in the vector, and then (because the function returns a vector) we get a dataframe in as output.
+In the next step, we do several important things at once.
+We will do these things using the `map_dfc` function from `purrr` (part of the tidyverse).
+This works a lot like the `apply` family of functions, but is somewhat faster.
+It maps a function to a vector and then, because of the `_dfc` suffix, column binds the results of the function together.
+In other words, we apply the function to each element in the vector, and then (because the function returns a vector) we get a dataframe in as output.
+Note that we could do something similar with `lapply`, but then we would get a list of vectors.
 
-#This is doing a few things: 
-#1. takes the temporary design matrix we created above, and then replaces *all* of the values of our predictor of interest with one value (taken from the hypothetical value vector).
-
-#2.
-We then matrix multiply this new design matrix, with all actual values of our predictor of interest replaced by one hypothetical value of our predictor interest and with all the other values held at their observed values, by our simulated betas.
-
-#3.
-We then take the mean across the columns (`margins = 2`) because different columns represent different predicted probabilities, for each of the observations in our temporary design matrix, when we use different betas and when our predictor of interest has been replaced with one value.
-
-#4. The call to `map_dfc` then returns a dataframe that has 1000 rows and 100 columns (1 column for each value of our hypothetical value vector).
-
+```{r}
 pp <- map_dfc(income_vec, function(j) {
@@
     })
+```
 
-#Pulling it all into a df and getting the mean of predicted values for each value of our hypothetical value vector, and then use the `quantile` with a `probs` argument equal to the bounds of our desired confidence interval (for example, if we want a 95% confidence interval, we will do `quantile(x, probs = c(0.025, 0.975))`)
+The function we apply to each element of our hypothetical predictor value vector does a few things:
+1. It takes the temporary design matrix we created above, and then replaces *all* of the values of our predictor of interest with one value (taken from the hypothetical value vector).
+This looks like the following:
+```{r}
+head(income_vec)
+
+temporary[, "income"] <- income_vec[1]
+
+head(temporary)
+```
+2.
+We then matrix multiply this new design matrix, with all actual values of our predictor of interest replaced by one hypothetical value of our predictor interest and with all the other values held at their observed values, by our simulated betas.
+This creates a matrix, which has as many rows as the rows of our data, and as many columns as there are samples of our betas.
+```{r}
+dim(invlogit(temporary %*% t(sim_betas_m1)))
+```
+3.
+We then take the mean across the columns (`margins = 2`) because different columns represent different predicted probabilities, for each of the observations in our temporary design matrix, when we use different betas and when our predictor of interest has been replaced with one value.
+This means that we then have a mean predicted probability for each different set of betas when our predictor of interest is at a certain value.
+
+The call to `map_dfc` then returns a dataframe that has 1000 rows and 100 columns (1 column for each value of our hypothetical value vector).
+Because we are interested in how the predicted probability changes according to the hypothetical values of our predictor of interest and because we currently have 1000 predicted probabilities for each of these hypothetical values (1 for each of our simulated betas), we once again use `apply` with `margins = 2`.
+We get the mean of predicted values for each value of our hypothetical value vector, and then use the `quantile` with a `probs` argument equal to the bounds of our desired confidence interval (for example, if we want a 95% confidence interval, we will do `quantile(x, probs = c(0.025, 0.975))`).
+
+```{r}
 
@@
 
-plotdat <- data.frame(plotdat, ______)
+plotdat <- data.frame(plotdat, income_vec)
+```
 
-#We can then use these data to plot the mean predicted probability and our confidence bands.
+We can then use these data to plot the mean predicted probability and our confidence bands.
+```{r}
+colnames(plotdat) <- c('pe', 'lo', 'hi', 'income')
+head(plotdat)
 
-#
-colnames(______) <- c('pe', 'lo', 'hi', 'income')
-
-ggplot()
-
+ggplot(plotdat, aes(x = income, y = pe, ymin = lo, ymax = hi)) +
+  geom_ribbon(aes(fill = '95% Confidence Interval (Simulation)'),
+              alpha = .5) +
+  geom_line(col = 'maroon') +
+  labs(x = 'Income', y = 'Predicted Probability of Voting') +
+  ylim(c(0, 1)) +
+  scale_fill_manual('', values = 'darkorange') +
+  theme_bw() +
+  theme(legend.title = element_text(), legend.position = 'bottom',
+        panel.grid.minor.x = element_blank(),
+        panel.grid.minor.y = element_blank())
 ```

### 2023-01-04 Lab 9  Multinom_orderedprob/Lab9_Multinom_OrderedProbit.Rmd
Add files via upload

---
title: "Lab 9: Multinomial and Ordered Probit Models"
author: "Eric Parajon"
output: pdf_document
---

Code has been adapted from Simon Hoellerbauer and Isabel Laterzo.

Load necessary packages:
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, messsage = FALSE)
rm(list=ls())

#Load in packages
pacman::p_load(tidyverse,
               haven,#for importing data
               labelled,
               MASS,
               arm)

```

# Ordered Probit

Data and resources for this section from <http://stats.idre.ucla.edu/r/dae/ordinal-logistic-regression/>.

We are going to expand an earlier example that we used for logistic regression: the data about application/acceptance into graduate school.

These data include our DV, 'a three level variable called apply, with levels unlikely, somewhat likely, and very likely, and three predictor variables: pared (dichotomous, does either parent have a graduate degree?) public (dichotomous, 0 or 1 indicating whether or not the students' undergrad institution is public) and gpa (continuous).

```{r}
grad <- read_dta("https://stats.idre.ucla.edu/stat/data/ologit.dta")
#when using read_dta, and we want to turn labelled vectors into factors,
#we have to use to_factor from labelled packaged
grad$apply <- to_factor(grad$apply)

```

Remember that before we even ran any tests for our logit model, we made a plot to understand the distribution of our variables of interest? We should do that again. Let's try a boxplot- or rather, a series of boxplots.

```{r}
ggplot(grad, aes(x = apply, y = gpa)) +
  geom_boxplot(size = 1) +
  geom_jitter(alpha = .25) +
  facet_grid(pared~public, margins = TRUE, labeller=label_both) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

Now, after looking at our data and checking for any unusual patterns, we can build a simple model.

For ordered outcomes, we need to use the MASS package (rather than base R). Otherwise, our syntax looks much the same.

```{r}
#First, a few checks to make sure our variables are in order: 
is.factor(grad$apply)
print(levels(grad$apply))

#Now, our model:
m1 <- polr(apply ~ pared + public + gpa, 
           data = grad, method = "probit", Hess = TRUE)
summary(m1)

#What is the effect of increasing each predictor by one unit?
#How do the odds of moving from a lower to adjacent higher category change?
coefs <- m1$coefficients[1:3]
#odds ratios
exp(coefs)

```

We can interpret these exponentiated coefficients in the same way as odds ratios in a logistic regression, but for the odds of each outcome category. Here, we can say that for pared, a one unit increase in parental education, i.e., going from 0 (Low) to 1 (High), the odds of very likely applying versus somewhat likely or unlikely applying combined are 1.81x greater, given that all of the other variables in the model are held constant.

We can (and should) also informally test the proportional odds assumption. What's this?

```{r}
m1_prop_1 <- glm(I(as.numeric(apply) >= 2) ~ pared, family = "binomial",
                 data = grad)

m1_prop_2 <- glm(I(as.numeric(apply) >= 3) ~ pared, family = "binomial",
                 data = grad)

#Pared = 0 
check_1 <- m1_prop_1$coef[1] - m1_prop_2$coef[1]
check_1
#Pared = 1
check_2 <- (m1_prop_1$coef[1] + m1_prop_1$coef[2]) - 
  (m1_prop_2$coef[1] + m1_prop_2$coef[2])
check_2

```

When pared is equal to no (0) the difference between the predicted value for apply greater than or equal to two and apply greater than or equal to three is roughly 2 (-0.378 - -2.440 = 2.062). For pared equal to yes the difference in predicted values for apply greater than or equal to two and apply greater than or equal to three is also roughly 2 (0.765 - -1.347 = 2.112).

We will also want to generate predicted probabilities and plot them. Let's first do so with a plot of the cumulative probability (so the probabilities of being in each subcategory sum to 1).

We can also check this first: 
```{r}
tau1 <- m1$zeta[1] #extract latent thresholds (calculated w polr) from our model!
tau2 <- m1$zeta[2]
p1 <- invlogit(tau1) # 1
p2 <- invlogit(tau2) - invlogit(tau1) # 2
p3 <- 1-invlogit(tau2)

p1 + p2 + p3
```

And now move on to plotting.

```{r}
# if we wanted to plot the range of gpa for all four combinations of pared and
#public, we could do the following:
#grad_2 <- expand.grid(pared = c(0:1), 
#                      public = c(0,1), 
#                      gpa = seq(from = 1.9, to = 4, length.out = 1000))
#We would then create an indicator variable for the four combinations, and then
#use facet_wrap in ggplot to create 4 different plots
#however, if we just want one easily interpretable plot, we can do the following:
table(grad$pared); table(grad$public)
grad_2 <- expand.grid(pared = 0, #modal value
                      public = 0, #modal value
                      gpa = seq(from = 1.9, to = 4, length.out = 1000))
colnames(grad_2)<- c("pared", "public", "gpa")
probs <- predict(m1, grad_2, type = "probs") #type = is different than for binomial model
head(rowSums(probs)) #just to double check
cumul_probs <- t(apply(probs, 1, cumsum)) #get the cumulative sum of the rows
#transpose into columns for plotting and merging with values of gpa
head(cumul_probs)
plot_data <- data.frame(grad_2$gpa, cumul_probs)
ggplot(data=plot_data, aes(x=grad_2.gpa))+
    geom_smooth(data=plot_data, aes(x=grad_2.gpa, y=unlikely), se=FALSE)+
    geom_smooth(data=plot_data, aes(x=grad_2.gpa, y=somewhat.likely), se=FALSE)+
    geom_smooth(data=plot_data, aes(x=grad_2.gpa, y=very.likely), se=FALSE)+
    geom_ribbon(data=plot_data, aes(ymin=0, ymax=unlikely, fill="Unlikely"))+
    geom_ribbon(data=plot_data, aes(ymin=unlikely, ymax=somewhat.likely,
                                    fill="Somewhat Likely"))+
    geom_ribbon(data=plot_data, aes(ymin=somewhat.likely, ymax=very.likely,
                                    fill="Very Likely"))+
    labs(x="Values of GPA", y="Cumulative Predicted Probability")+
    scale_fill_discrete(name="Applicant Response")+
    theme_bw()
```

And plotting a non-cumulative probability example.

```{r}
#data frame
grad_3 <- expand.grid(
  pared = 1, #assume parents education =1
  public = c(1,0), #vary values of public
  gpa = mean(grad$gpa)) #average gpa

library(MASS)
set.seed(831213)

coef_samp <- mvrnorm(1e3,coef(m1)
                     ,vcov(m1)[1:3,1:3]) #keep relevant vals of vcov matrix

## Get linear predictors
hyp_X <- model.matrix(~pared+public+gpa
                      ,data = grad_3)[,-1] #no intercept 
eta <- hyp_X %*% t(coef_samp) 
summary(t(eta)) 
## Get pred probs samples 
taus <- m1$zeta #extract thresholds 
pred_prob_diffs <- apply(eta,c(1,2)
                    ,function(x, J, tau){
                      prob_vec <- array(NA,J) #J categories - array() is generalizable
                      prob_vec[1] <- plogis(tau[1]-x) #prob of being in 1st category
                      for(i in 2:(J-1)) #move to next category
                        prob_vec[i] <- plogis(tau[i]-x) - plogis(tau[i-1]-x)
                        #probability of being in middle category
                      prob_vec[J] <- 1 - plogis(tau[J-1]-x)
                        #probability of being in top category (sums to 1)
                      return(prob_vec[J]-prob_vec[1]) 
                      # Diff. in prob from top to bottom cat.
                    }
                    ,J = 3 #3 possible outcomes
                    ,tau = taus)

## Obtain 90% confidence interval on probabilities

fx_on_prob_ci <- t(apply(pred_prob_diffs,1
                       ,quantile,probs=c(0.05,0.5,0.95)))

#find difference in predicted probabilities and confidence around those differences
public <- c("Yes", "No")
#for easy plotting
fx_on_prob_ci <- as.data.frame(cbind(fx_on_prob_ci, public))
colnames(fx_on_prob_ci)<-c("low", "pe", "high", "public")

fx_on_prob_ci$low <- as.numeric(fx_on_prob_ci$low) #fix issues with factors
fx_on_prob_ci$pe <- as.numeric(fx_on_prob_ci$pe)
fx_on_prob_ci$high <- as.numeric(fx_on_prob_ci$high)

## Produce plot
ggplot(fx_on_prob_ci, aes(x=pe, y=as.factor(public))) +
  labs(x="Change in Probability of Admission", y="") +
  ggtitle("Change in Probability for Public Vs Non Public")+
  geom_point(aes(x=pe)) +
  geom_errorbarh(aes(xmin=low, xmax=high), height=0.1) + #errorbar for confidence
  theme_bw()

#not much of an effect, as our model told us originally!

```

### Multinomial and Conditional Logit
```{r, message = FALSE, warning = FALSE}
#Loading additional necessary packages

pacman::p_load(nnet,#for multinom()
               mlogit,#for conditional mnl
               cowplot)#we will use it to put plots together (an extension to ggplot2)

```

For this example, we'll use data from the 2017/18 round of LAPOP in Brazil. Ths data has a variable called "vote_choice" which is a categorical variable which we'll treat as unordered. In particular, we'll examine who individuals voted for in Brazil in the 2018 national election (note there were MANY candidates, we'll only examine three - Bolsonaro, Haddad, and Gomes).

```{r, message=F, warning=F, results = 'hide'}
brazil_data <- readRDS("BrazilLAPOP_2018.rds")
#A bit of housekeeping:

#vote choice variable
brazil_data <- subset(brazil_data, vote_choice == "1501" |
                        vote_choice == "1502" |
                        vote_choice == "1503")
brazil_data$vote_choice <- as.factor(brazil_data$vote_choice)

recode(brazil_data$vote_choice, "1501" = "Bolsonaro", "1502" = "Haddad",
       "1503" = "Gomes")

levels(brazil_data$vote_choice) <- c("Bolsonaro", "Haddad", "Gomes")

#capital punishment variable
brazil_data$capital_pun <- ifelse(brazil_data$capital_pun == 1, 1, 
                                  ifelse(brazil_data$capital_pun == 2, 0, NA))

#sex variable
#male is 1, female is 0
brazil_data$sex <- ifelse(brazil_data$sex == 1, 1, 
                                  ifelse(brazil_data$sex == 2, 0, NA))
```

Let's make a descriptive plot of our DV. This is good practice - something to consider for your replication papers!

```{r, message=F, warning=F}
ggplot(data=brazil_data)+ 
  geom_bar(aes(x=vote_choice), fill="springgreen4")+
  scale_x_discrete(name="Vote Choice")+
  theme_bw()

```

Let's run our first model.

```{r, message=F, warning=F, results = 'hide'}
multi_1 <- multinom(vote_choice ~ sex + age + capital_pun,
                    data=brazil_data, Hess=TRUE)
summary(multi_1)

#p values
z <- summary(multi_1)$coefficients/summary(multi_1)$standard.errors
# 2-tailed Wald z tests to test significance of coefficients
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p

#variance covariance
vcov(multi_1)
```

We can change the reference category if we wish, and re-run:

```{r, message=F, warning=F, echo=F}
brazil_data$vote_choice <- relevel(brazil_data$vote_choice, ref = "Haddad")

multi_2 <- multinom(vote_choice ~ sex + age + capital_pun,
                    data=brazil_data, Hess=TRUE)
summary(multi_2)

```

We can also, as you may recall, run this as a series of binary logits (after some minor data recoding):

```{r, message=F, warning=F, results= 'hide'}
#filter data
brazil_data_2<-subset(brazil_data[which(brazil_data$vote_choice=="Bolsonaro" |
                                          brazil_data$vote_choice=="Haddad"),])

brazil_data_3<-subset(brazil_data[which(brazil_data$vote_choice=="Haddad" | 
                                          brazil_data$vote_choice=="Gomes"),])

#recode values
brazil_data_2$vote_choice<-ifelse(brazil_data_2$vote_choice=="Haddad", 0,1) 
#votes for Bolsonaro
brazil_data_3$vote_choice<-ifelse(brazil_data_3$vote_choice=="Haddad", 0,1) 
#votes for Gomes

#check it out
head(brazil_data_2$vote_choice)
head(brazil_data_3$vote_choice)

#models!

glm_1<-glm(vote_choice~sex + age + capital_pun,
           data=brazil_data_2,
           family=binomial (link="logit"))

glm_2<-glm(vote_choice~sex + age + capital_pun,
           data=brazil_data_3,
           family=binomial (link="logit"))

coefs<-rbind(coef(glm_1), coef(glm_2), coef(multi_2))
coefs
```

Finally, we can run this as a slightly different model using a multinomial choice model and another package. To do so, we need to take a few extra steps. This allows us to change our data from a single observation per individual that records their vote intent to *multiple observations for an individual that encompass all of their choices.*

This means we can have information about alternatives that would be constant across individuals (ie policy position for each of these candidates) *and* information about individuals that is constant across alternatives (like gender or age of an individual).

In *these particular data* we only have individual-level information. However, we can set up these data in the same way we would a model that allows us alternative choices.

```{r, message=F, warning=F, results = 'hide'}

#First, a look at our data 
brazil_data[1:10,]
levels(brazil_data$vote_choice)

#Reshape- we can do this with a function in the mlogit package

brazil_data <- as.data.frame(na.omit(brazil_data))

brazil_long <- mlogit.data(brazil_data,
                             shape = "wide", 
                             choice = "vote_choice",
                             alt.levels = c("Haddad", "Bolsonaro", "Gomes"))
                            #These match the levels above

#Check our data now
head(brazil_long)
brazil_long[1:10,]

#Replicate our multinomial logit and check with the 2nd model, above:
multi_3 <- mlogit(vote_choice ~ 0|sex+age+capital_pun,
                  data=brazil_long, reflevel="Haddad")
# 0| is used in this case because we have no 'alternative-level variables'

head(model.matrix(multi_3))

check<-list(summary(multi_3), summary(multi_2))
check #look at the intercept for multi_3 and compare to multi_2

```

Let's return to our multinomial model (multi_2) and generate some QIs (predicted probabilities). We'll take a random sample of our data and then use our expand.grid() to see the effect of different combinations of our explanatory variables.

```{r, message=F, warning=F, results = 'hide'}

set.seed(2876)
brazil_test<-brazil_data[,c("vote_choice", "sex", "age", "capital_pun")]
brazil_probs<-brazil_test[sample(nrow(brazil_test), 50),]
#expand.grid gives us ALL possible combinations of the values in our data
#we randomly sampled 50 rows to make the computation easier.
brazil_d<-expand.grid(brazil_probs)

pred<-predict(multi_2, newdata=brazil_d, type="probs")
#'probs' gives us information about the predicted probability of each category
# in our DV

#What should this look like?
head(rowSums(pred)) #looks good!

plotdat<-as.data.frame(cbind(pred, brazil_d$sex,
                             brazil_d$age, brazil_d$capital_pun))

```

And we can plot, with a bit more detail:

```{r, message=F, warning=F}

#Let's separate out our predicted probabilities for each candidate by gender:
set.seed(12)
m_pred <-plotdat[which(plotdat$V4==1),]
m_pred <- m_pred[sample(nrow(m_pred), 1000), ]
f_pred <-plotdat[which(plotdat$V4==0),]
f_pred <- f_pred[sample(nrow(f_pred), 1000), ]

mplot<-ggplot(data=m_pred)+
  geom_smooth(data=m_pred, aes(x=V5, y=Haddad, colour="Haddad"), se=FALSE)+
  geom_smooth(data=m_pred, aes(x=V5, y=Bolsonaro, colour="Bolsonaro"), se=FALSE)+
  geom_smooth(data=m_pred, aes(x=V5, y=Gomes, colour="Gomes"), se=FALSE)+
  labs(x="Age", y="Predicted Probability of Voter Preference")+
  scale_colour_manual(name="Candidate",
    values=c(Haddad="red", Bolsonaro="blue", Gomes="green"))+
  theme_bw()

fplot<-ggplot(data=f_pred)+
  geom_smooth(data=f_pred, aes(x=V5, y=Haddad, colour="Haddad"), se=FALSE)+
  geom_smooth(data=f_pred, aes(x=V5, y=Bolsonaro, colour="Bolsonaro"), se=FALSE)+
  geom_smooth(data=f_pred, aes(x=V5, y=Gomes, colour="Gomes"), se=FALSE)+
  labs(x="Age", y="Predicted Probability of Voter Preference")+
  scale_colour_manual(name="Candidate",
    values=c(Haddad="red", Bolsonaro="blue", Gomes="green"))+
  theme_bw()

both <- plot_grid(mplot, fplot, labels = c('Male', 'Female'))
both

#Now let's separate out by support for capital punishment

cap_pred <-plotdat[which(plotdat$V6==1),]
cap_pred <- cap_pred[sample(nrow(cap_pred), 1000), ]
nocap_pred <-plotdat[which(plotdat$V6==0),]
nocap_pred <- nocap_pred[sample(nrow(nocap_pred), 1000), ]

cap_plot<-ggplot(data=cap_pred)+
  geom_smooth(data=cap_pred, aes(x=V5, y=Haddad, colour="Haddad"), se=FALSE)+
  geom_smooth(data=cap_pred, aes(x=V5, y=Bolsonaro, colour="Bolsonaro"), se=FALSE)+
  geom_smooth(data=cap_pred, aes(x=V5, y=Gomes, colour="Gomes"), se=FALSE)+
  labs(x="Age", y="Predicted Probability of Voter Preference")+
  scale_colour_manual(name="Candidate",
    values=c(Haddad="red", Bolsonaro="blue", Gomes="green"))+
  theme_bw()

nocap_plot<-ggplot(data=nocap_pred)+
  geom_smooth(data=nocap_pred, aes(x=V5, y=Haddad, colour="Haddad"), se=FALSE)+
  geom_smooth(data=nocap_pred, aes(x=V5, y=Bolsonaro, colour="Bolsonaro"), se=FALSE)+
  geom_smooth(data=nocap_pred, aes(x=V5, y=Gomes, colour="Gomes"), se=FALSE)+
  labs(x="Age", y="Predicted Probability of Voter Preference")+
  scale_colour_manual(name="Candidate",
    values=c(Haddad="red", Bolsonaro="blue", Gomes="green"))+
  theme_bw()

both2 <- plot_grid(cap_plot, nocap_plot, labels = c('Cap: Yes', 'Cap: No'))
both2

```

# Lab Assignment

## Ordered Outcome Practice

Use the data below. Estimate an ordered logit with `warm' as the DV, using male, white, age, education, and prestige as independent variables.

Check the proportional odds assumption for at least one of your variables - does it hold?

Generate a cumulative predicted probability plot.

Data and description from Christopher Adolph: 
Example data from 1977, 1989 GSS:  Attitudes towards working mothers: "A working mother can establish just as warm and secure of a relationship with her child as a mother who does not work." SD, D, A, SA
Covariates: male respondent; white respondent; age of respondent;
years of education of respondent;
prestige of respondent's occupation (% considering prestigious)

```{r, eval = F}
hw <- read.csv(url("http://faculty.washington.edu/cadolph/mle/ordwarm2.csv"))

```

### 2023-01-04 Lab 9  Multinom_orderedprob/Lab9_Multinom_OrderedProbit_Answers.Rmd
Add files via upload

--- 2023-01-04 Lab 9  Multinom_orderedprob/Lab9_Multinom_OrderedProbit.Rmd
@@
 
+#Cleaning
+
+hw<-hw %>% 
+   mutate(warm = case_when(.$warm == 1 ~ "Strongly disagree", 
+                          .$warm==2 ~ "Disagree", 
+                          .$warm==3 ~ "Agree", 
+                          .$warm==4 ~ "Strongly agree")) %>%
+    mutate(warm = factor(warm, levels=c("Strongly disagree","Disagree","Agree","Strongly agree")))
+
+#Create the model
+warm_model <- polr(warm~male+white+age+ed+prst, 
+           data = hw, method = "probit", Hess = TRUE)
+summary(warm_model)
+
+#What is the effect of increasing each predictor by one unit?
+#How do the odds of moving from a lower to adjacent higher category change?
+coefs <- warm_model$coefficients[1:5]
+#odds ratios
+exp(coefs)
+
+#Check the proportional odds assumption for male
+m1_prop_1 <- glm(I(as.numeric(warm) >= 2) ~ male, family = "binomial",
+                 data = hw)
+
+m1_prop_2 <- glm(I(as.numeric(warm) >= 3) ~ male, family = "binomial",
+                 data = hw)
+
+#male=0
+check_1 <- m1_prop_1$coef[1] - m1_prop_2$coef[1]
+check_1
+
+#male=1
+check_2 <- (m1_prop_1$coef[1] + m1_prop_1$coef[2]) - 
+  (m1_prop_2$coef[1] + m1_prop_2$coef[2])
+check_2
+
+#Generate a cumulative predicted probability plot.
+table(hw$male); table(hw$white); mean(hw$ed)
+hw_2 <- expand.grid(male = 1, #modal value
+                      white = 1, #modal value
+                      ed=mean(hw$ed),
+                      age=mean(hw$age),
+                      prst = seq(from = min(hw$prst), to = max(hw$prst), length.out = 2293))
+
+probs <- predict(warm_model, hw_2, type = "probs") #type = is different than for binomial model
+head(rowSums(probs)) #just to double check
+cumul_probs <- t(apply(probs, 1, cumsum)) #get the cumulative sum of the rows
+#transpose into columns for plotting and merging with values of gpa
+head(cumul_probs)
+plot_data <- data.frame(hw$prst, cumul_probs)
+ggplot(data=plot_data, aes(x=hw.prst))+
+    geom_smooth(data=plot_data, aes(x=hw.prst, y=Strongly.disagree), se=FALSE)+
+    geom_smooth(data=plot_data, aes(x=hw.prst, y=Disagree), se=FALSE)+
+    geom_smooth(data=plot_data, aes(x=hw.prst, y=Agree), se=FALSE)+
+      geom_smooth(data=plot_data, aes(x=hw.prst, y=Strongly.agree), se=FALSE)+
+    geom_ribbon(data=plot_data, aes(ymin=0, ymax=Strongly.disagree, fill="Strongly.disagree"))+
+    geom_ribbon(data=plot_data, aes(ymin=Strongly.disagree, ymax=Disagree,
+                                    fill="Disagree"))+
+    geom_ribbon(data=plot_data, aes(ymin=Disagree, ymax=Agree,
+                                    fill="Agree"))+
+         geom_ribbon(data=plot_data, aes(ymin=Agree, ymax=Strongly.agree,
+                                    fill="Strongly.agree"))+
+    labs(x="Values of prestige of occupation", y="Cumulative Predicted Probability")+
+    scale_fill_discrete(name="Hard Work")+
+    theme_bw()
+
 ```

### 2023-01-04 Lab 14  RDD/Lab14_RDD.Rmd
Add files via upload

---
title: "Lab 14 - Regression Discontinuity (RDD) and Instrumental variables (IV)"
author: "Eric Parajon, with code adapted from Colin Case and Sean Norton"
output: html_document
bibliography: rdd.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, message = FALSE, warning = FALSE)
```

# Regression Discontinuity

Regression discontinuity designs (RDD) exploits the structure of non-random treatment assignment to treatment to give us *as-if random* assignment to treatment. It does this by exploiting treatments that are assigned with respect to the level of another variable.

In less abstract terms, let's consider the motivating example of the first RDD. Thistlewaite and Campbell(1960) study the effect of a merit scholarship on students' academic performance. Of course, the high-performing students were most likely to receive the scholarship, so would do better anyways. There's no existing counterfactual for us to compare to in a standard set-up, such as OLS or difference-in-differences.

However, let's say the scholarship is assigned (either purposefully or unintentionally) based on SAT scores, with a certain score serving as the cutoff between those who received the scholarship and those who didn't. People in the area right around that cutoff are likely to be pretty similar in terms of academic performance; missing the cutoff by just a few points could plausibly be due to random influences like what you ate for breakfast, a mistake on the scantron, etc. In other words, within some small region of the cutoff, assignment to treatment is *as-if* random, and those who didn't receive the scholarship will be a plausible counterfactual. We call the variable that the cutoff exists on the *running variable* or the *forcing variable* - in this case, SAT scores.

There are two types of RDD:

1. *Fuzzy*: the running variable probabilistically increases or decreases assignment to treatment along its range.
2. *Sharp*: has a strict, binary division between treatment and control

We'll be working with sharp RDD in this lab because a) it's simpler and b) the fuzzy RDD is just a different formulation of instrumental variables.

## RDD Example: The Incumbency Effect

We'll be working with a dataset from the creators of the `rdrobust` package on Senate races. The outcome variable is `vote`, which is the vote share for the Democratic candidate in that election. Our forcing variable is `margin`, which is that candidate's margin of victory in the *previous election*.

We want to identify the effect of having won the previous election on the subsequent election, i.e the incumbency effect. The obvious problem with this is that states with long-serving Democratic senators are likely to be Democratic strongholds.

However, senators who *barely won* their last election are probably in more competitive state, so by comparing these senators to those who barely lost we can recover a causal effect for experience.

Let's download the data and visualize it.

```{r}
# Clear Environment
rm(list = ls())
# Set seed 
set.seed(1996)
# Load packages
pacman::p_load(haven,rdrobust,rdd,readr,ggplot2,AER)

# Load data
senate <- read_csv('https://raw.githubusercontent.com/rdpackages/rdrobust/master/R/rdrobust_senate.csv')

# Plot  previous vote margin (x) versus vote for current election (y) by Senate class (color)
ggplot(_____, aes(x = _____, y = _____, color = ______) +
  geom_point() +
  theme_bw() +
  scale_color_discrete(name = 'Senate \n Class', labels = c('1', '2', '3')) +
  geom_vline(xintercept = ____) +
  theme(legend.position = 'right',
        plot.background = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_blank())
```

As you can see, we have quite a few observations around the cutoff of 0. This is what you  want to see - obviously, observations made on few data points won't be very robust.

To get a sense of if the cutoff actually makes sense, we can use the `rdplot` function.

Consult the documentation and use `rdplot` to visualize how the outcome varies around the  cutoff. All of the default arguments that control the bins and the kernel used to fit a line to the data are fine.

```{r}
# Use Rdplot to see how outcome varies on either side of cutoff
rdplot(senate$____, senate$______)
```

Seems like there's definitely a jump after the cutoff - good! If you didn't really see anything here, your cutoff is probably incorrect.

Now let's use a simple model, where the outcome is only a function of the running variable. Use the `rdrobust()` function to do so. Summarize it.

```{r}
# Fit model 
fit_rd <- rdrobust(senate$_____, senate$______)
summary(fit_rd)
```

We'll ignore most this output in order to move through the lab - but I encourage you to read through the [paper introducing the package](https://journal.r-project.org/archive/2015/RJ-2015-004/RJ-2015-004.pdf) to understand what's going on under the hood.

What's important for now is the the estimate. In this case the conventional and robust estimate are the same, with the robust estimate having a more conservative z-score. We can interpret this as the effect of being over the cutoff increasing the following election's vote share by 7.4%. In other words, we can determine that having won the past election (even barely) increases the Democrat's vote share in the following election. We were able to recover this estimate despite the fact that Democrats were more likely to win in some states/years than others.

What about covariates though? Obviously there are some things that affect vote share other than previous margin of victory. Let's include those in the RDD estimation.

You'll use the `covs` argument to do so. Includes `termssenate`, `termshouse`, and `class`. Summarize it, and compare it to the estimate with no covariates.

(`covs` expects a matrix. As a hint, you can subset specific named columns of a dataframe or matrix using a character vector and `[]` notation. e.g.: `dataframe[ , c('variable', 'names')]`)

```{r}
# Use rdrobust to estimtate the model with covs argument specified
fit_rd_covs <- rdrobust(senate$____, senate$______, 
                        covs=_____________________)

summary(fit_rd_covs)
```

That accounted for some of the effect of an additional term, but the effect still exists and is quite large. We can't actually get estimates for the covariates, because `rdrobust` is using a non-parametric kernel regression under-the-hood.

The idea is intuitive as well, which is always a plus when arguing your model can recover causality. A word of caution though: *this only works if the cutoff is exogenous!* Actors are perfectly capable of manipulating cutoffs or scores on the running variable to their advantage, e.g. low margins may reflect electoral fraud in some contexts. Additionally, if you aren't careful in how you define the running variable, you may violate the assumption of as-if random assignment. See [Erikson and Rader (2017)](https://pdfs.semanticscholar.org/2fd6/08c0afe3f83959bc2fb48904be18d9c4537c.pdf) for a fun discussion of how the paper this data is from is wrong.[Marshall (2022)](https://doi.org/10.1111/ajps.12741) also just published a paper in AJPS that is somewhat critical of this design for certain applications and how the assumptions can be violated.

## The McCrary Test

One of the key assumptions about the RD design is that units cannot "choose" where they are vis-a-vis the running variable - i.e. respondents can't choose to be before the running variable, or after - so they can't "select" themselves to the treatment/ control conditions. This can be an issue with as-if random experiments. For example, take Germany's 5% representation threshold as an example - that parties will only get a seat in the Bundestag if they get 5% of votes nationally. Since the institution is well-known to the players of the game, they may try in every way to get into the parliament, so in that case the treatment is not randomly assigned.

To test this assumption, we can conduct what we call a McCrary Test (aka Density Test). The McCrary test plots a histogram of the running variable. With that histogram we plot the density of the distribution of the running variable. We then examin the density around the cut-off. We are looking at whether the respondents could somehow get through the cut-off. If we see that around the cut-off, i.e. if there's a sudden decrease on either side of the cut-off, then there's some evidence suggesting that the respondents do have an influence on where they are in the running variable. In other words, the cut-off assignment is not entirely exogeneous to the outcome, and thus the Regression Discontinuity Design could be violated.

```{r}
# Density test
DCdensity(senate$_____, ___)
```

It seems like in this example there's a tiny break between those that are below the cutoff and higher than the cutoff of 0, although the break is not large. The p-value from the function also suggests that we did not reject the null of no non-random assignment (p=.39). However, this could be a sign that the treatment assignment is not entirely random.

# Instrumental variables (IV) regression:

To perform IV regression, we use a two-stage least squares (TSLS) function (ivreg()) from the aer package.
The formula is relatively straight forward. Assume that we wish to estimate the model

<div align="left">$Y_i=\beta_0+\beta_1 X_{1 i}+\beta_2 X_{2 i}+W_{1 i}+u_i$</div>

Where $X{1 i}$ and $X{2 i}$ are endogenous regressors instrumented by $Z{1 i}$, $Z{2 i}$, and $Z{3 i}$ and $W_{1 i}$ is an exogenous regressor.

The correct formula in ivreg() would be:

y ~ x1 + x2 + w1 | w1 + z1 + z2 + z3

## Example
Below I model an IV model regressing education (educ), age age, actual labor market experience (exper) and its squared term expersq on log wages lwage as the dependent variable. I also use the parents’ education (fatheduc and motheduc) as instruments for education. I then conduct an F-test of relevance and a Sargan test for validity.

```{r}
#Reading in the data
women<- read.csv("http://eclr.humanities.manchester.ac.uk/images/5/5f/Mroz.csv")

#Removing null values for log wages and converting to numeric
women<- subset(women, lwage != ".")
women$lwage<-as.numeric(women$lwage)

#Running iv model using ivreg from the aer package
reg_iv1 <- ivreg(lwage~ educ+age+exper+expersq| 
                   .-educ+fatheduc+motheduc, data=women)

#Calling summary on model
summary(reg_iv1)

#Diagnostics: 
#Conducting an F-test
#Running first_stage model
first_stage <- lm(educ~fatheduc+motheduc,data=women) #regress X on Z 
#Running null model
null <- lm(educ~1,data=women)
relevance_test <- waldtest(first_stage, null)
relevance_test

#Sargan test
sargan <- lm(reg_iv1$residuals ~ age+exper+expersq+fatheduc+motheduc,data=women)
r.squared<-summary(sargan)$r.squared
sargan_test <- r.squared*nrow(women)

(1-pchisq(sargan_test,1))  #prints p-value

#The null hypothesis of the Sargan test implies all instruments are valid. 
#So the high p value confirms that we can accept H0, that is all instruments are valid.

#Based on the output, I find that the instruments are both relevant (confirmed by F-test) and valid (confirmed by Sargan test) to education, and that higher levels of education slightly increases a women's wage (significant at the 0.1 level).

```

# Exercise: Do Voters change the mind of their representatives? Evidence from the U.S. House.

For the exercise we are going to replicate the paper by @lmb2004 : \textit{Do Voters Affect or Elect Policies? Evidence from the U.S. House} in the QJE (referred to as LMB below). You don't have to read the paper, but essentially in their paper they are trying to test which model of public policy is more plausible: Do elections push elected officials somewhere to the middle (i.e. following the Downsian Model of voting), such that the resultant policy outcomes are all clustered around the median voter? Or do elections push the politician to a specific direction? If politicians cannot make credible promises to moderate their policies, then politicians would "change their minds" after an election, to follow the general ideological direction expressed by the electorate in the election.

To test this proposition, LMB fitted two simple regression discontinuity models. They used the ADA Score and the DW-Nominate Scores (figures I and VI respectively in the paper), both are quantitative measures of the ideological position of representatives. They argue that the running variable to the Democratic Vote Share at time $t$ - So if democrats receive more than half the share of votes, they would win in that district and a democratic representative would be elected. If elections can coerce the politician to change their policy predispositions, then at time $t+1$ we should observe a movement in the ADA Score and DW-Nominate Score (Increase in the former - converging to 100; and decrease in the latter, converging to -1).

Please replicate the regression discontinuity analysis they performed - fit the model and plot the results. `realada` is the ADA score and `dwnom1` is the DW-Nominate Score. The running variable is the lagged vote share of democrats `lagdemvoteshare`. Note that in the original paper they reported cluster-robust SEs, so you would have to do the same with `clusterid`. (Hint: check out the `cluster` parameter in the call to `rdrobust()`).

Then, conduct the McCrary Test and verify that the House Representatives did not get to "choose" whether they would win the election or lose it (which is quite a hard test!).

## Solution

```{r}
# Load Data
enricoall4 <- read_dta("enricoall4.dta")

# Fit RDD Model specifying clustered standard errors for ADA as DV

# Plot RDD

```

```{r}
# Fit RDD Model specifying clustered standard errors for DW-Nom as DV

# Plot RDD

```

Yes, we have successfully replicated their results found in the paper.

## McCrary Test

```{r}
#McCrary Test

```
The McCrary Test does not show sharp jumps at around the threshold (c=0.5), but it does alert us that there are heavy tails in the distribution of the running variable. The heavy tails come from Republican strongholds (Democrats have no chance of winning) and Democrat strongholds (Democrats always win). You can also see this above with the lack of observations around the cutpoint. This does lead to some concerns about the robustness of the RD effect we found, and particularly in the sudden reversal of RD trends at the upper tail of the running variable. Substantively this makes sense - if Democrats always win a district, then there's no point to adjust their policy predispositions to election results!

# Bibliography

### 2023-01-04 Lab 14  RDD/Lab14_rdd_Answers.Rmd
Add files via upload

--- 2023-01-04 Lab 14  RDD/Lab14_RDD.Rmd
@@
 ---
-title: "Lab 14 - Regression Discontinuity (RDD) and Instrumental variables (IV)"
+title: "Lab 14 - Regression Discontinuity (RDD) and Instrumental variables (IV):Answers"
 author: "Eric Parajon, with code adapted from Colin Case and Sean Norton"
-output: html_document
-bibliography: rdd.bib
+output:
+  pdf_document: default
+  html_document: default
 ---
@@
 # Plot  previous vote margin (x) versus vote for current election (y) by Senate class (color)
-ggplot(_____, aes(x = _____, y = _____, color = ______) +
+ggplot(senate, aes(x = margin, y = vote, color = as.factor(class))) +
   geom_point() +
@@
   scale_color_discrete(name = 'Senate \n Class', labels = c('1', '2', '3')) +
-  geom_vline(xintercept = ____) +
+  geom_vline(xintercept = 0) +
   theme(legend.position = 'right',
@@
 # Use Rdplot to see how outcome varies on either side of cutoff
-rdplot(senate$____, senate$______)
+rdplot(senate$vote, senate$margin)
 ```
@@
 # Fit model 
-fit_rd <- rdrobust(senate$_____, senate$______)
+fit_rd <- rdrobust(senate$vote, senate$margin)
 summary(fit_rd)
@@
 # Use rdrobust to estimtate the model with covs argument specified
-fit_rd_covs <- rdrobust(senate$____, senate$______, 
-                        covs=_____________________)
+fit_rd_covs <- rdrobust(senate$vote, senate$margin, covs=senate[,c("termssenate","termshouse","class")])
 
@@
 # Density test
-DCdensity(senate$_____, ___)
+DCdensity(senate$margin, 0)
 ```
@@
 ```{r}
+
 #Reading in the data
@@
 ```
-
 # Exercise: Do Voters change the mind of their representatives? Evidence from the U.S. House.
@@
 # Load Data
-enricoall4 <- read_dta("enricoall4.dta")
+enricoall4 <- read_dta("enricoall4_small.dta")
 
 # Fit RDD Model specifying clustered standard errors for ADA as DV
+lmb.fit <- rdrobust(enricoall4$realada,
+                    enricoall4$lagdemvoteshare,
+                    cluster=enricoall4$clusterid,
+                    c=0.5)
 
 # Plot RDD
-
+rdplot(enricoall4$realada,
+       enricoall4$lagdemvoteshare,
+       x.label="Democrat Vote Share at time t",
+       y.label="ADA Score at time t+1",
+       c=0.5)
 ```
@@
 # Fit RDD Model specifying clustered standard errors for DW-Nom as DV
-
+lmb.fit2 <- rdrobust(enricoall4$dwnom1,
+                     enricoall4$lagdemvoteshare,
+                     cluster=enricoall4$clusterid,
+                     c=0.5)
 # Plot RDD
-
+rdplot(enricoall4$dwnom1,
+         enricoall4$lagdemvoteshare,
+       x.label="Democrat Vote Share at time t",
+       y.label="DW-Nominate Score at time t",
+         c=0.5)
 ```
@@
 #McCrary Test
-
+DCdensity(enricoall4$lagdemvoteshare, 0.5)
 ```
@@
 Substantively this makes sense - if Democrats always win a district, then there's no point to adjust their policy predispositions to election results!
-
-# Bibliography

### 2023-01-13 Lab 1  R Review/Lab1_R_Review_Answers.Rmd
Add files via upload

--- 2023-01-04 Lab1_R_Review_Answers.Rmd
@@
 output:
+  html_document:
+    df_print: paged
   pdf_document:
     keep_tex: TRUE
-  html_document:
-    df_print: paged
 ---
@@
 
-For a refresher on how to use R Markdown, please see this cheat sheet: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf.
+For a refresher on how to use R Markdown, please see this cheat sheet: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
 
@@
 
-#R Review
+# R Review
 \medskip
+
 1.
@@
 \bigskip
-2.
-First, create a vector of ten 5s using the function `rep()`.
-Second, create a vector that counts from 1 to 10 - do this in two ways, one using `c()` and one using `seq()`.
-Third, create a vector of length 10 filled with NAs.
-Finally, uses `cbind()` to bind theses four vectors together to create a 10 x 14 (10 rows, 4 columns) matrix.
-Remember, if you don't know how to use a function, type ?
-function name into your console to see the help file (e.g., `?seq()`).
 
-Try binding the vectors together with `rbind()` instead. What happens?
+\bigskip
+
+2. To start, create a vector of ten 5s using the function `rep()`.
+
++ Second, create a vector that counts from 1 to 10 - do this in two ways, one using `c()` and one using `seq()`.
+
++ Third, create a vector of length 10 filled with NAs. Finally, uses `cbind()` to bind theses four vectors together to create a 10 x 14 (10 rows, 4 columns) matrix.
+
++ Remember, if you don't know how to use a function, type ? function name into your console to see the help file (e.g., `?seq()`).
+
++ Try binding the vectors together with `rbind()` instead. What happens?
 
@@
 \bigskip
+
 3. Set your seed to 425. Simulate 100 draws from a Normal distribution with mean 2 and standard deviation 4, assigning the result to an object.
 
-What proportion of the draws is above 3.1 in our sample? Think about using the functions `sum()` and `length()` to solve this, although there are other ways as well.
++ What proportion of the draws is above 3.1 in our sample? Think about using the functions `sum()` and `length()` to solve this, although there are other ways as well.
 
-What is the expected proportion of draws above 3.1? Hint, look into the function `pnorm()`
++ What is the expected proportion of draws above 3.1? Hint, look into the function `pnorm()`
 
-Repeat this process, but this time take 10,000 draws from the same Normal distribution. How does our result compare with the expected proportion of draws above 3.1 now?
++ Repeat this process, but this time take 10,000 draws from the same Normal distribution. How does our result compare with the expected proportion of draws above 3.1 now?
 
-Thinking back to last semester, what did we just calculate using simulation?
++ Thinking back to last semester, what did we just calculate using simulation?
 
@@
 \bigskip
+
 4. Another important concept in R is a `for` loop. If you need a refresher on these, visit: https://www.r-bloggers.com/2015/12/how-to-write-the-first-for-loop-in-r/.
 
-Using the 100 draws from the Normal distribution from Q3, create a new vector that is 1 when the vector of draws is greater than 3.1 and 0 when it is less than or equal to 3.1.
++ Using the 100 draws from the Normal distribution from Q3, create a new vector that is 1 when the vector of draws is greater than 3.1 and 0 when it is less than or equal to 3.1.
 First, do this using a `for` loop with an embedded `if` `else` sequence.
@@
 
-In addition to loops, `ifelse()` is also an important function. Look at the documentation of this function and repeat the above activity using `ifelse()`.
++ In addition to loops, `ifelse()` is also an important function. Look at the documentation of this function and repeat the above activity using `ifelse()`.
 
@@
 \bigskip
-5.
-Load in the `swiss` dataset.
-A description for this data can be found here: https://stat.ethz.ch/R-manual/R-patched/library/datasets/html/swiss.html How many variables does it have?
-How many observations?
 
-The variable `Catholic` is the \% of Catholic individuals, as opposed to Protestant in each province.
+5. Practice: Load in the `swiss` dataset. A description for this data can be found here: https://stat.ethz.ch/R-manual/R-patched/library/datasets/html/swiss.html
+
++ How many variables does it have?
+
++ How many observations?
+
++ The variable `Catholic` is the \% of Catholic individuals, as opposed to Protestant in each province.
 First look at the summary of this variable.
@@
 
-What is the mean percentage of this new variable?
++ What is the mean percentage of this new variable?
 
@@
 \bigskip
+
 6.
@@
 
-Create a matrix filled with random values that has 4 rows and 2 columns. Use `length()` on this matrix. Why might this be? Use `?` to read the documentation for `length()`.
++ Create a matrix filled with random values that has 4 rows and 2 columns. Use `length()` on this matrix. Why might this be? Use `?` to read the documentation for `length()`.
+
 ```{r}
@@
 \bigskip
+
 7.
@@
 
-Some useful functions here are `vector()`, `seq_along()`, and `names()`. Remember to use the help function `?` for further information about these functions.
++ Some useful functions here are `vector()`, `seq_along()`, and `names()`. Remember to use the help function `?` for further information about these functions.
 ```{r}
@@
 \bigskip
+
 8. Look into the three functions `apply()`, `sapply()`, and `lapply()`. What are the differences between these functions?
 
-Repeat 7, but this time use the three above functions.
++ Repeat what you did in 7, but this time use the three above functions.
 Using each of these functions, examine what type of object they produce.
@@
 \bigskip
+
 9.
-Now we will explore `ggplot2`, which is a great (but sometimess tedious) package for plotting data and results.
+Now we will explore `ggplot2`, which is a great package for plotting data and results.
 You will probably grow to have a love/hate relationship with `ggplot2`.
-For help with this package, reference this cheat sheet: https://rstudio.com/wp-content/uploads/2016/11/ggplot2-cheatsheet-2.1.pdf
+For help with this package, reference this cheat sheet: https://res.cloudinary.com/dyd911kmh/image/upload/v1666806657/Marketing/Blog/ggplot2_cheat_sheet.pdf.
+Here is another useful link to a variety of graphs created using ggplot2: 
+http://r-graph-gallery.com/
 
-Using `ggplot`, create a scatterplot with the `swiss` dataset that plots `Education` against `Examination`.
-Add axes labels and a title.
-Does it look like there might be a relationship between these two variables?
++ Reminder, load a package using library().
+
++ Using `ggplot`, create a scatterplot with the `swiss` dataset that plots `Education` against `Examination`. Add axes labels and a title.
+
++ Does it look like there might be a relationship between these two variables?
 ```{r}
@@
 \bigskip
+
 10.
@@
 
-For example, if I wanted to write a function that evaluated the equation \[f(x, y) = xy \] I could write one of the following options:
++ For example, if I wanted to write a function that evaluated the equation \[f(x, y) = xy \] I could write one of the following options:
 
@@
 
-Now your turn. Write a function that evaluates the following equation:
++ Now your turn. Write a function that evaluates the following equation:
 \[f(x, y, z) = x^2 + 3xy + \sin(z)
 \]
-Have this function takes a vector as its single argument (see example 1), made up of the three parameters ($x, y, z$). Evaluate the function at $x = 3/4, y = 13, z = 3.4$.
+
++ Have this function takes a vector as its single argument (see example 1), made up of the three parameters ($x, y, z$). Evaluate the function at $x = 3/4, y = 13, z = 3.4$.
 
@@
 \bigskip
-11. Let's try out some linear algebra. Below are 2 vectors, a and b, and two matrices A and B. Use them for this section.
-Here is a link to a linear algebra cheat sheet in R (and other languages): 
+
+11. **Exercise**: Let's try out some linear algebra. Below are 2 vectors, a and b, and two matrices A and B. Use them for this section.
+
++ Here is a link to a linear algebra cheat sheet in R (and other languages): 
 https://github.com/scalanlp/breeze/wiki/Linear-Algebra-Cheat-Sheet
+
 ```{r, include = TRUE}
@@
 
-What is the dot product of $\mathbf{a}$ and $\mathbf{b}$? 
-Does the result change if we take the dot product of $\mathbf{b}$ and $\mathbf{a}$?
+1. What is the dot product of $\mathbf{a}$ and $\mathbf{b}$?
 
-What are the dimensions for the matrix $\mathbf{A}\mathbf{B}$? Is this new matrix invertible?
+2. Does the result change if we take the dot product of $\mathbf{b}$ and $\mathbf{a}$?
 
-What are the eigenvectors and eigenvalues of this new matrix?
+3. What are the dimensions for the matrix $\mathbf{A}\mathbf{B}$? Is this new matrix invertible?
 
-Verify that the values produced are actually eigenvalues.
+4.
+What are the eigenvectors (vector whose direction remains unchanged when a linear transformation is applied to it) and eigenvalues (scalars associated with a linear system of equations) of this new matrix?
+
+6. Verify that the values produced are actually eigenvalues.
 
@@
 
-#verifying that they are eigenvalues
+#verifying that they are eigenvalues by checking that det(AB−λiI)=0 for each value
 sapply(eigen_AB$values, function(x) det(AB - diag(2)*x))

### 2023-01-13 Lab 1  R Review/Lab1_R_Review.Rmd
Add files via upload

--- 2023-01-04 Lab1_R_Review.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(cache = TRUE, include = FALSE)
+knitr::opts_chunk$set(cache = TRUE, include = TRUE, echo=TRUE, eval=FALSE)
 ```
@@
 \medskip
+
 1.
@@
 
-rm(list=ls()) #clear global environ
+rm(list=ls()) #clear global environment
 ```
@@
 \bigskip
-2. First, create a vector of ten 5s using the function `rep()`. 
-Second, create a vector that counts from 1 to 10 - do this in two ways, one using `c()` and one using `seq()`.
 
-Third, create a vector of length 10 filled with NAs. Finally, uses `cbind()` to bind theses four vectors together to create a 10 x 14 (10 rows, 4 columns) matrix.
+2. To start, create a vector of ten 5s using the function `rep()`.
 
-Remember, if you don't know how to use a function, type ? function name into your console to see the help file (e.g., `?seq()`).
++ Second, create a vector that counts from 1 to 10 - do this in two ways, one using `c()` and one using `seq()`.
 
-Try binding the vectors together with `rbind()` instead. What happens?
++ Third, create a vector of length 10 filled with NAs. Finally, uses `cbind()` to bind theses four vectors together to create a 10 x 14 (10 rows, 4 columns) matrix.
+
++ Remember, if you don't know how to use a function, type ? function name into your console to see the help file (e.g., `?seq()`).
+
++ Try binding the vectors together with `rbind()` instead. What happens?
 
@@
 \bigskip
+
 5. Practice: Load in the `swiss` dataset. A description for this data can be found here: https://stat.ethz.ch/R-manual/R-patched/library/datasets/html/swiss.html
 
-How many variables does it have? How many observations?
++ How many variables does it have?
 
-The variable `Catholic` is the \% of Catholic individuals, as opposed to Protestant in each province.
++ How many observations?
+
++ The variable `Catholic` is the \% of Catholic individuals, as opposed to Protestant in each province.
 First look at the summary of this variable.
@@
 
-Create a matrix filled with random values that has 4 rows and 2 columns. Use `length()` on this matrix. Why might this be? Use `?` to read the documentation for `length()`.
++ Create a matrix filled with random values that has 4 rows and 2 columns. Use `length()` on this matrix. Why might this be? Use `?` to read the documentation for `length()`.
 
@@
 
-#using apply (2 is required because apply needs an arguement to margin (2 indicates what?))
+#using apply (2 is required because apply needs an argument to margin (2 indicates what?))
 medians2 <- ____(swiss, 2, median)
@@
 \bigskip
+
 9.
@@
 
-Reminder, load a package using library().
++ Reminder, load a package using library().
 
-Using `ggplot`, create a scatterplot with the `swiss` dataset that plots `Education` against `Examination`. Add axes labels and a title.
++ Using `ggplot`, create a scatterplot with the `swiss` dataset that plots `Education` against `Examination`. Add axes labels and a title.
 
-Does it look like there might be a relationship between these two variables?
++ Does it look like there might be a relationship between these two variables?
 ```{r}
@@
 
-Now your turn. Write a function that evaluates the following equation:
++ Now your turn. Write a function that evaluates the following equation:
 \[f(x, y, z) = x^2 + 3xy + \sin(z)
@@
 
-Have this function takes a vector as its single argument (see example 1), made up of the three parameters ($x, y, z$). Evaluate the function at $x = 3/4, y = 13, z = 3.4$.
++ Have this function takes a vector as its single argument (see example 1), made up of the three parameters ($x, y, z$). Evaluate the function at $x = 3/4, y = 13, z = 3.4$.
 
-```{r, echo = FALSE}
+```{r}
 
@@
 \bigskip
-11. Let's try out some linear algebra. Below are 2 vectors, a and b, and two matrices A and B. Use them for this section.
 
-Here is a link to a linear algebra cheat sheet in R (and other languages): 
+11. **Exercise**: Let's try out some linear algebra. Below are 2 vectors, a and b, and two matrices A and B. Use them for this section.
+
++ Here is a link to a linear algebra cheat sheet in R (and other languages): 
 https://github.com/scalanlp/breeze/wiki/Linear-Algebra-Cheat-Sheet
 
-```{r, include = TRUE}
+```{r}
 #a
@@
 
-What is the dot product of $\mathbf{a}$ and $\mathbf{b}$? 
-Does the result change if we take the dot product of $\mathbf{b}$ and $\mathbf{a}$?
+1. What is the dot product of $\mathbf{a}$ and $\mathbf{b}$?
 
-What are the dimensions for the matrix $\mathbf{A}\mathbf{B}$? Is this new matrix invertible?
+2. Does the result change if we take the dot product of $\mathbf{b}$ and $\mathbf{a}$?
 
-What are the eigenvectors(vector whose direction remains unchanged when a linear transformation is applied to it) and eigenvalues (scalars associated with a linear system of equations) of this new matrix?
+3. What are the dimensions for the matrix $\mathbf{A}\mathbf{B}$? Is this new matrix invertible?
 
-Verify that the values produced are actually eigenvalues.
+4. What are the eigenvectors and eigenvalues of this new matrix?
 
-```{r, echo = FALSE}
+6. Verify that the values produced are actually eigenvalues.
+
+```{r}
 #dot product

### 2023-01-13 Lab 2  Tidyverse Review/Lab2_Tidyverse_Review_Answers.Rmd
Add files via upload

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(cache = TRUE, include = FALSE)
+knitr::opts_chunk$set(cache = TRUE, include = TRUE, echo=TRUE, eval=TRUE)
 ```
@@
 
-2.The tidyverse allows us to use the piping operator `%>%`, which is a useful piece of code (base R can do this too with |>) .
+2. The tidyverse allows us to use the piping operator `%>%`, which is a useful piece of code (base R can do this too with |>) .
 
-It pipes an object (either by itself or created by another function) into the next function in the sequence, as the first argument. This makes the code much easier to read.
++ It pipes an object (either by itself or created by another function) into the next function in the sequence, as the first argument. This makes the code much easier to read.
 
-For example, we could calculate the standard deviation of the `mpg` variable in the `mtcars` dataset in the following ways:
++ For example, we could calculate the standard deviation of the `mpg` variable in the `mtcars` dataset in the following ways:
 
@@
 ```
-That is a very simple example, of course, and using a pipe in this case does not really gain you much.
++ That is a very simple example, of course, and using a pipe in this case does not really gain you much.
 However, pipes and the myriad of tools in the `tidyverse` can make our work much, much easier.
@@
 
-`dplyr` allows us to manipulate our data in a structured way.
++ `dplyr` allows us to manipulate our data in a structured way.
 The key functions are `mutate()`, which creates new variables, `select()`, which allows us to pick variable by characteristics of their names, `filter()`, which allows us to choose rows according to criteria we define, `summarise()`, which allows us to "reduce multiple values down to a single summary", and `arrange()`, which allows us to order the observations according values of 1 or more variables.
 
-RStudio makes useful cheat sheets for their packages, including [`dplyr`](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf).
++ RStudio makes useful cheat sheets for their packages, including [`dplyr`](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf).
 
 ### Star Wars example
-```{r}
+
++ Next, we'll use dplyr to conduct some data manipulation!
++ We start by making our data into a tibble.
+Tibbles have some valuable characteristics and are an evolution of data.frames.
+Tibbles do not match partial variable names, do not convert character vectors to factors, always returns a tibble when [ is used and allows readable variable names.
+```{r, results="hide"}
 #1. Star Wars example
@@
 
-#Tibbles are evolution of data.frame
-#Inherits nominally, but not strictly!
-  #Does not match partial variable names
-  #Does not convert character vectors to factors
-  #Always returns a tibble when [ is used
-  #Allows readable variable names
-
 dplyr::as_tibble(starwars)
-
-  #Now that we have our data imported, we should start exploring it to see exactly what we have
+```
++ Now that we have our data imported, we should start exploring it to see exactly what we have
+```{r}
 
@@
 ```
-## Joins
+
+## Functions
+
++ It is also straightforward to write your own functions using tidyverse principles.
+Below is an example of a function I use quite often that pipes various operators together to summarize data.
+
++ You may also notice that the functions use ...
+(aka dots).
+The dots operator allows you to pass an arbitrary number of additional arguments that aren't explicitly defined in the function (for example, additional arguments to functions nested within the function).
+
++ The single dot (.) refers to whatever is on the left hand side of the pipe at that moment.
 
 ```{r}
+#Creating a function to easily summarize data
+pctgroup <- function(data,...,rd = 1) {
+  dplyr::select(data,...) %>%
+    dplyr::filter(complete.cases(.)) %>% #drop any missing data in this case the . is referring to the dataframe
+    dplyr::group_by(...) %>% #group by
+    dplyr::summarise(n = n()) %>% #summarize the data
+    dplyr::mutate(Percentage = round(100*n/sum(n),rd)) #create a new column Percentage
+}
 
-# Load data 1
-(dates <- read_csv("https://bit.ly/DM22_dates"))
+#Application 
+pctgroup(diamonds, cut) %>% 
+  head(5)
 
-# Load data 2
-(works <- read_csv("https://bit.ly/DM22_works"))
+#Creating a function to first group by then calculate mean of a given variable x
+#enquo() and enquos() are defuse function arguments that are useful when writing nested functions like this.
 
-#Inner join
-#With an inner_join(x, y), only observations that appear in both x and y datasets are returned
+#!! is a bang bang operator it is useful to modify a piece of R code before R evaluates it. In this case the code line is being told that we are using an encoded variable (mean_col) in the summarise.
 
-# Do inner join
-dates_works <- inner_join(dates, works) 
-(dates_works)
+grouped_means <- function(df, x, ...){
+  mean_col <- enquo(x)
+  df %>% 
+    group_by(...) %>% 
+    summarise(mean = mean(!! mean_col))
+}
 
-#Left & Right Join
-#A left_join(x, y) call returns all observations in x, and drops observations that only appear in y
-#Variables in y that do not exist in x are appended to x in the returned tibble.
-
-#The same is true for a right_join(x, y), but with the roles of x and y reversed.
-
-dates_someworks <- left_join(dates, works) 
-(dates_someworks)
-
-dates_allworks <- right_join(dates, works) 
-print(dates_allworks, width=70)
-
-#Full join
-#Finally, a full_join(x, y) returns all observations and all variables in either in x or in y (filled with NAs)
-
-alldates_allworks <- full_join(dates, works) 
-print(alldates_allworks, width=70)
+#Grouping by cut, and color then calculating the mean price
+grouped_means(diamonds, price, cut, color) %>% 
+  head(5)
 
@@
 
-## Pivoting
-Pivoting: Wide -> Long
-With pivot_longer(), you can take values of a variable that are stored as variable names pivot them onto the cells of a single (implicit) variable column
+## Joins
+
++ Often you may wish to join two datasets together.
 
 ```{r}
+library(nycflights13)
+
+# Load data 1
+(band_members)
+
+# Load data 2
+(band_instruments)
+
+```
++ Inner join 
+With an inner_join(x, y), only observations that appear in both x and y datasets are returned
+```{r}
+# Do inner join
+members_instruments <- inner_join(band_members, band_instruments) 
+(members_instruments)
+
+```
+
++ Left & Right Join
++ A left_join(x, y) call returns all observations in x, and drops observations that only appear in y. Variables in y that do not exist in x are appended to x in the returned tibble.
+
++ The same is true for a right_join(x, y), but with the roles of x and y reversed.
+
+```{r}
+members_someinstruments <- left_join(band_members, band_instruments) 
+(members_someinstruments)
+
+members_allinstruments <- right_join(band_members, band_instruments) 
+(members_allinstruments)
+
+```
++ Full join
++ Finally, a full_join(x, y) returns all observations and all variables in either in x or in y (filled with NAs)
+
+```{r}
+allmembers_allinstruments <- full_join(band_members, band_instruments) 
+(allmembers_allinstruments)
+```
+
+## Pivoting
++ Pivoting: Wide -> Long
++ With pivot_longer(), you can take values of a variable that are stored as variable names pivot them onto the cells of a single (implicit) variable column
+
+```{r, results="hide"}
 (relig_income)
@@
 ```
-Pivoting: Long -> Wide
++ Pivoting: Long -> Wide
 
-With pivot_wider(), we move in the opposite direction: your units are distributed over multiple rows, so pivot until you have a single case per row of your tibble
++ With pivot_wider(), we move in the opposite direction: your units are distributed over multiple rows, so pivot until you have a single case per row of your tibble
 
-```{r}
+```{r, results="hide"}
 (fish_encounters) 
@@
 ```
-#Part 2: ggplot
+# Part 2: ggplot
 
-`ggplot` is very important, and requires some time and patience to get to know well.
++ `ggplot` is very important, and requires some time and patience to get to know well.
 We won't have time to dive fully into it, but we will begin reviewing the following code so you become more familiar.
@@
 
-For a comprehensive 'cheat sheet', see: https://www.maths.usyd.edu.au/u/UG/SM/STAT3022/r/current/Misc/data-visualization-2.1.pdf
++ For a comprehensive 'cheat sheet', see: https://www.maths.usyd.edu.au/u/UG/SM/STAT3022/r/current/Misc/data-visualization-2.1.pdf
 
@@
 
-To begin, we can use the 'qplot' or 'ggplot' commands. I use ggplot. It's more customization.
++ To begin, we can use the 'qplot' or 'ggplot' commands. I use ggplot. It's more customization.
 
@@
 
-What are some ways we can make this plot clearer?
++ What are some ways we can make this plot clearer?
 
@@
 
-You can also assign this plot to an object and save that object as a .pdf file.
++ You can also assign this plot to an object and save that object as a .pdf file.
 
@@
 ```
+
 ## Themes
-As you've hopefully seen ggplot is highly customization.
+
++ As you've hopefully seen ggplot is highly customization.
 One way to standardize your graphs and avoid repeating your code to control ancillary parameters (e.g.
@@
 
-You can also create custom themes yourself!
++ You can also create custom themes yourself!
 
@@
 ```
+
 ## Facets
-Using facets lets you create multi-panel plots based off a given variable.
-There are two different ggplot functions to use to facet. 
-```{r}
-#1.
+
++ Using facets lets you create multi-panel plots based off a given variable. There are two different ggplot functions to use to facet.
+
+1.
 facet_wrap() "wraps" a 1d ribbon of panels into 2d.
@@
 This is most useful when you want to facet by a single variable (or two)
+
+```{r}
 
@@
   facet_wrap(~ am)
+```
 
-#2. facet_grid() forms a matrix of panels defined by row and column faceting variables, creating panels based on observed combinations of values in faceting variables
+2. facet_grid() forms a matrix of panels defined by row and column faceting variables, creating panels based on observed combinations of values in faceting variables.
 
+```{r}
 #facet_grid(.~variable) will return facets equal to the levels of variable distributed horizontally
@@
 
-What are other types of common geoms?
+3. What are other types of common geoms?
 
-geom_histogram(): plots a histogram of single value; usually needs modification to number of bins.
++ geom_histogram(): plots a histogram of single value
 
-geom_density(): density plot of single variable.
++ geom_density(): density plot of single variable.
 
-geom_bar(): vertical barplot by categorical variable
++ geom_bar(): vertical barplot by categorical variable
 
-geom_point(): typical two-way scatterplot
++ geom_point(): typical two-way scatterplot
 
-geom_smooth(): simple line interpolator
++ geom_smooth(): simple line interpolator
 
-#Part 3: Practice
-## Tidyverse practice
+# Part 3: Exercise
 
-Use the data below to first split the complete dataset into individual planes and then summarise each plane 
-by counting the number of flights (count = n()) and computing the mean distance flown.
+1. Use the data below to first split the complete dataset into individual planes and then summarise each plane by counting the number of flights (count = n()) and computing the mean distance flown.
 
-Next, determine what city is flown to the most often from NYC.
+2. Next, determine what city is flown to the most often from NYC.
 
-Finally, create a plot to create a barplot visualizing if flights are on time or delayed (defined as later than 5 minutes) by month. Additionally, ensure your plot has meaningful labels.
+3.
+Finally, use ggplot to create a plot to create a barplot visualizing if flights are on time or delayed (defined as later than 5 minutes) by month.
+Additionally, ensure your plot has meaningful labels.
 
-```{r}
+```{r, results="hide"}
 #First, we need to install our data. This dataset includes airline on-time data for all flights departing NYC in 2013. 
@@
 
+plot1
+
 ```

### 2023-01-13 Lab 2  Tidyverse Review/Lab2_Tidyverse_Review.Rmd
Add files via upload

@@
 
-`dplyr` allows us to manipulate our data in a structured way.
++ `dplyr` allows us to manipulate our data in a structured way.
 The key functions are `mutate()`, which creates new variables, `select()`, which allows us to pick variable by characteristics of their names, `filter()`, which allows us to choose rows according to criteria we define, `summarise()`, which allows us to "reduce multiple values down to a single summary", and `arrange()`, which allows us to order the observations according values of 1 or more variables.
 
-RStudio makes useful cheat sheets for their packages, including [`dplyr`](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf).
++ RStudio makes useful cheat sheets for their packages, including [`dplyr`](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf).
 
@@
 
++ Next, we'll use dplyr to conduct some data manipulation!
++ We start by making our data into a tibble.
+Tibbles have some valuable characteristics and are an evolution of data.frames.
+Tibbles do not match partial variable names, do not convert character vectors to factors, always returns a tibble when [ is used and allows readable variable names.
 ```{r}
@@
   #Let's make our data into a tibble
-
-#Tibbles are evolution of data.frame
-#Inherits nominally, but not strictly!
-  #Does not match partial variable names
-  #Does not convert character vectors to factors
-  #Always returns a tibble when [ is used
-  #Allows readable variable names
 
@@
   group_by(species) %>%
-  summarise(avg = mean(mass, na.rm=TRUE)) %>%
+  summarise(avg = mean(____, na.rm=TRUE)) %>%
   arrange(desc(avg))
@@
 Human_BMI<-Star_BMI %>% 
-  filter(species=="Human")
+  filter(____)
 
@@
 starwars %>%
-  unite(name_homeworld, name, homeworld) %>% 
+  unite(name_homeworld, ____) %>% 
   select(name_homeworld)
+
+```
+
+## Functions
+
++ It is also straightforward to write your own functions using tidyverse principles.
+Below is an example of a function I use quite often that pipes various operators together to summarize data.
+
++ You may also notice that the functions use ...
+(aka dots).
+The dots operator allows you to pass an arbitrary number of additional arguments that aren't explicitly defined in the function (for example, additional arguments to functions nested within the function).
+
++ The single dot (.) refers to whatever is on the left hand side of the pipe at that moment.
+```{r}
+#Creating a function to easily summarize data
+pctgroup <- function(data,...,rd = 1) {
+  dplyr::select(data,...) %>%
+    dplyr::filter(complete.cases(.)) %>% #drop any missing data in this case the . is referring to the dataframe
+    dplyr::group_by(...) %>% #group by
+    dplyr::summarise(n = n()) %>% #summarize the data
+    dplyr::mutate(Percentage = round(100*n/sum(n),rd)) #create a new column Percentage
+}
+
+#Application 
+pctgroup(diamonds, cut) %>% 
+  head(5)
+
+#Creating a function to first group by then calculate mean of a given variable x
+#enquo() and enquos() are defuse function arguments that are useful when writing nested functions like this.
+
+#!! is a bang bang operator it is useful to modify a piece of R code before R evaluates it. In this case the code line is being told that we are using an encoded variable (mean_col) in the summarise.
+
+grouped_means <- function(df, x, ...){
+  mean_col <- enquo(x)
+  df %>% 
+    group_by(...) %>% 
+    summarise(mean = mean(!! mean_col))
+}
+
+#Grouping by cut, and color then calculating the mean price
+grouped_means(diamonds, price, cut, color) %>% 
+  head(5)
 
@@
 
++ Often you may wish to join two datasets together.
+
 ```{r}
+library(nycflights13)
 
 # Load data 1
-(dates <- read_csv("https://bit.ly/DM22_dates"))
+(band_members)
 
 # Load data 2
-(works <- read_csv("https://bit.ly/DM22_works"))
+(band_instruments)
 
-#Inner join
-#With an inner_join(x, y), only observations that appear in both x and y datasets are returned
-
+```
++ Inner join 
+With an inner_join(x, y), only observations that appear in both x and y datasets are returned
+```{r}
 # Do inner join
-dates_works <- inner_join(dates, works) 
-(dates_works)
-
-#Left & Right Join
-#A left_join(x, y) call returns all observations in x, and drops observations that only appear in y
-#Variables in y that do not exist in x are appended to x in the returned tibble.
-
-#The same is true for a right_join(x, y), but with the roles of x and y reversed.
-
-dates_someworks <- left_join(dates, works) 
-(dates_someworks)
-
-dates_allworks <- right_join(dates, works) 
-print(dates_allworks, width=70)
-
-#Full join
-#Finally, a full_join(x, y) returns all observations and all variables in either in x or in y (filled with NAs)
-
-alldates_allworks <- full_join(dates, works) 
-print(alldates_allworks, width=70)
+members_instruments <- inner_join(band_members, band_instruments) 
+(members_instruments)
 
@@
 
++ Left & Right Join
++ A left_join(x, y) call returns all observations in x, and drops observations that only appear in y. Variables in y that do not exist in x are appended to x in the returned tibble.
+
++ The same is true for a right_join(x, y), but with the roles of x and y reversed.
+
+```{r}
+members_someinstruments <- left_join(band_members, band_instruments) 
+(members_someinstruments)
+
+members_allinstruments <- right_join(band_members, band_instruments) 
+(members_allinstruments)
+
+```
++ Full join
++ Finally, a full_join(x, y) returns all observations and all variables in either in x or in y (filled with NAs)
+
+```{r}
+allmembers_allinstruments <- full_join(band_members, band_instruments) 
+(allmembers_allinstruments)
+```
+
 ## Pivoting
-Pivoting: Wide -> Long
-With pivot_longer(), you can take values of a variable that are stored as variable names pivot them onto the cells of a single (implicit) variable column
++ Pivoting: Wide -> Long
++ With pivot_longer(), you can take values of a variable that are stored as variable names pivot them onto the cells of a single (implicit) variable column
 
@@
 ```
-Pivoting: Long -> Wide
++ Pivoting: Long -> Wide
 
-With pivot_wider(), we move in the opposite direction: your units are distributed over multiple rows, so pivot until you have a single case per row of your tibble
++ With pivot_wider(), we move in the opposite direction: your units are distributed over multiple rows, so pivot until you have a single case per row of your tibble
 
@@
 
-`ggplot` is very important, and requires some time and patience to get to know well.
++ `ggplot` is very important, and requires some time and patience to get to know well.
 We won't have time to dive fully into it, but we will begin reviewing the following code so you become more familiar.
@@
 ## Themes
-As you've hopefully seen ggplot is highly customization.
++ As you've hopefully seen ggplot is highly customization.
 One way to standardize your graphs and avoid repeating your code to control ancillary parameters (e.g.
@@
 ```
-As you can see there are tons of options that you can set yourself. Alternatively you can use existing gplot_themes. Here are some of the most popular
++ As you can see there are tons of options that you can set yourself. Alternatively you can use existing gplot_themes. Here are some of the most popular
 ```{r}
@@
 ## Facets
-Using facets lets you create multi-panel plots based off a given variable.
-There are two different ggplot functions to use to facet. 
-```{r}
-#1.
++ Using facets lets you create multi-panel plots based off a given variable. There are two different ggplot functions to use to facet.
+
+1.
 facet_wrap() "wraps" a 1d ribbon of panels into 2d.
@@
 
-What are other types of common geoms?
+3. What are other types of common geoms?
 
-geom_histogram(): plots a histogram of single value; usually needs modification to number of bins.
++ geom_histogram(): plots a histogram of single value
 
-geom_density(): density plot of single variable.
++ geom_density(): density plot of single variable.
 
-geom_bar(): vertical barplot by categorical variable
++ geom_bar(): vertical barplot by categorical variable
 
-geom_point(): typical two-way scatterplot
++ geom_point(): typical two-way scatterplot
 
-geom_smooth(): simple line interpolator
++ geom_smooth(): simple line interpolator
 
-# Part 3: Practice
+# Part 3: Exercise
 ## Tidyverse practice
 
-Use the data below to first split the complete dataset into individual planes and then summarise each plane 
-by counting the number of flights (count = n()) and computing the mean distance flown.
+1. Use the data below to first split the complete dataset into individual planes and then summarise each plane by counting the number of flights (count = n()) and computing the mean distance flown.
 
-Next, determine what city is flown to the most often from NYC.
+2. Next, determine what city is flown to the most often from NYC.
 
-Finally, create a plot to create a barplot visualizing if flights are on time or delayed (defined as later than 5 minutes) by month. Additionally, ensure your plot has meaningful labels.
+3.
+Finally, use ggplot to create a plot to create a barplot visualizing if flights are on time or delayed (defined as later than 5 minutes) by month.
+Additionally, ensure your plot has meaningful labels.
 

### 2023-01-13 Lab 3  Fitting LMs/Lab3_LM.Rmd
Add files via upload

@@
 ---
-title: "Lab 3: The Linear Model in R"
+title: 'Lab 3: The Linear Model in R'
 author: "Eric Parajon"
-output: pdf_document
+output:
+  html_document:
+    df_print: paged
+  pdf_document: default
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(cache = TRUE, include = TRUE, eval=FALSE)
+knitr::opts_chunk$set(cache = TRUE, include = TRUE, echo=TRUE, eval=FALSE)
 
@@
 #Installing and loading packages for use
-pacman::p_load(xtable,stargazer,palmerpenguins,tidyverse)
+pacman::p_load(xtable,stargazer,palmerpenguins,tidyverse,texreg)
 
@@
 
-First, we'll simulate some data to use. This will be a very useful skill in the future, so pay attention to what we do here.
++ First, we'll simulate some data to use. This will be a very useful skill in the future, so pay attention to what we do here.
 
-We first create our variables and then stipulate what our effects should be.
++ We first create our variables and then stipulate what our effects should be.
 
-Let's set our seed to 123.
++ Let's set our seed to 123.
 
-For our $\mathbf{x}_1$ variable, let's take 20 draws from a $\mathcal{N}(1.3, 2.3)$ (using the `R` parameterization, where the second parameter is the standard deviation).
++ For our $\mathbf{x}_1$ variable, let's take 20 draws from a $\mathcal{N}(1.3, 2.3)$ (using the `R` parameterization, where the second parameter is the standard deviation).
 
-Let's set $\beta_0$ to be 17, and $\beta_1$ to be 1.3. We also can't forget to simulate $\boldsymbol{\epsilon}$! Let's $\boldsymbol{\epsilon} \sim \mathcal{N}(0, 3.4)$.
++ Let's set $\beta_0$ to be 17, and $\beta_1$ to be 1.3. We also can't forget to simulate $\boldsymbol{\epsilon}$! Let's $\boldsymbol{\epsilon} \sim \mathcal{N}(0, 3.4)$.
 
@@
 
-What would we do in order to simulate a linear relationship with the variables and effects above? Create the outcome variable `y`.
++ What would we do in order to simulate a linear relationship with the variables and effects above? Create the outcome variable `y`.
 ```{r, eval = FALSE}
@@
 
-What do $\mathbf{x}_1$ and $\mathbf{y}$ look like together? Use the `plot` function to plot a scatter plot of $\mathbf{x}_1$ against $\mathbf{y}$.
++ What do $\mathbf{x}_1$ and $\mathbf{y}$ look like together? Use the `plot` function to plot a scatter plot of $\mathbf{x}_1$ against $\mathbf{y}$.
 
@@
 
-Now that we have simulated explanatory variables and set up a 'true' relationship between our explanatory variable (x1) and outcome variable, we can use R's built-in functions to evaluate our linear model.
++ Now that we have simulated explanatory variables and set up a 'true' relationship between our explanatory variable (x1) and outcome variable, we can use R's built-in functions to evaluate our linear model.
 For a linear model, we will want to use the `lm()` function.
@@
 
-Use `summary()` on the model object (`m1` in this case) in order to see some of the most interesting information about the fitted model.
++ Use `summary()` on the model object (`m1` in this case) in order to see some of the most interesting information about the fitted model.
 
@@
 
-We can also combine what we learned about using the tidyverse with linear modeling!
++ We can also combine what we learned about using the tidyverse with linear modeling!
 
-As we dive into learning linear models in `R`, we're going to use some data from the `mtcars`.
++ As we dive into learning linear models in `R`, we're going to use some data from the `mtcars`.
 But first, lets use the `tidyverse` to clean it up a bit.
@@
 
-Great.
++ Great.
 Let's say we want to analyze the role that `disp` and `cyl` play in influencing `mpg`.
@@
 
-Hint: the `dplyr` functions you are interested in here are `filter()` and `select()` - don't forget to call their help files if you need more information (e.g., `?filter()` in your console).
++ Hint: the `dplyr` functions you are interested in here are `filter()` and `select()` - don't forget to call their help files if you need more information (e.g., `?filter()` in your console).
 ```{r, eval = FALSE}
@@
 
-Now that we have our data ready, we can use `lm()`.
++ Now that we have our data ready, we can use `lm()`.
 Fill in the blank in the chunk below where Y is `mpg`, X1 is `cyl`, and X2 is `disp`.
@@
 
-Remember, our data is `new_data`.
++ Remember, our data is called `new_data`.
 
@@
 
-The `summary` function provides lots of valuable information about the model we've created.
++ The `summary` function provides lots of valuable information about the model we've created.
 We can also extract different information if we need, for example, just the coefficients, or just the residuals.
@@
 
-We can also use indexing to get to parts of these vectors - the output from a model is a list object, which means we can use brackets with $ indexing.
++ We can also use indexing to get to parts of these vectors - the output from a model is a list object, which means we can use brackets with $ indexing.
 
@@
 
-What if we wanted to include an interaction term? Try interacting `disp` and `cyl`. 
++ What if we wanted to include an interaction term? Try interacting `disp` and `cyl`. 
 ```{r, eval = FALSE}
@@
 
-We don't actually have to write all of the variables two times; the `lm` function will include the variables involved in the interaction even if we don't specify them individually in the formula.
++ We don't actually have to write all of the variables two times; the `lm` function will include the variables involved in the interaction even if we don't specify them individually in the formula.
 Try this out below, creating model object `m4` and compare it to `m3`.
@@
 
-How might we present these results in a table?
-Both `xtable()` and `stargazer()` are functions that produce LaTeX code that could be copy-pasted into a .tex document or included in an Rmarkdown document.
-Try both of these on one of the model object `m4` we created above and see what happens.
++ How might we present these results in a table?
+Both `xtable()`, `stargazer()`, and `texreg()` are packages that produce LaTeX code that could be copy-pasted into a .tex document or included in an Rmarkdown document.
+Try these on one of the model object `m4` we created above and see what happens.
 Note, these are VERY flexible and can be customized to make very nice tables, the ones we are generating here are just basic.
@@
 stargazer(_______)
+texreg(_______)
 
@@
 
+2.
 When presenting our results, we might be largely concerned with prediction: if I have a model and provide it with 'new' data, what should I expect?
@@
 
-Check out `predict.lm` using ?, as that is the method that `predict` invokes when we feed it an lm object.
++ Check out `predict.lm` using ?, as that is the method that `predict` invokes when we feed it an lm object.
 
@@
 
-Note that when using predict, it is very important that our new observations have variables that bear the same name as in the formula. Otherwise `predict` won't know what to do! Try out the following:
++ Note that when using predict, it is very important that our new observations have variables that bear the same name as in the formula.
+Otherwise `predict` won't know what to do!
+Try out the following:
 ```{r, eval = FALSE}
@@
 
-# Part 3: LM practice
-Using the palmerpenguines dataset, create an OLS model using multiple predictors and then display the results in a publication quality table (using stargazer).
+# Part 3 Exercise: LM practice
+
++ Using the palmerpenguines dataset, create a linear model model using multiple predictors of your choice and then display the results in a publication quality table (using stargazer).
 ```{r}
@@
 
-#Create table
-
+#Create nicely formatted table
 stargazer(____)

### 2023-01-13 Lab 3  Fitting LMs/Lab3_LM_Answers.Rmd
Add files via upload

@@
 ---
-title: "Lab 3: The Linear Model in R"
+title: 'Lab 3: The Linear Model in R: Answer key'
 author: "Eric Parajon"
-output: pdf_document
+output:
+  html_document:
+    df_print: paged
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(cache = TRUE, include = FALSE, echo=TRUE)
+knitr::opts_chunk$set(cache = TRUE, include = TRUE, echo=TRUE, eval=TRUE)
 
@@
 
-What would we do in order to simulate a linear relationship with the variables and effects above? Create the outcome variable `y`.
++ What would we do in order to simulate a linear relationship with the variables and effects above? Create the outcome variable `y`.
 ```{r}
@@
 # Part 2: Combining LM and the tidyverse
++ We can also combine what we learned about using the tidyverse with linear modeling!
+
++ As we dive into learning linear models in `R`, we're going to use some data from the `mtcars`.
+But first, lets use the `tidyverse` to clean it up a bit.
+First, take a look at the mtcars data (its readily available so you don't really need to load it in!).
 
 ```{r}
+library(tidyverse)
 #examine data
@@
 
-Hint: the `dplyr` functions you are interested in here are `filter()` and `select()` - don't forget to call their help files if you need more information (e.g., `?filter()` in your console).
++ Hint: the `dplyr` functions you are interested in here are `filter()` and `select()` - don't forget to call their help files if you need more information (e.g., `?filter()` in your console).
 
@@
 
-Now that we have our data ready, we can use R's built-in functions to evaluate our linear model.
-For a linear model, we will want to use the `lm()` function.
-The basic syntax for this function is `lm(Y ~ X1 + X2)`.
++ Now that we have our data ready, we can use `lm()`.
 Fill in the blank in the chunk below where Y is `mpg`, X1 is `cyl`, and X2 is `disp`.
-Use `summary()` on the model object (`m1` in this case) in order to see some of the most interesting information about the fitted model.
+Use `summary()` on the model object (`m2` in this case) in order to see some of the most interesting information about the fitted model.
 
-Remember, our data is `new_data`.
++ Remember, our data is called `new_data`.
 
@@
 
-What if we wanted to include an interaction term?
++ What if we wanted to include an interaction term?
 ```{r}
@@
 
-How might we present these results in a table?
-Both `xtable()` and `stargazer()` are functions that produce LaTeX code that could be copy-pasted into a .tex document or included in an Rmarkdown document.
-Try both of these on one of the model object `m4` we created above and see what happens.
++ How might we present these results in a table?
+Both `xtable()`, `stargazer()`, and `texreg()` are packages that produce LaTeX code that could be copy-pasted into a .tex document or included in an Rmarkdown document.
+Try these on one of the model object `m4` we created above and see what happens.
 Note, these are VERY flexible and can be customized to make very nice tables, the ones we are generating here are just basic.
-```{r, eval=FALSE, echo=TRUE}
-
+```{r, results='asis'}
+#These produce code for a .pdf that we can copy and paste
 xtable(m4)
-stargazer(m4)
+stargazer(m4,type="html")
+htmlreg(m4)
 
@@
 #check out guides online!
+
 ```
 
-When presenting our results, we might be largely concerned with prediction: if I have a model and provide it with 'new' data, what should I expect?
++ When presenting our results, we might be largely concerned with prediction: if I have a model and provide it with 'new' data, what should I expect?
 How confident are we in that expectation?
@@
 
-# Part 3: LM practice
-Using the palmerpenguines dataset, create an OLS using multiple predictors and then display the results in a publication quality table (using stargazer).
-```{r}
+# Part 3 Exercise: LM practice
+
++ Using the palmerpenguines dataset, create a linear model model using multiple predictors of your choice and then display the results in a publication quality table (using stargazer).
+```{r, results='hide'}
 library(palmerpenguins)
@@
 
-#Putting results in table
+#Putting results in table (latex version)
 stargazer(bill_model,title="Effect of Predictors on Bill Length",digits=3,star.cutoffs = c(0.1,0.05,0.01),
@@
                       "\\textbf{Body mass (grams)}"),
-          notes="\\parbox[t]{12cm}{Coefficients reported from OLS regression models. Significance codes:*$<$ 0.10,**p$<$ 0.05, ***p$<$ 0.01, two-tailed tests.}",
+          notes="\\parbox[t]{12cm}{Coefficients reported from an OLS regression model. Significance codes:*$<$ 0.10,**p$<$ 0.05, ***p$<$ 0.01, two-tailed tests.}",
           notes.append=FALSE,
@@
           no.space = TRUE)
+```
+
+```{r, results='asis'}
+#Putting results in table (html version)
+stargazer(bill_model,title="Effect of Predictors on Bill Length",digits=3,star.cutoffs = c(0.1,0.05,0.01),
+          label="bill_model",
+          covariate.labels = c("\\textbf{Species} Chinstrap",
+                      "\\textbf{Species} Gentoo",
+                    "\\textbf{Male} (reference= Female)",
+                      "\\textbf{Body mass (grams)}"),
+          notes="\\parbox[t]{12cm}{Coefficients reported from an OLS regression model. Significance codes:*$<$ 0.10,**p$<$ 0.05, ***p$<$ 0.01, two-tailed tests. Species reference= Adélie}",
+          notes.append=FALSE,
+          style="apsr",
+          dep.var.caption = c(""),
+          dep.var.labels.include = FALSE,
+          table.layout ="-c-!t-s-n",
+          header=FALSE,
+          notes.align= "l",
+          column.sep.width = "1pt",
+          font.size = "scriptsize",
+          no.space = TRUE,
+          type="html")
 
 ```
+
+```

### 2023-01-13 Lab 4  Graphical Interpretation of LMs/Lab4_Graphical_interpretation_LM.Rmd
Add files via upload

@@
 ---
-title: "Lab 4: Graphical interpretation of LMs"
+title: 'Lab 4: Graphical interpretation of LMs'
 author: "Eric Parajon"
-output: pdf_document
+output:
+  html_document:
+    df_print: paged
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(cache = TRUE, include = TRUE)
+knitr::opts_chunk$set(cache = TRUE, include = TRUE, echo=TRUE, eval=FALSE)
 
@@
 
-#Part 1: Ggplot visualizations
+# Part 1: Ggplot visualizations
+
 Lets begin by returning to the mtcars data set and creating a basic scatterplot of displacement (disp) vs mpg.
@@
 
-Okay great, but what if we wanted to plot the linear relationship between our x and y?
++ Okay great, but what if we wanted to plot the linear relationship between our x and y?
 
@@
 
-Right now, we're using color as the aesthetic mapped to a specific explanatory variable (color represents the number of cylinders). We can play around with this:
++ Right now, we're using color as the aesthetic mapped to a specific explanatory variable (color represents the number of cylinders). We can play around with this:
 
@@
 
-What are the benefits/drawbacks of each of these strategies? Think about how well the plot communicates information about the types of variables plotted as well as the relationship between them.
++ What are the benefits/drawbacks of each of these strategies? Think about how well the plot communicates information about the types of variables plotted as well as the relationship between them.
 
-The above are great places to start out visualizing any potential relationships between your variables.
++ The above are great places to start out visualizing any potential relationships between your variables.
 
-Now let's jump back into the palmerpenguins dataset to perform some more in-depth visualization of regression models.
+2. Now let's jump back into the palmerpenguins dataset to perform some more in-depth visualization of regression models.
 
-Here we will evaluate the association between penguin body mass (measured in grams) and flipper length (measured in mm).
++ Here we will evaluate the association between penguin body mass (measured in grams) and flipper length (measured in mm).
 
-Let's suppose our main research question is to determine whether body mass was associated with flipper length We will set the body mass as our dependent variable and flipper length as our independent variable (or predictor of interest).
++ Let's suppose our main research question is to determine whether body mass was associated with flipper length We will set the body mass as our dependent variable and flipper length as our independent variable (or predictor of interest):
 
-<div align="left">$\large E[Body_mass_{i} | Flipper_length_{i}] = \beta_0 + \beta_1  Flipper_length_{i} + \epsilon,$</div>
+<div align="left">$\large E[Body\_mass_{i} | Flipper\_length_{i}] = \beta_0 + \beta_1  Flipper\_length_{i} + \epsilon,$</div>
 
-where $E[Body_mass_{i} |  Flipper_length_{i}]$ denotes the expected body mass for penguin $i$ given the flipper length of penguin $i$.
++ where $E[Body\_mass_{i} |  Flipper\_length_{i}]$ denotes the expected body mass for penguin $i$ given the flipper length of penguin $i$.
 
@@
 
-Now, we can construct a linear regression model with body mass as the dependent variable and flipper length. as the independent variable (or predictor of interest).
++ Now, we can construct a linear regression model with body mass as the dependent variable and flipper length. as the independent variable (or predictor of interest).
 
@@
 ```
-Lets again plot it but this time break out the results by species.
++ Lets again plot it but this time break out the results by species.
 
@@
 
-where $E[Body\_mass_{i} |  Flipper\_length_{i}], Species{i}]$ denotes the expected body mass for penguin $i$ given the flipper length of penguin $i$ and controlling for species of the penguin $i$.
++ where $E[Body\_mass_{i} |  Flipper\_length_{i}], Species{i}]$ denotes the expected body mass for penguin $i$ given the flipper length of penguin $i$ and controlling for species of the penguin $i$.
 
@@
 
-Okay great, but what if we want to visualize substantive effects? For this we can use our old friend predict()!
++ Okay great, but what if we want to visualize substantive effects? For this we can use our old friend predict()!
 
@@
        color = "Penguin species",
-       shape = "Penguin species")
-subeffects_plot+
+       shape = "Penguin species")+
   plot_theme
@@
 ## Interaction analysis
-Additionally, in other models we may be interested in exploring the interaction between variables. Often these interaction effects are difficult to interpret without graphic aids.
++ Additionally, in other models we may be interested in exploring the interaction between variables. Often these interaction effects are difficult to interpret without graphic aids.
 
-In this example, suppose we are interested in knowing if the effect of body mass on bill length is attenuated by the sex of the penguin.
++ In this example, suppose we are interested in knowing if the effect of body mass on bill length is attenuated by the sex of the penguin.
 We begin by plotting the data and then perform interaction analysis.
 
-Our regression equation looks like this:
++ Our regression equation looks like this:
 
@@
 penguins %>%  
-  # drop rows with missing data
-  drop_na(sex) %>% 
+  drop_na(sex) %>%  # drop rows with missing data
   ggplot(aes(x = body_mass_g, y = bill_length_mm, color = sex)) +
@@
 
-#Part 3: LM practice
-Use the msleep dataset to model how body weight (bodywt) and brain weight (brainwt) influences total REM sleep (sleep_rem).
-Then create a model including an additional predictor (of your choice).
-Create a coefficient plot of the results of the two models and compare them.
+# Part 3: **Exercise** LM practice
+
+1.  Use the msleep dataset to model how body weight (bodywt) and brain weight (brainwt) influences total REM sleep (sleep_rem). 
+2.  Then create a model including an additional predictor (of your choice). 
+3. Create a coefficient plot of the results of the two models and compare them.
 
@@
 
-#Interpret 
+#Interpret the graph
 ```

### 2023-01-13 Lab 4  Graphical Interpretation of LMs/Lab4_Graphical_interpretation_LM_Answers.Rmd
Add files via upload

@@
 ---
-title: "Lab 4: Graphical interpretation of LMs"
+title: 'Lab 4: Graphical interpretation of LMs: Answer Key'
 author: "Eric Parajon"
-output: pdf_document
+output:
+  html_document:
+    df_print: paged
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(cache = TRUE, include = TRUE)
+knitr::opts_chunk$set(cache = TRUE, include = TRUE, echo=TRUE, eval=TRUE)
 
@@
 
-#Part 1: Ggplot visualizations
+# Part 1: Ggplot visualizations
 Lets begin by returning to the mtcars data set and creating a basic plot of Displacement vs mpg.
@@
 carsplot + geom_smooth(method = "lm", se = FALSE, 
-                       color = "darkgrey", size = 0.75)
+                       color = "darkgrey", linewidth = 0.75)
 #geom_smooth is for fitted lines.
@@
 
+Penguin_results
 #Okay great so huge effect! But what about confounders? Perhaps the penguins species is also important.
@@
 
+Penguin_results_2
 ```
 
-Okay great, but what if we want to visualize substantive effects? For this we can use our old friend predict()!
++ Okay great, but what if we want to visualize substantive effects? For this we can use our old friend predict()!
 
@@
        color = "Penguin species",
-       shape = "Penguin species")
-subeffects_plot+
+       shape = "Penguin species")+
   plot_theme
 
+subeffects_plot
 ```
@@
 ```
-#Part 3: LM practice
-Use the msleep dataset to model how body weight (bodywt) and brain weight (brainwt) influences total REM sleep (sleep_rem).
-Then create a model including an additional predictor (of your choice).
-Create a coefficient plot of the results of the two models and compare them.
+# Part 3: **Exercise** LM practice
+
+1.  Use the msleep dataset to model how body weight (bodywt) and brain weight (brainwt) influences total REM sleep (sleep_rem). 
+2.  Then create a model including an additional predictor (of your choice). 
+3. Create a coefficient plot of the results of the two models and compare them.
 
@@
     geom_hline(yintercept=0, color="red", size=.5) +
-    geom_errorbar(aes(ymin=conf.low, ymax=conf.high, width=0), size=.5, position=position_dodge(width=0.5)) +
+    geom_errorbar(aes(ymin=conf.low, ymax=conf.high, width=0), linewidth=.5, position=position_dodge(width=0.5)) +
     geom_point(aes(y=estimate), size=1.75, position = position_dodge(width=0.5)) +
@@
 
+sleep_results_plot
 ```

### 2023-01-13 Lab 10 PoissonNBmodels/Lab10_PoissonNBmodels_AnswerKey.Rmd
Add files via upload

@@
 
-#Predicted Counts
+# Predicted Counts
 
@@
 
+# Exercise
+
+To do on your own:
+
+Below, try using ggplot to plot this information we just generated. Follow these guidelines:
+- X axis: Average number of Democratic Cliques
+- Y axis: Predicted count of Intl. Wars
+- Include a geom_line() for Dem_Cliques
+- Include a geom_ribbon() with CIs (80% confidence interval)
+- Include a legend which specifies what the line is and what the CIs are, try putting that legend on the bottom
+
 ```{r}

### 2023-01-13 Lab 11 Missing data/Lab11_MissingData.Rmd
Add files via upload

@@
 pacman::p_load(mice,
-               VIM) #Visualization
+               VIM,
+               texreg) #Visualization and tables
 
@@
 summary(pool(model_mice))
+
+#Using texreg package to display results in a table
+texreg(pool(model_mice))
+
 ```
@@
 pacman::p_load(Amelia,
-               lubridate,
-               xtable)#For tables
+               lubridate)
 
-#create running time variable, lets just assume we're in 2021
+#create running time variable, lets just assume we're in 2023
 data_miss$time <- make_datetime(month = data_miss$Month,
@@
 
-#Pulling into DF
-pooled_df <- data.frame(summary(pool(models_amelia)))
-
-#Displaying in table
-xtable(pooled_df)
+#Putting into table
+texreg(pool(models_amelia))
 

### 2023-01-13 Lab 11 Missing data/Lab11_MissingData_AnswerKey.Rmd
Add files via upload

@@
 pacman::p_load(mice,
-               VIM) #Visualization
+               VIM,
+               texreg) #Visualization
 
@@
 #try this out with 25 iterations and 10 imputations
-#select animputation method (try pmm) - what does it do differently?
+#select an imputation method (try pmm) - what does it do differently?
 mice_data_2 <- mice(data_miss, #call our data
@@
 summary(mice_data_2)
-
 ```
@@
 summary(pool(model_mice))
+
+#Using texreg package to display
+texreg(pool(model_mice))
 
@@
 
-#create running time variable, lets just assume we're in 2021
+#create running time variable, lets just assume we're in 2023
 data_miss$time <- make_datetime(month = data_miss$Month,
@@
 #let's run the model again, but this time using a for loop
-models_amelia <- list() #create with as many elements that you know exist
-#create a vector of type list of size five
+models_amelia <- vector("list", 5)  #vector for our 5 newly imputed data sets
 
@@
 summary(pool(models_amelia))
+
+#Putting into table
+texreg(pool(models_amelia))
 
@@
 
-**On your own:**
+#Exercise **On your own:**
 
@@
 
+#Putting into table
+texreg(pool(models_africa))
+
 #alternatively!

### 2023-01-13 Lab 6  Diagnostics and simulation based inferences/Lab6_Simulation_Uncertainty.Rmd
Add files via upload

@@
 ```{r, message = FALSE}
-#clear global environ
+#clear global environment
 rm(list=ls())
@@
 
+And last but not least, you can use the performance package (and the see package) to generate really nicely formatted regression assumption plots!
+```{r}
+pacman::p_load(see,
+               performance)
+
+#using the performance package to generate plots (zoom out if you get an error about plot size)
+check_model(m1)
+
+# performance package also has functions to help check assumptions
+check_normality(m1)
+?check_normality
+
+check_heteroskedasticity(m1)
+?check_heteroskedasticity
+
+```
+
 # Practice

### 2023-01-13 Lab 6  Diagnostics and simulation based inferences/Lab6_Simulation_Uncertainty_Answers.Rmd
Add files via upload

@@
 
+And last but not least, you can use the performance package (and the see package) to generate really nicely formatted regression assumption plots!
+```{r}
+pacman::p_load(see,
+               performance)
+
+#using the performance package to generate plots (zoom out if you get an error about plot size)
+check_model(m1)
+
+# performance package also has functions to help check assumptions
+check_normality(m1)
+?check_normality
+
+check_heteroskedasticity(m1)
+?check_heteroskedasticity
+
+```
 # To do on your own:

### 2023-01-13 Lab 8  BinomialModels/Lab8_BinomialModels.Rmd
Add files via upload

@@
 
-# Assignment
+# Exercise
 
-Read in the data 'turnout.csv.' Fit a logit model that predicts voting based on race, education and income.
++ Read in the data 'turnout.csv.' Fit a logit model that predicts voting based on race, education and income.
 
-Now, fit a second model that predicts voting based on race, education, income, and age.
++ Now, fit a second model that predicts voting based on race, education, income, and age.
 Assume that you've hypothesized that age has a curvilinear relationship with voting turnout, such that lower ages and higher ages are less likely to vote.
@@
 
-Compare the model fit of these two options using lrtest, AIC, and BIC. Which do you choose as the better fitting model?
++ Compare the model fit of these two options using lrtest, AIC, and BIC. Which do you choose as the better fitting model?
 
-Simulate and plot predicted probabilities over the range of age, holding all other variables at their mean, for the second model. Plot this.
++ Average-Case Approach: Simulate and plot predicted probabilities over the range of age, holding all other variables at their mean, for the second model. Plot this.
 
-Generate an average predicted effect over the observed values of income for the first model. Plot this too.
++ Observed-Case Approach: Generate an average predicted effect over the observed values of income for the first model. Plot this too.
 
@@
 
+#Which is better?
+
 # Average-Case Approach
+
 #When we combine the average-case approach with simulation, we create a hypothetical design matrix, where we hold all continuous variables at their mean or another justifiable value, and then let the predictor of interest vary along some range.
@@
 # Observed-Case Approach
+
 #When taking the observed-case approach (see Hanmer and Kalkan (2013) for why this is the preferable approach) and using simulation to get confidence intervals, we have to do things slightly differently.
@@
 
-#
 colnames(______) <- c('pe', 'lo', 'hi', 'income')
 
-ggplot()
+ggplot(______)
 

### 2023-01-13 Lab 8  BinomialModels/Lab8_BinomialModels_AnswerKey.Rmd
Add files via upload

@@
 
-Compare the model fit of these two options using lrtest, AIC, and BIC. Which do you choose as the better fitting model?
++ Compare the model fit of these two options using lrtest, AIC, and BIC. Which do you choose as the better fitting model?
 
-Simulate and plot predicted probabilities over the range of age, holding all other variables at their mean, for the second model. Plot this.
++ Average-Case Approach: Simulate and plot predicted probabilities over the range of age, holding all other variables at their mean, for the second model. Plot this.
 
-Generate an average predicted effect over the observed values of income for the first model. Plot this.
++ Observed-Case Approach: Generate an average predicted effect over the observed values of income for the first model. Plot this too.
 

### 2023-01-17 Lab 4  Graphical Interpretation of LMs/Lab4_Graphical_interpretation_LM.Rmd
Add files via upload

@@
 #Installing and loading packages for use
-pacman::p_load(tidyverse,palmerpenguins,ggeffects)
+pacman::p_load(tidyverse,palmerpenguins,ggeffects,ggthemes)
 
@@
   labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
+  scale_size_continuous(name = "Cylinders")
+
+ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
+  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
+  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
+  scale_size_continuous(name = "Cylinders") 
+```
+
++ What are the benefits/drawbacks of each of these strategies? Think about how well the plot communicates information about the types of variables plotted as well as the relationship between them.
+
++ The above are great places to start out visualizing any potential relationships between your variables.
+
+## Brief divergence 1:Themes
+
++ As you've hopefully seen ggplot is highly customization.
+One way to standardize your graphs and avoid repeating your code to control ancillary parameters (e.g.
+color palettes, axes labels, etc.) is to use a ggplot() theme that includes bundles of code.
+This is what we were doing using theme_bw() above.
+
++ You can also create custom themes yourself!
+
+```{r}
+#Here is a theme I use often
+
+plot_theme = theme(text = element_text(size=11),  #set text size to 11
+                   panel.grid.minor = element_blank(), #these next two lines set the grid lines to null
+                   panel.grid.major = element_blank(),
+                   panel.background = element_blank(),#these lines make the background of the plot blank
+                   plot.background = element_rect(fill = "transparent",colour = NA),
+                   axis.text = element_text(color="black"), #setting axis text to black 
+                   axis.line = element_line(colour = "black"),#setting axis lines to black 
+                   strip.text.y = element_text(angle = 0), axis.title.y = element_text(angle=90, hjust=.5, vjust=.5),#adjusting the angle of text
+                   axis.title = element_text(lineheight=1.1), #increasing height of axis title
+                   legend.position = "bottom",  #Placing the legend at the bottom of the figure
+                   legend.key = element_rect(colour = "transparent",  fill = "transparent") #setting legend to transparent
+)
+
+#Using the theme
+ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
+  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
+  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
   scale_size_continuous(name = "Cylinders") +
-  theme_bw()
+  plot_theme
 
+#As you can see there are tons of options that you can set yourself. Alternatively you can use existing gplot_themes. Here are some of the most popular
+#bw
 ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
@@
   theme_bw()
+
+#minimal
+ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
+  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
+  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
+  scale_size_continuous(name = "Cylinders") +
+  theme_minimal()
+
+#classic
+ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
+  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
+  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
+  scale_size_continuous(name = "Cylinders") +
+  theme_classic()
+
+#If you want even more themes you can use the ggthemes library!
+#install.packages("ggthemes")
+library(ggthemes)
+
+#WSJ
+ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
+  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
+  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
+  scale_size_continuous(name = "Cylinders") +
+  theme_wsj()
+
+#538
+ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
+  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
+  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
+  scale_size_continuous(name = "Cylinders") +
+  theme_fivethirtyeight()
+
+#The Economist
+ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
+  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
+  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
+  scale_size_continuous(name = "Cylinders") +
+  theme_economist()
+
 ```
-
-+ What are the benefits/drawbacks of each of these strategies? Think about how well the plot communicates information about the types of variables plotted as well as the relationship between them.
-
-+ The above are great places to start out visualizing any potential relationships between your variables.
 
@@
 
+## Brief divergence 2: Facets
+
++ Using facets lets you create multi-panel plots based off a given variable. There are two different ggplot functions to use to facet.
+
+1.
+facet_wrap() "wraps" a 1d ribbon of panels into 2d.
+Essentially, creating a square grid of panels, with rows and columns defined by combinations of two categorical variables (even if some combinations of faceting values do not exist in data).
+This is most useful when you want to facet by a single variable (or two)
+
+```{r}
+
+carsplot_facet <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
+  geom_point(aes(colour = cyl), size = 2) +
+  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
+  scale_colour_continuous(name = "Cylinders")+
+  theme_bw()+
+  facet_wrap(~ am)
+```
+
+2. facet_grid() forms a matrix of panels defined by row and column faceting variables, creating panels based on observed combinations of values in faceting variables.
+
+```{r}
+#facet_grid(.~variable) will return facets equal to the levels of variable distributed horizontally
+#facet_grid(variable~.) will return facets equal to the levels of variable distributed vertically.
+
+carsplot_facet_2 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
+  geom_point(aes(colour = cyl), size = 2) +
+  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
+  scale_colour_continuous(name = "Cylinders")+
+  theme_bw()+
+  facet_grid(vs ~ am)
+
+carsplot_facet_3 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
+  geom_point(aes(colour = cyl), size = 2) +
+  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
+  scale_colour_continuous(name = "Cylinders")+
+  theme_bw()+
+  facet_grid(am ~ .)
+
+carsplot_facet_4 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
+  geom_point(aes(colour = cyl), size = 2) +
+  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
+  scale_colour_continuous(name = "Cylinders")+
+  theme_bw()+
+  facet_grid(. ~ am)
+```
+
 ## Interaction analysis
-+ Additionally, in other models we may be interested in exploring the interaction between variables. Often these interaction effects are difficult to interpret without graphic aids.
++ Additionally, in other models we may be interested in exploring the interaction between variables. Often these interaction effects are difficult to interpret without graphic aids (like facets).
 

### 2023-01-17 Lab 4  Graphical Interpretation of LMs/Lab4_Graphical_interpretation_LM_Answers.Rmd
Add files via upload

@@
   labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
+  scale_size_continuous(name = "Cylinders")
+
+ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
+  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
+  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
+  scale_size_continuous(name = "Cylinders") 
+```
+
++ What are the benefits/drawbacks of each of these strategies? Think about how well the plot communicates information about the types of variables plotted as well as the relationship between them.
+
++ The above are great places to start out visualizing any potential relationships between your variables.
+
+## Themes
+
++ As you've hopefully seen ggplot is highly customization.
+One way to standardize your graphs and avoid repeating your code to control ancillary parameters (e.g.
+color palettes, axes labels, etc.) is to use a ggplot() theme that includes bundles of code.
+This is what we were doing using theme_bw() above.
+
++ You can also create custom themes yourself!
+
+```{r}
+#Here is a theme I use often
+
+plot_theme = theme(text = element_text(size=11),  #set text size to 11
+                   panel.grid.minor = element_blank(), #these next two lines set the grid lines to null
+                   panel.grid.major = element_blank(),
+                   panel.background = element_blank(),#these lines make the background of the plot blank
+                   plot.background = element_rect(fill = "transparent",colour = NA),
+                   axis.text = element_text(color="black"), #setting axis text to black 
+                   axis.line = element_line(colour = "black"),#setting axis lines to black 
+                   strip.text.y = element_text(angle = 0), axis.title.y = element_text(angle=90, hjust=.5, vjust=.5),#adjusting the angle of text
+                   axis.title = element_text(lineheight=1.1), #increasing height of axis title
+                   legend.position = "bottom",  #Placing the legend at the bottom of the figure
+                   legend.key = element_rect(colour = "transparent",  fill = "transparent") #setting legend to transparent
+)
+
+#Using the theme
+ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
+  geom_point(aes(shape = factor(cyl)), colour = "red", size = 2) +
+  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
   scale_size_continuous(name = "Cylinders") +
-  theme_bw()
+  plot_theme
 
+#As you can see there are tons of options that you can set yourself. Alternatively you can use existing gplot_themes. Here are some of the most popular
+#bw
 ggplot(data=mtcars, aes(x = disp, y = mpg)) + 
@@
 Penguin_results <- ggplot(data=temp_lm, aes(x=term, y=estimate)) + 
-    geom_hline(yintercept=0, color="red", size=.5) +
+    geom_hline(yintercept=0, color="red", linewidth=.5) +
     geom_errorbar(aes(ymin=conf.low, ymax=conf.high, width=0), size=.5, position=position_dodge(width=0.5)) +
@@
 
+## Facets
+
++ Using facets lets you create multi-panel plots based off a given variable. There are two different ggplot functions to use to facet.
+
+1.
+facet_wrap() "wraps" a 1d ribbon of panels into 2d.
+Essentially, creating a square grid of panels, with rows and columns defined by combinations of two categorical variables (even if some combinations of faceting values do not exist in data).
+This is most useful when you want to facet by a single variable (or two)
+
+```{r}
+
+carsplot_facet <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
+  geom_point(aes(colour = cyl), size = 2) +
+  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
+  scale_colour_continuous(name = "Cylinders")+
+  theme_bw()+
+  facet_wrap(~ am)
+```
+
+2. facet_grid() forms a matrix of panels defined by row and column faceting variables, creating panels based on observed combinations of values in faceting variables.
+
+```{r}
+#facet_grid(.~variable) will return facets equal to the levels of variable distributed horizontally
+#facet_grid(variable~.) will return facets equal to the levels of variable distributed vertically.
+
+carsplot_facet_2 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
+  geom_point(aes(colour = cyl), size = 2) +
+  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
+  scale_colour_continuous(name = "Cylinders")+
+  theme_bw()+
+  facet_grid(vs ~ am)
+
+carsplot_facet_3 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
+  geom_point(aes(colour = cyl), size = 2) +
+  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
+  scale_colour_continuous(name = "Cylinders")+
+  theme_bw()+
+  facet_grid(am ~ .)
+
+carsplot_facet_4 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
+  geom_point(aes(colour = cyl), size = 2) +
+  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
+  scale_colour_continuous(name = "Cylinders")+
+  theme_bw()+
+  facet_grid(. ~ am)
+```
+
 ## Interaction analysis
-+ Additionally, in other models we may be interested in exploring the interaction between variables. Often these interaction effects are difficult to interpret without graphic aids.
++ Additionally, in other models we may be interested in exploring the interaction between variables. Often these interaction effects are difficult to interpret without graphic aids (like facets).
 
@@
   coord_flip()+
+  geom_text(aes(y=estimate,label=round(estimate, digits=2)), size=2.5, vjust=-1, position = position_dodge(width=.2))+
   plot_theme

### 2023-01-19 Lab 2  Tidyverse Review/Lab2_Tidyverse_Review_Answers.Rmd
Add files via upload

@@
 
-## Themes
++ Simple barplot example
+```{r, message=FALSE, warning=FALSE}
 
-+ As you've hopefully seen ggplot is highly customization.
-One way to standardize your graphs and avoid repeating your code to control ancillary parameters (e.g.
-color palettes, axes labels, etc.) is to use a ggplot() theme that includes bundles of code.
-This is what we were doing using theme_bw() above.
+ggplot(mtcars, aes(x=as.factor(cyl) )) + 
+     geom_bar() + 
+    labs(title = "Count of Cars By Cylinders", x = "Number of Cylinders", y = "Count")
 
-+ You can also create custom themes yourself!
+ggplot(mtcars, aes(x=as.factor(cyl) )) + 
+    geom_bar() + 
+    labs(title = "Count of Cars By Cylinders (Horizontal)", x = "Number of Cylinders", y = "Count")+
+     coord_flip()
 
-```{r}
-#Here is a theme I use often
+#Grouped barplot
+#Calculate percentage (and counts) of cyl by gear type
+  d1<-pctgroup(mtcars,cyl,gear)
+#Graph
+ggplot(d1, aes(x=cyl,y=Percentage, fill=factor(gear),label=round(Percentage,digits=2))) + 
+      geom_bar(position="dodge", stat="identity")+
+      labs(title = "Percentage of Cars By Cylinders and Gear Type", x = "Number of Cylinders", y = "Percentage",fill="Gear")
 
-plot_theme = theme(text = element_text(size=11),  #set text size to 11
-                   panel.grid.minor = element_blank(), #these next two lines set the grid lines to null
-                   panel.grid.major = element_blank(),
-                   panel.background = element_blank(),#these lines make the background of the plot blank
-                   plot.background = element_rect(fill = "transparent",colour = NA),
-                   axis.text = element_text(color="black"), #setting axis text to black 
-                   axis.line = element_line(colour = "black"),#setting axis lines to black 
-                   strip.text.y = element_text(angle = 0), axis.title.y = element_text(angle=90, hjust=.5, vjust=.5),#adjusting the angle of text
-                   axis.title = element_text(lineheight=1.1), #increasing height of axis title
-                   legend.position = "bottom",  #Placing the legend at the bottom of the figure
-                   legend.key = element_rect(colour = "transparent",  fill = "transparent") #setting legend to transparent
-)
-
-#Using the theme
-ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  plot_theme
-
-#As you can see there are tons of options that you can set yourself. Alternatively you can use existing gplot_themes. Here are some of the most popular
-#bw
-ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_bw()
-
-#minimal
-ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_minimal()
-
-#classic
-ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_classic()
-
-#If you want even more themes you can use the ggthemes library!
-#install.packages("ggthemes")
-library(ggthemes)
-
-#WSJ
-ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_wsj()
-
-#538
-ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_fivethirtyeight()
-
-#The Economist
-ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_economist()
-
-```
-
-## Facets
-
-+ Using facets lets you create multi-panel plots based off a given variable. There are two different ggplot functions to use to facet.
-
-1.
-facet_wrap() "wraps" a 1d ribbon of panels into 2d.
-Essentially, creating a square grid of panels, with rows and columns defined by combinations of two categorical variables (even if some combinations of faceting values do not exist in data).
-This is most useful when you want to facet by a single variable (or two)
-
-```{r}
-
-carsplot_facet <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_bw()+
-  facet_wrap(~ am)
-```
-
-2. facet_grid() forms a matrix of panels defined by row and column faceting variables, creating panels based on observed combinations of values in faceting variables.
-
-```{r}
-#facet_grid(.~variable) will return facets equal to the levels of variable distributed horizontally
-#facet_grid(variable~.) will return facets equal to the levels of variable distributed vertically.
-
-carsplot_facet_2 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_bw()+
-  facet_grid(vs ~ am)
-
-carsplot_facet_3 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_bw()+
-  facet_grid(am ~ .)
-
-carsplot_facet_4 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_bw()+
-  facet_grid(. ~ am)
 ```

### 2023-01-19 Lab 2  Tidyverse Review/Lab2_Tidyverse_Review.Rmd
Add files via upload

@@
 glimpse(starwars)
+
  #What if we want to pull a specific piece of data?
@@
   arrange(desc(avg))
-#Hutts are the heaviest!
 
@@
 #This is easy using tidyr (another package in tidyverse suite)
+?unite
+
 starwars %>%
@@
 ```{r}
-library(nycflights13)
-
 # Load data 1
@@
 ```
++ Simple barplot example
+```{r, message=FALSE, warning=FALSE}
 
-## Themes
-+ As you've hopefully seen ggplot is highly customization.
-One way to standardize your graphs and avoid repeating your code to control ancillary parameters (e.g.
-color palettes, axes labels, etc.) is to use a ggplot() theme that includes bundles of code.
-This is what we were doing using theme_bw() above.
+ggplot(mtcars, aes(x=as.factor(cyl) )) + 
+     geom_bar() + 
+    labs(title = "Count of Cars By Cylinders", x = "Number of Cylinders", y = "Count")
 
-+ You can also create custom themes yourself!
+ggplot(mtcars, aes(x=as.factor(cyl) )) + 
+    geom_bar() + 
+    labs(title = "Count of Cars By Cylinders (Horizontal)", x = "Number of Cylinders", y = "Count")+
+     coord_flip()
 
-```{r}
-#Here is a theme I use often
+#Grouped barplot
+#Calculate percentage (and counts) of cyl by gear type
+  d1<-pctgroup(mtcars,cyl,gear)
+#Graph
+ggplot(d1, aes(x=cyl,y=Percentage, fill=factor(gear),label=round(Percentage,digits=2))) + 
+      geom_bar(position="dodge", stat="identity")+
+      labs(title = "Percentage of Cars By Cylinders and Gear Type", x = "Number of Cylinders", y = "Percentage",fill="Gear")
 
-plot_theme = theme(text = element_text(size=11),  #set text size to 11
-                   panel.grid.minor = element_blank(), #these next two lines set the grid lines to null
-                   panel.grid.major = element_blank(),
-                   panel.background = element_blank(),#these lines make the background of the plot blank
-                   plot.background = element_rect(fill = "transparent",colour = NA),
-                   axis.text = element_text(color="black"), #setting axis text to black 
-                   axis.line = element_line(colour = "black"),#setting axis lines to black 
-                   strip.text.y = element_text(angle = 0), axis.title.y = element_text(angle=90, hjust=.5, vjust=.5),#adjusting the angle of text
-                   axis.title = element_text(lineheight=1.1), #increasing height of axis title
-                   legend.position = "bottom",  #Placing the legend at the bottom of the figure
-                   legend.key = element_rect(colour = "transparent",  fill = "transparent") #setting legend to transparent
-)
-
-#Using the theme
-ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  plot_theme #no () after the custom theme call
-```
-+ As you can see there are tons of options that you can set yourself. Alternatively you can use existing gplot_themes. Here are some of the most popular
-```{r}
-#bw
-ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_bw()
-
-#minimal
-ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_minimal()
-
-#classic
-ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_classic()
-
-#If you want even more themes you can use the ggthemes library!
-#install.packages("ggthemes")
-library(ggthemes)
-
-#WSJ
-ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_wsj()
-
-#538
-ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_fivethirtyeight()
-
-#The Economist
-ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_economist()
-
-```
-## Facets
-+ Using facets lets you create multi-panel plots based off a given variable. There are two different ggplot functions to use to facet.
-
-1.
-facet_wrap() "wraps" a 1d ribbon of panels into 2d.
-Essentially, creating a square grid of panels, with rows and columns defined by combinations of two categorical variables (even if some combinations of faceting values do not exist in data).
-This is most useful when you want to facet by a single variable (or two)
-
-```{r}
-
-carsplot_facet <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_bw()+
-  facet_wrap(~ am)
-```
-
-2. facet_grid() forms a matrix of panels defined by row and column faceting variables, creating panels based on observed combinations of values in faceting variables.
-
-```{r}
-#facet_grid(.~variable) will return facets equal to the levels of variable distributed horizontally
-#facet_grid(variable~.) will return facets equal to the levels of variable distributed vertically.
-
-carsplot_facet_2 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_bw()+
-  facet_grid(vs ~ am)
-
-carsplot_facet_3 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_bw()+
-  facet_grid(am ~ .)
-
-carsplot_facet_4 <- ggplot(data = mtcars, aes(x = disp, y = mpg)) + 
-  geom_point(aes(colour = cyl), size = 2) +
-  labs(title = "Car Stuff", x = "Displacement (cc)", y = "Miles per Gallon") +
-  scale_colour_continuous(name = "Cylinders")+
-  theme_bw()+
-  facet_grid(. ~ am)
 ```

### 2023-01-21 Lab 2  Tidyverse Review/Lab2_Tidyverse_Review_Answers.Rmd
Add files via upload

@@
       geom_bar(position="dodge", stat="identity")+
+      geom_text(size=4,vjust=-0.5,hjust=0.75,position = position_dodge(width = 1))+
       labs(title = "Percentage of Cars By Cylinders and Gear Type", x = "Number of Cylinders", y = "Percentage",fill="Gear")
@@
 
-plot1<-ggplot(data=flight_delays, aes(x = month,y=n, fill = dep_type)) +
+plot1<-ggplot(data=flight_delays, aes(x = factor(month),y=n, fill = dep_type)) +
   geom_bar(stat="identity")+
-  labs(main="Frequency of On Time or Delayed Arrivals by Month", y="Count", fill="Type")
+  labs(title="Frequency of On Time or Delayed Arrivals by Month", y="Count", fill="Type", x="Month")
 

### 2023-02-15 Lab 5  Optim/Lab5_Optim.Rmd
Add files via upload

@@
 ---
-title: "Lab 5: Optimization"
+title: 'Lab 5: Optimization'
 author: "Eric Parajon"
-output: pdf_document
+output:
+  html_document:
+    df_print: paged
 ---
@@
 
-pacman::p_load(tidyverse,broom)
 ```
@@
 Let's say we have a function $f(x)$ - we want to know what value of $x$ will maximize or minimize that function.
-In other words, let us say that the local maximum (or minimum) of a functtion is called $x^*$ (because it's the star on top of the plot of the function).
+In other words, let us say that the local maximum (or minimum) of a function is called $x^*$ (because it's the star on top of the plot of the function).
 Then we could say, for maximization, that:
@@
 Why?
-(Think of Zeno's Paradox.) Thus, we need to either (or both) cap the number of interations or set some convergence criteria.
+(Think of Zeno's Paradox and asymptotes.) Thus, we need to either (or both) cap the number of iterations or set some convergence criteria.
 For simplicity's sake, we'll just cap the maximum number of iterations here.
@@
 Don't worry so much about this code, but look at the plot.
-```{r, include = TRUE, message = FALSE, warning = FALSE}
+
+```{r, include=TRUE}
+rm(list=ls())
+
+#Installing and loading packages
+pacman::p_load(tidyverse,broom)
+```
+
+```{r, include = TRUE, message = FALSE, warning = FALSE,fig.width=11, fig.height=5}
 
@@
 
-#Using `optim()` to maximize the log likelihood
+# Using `optim()` to maximize the log likelihood
 
@@
 
-#just another way to simulate data
+#just another way to simulate data using matrix multiplication
 y <- rnorm(100, mean = X%*%theta_true[1:2], sd = theta_true[3])

### 2023-02-15 Lab 5  Optim/Lab5_Optim_AnswerKey.Rmd
Add files via upload

@@
 ---
-title: "Lab 5: Optimization: Answers"
+title: 'Lab 5: Optimization: Answers'
 author: "Eric Parajon"
-output: pdf_document
+output:
+  html_document:
+    df_print: paged
 ---
-
 ```{r setup, include=FALSE}
@@
 
-#Installing and loading packages for use
-pacman::p_load(tidyverse,broom)
 ```
@@
 
-```{r, include = TRUE, message = FALSE, warning = FALSE}
+```{r, include=TRUE}
+rm(list=ls())
+
+#Installing and loading packages
+pacman::p_load(tidyverse,broom)
+```
+
+```{r, include = TRUE, message = FALSE, warning = FALSE,fig.width=11, fig.height=5}
 

### 2023-02-15 Lab 11 Missing data/Lab11_MissingData.Rmd
Add files via upload

@@
 ```{r, eval = FALSE}
-#true model
-summary(model_true)
+# Estimate models
+mod <- list()
+mod[['True']] <- model_true
+mod[['Listwise deletion']] <- model_listwise
+mod[['Recoding']] <- model_recode
+mod[['Mice']] <- pool(model_mice)
+mod[['Amelia']] <- pool(models_amelia)
 
-#listwise
-summary(model_listwise)
+texreg(mod)
 
-#recoding
-summary(model_recode)
-
-#mice
-summary(pool(model_mice))
-
-#Amelia
-summary(pool(models_amelia))
 ```

### 2023-02-15 Lab 11 Missing data/Lab11_MissingData_AnswerKey.Rmd
Add files via upload

@@
 texreg(pool(model_mice))
-
 ```
@@
 ```{r}
-#true model
-summary(model_true)
 
-#listwise
-summary(model_listwise)
+# Estimate models
+mod <- list()
+mod[['True']] <- model_true
+mod[['Listwise deletion']] <- model_listwise
+mod[['Recoding']] <- model_recode
+mod[['Mice']] <- pool(model_mice)
+mod[['Amelia']] <- pool(models_amelia)
 
-#recoding
-summary(model_recode)
-
-#mice
-summary(pool(model_mice))
-
-#Amelia
-summary(pool(models_amelia))
+texreg(mod)
 

### 2023-02-15 Lab 11 Missing data/Lab11_MissingData.Rmd
Add files via upload

--- 2023-01-13 Lab 11 Missing data/Lab11_MissingData.Rmd
@@
 ---
-title: 'Lab 11 - Missing Data and Imputation'
+title: "Lab 11 - Missing Data and Imputation"
 author: "Eric Parajon"
@@
 - \usepackage{graphicx}
-output: pdf_document
+output:
+  html_document:
+    df_print: paged
 ---
@@
 
-#Using texreg package to display results in a table
-texreg(pool(model_mice))
+#Using htmlreg package to display results in a table
+htmlreg(pool(model_mice))
 
@@
 
-```{r, message = FALSE}
+```{r, message = FALSE, results='asis'}
 
@@
 pacman::p_load(Amelia,
-               lubridate)
+               lubridate,
+               mice,
+               texreg)
 
@@
 
-#Pooling models (using mice)
 summary(pool(models_amelia))
@@
 #Putting into table
-texreg(pool(models_amelia))
+htmlreg(pool(models_amelia))
 
@@
 Compare how models from mice, Amelia, and the above methods perform to our true data:
-```{r, eval = FALSE}
-#true model
-summary(model_true)
+```{r, eval = FALSE, results='asis'}
+# Estimate models
+mod <- list()
+mod[['True']] <- model_true
+mod[['Listwise deletion']] <- model_listwise
+mod[['Recoding']] <- model_recode
+mod[['Mice']] <- pool(model_mice)
+mod[['Amelia']] <- pool(models_amelia)
 
-#listwise
-summary(model_listwise)
+htmlreg(mod)
 
-#recoding
-summary(model_recode)
-
-#mice
-summary(pool(model_mice))
-
-#Amelia
-summary(pool(models_amelia))
 ```
 
-**On your own:**
+# Exercise **On your own:**
 
@@
 #use Amelia to generate 5 imputed datasets
-#use ts and cs so imputation algerithem knows which variables represent countries and years
+#use ts and cs so imputation algorithm knows which variables represent countries and years
 
@@
 #report results, including standard errors
+htmlreg(pool(___))
+
+#or if knitting to pdf
+#texreg(pool(___))
 

### 2023-02-15 Lab 11 Missing data/Lab11_MissingData_AnswerKey.Rmd
Add files via upload

@@
 ---
-title: 'Lab 11 - Missing Data and Imputation:Answers'
+title: "Lab 11 - Missing Data and Imputation:Answers"
 author: "Eric Parajon"
@@
 ```{r}
-#clear global environ
+#clear global environment
 rm(list=ls())
@@
 
-```{r, eval = FALSE}
+```{r, eval = TRUE, results='asis'}
 #Load in packages
@@
 #what is this showing us?
+```
 
+```{r, eval = TRUE, results='hide'}
 #MICE TIME
-mice_data <- mice(data_miss, #call our data
+mice_data <- mice::mice(data_miss, #call our data
                   m=5, #number of multiple imputations
                   maxit=50, #number of iterations
-                  # method = ? mice will decide the most appropriate method based on your data
                   seed=500)
@@
 So this is great, but then what do we do if we want to run a model? We have five data sets. One option is to pool the data this way:
-```{r}
+```{r, eval = TRUE, results='asis'}
 #pooling
 model_mice <- with(mice_data,lm(Temp~ Ozone + Solar.R + Wind))
+
 summary(pool(model_mice))
@@
 #Using texreg package to display
-texreg(pool(model_mice))
+htmlreg(pool(model_mice))
 ```
@@
 
-```{r}
+```{r, eval = TRUE, results='asis'}
 #Load in packages
@@
 Compare how models from mice, Amelia, and the above methods perform to our true data:
-```{r}
+```{r, eval = TRUE, results='asis'}
 
@@
 
-texreg(mod)
+htmlreg(mod)
 
@@
 
-#Exercise **On your own:**
+# Exercise **On your own:**
 
@@
 
-```{r}
+```{r, eval = TRUE, results='asis'}
 library(Amelia)
@@
 #use Amelia to generate 5 imputed datasets
-#use ts and cs so imputation algs knows countries and years
+#use ts and cs so imputation algorithm also knows countries and years
 
@@
 #Putting into table
-texreg(pool(models_africa))
+
+coefficient.names <-c("Intercept","Inflation","Trade","Civil Liberty","Population")
+
+htmlreg(pool(models_africa),custom.coef.names = coefficient.names)
 

### 2023-02-15 Lab 5  Optim/Lab5_Optim.Rmd
Add files via upload

@@
 ```{r, include = TRUE, message = FALSE, warning = FALSE,fig.width=11, fig.height=5}
+#first generate our data (remember to set the seed)
+set.seed(123)
+x <- rpois(100, 5) #sample from a poisson
+xbar <- mean(x) #mean of the poisson samples
+n <- length(x) #n obs
 
-#first generate our data (remember to set the seed)
-
-set.seed(123)
-x <- rpois(100, 5)
-xbar <- mean(x)
-n <- length(x)
+#Here we will create several nested functions
 
@@
 
-Funcall <- function(f, ...) f(...)
+Funcall <- function(f, ...) f(...) #Remember the dots! The dots operator allows you to pass an arbitrary number of additional arguments that aren't explicitly defined in the function
+
 Iterate <- function(f, n = 1) {
@@
 g <- Vectorize(g, "n") #vectorize our iterator with respect to n
+
+#calling the function g that we created above to find g(x)
 iterates <- g(10, 1:7)
@@
 The algorithm continues (for as many iterations as you specify) until the parameter values stabilize, or $\theta_{n+1}$ and $\theta_{n}$ become approximately equal, and the root is approximated.
-If this doesn't make sense to you, watch \href{https://www.youtube.com/watch?v=-5e2cULI3H8}{this video}.
+If this doesn't make sense to you, watch [this video](https://www.youtube.com/watch?v=-5e2cULI3H8).
 
@@
 \newpage
-#Practice on your own:
+# Practice on your own:
 Now, on your own, try doing this with `mtcars` data, instead of generated data. Let's say we want to set our $y$ value as `mpg`, $x1$ as `wt`, and $x2$ as `gear`.

### 2023-02-15 Lab 5  Optim/Lab5_Optim_AnswerKey.Rmd
Add files via upload

@@
 
+#Here we will create several nested functions
+
 #our score function (first derivative of the log likelihood)
@@
 The algorithm continues (for as many iterations as you specify) until the parameter values stabilize, or $\theta_{n+1}$ and $\theta_{n}$ become approximately equal, and the root is approximated.
-If this doesn't make sense to you, watch \href{https://www.youtube.com/watch?v=-5e2cULI3H8}{this video}.
+If this doesn't make sense to you,watch [this video](https://www.youtube.com/watch?v=-5e2cULI3H8).
+
 We can see that by the 7th NR iteration, we are quite close to the root (5.1).This method is good for finding local optima as well, if you have a multimodal case for example.

### 2023-02-17 Lab 5  Optim/Lab5_Optim.Rmd
Add files via upload

@@
 #first generate our data (remember to set the seed)
-set.seed(123)
+set.seed(1017)
 x <- rpois(100, 5) #sample from a poisson

### 2023-02-22 Lab 6  Diagnostics and simulation based inferences/Lab6_Simulation_Uncertainty.Rmd
Add files via upload

@@
 author: "Eric Parajon"
-output: pdf_document
+output:
+  html_document:
+    df_print: paged
 ---
@@
 
-Today, we will use simulation to help us assess the uncertainty around our parameter estimates and will then incorporate this uncertainty into the predicted values of our outcome variable.
+Today, we will use **simulation** to help us assess the uncertainty around our parameter estimates and will then incorporate this uncertainty into the predicted values of our outcome variable.
 
-We will be examining **Regression Diagnostics**, or ways to examine if our regression meets necessary assumptions, and if anything weird (e.g., outliers, etc.) is going on.
+We will be also be examining **Regression Diagnostics**, or ways to examine if our regression meets necessary assumptions, and if anything weird (e.g., outliers, etc.) is going on.
 Remember the four key assumptions of a linear regression:
@@
 
-Download the "aid_data_95_ee.csv" in the Labs folder on Sakai and then read it into R:
+Download the "aid_data_95_ee.csv" in the Labs folder on Canvas and then read it into R:
 ```{r, warning = F, message = F}
 #remember to set your working directory!
-setwd("______")
+#setwd("______")
 
@@
 # Regression Diagnostics
+Next we turn to another important topic evaluating your regressions!
 
@@
 And last but not least, you can use the performance package (and the see package) to generate really nicely formatted regression assumption plots!
-```{r}
+```{r,fig.width=10, fig.height=7}
 pacman::p_load(see,
@@
 
-#confidence intervals
+#confidence intervals (showing both methods)
 t(apply(____, 2, quantile, prob = c(.1, .9)))
@@
                    ))
+ ## Compute point estimates (estimation uncertainty)
 
@@
 
+#Creating prediction df
 pred_df_guns <- data.frame(PP = robbery_hat_quant[2, ],
@@
 
+#Plotting
 ggplot(data =____) +
   geom_ribbon(aes(____fill = Law), alpha = 0.5) +
-  geom_line()
+  geom_line()+
+  xlab("____") +
+  ylab("____") +
+  theme_bw()
 

### 2023-02-22 Lab 6  Diagnostics and simulation based inferences/Lab6_Simulation_Uncertainty_Answers.Rmd
Add files via upload

@@
 title: "Lab 6 - Diagnostics and simulation based inferences: Answers"
-author: "TA: Eric Parajon"
-output: pdf_document
+author: "Eric Parajon"
+output:
+  html_document:
+    df_print: paged
 ---
@@
 
-Today, we will use simulation to help us assess the uncertainty around our parameter estimates and will then incorporate this uncertainty into the predicted values of our outcome variable.
+Today, we will use **simulation** to help us assess the uncertainty around our parameter estimates and will then incorporate this uncertainty into the predicted values of our outcome variable.
 
-```{r}
+We will be also be examining **Regression Diagnostics**, or ways to examine if our regression meets necessary assumptions, and if anything weird (e.g., outliers, etc.) is going on.
+Remember the four key assumptions of a linear regression:
+
+1) Linearity of the data - the relationship between x and y is linear
+
+2) Normality of residuals - the residual errors are normally distributed
+
+3) Homogeneity of residual variance - the residuals have CONSTANT variance (homoscedasticity)
+
+4) Independence of residual error terms
+
+The following tests will help us see if these assumptions hold, and will examine a number of possible issues.
+These include (but are not limited to) non-linearity of the data, heteroscedasticity (non-constant variance in the errors), and presence of influential values (outliers and high-leverage points).
+
+```{r, message = FALSE}
 #clear global environ
@@
 
-Download the "aid_data_95_ee.csv" in the Labs folder on Sakai and then read it into R:
+Download the "aid_data_95_ee.csv" in the Labs folder on Canvas and then read it into R:
 ```{r, warning = F, message = F}
@@
 
-set.seed(53)
+set.seed(123)
 
@@
 #confidence intervals
+  #Method 1
 t(apply(samp_beta_guns, 2, quantile, prob = c(.1, .9)))
-confint(guns_model, level = .8)
+  #Method 2
+CI<-confint(guns_model, level = .8)
+
+CI
+
+#Now doing the Monte Carlo inference for average prediction
+# to get the predicted robbery rate as the population density varies from its 10th percentile to its 90th percentile dependent on if a shall carry law is in effect
 
@@
 
+ ## Compute point estimates (estimation uncertainty)
+
 robbery_hat <- X_sim_guns %*% t(samp_beta_guns)
@@
 
+#Creating prediction df
+
 pred_df_guns <- data.frame(PP = robbery_hat_quant[2, ],
@@
                       Density = X_sim_guns[, "density"],
-                      Law = c("No", "Yes")[X_sim_guns[, "law"] + 1]) #why can we do this?  `
+                      Law = c("No", "Yes")[X_sim_guns[, "law"] + 1]) 
+#Plotting
 
@@
   geom_ribbon(aes(ymin = LB, ymax = UB, fill = Law), alpha = 0.5) +
-  geom_line() + #what do you think this does?
+  geom_line() + 
   xlab("Population Density") +

### 2023-03-07 Lab 8  BinomialModels/Lab8_BinomialModels.Rmd
Add files via upload

@@
 ---
-title: 'Lab 7: Missing Data and Imputation:Answers'
+title: 'Lab 8: Binomiaal Models in R'
 author: "Eric Parajon"
@@
 #setwd("____") # change to wherever you saved the data
-set.seed(18765) # set seed for replication
+set.seed(1234) # set seed for replication
 
@@
 
-Today, we're going to simulate predicted probabilities for a model with an interaction term and plot our results. We will use both the average case and the observed-case approaches.
+Today, we're going to simulate predicted probabilities for a model with an interaction term, use the distribution of these simulated probabilities to define confidence intervals, and plot our results, using both the average case and the observed-case approaches.
 
@@
 
-+ Average-Case Approach: Simulate and plot predicted probabilities over the range of age, holding all other variables at their mean, for the second model. Plot this.
++ Average-Case Approach: Simulate and plot predicted probabilities over the range of **age**, holding all other variables at their mean, for the second model. Plot this.
 
-+ Observed-Case Approach: Generate an average predicted effect over the observed values of income for the first model. Plot this too.
++ Observed-Case Approach: Generate an average predicted effect over the observed values of **income** for the first model. Plot this too.
 
@@
 
-sim_betas <- mvrnorm(______)
+sim_betas <- mvrnorm(______,______, vcov(m2))
 
@@
 sim_gg <- data.frame(age = ______, pe = apply(sim_pp, 1, mean),
-                     lo = ______,
+                     lo = apply(sim_pp, 1, quantile, probs = .025),
                      hi = ______)
 
-ggplot()
+ggplot(sim_gg, aes(x = age, y = pe, ymin = lo, ymax = hi)) +
+        ______
 
@@
 #getting sim betas for m1
-sim_betas_m1 <- mvrnorm(______)
+sim_betas_m1 <- mvrnorm(______,______,vcov(m1))
 
 #getting range of income
-income_vec <- seq(from = ______,
+income_vec <- seq(from = min(turnout$income),to=____),
                   length.out = 100)
@@
 
-ggplot(______)
+ggplot(plotdat, aes(x = income, y = pe, ymin = lo, ymax = hi)) +
+          ______
 

### 2023-03-07 Lab 8  BinomialModels/Lab8_BinomialModels_AnswerKey.Rmd
Add files via upload

@@
 ---
-title: 'Lab 8: Binomiaal Models in R'
-author: "TA: Eric Parajon"
-output: pdf_document
+title: 'Lab 8: Binomiaal Models in R: Answers'
+author: 'TA: Eric Parajon'
+output:
+  html_document:
+    df_print: paged
 ---
@@
 
-We talked about simulations in Lab 4, when we talked about OLS. This time around, we will do similar things, but with non-Normal outcomes, where simulations are even more helpful!
+Simulation strikes again! This time around, we will do similar things, but with non-Normal outcomes, where simulations are even more helpful.
 

### 2023-03-08 Lab 8  BinomialModels/Lab8_BinomialModels.Rmd
Add files via upload

@@
 ---
-title: 'Lab 8: Binomiaal Models in R'
+title: 'Lab 8: Binomial Models in R'
 author: "Eric Parajon"
@@
 Today's lab has been adapted from code by Chelsea Estancona, Isabel Laterzo, and Simon Hoellerbauer.
-```{r setup, include=FALSE}
+```{r setup, include=T}
 knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
@@
 pacman::p_load(tidyverse,
-               rio,#for importing data
-               arm)#inverse link function
+               rio)#for importing data
 
@@
   geom_bar() +
-  labs(x = 'AUC Presence') +
+  labs(x = 'Autodefensas Unidas de Colombia Presence') +
   theme_bw()
@@
 linpred <- as.numeric(with(colombia, coef(m1)[1] +
-                             coef(m1)[2] * mean(tribut_2_log, na.rm = T) +
-                             coef(m1)[3] * mean(tribut_2_log, na.rm = T)^2 +
-                             coef(m1)[4] * mean(pop_attacks, na.rm = T) +
-                             coef(m1)[5] * mean(disbogota, na.rm = T) +
+                             coef(m1)[2] * mean(tribut_2_log, na.rm = TRUE) +
+                             coef(m1)[3] * mean(tribut_2_log, na.rm = TRUE)^2 +
+                             coef(m1)[4] * mean(pop_attacks, na.rm = TRUE) +
+                             coef(m1)[5] * mean(disbogota, na.rm = TRUE) +
                              coef(m1)[6] * (1) + #could set this to its mode
-                             coef(m1)[7] * mean(pop_log, na.rm = T)))
+                             coef(m1)[7] * mean(pop_log, na.rm = TRUE)))
 
 #let's take a look!
-invlogit #no parentheses from the arm package
+stats::plogis
 
-## invlogit lets us transform our log-odds to the range (0,1) - now interpretable
+## plogis lets us transform our log-odds to the range (0,1) - now interpretable
 ## as a standard probability
-invlogit(linpred)
+
+plogis(linpred)
 ```
@@
 ## create a 'new data' frame: we want to predict the probability for a hypothetical situation.
-new_data <- with(colombia, data.frame(tribut_2_log = mean(tribut_2_log, na.rm = T),
-                                      tribut_2_log_sq = mean(tribut_2_log, na.rm = T)^2,
-                                      pop_attacks = mean(pop_attacks, na.rm = T),
-                                      disbogota = mean(disbogota, na.rm = T),
+new_data <- with(colombia, data.frame(tribut_2_log = mean(tribut_2_log, na.rm = TRUE),
+                                      tribut_2_log_sq = mean(tribut_2_log, na.rm = TRUE)^2,
+                                      pop_attacks = mean(pop_attacks, na.rm = TRUE),
+                                      disbogota = mean(disbogota, na.rm = TRUE),
                                       drugs = c(0, 1),
-                                      pop_log = mean(pop_log, na.rm = T)))
+                                      pop_log = mean(pop_log, na.rm = TRUE)))
 ## note above that I include both possible values for drugs. the 'data.frame()'
@@
 ```{r}
-min_tax <- min(colombia$tribut_2_log, na.rm = T)
-max_tax <- max(colombia$tribut_2_log, na.rm = T)
-min_tax_2 <- min(colombia$tribut_2_log_sq, na.rm = T)
-max_tax_2 <- max(colombia$tribut_2_log_sq, na.rm = T)
+min_tax <- min(colombia$tribut_2_log, na.rm = TRUE)
+max_tax <- max(colombia$tribut_2_log, na.rm = TRUE)
+min_tax_2 <- min(colombia$tribut_2_log_sq, na.rm = TRUE)
+max_tax_2 <- max(colombia$tribut_2_log_sq, na.rm = TRUE)
 
@@
                                                   length.out = 100),
-                            pop_attacks = mean(pop_attacks, na.rm = T),
-                            disbogota = mean(disbogota, na.rm = T),
-                            drugs = 1, pop_log = mean(pop_log, na.rm = T)))
+                            pop_attacks = mean(pop_attacks, na.rm = TRUE),
+                            disbogota = mean(disbogota, na.rm = TRUE),
+                            drugs = 1, pop_log = mean(pop_log, na.rm = TRUE)))
 
@@
                                                   length.out = 100),
-                            pop_attacks = mean(pop_attacks, na.rm = T),
-                            disbogota = mean(disbogota, na.rm = T),
-                            drugs = 0, pop_log = mean(pop_log, na.rm = T)))
+                            pop_attacks = mean(pop_attacks, na.rm = TRUE),
+                            disbogota = mean(disbogota, na.rm = TRUE),
+                            drugs = 0, pop_log = mean(pop_log, na.rm = TRUE)))
 
@@
   labs(x = 'Tax Revenue', y = 'Predicted Probability') +
-  theme_bw()
+  theme_bw()+
+  geom_rug(data = colombia, aes(x = tribut_2_log), 
+           size = 1, inherit.aes = F) #adding rug plot what is this showing us? Why is there a divergence?
+
 ```
@@
 ## do this for the (tax^2) variable, too!
-tax_min <- with(colombia, min(tribut_2_log, na.rm = T))
-tax_max <- with(colombia, max(tribut_2_log, na.rm = T))
+tax_min <- with(colombia, min(tribut_2_log, na.rm = TRUE))
+tax_max <- with(colombia, max(tribut_2_log, na.rm = TRUE))
 tax_vec <- seq(from = tax_min, to = tax_max, length = 100)
@@
                                  tribut_2_log_sq = trans_vec,
-                                 pop_attacks = mean(pop_attacks, na.rm = T),
-                                 disbogota = mean(disbogota, na.rm = T),
+                                 pop_attacks = mean(pop_attacks, na.rm = TRUE),
+                                 disbogota = mean(disbogota, na.rm = TRUE),
                                  drugs = 1,
-                                 pop_log = mean(pop_log, na.rm = T)))
+                                 pop_log = mean(pop_log, na.rm = TRUE)))
 
@@
 
-  pp_sims[i, ] <- invlogit(as.matrix(hyp) %*% beta_sims[i, ])
+  pp_sims[i, ] <- plogis(as.matrix(hyp) %*% beta_sims[i, ])
 
@@
 #we could also do this with less code:
-pp_sims2 <- invlogit(as.matrix(hyp) %*% t(beta_sims))
+pp_sims2 <- plogis(as.matrix(hyp) %*% t(beta_sims))
 
@@
 
-    pp <- invlogit(temporary %*% t(beta_sims))
+    pp <- plogis(temporary %*% t(beta_sims))
 
@@
 
-sim_pp <- apply(sim_betas, 1, function(x) invlogit(as.matrix(sim_dat) %*% x))
+sim_pp <- apply(sim_betas, 1, function(x) plogis(as.matrix(sim_dat) %*% x))
 
@@
 
-    pp <- invlogit(temporary %*% t(sim_betas_m1))
+    pp <- plogis(temporary %*% t(sim_betas_m1))
 

### 2023-03-08 Lab 8  BinomialModels/Lab8_BinomialModels_AnswerKey.Rmd
Add files via upload

@@
 ---
-title: 'Lab 8: Binomiaal Models in R: Answers'
+title: 'Lab 8: Binomial Models in R: Answers'
 author: 'TA: Eric Parajon'
@@
 Today's lab has been adapted from code by Chelsea Estancona, Isabel Laterzo, and Simon Hoellerbauer.
-```{r setup, include=FALSE}
+```{r setup, include=T}
 knitr::opts_chunk$set(echo = TRUE)
@@
 pacman::p_load(tidyverse,
-               rio,
-               arm)#for importing data
+               rio)#for importing data
 ```
@@
 linpred <- as.numeric(with(colombia, coef(m1)[1] +
-                             coef(m1)[2] * mean(tribut_2_log, na.rm = T) +
-                             coef(m1)[3] * mean(tribut_2_log, na.rm = T)^2 +
-                             coef(m1)[4] * mean(pop_attacks, na.rm = T) +
-                             coef(m1)[5] * mean(disbogota, na.rm = T) +
+                             coef(m1)[2] * mean(tribut_2_log, na.rm = TRUE) +
+                             coef(m1)[3] * mean(tribut_2_log, na.rm = TRUE)^2 +
+                             coef(m1)[4] * mean(pop_attacks, na.rm = TRUE) +
+                             coef(m1)[5] * mean(disbogota, na.rm = TRUE) +
                              coef(m1)[6] * (1) + #could set this to its mode
-                             coef(m1)[7] * mean(pop_log, na.rm = T)))
+                             coef(m1)[7] * mean(pop_log, na.rm = TRUE)))
 
 #let's take a look!
-arm::invlogit #no parentheses from the arm package
+stats::plogis
 
-## invlogit lets us transform our log-odds to the range (0,1) - now interpretable
+## plogis lets us transform our log-odds to the range (0,1) - now interpretable
 ## as a standard probability
-invlogit(linpred)
+
+plogis(linpred)
 ```
@@
 
-```{r}
-min_tax <- min(colombia$tribut_2_log, na.rm = T)
-max_tax <- max(colombia$tribut_2_log, na.rm = T)
-min_tax_2 <- min(colombia$tribut_2_log_sq, na.rm = T)
-max_tax_2 <- max(colombia$tribut_2_log_sq, na.rm = T)
+```{r,results="hide"}
+min_tax <- min(colombia$tribut_2_log, na.rm = TRUE)
+max_tax <- max(colombia$tribut_2_log, na.rm = TRUE)
+min_tax_2 <- min(colombia$tribut_2_log_sq, na.rm = TRUE)
+max_tax_2 <- max(colombia$tribut_2_log_sq, na.rm = TRUE)
 
@@
 
+```
+
+```{r}
 #we can calculate the logit predicted probs by hand by doing 1/(1 + exp(-linpred))
@@
 
-```{r, cache = TRUE}
+```{r,message=FALSE, results='hide'}
 ## create a 'temporary' dataset - bind a 1 for the intercept and omit missingness
@@
     })
+
+```
+
+```{r}
 
@@
 
-```{r}
+```{r,message=FALSE, results='hide'}
 pp <- map_dfc(income_vec, function(j) {
@@
 ```{r}
-dim(invlogit(temporary %*% t(sim_betas_m1)))
+dim(plogis(temporary %*% t(sim_betas_m1)))
 ```

### 2023-03-08 Lab 9  Multinom_orderedprob/Lab9_Multinom_OrderedProbit.Rmd
Add files via upload

@@
 ---
-title: "Lab 9: Multinomial and Ordered Probit Models"
+title: 'Lab 9: Multinomial and Ordered Probit Models'
 author: "Eric Parajon"
-output: pdf_document
+output:
+  html_document:
+    df_print: paged
 ---
@@
 
-# Lab Assignment
+# Exercise
 
@@
 
-Generate a cumulative predicted probability plot.
+Generate a cumulative predicted probability plot varying the prestige of the respondent's occupation.
 
@@
 
+#Cleaning the data
+hw<-hw %>% 
+   mutate(warm = case_when(____)) %>%
+    mutate(warm = factor(warm, levels=c("Strongly disagree","Disagree","Agree","Strongly agree")))
+
+#Create the model
+warm_model <- polr(____, Hess = TRUE)
+summary(warm_model)
+
+#What is the effect of increasing each predictor by one unit?
+#How do the odds of moving from a lower to adjacent higher category change?
+coefs <- warm_model$_____
+#odds ratios
+_____(coefs)
+
+#Check the proportional odds assumption for a variable
+
+#Generate a cumulative predicted probability plot.
+table(hw$male); table(hw$white); mean(hw$ed)
+
+hw_2 <- expand.grid(_____,
+                      prst = seq(from = min(hw$prst), to = max(hw$prst), length.out = 2293))
+
+probs <- predict(____, type = "probs") #type = is different than for binomial model
+
+cumul_probs <- t(apply(probs, 1, cumsum)) #get the cumulative sum of the rows
+#transpose into columns for plotting and merging with values of gpa
+
+plot_data <- data.frame(hw$prst, cumul_probs)
+
+#Generate a cumulative predicted probability plot.
+ggplot(data=plot_data, aes(x=hw.prst))+
+      ______
+
 ```

### 2023-03-08 Lab 9  Multinom_orderedprob/Lab9_Multinom_OrderedProbit_Answers.Rmd
Add files via upload

--- 2023-01-04 Lab 9  Multinom_orderedprob/Lab9_Multinom_OrderedProbit.Rmd
@@
 ---
-title: "Lab 9: Multinomial and Ordered Probit Models"
+title: 'Lab 9: Multinomial and Ordered Probit Models: Answers'
 author: "Eric Parajon"
-output: pdf_document
+output:
+  html_document:
+    df_print: paged
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, messsage = FALSE)
+knitr::opts_chunk$set(echo = TRUE, messsage = FALSE, warning=FALSE)
 rm(list=ls())
@@
 
-```{r, eval = F}
+```{r}
 hw <- read.csv(url("http://faculty.washington.edu/cadolph/mle/ordwarm2.csv"))
 
+#Cleaning
+
+hw<-hw %>% 
+   mutate(warm = case_when(.$warm == 1 ~ "Strongly disagree", 
+                          .$warm==2 ~ "Disagree", 
+                          .$warm==3 ~ "Agree", 
+                          .$warm==4 ~ "Strongly agree")) %>%
+    mutate(warm = factor(warm, levels=c("Strongly disagree","Disagree","Agree","Strongly agree")))
+
+#Create the model
+warm_model <- polr(warm~male+white+age+ed+prst, 
+           data = hw, method = "probit", Hess = TRUE)
+summary(warm_model)
+
+#What is the effect of increasing each predictor by one unit?
+#How do the odds of moving from a lower to adjacent higher category change?
+coefs <- warm_model$coefficients[1:5]
+#odds ratios
+exp(coefs)
+
+#Check the proportional odds assumption for male
+m1_prop_1 <- glm(I(as.numeric(warm) >= 2) ~ male, family = "binomial",
+                 data = hw)
+
+m1_prop_2 <- glm(I(as.numeric(warm) >= 3) ~ male, family = "binomial",
+                 data = hw)
+
+#male=0
+check_1 <- m1_prop_1$coef[1] - m1_prop_2$coef[1]
+check_1
+
+#male=1
+check_2 <- (m1_prop_1$coef[1] + m1_prop_1$coef[2]) - 
+  (m1_prop_2$coef[1] + m1_prop_2$coef[2])
+check_2
+
+#Generate a cumulative predicted probability plot.
+table(hw$male); table(hw$white); mean(hw$ed)
+hw_2 <- expand.grid(male = 1, #modal value
+                      white = 1, #modal value
+                      ed=mean(hw$ed),
+                      age=mean(hw$age),
+                      prst = seq(from = min(hw$prst), to = max(hw$prst), length.out = 2293))
+
+probs <- predict(warm_model, hw_2, type = "probs") #type = is different than for binomial model
+head(rowSums(probs)) #just to double check
+cumul_probs <- t(apply(probs, 1, cumsum)) #get the cumulative sum of the rows
+#transpose into columns for plotting and merging with values of gpa
+head(cumul_probs)
+plot_data <- data.frame(hw$prst, cumul_probs)
+
+#plot data
+ggplot(data=plot_data, aes(x=hw.prst))+
+    geom_smooth(data=plot_data, aes(x=hw.prst, y=Strongly.disagree), se=FALSE)+
+    geom_smooth(data=plot_data, aes(x=hw.prst, y=Disagree), se=FALSE)+
+    geom_smooth(data=plot_data, aes(x=hw.prst, y=Agree), se=FALSE)+
+      geom_smooth(data=plot_data, aes(x=hw.prst, y=Strongly.agree), se=FALSE)+
+    geom_ribbon(data=plot_data, aes(ymin=0, ymax=Strongly.disagree, fill="Strongly.disagree"))+
+    geom_ribbon(data=plot_data, aes(ymin=Strongly.disagree, ymax=Disagree,
+                                    fill="Disagree"))+
+    geom_ribbon(data=plot_data, aes(ymin=Disagree, ymax=Agree,
+                                    fill="Agree"))+
+         geom_ribbon(data=plot_data, aes(ymin=Agree, ymax=Strongly.agree,
+                                    fill="Strongly.agree"))+
+    labs(x="Values of prestige of occupation", y="Cumulative Predicted Probability")+
+    scale_fill_discrete(name="Hard Work")+
+    theme_bw()
+
 ```

### 2023-03-21 Lab 9  Multinom_orderedprob/Lab9_Multinom_OrderedProbit.Rmd
Add files via upload

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, messsage = FALSE)
+knitr::opts_chunk$set(echo = TRUE, messsage = FALSE,warning = FALSE)
 rm(list=ls())
@@
                labelled,
-               MASS,
+               ordinal,
                arm)
@@
 
-For ordered outcomes, we need to use the MASS package (rather than base R). Otherwise, our syntax looks much the same.
+For ordered outcomes, we need to use the ordinal package (rather than base R). Otherwise, our syntax looks much the same.
 
@@
 #Now, our model:
-m1 <- polr(apply ~ pared + public + gpa, 
-           data = grad, method = "probit", Hess = TRUE)
+m1 <- clm(apply ~ pared + public + gpa, 
+           data = grad, link = "probit")
 summary(m1)
@@
 #How do the odds of moving from a lower to adjacent higher category change?
-coefs <- m1$coefficients[1:3]
+coefs <- m1$coefficients[3:5]
 #odds ratios
@@
 ```{r}
-tau1 <- m1$zeta[1] #extract latent thresholds (calculated w polr) from our model!
-tau2 <- m1$zeta[2]
+tau1 <- m1$alpha[1] #extract latent thresholds (calculated w polr) from our model!
+tau2 <- m1$alpha[2]
 p1 <- invlogit(tau1) # 1
@@
 colnames(grad_2)<- c("pared", "public", "gpa")
-probs <- predict(m1, grad_2, type = "probs") #type = is different than for binomial model
-head(rowSums(probs)) #just to double check
-cumul_probs <- t(apply(probs, 1, cumsum)) #get the cumulative sum of the rows
+probs <- predict(m1, grad_2, type = "prob") #type = is different than for binomial model
+
+#?predict.clm
+
+head(rowSums(probs$fit)) #just to double check
+cumul_probs <- t(apply(probs$fit, 1, cumsum)) #get the cumulative sum of the rows
 #transpose into columns for plotting and merging with values of gpa
@@
 
-coef_samp <- mvrnorm(1e3,coef(m1)
-                     ,vcov(m1)[1:3,1:3]) #keep relevant vals of vcov matrix
+coef_samp <- mvrnorm(1e3,coef(m1)[3:5]
+                     ,vcov(m1)[3:5,3:5]) #keep relevant columns
 
@@
 ## Get pred probs samples 
-taus <- m1$zeta #extract thresholds 
+taus <- m1$alpha #extract thresholds 
 pred_prob_diffs <- apply(eta,c(1,2)
@@
                mlogit,#for conditional mnl
-               cowplot)#we will use it to put plots together (an extension to ggplot2)
+               patchwork)#we will use it to put plots together (an extension to ggplot2)
 
@@
 
-both <- plot_grid(mplot, fplot, labels = c('Male', 'Female'))
+#Combining plot using patchwork (could also use cowplot())
+both<-mplot+fplot
+
+#Adding annotations
+both<-both + plot_annotation(tag_levels = 'a')
+
+both<-both + plot_annotation(tag_levels = list(c("Male","Female")))& 
+  theme(plot.tag.position = c(0, 1),
+        plot.tag = element_text(hjust = -.5, vjust = -0.20))
+
 both
@@
 
-both2 <- plot_grid(cap_plot, nocap_plot, labels = c('Cap: Yes', 'Cap: No'))
+#combining
+both2<-cap_plot+nocap_plot
+
+#Adding annotations
+both2<-both2 + plot_annotation(tag_levels = 'a')
+
+both2<-both2 + plot_annotation(tag_levels = list(c("Cap: Yes","Cap: No")))& 
+  theme(plot.tag.position = c(0, 1),
+        plot.tag = element_text(hjust = -.5, vjust = -0.20))
+
 both2
@@
 #Create the model
-warm_model <- polr(____, Hess = TRUE)
+warm_model <- clm(____,)
 summary(warm_model)
@@
 #How do the odds of moving from a lower to adjacent higher category change?
-coefs <- warm_model$_____
+coefs <- warm_model$coefficients[5:8]
 #odds ratios
@@
 
-probs <- predict(____, type = "probs") #type = is different than for binomial model
+probs <- predict(____, type = "prob") #type = is different than for binomial model
 
-cumul_probs <- t(apply(probs, 1, cumsum)) #get the cumulative sum of the rows
+cumul_probs <- t(apply(probs$fit, 1, cumsum)) #get the cumulative sum of the rows
 #transpose into columns for plotting and merging with values of gpa

### 2023-03-21 Lab 9  Multinom_orderedprob/Lab9_Multinom_OrderedProbit_Answers.Rmd
Add files via upload

@@
 ```{r}
-tau1 <- m1$zeta[1] #extract latent thresholds (calculated w polr) from our model!
-tau2 <- m1$zeta[2]
+tau1 <- m1$alpha[1] #extract latent thresholds (calculated w clm) from our model!
+tau2 <- m1$alpha[2]
 p1 <- invlogit(tau1) # 1
@@
 #Create the model
-warm_model <- polr(warm~male+white+age+ed+prst, 
-           data = hw, method = "probit", Hess = TRUE)
+warm_model <- clm(warm~male+white+age+ed+prst, 
+           data = hw, link = "probit")
 summary(warm_model)
@@
 #How do the odds of moving from a lower to adjacent higher category change?
-coefs <- warm_model$coefficients[1:5]
+coefs <- warm_model$coefficients[5:8]
 #odds ratios
@@
 
-probs <- predict(warm_model, hw_2, type = "probs") #type = is different than for binomial model
-head(rowSums(probs)) #just to double check
-cumul_probs <- t(apply(probs, 1, cumsum)) #get the cumulative sum of the rows
+probs <- predict(warm_model, hw_2, type = "prob") #type = is different than for binomial model
+head(rowSums(probs$fit)) #just to double check
+cumul_probs <- t(apply(probs$fit, 1, cumsum)) #get the cumulative sum of the rows
 #transpose into columns for plotting and merging with values of gpa

### 2023-03-23 Lab 9  Multinom_orderedprob/Lab9_Multinom_OrderedProbit.Rmd
Add files via upload

@@
 
-Generate a cumulative predicted probability plot varying the prestige of the respondent's occupation.
+Generate a cumulative predicted probability plot varying the age of the respondent.
 
@@
 
-#Generate a cumulative predicted probability plot.
-table(hw$male); table(hw$white); mean(hw$ed)
+#Compute age quartiles and create hypothetical age values between min and max
+age_quarts <- with(hw,c(min(age,na.rm = TRUE), max(age,na.rm=TRUE)))
 
-hw_2 <- expand.grid(_____,
-                      prst = seq(from = min(hw$prst), to = max(hw$prst), length.out = 2293))
+age_rug <- with(hw,seq(age_quarts[1], age_quarts[2], length.out=100))
+
+#Finding modal values
+table(hw$male); table(hw$white); mean(hw$ed); mean(hw$prst)
+hw_2 <- expand.grid(male = 1, #modal value
+                      white = 1, #modal value
+                      ed=mean(hw$ed),
+                      prst=mean(hw$prst),
+                      age = age_rug)
 
@@
 
-plot_data <- data.frame(hw$prst, cumul_probs)
+plot_data <- data.frame(age_rug, cumul_probs)
 
-#Generate a cumulative predicted probability plot.
-ggplot(data=plot_data, aes(x=hw.prst))+
+#Generate a cumulative predicted probability plot varying by age
+ggplot(data=plot_data, aes(x=age_rug))+
       ______

### 2023-03-23 Lab 9  Multinom_orderedprob/Lab9_Multinom_OrderedProbit_Answers.Rmd
Add files via upload

@@
 #Generate a cumulative predicted probability plot.
-table(hw$male); table(hw$white); mean(hw$ed)
+
+#Compute age quartiles and create hypothetical age values between min and max
+age_quarts <- with(hw,c(min(age,na.rm = TRUE), max(age,na.rm=TRUE)))
+
+age_rug <- with(hw,seq(age_quarts[1], age_quarts[2], length.out=100))
+
+#Finding modal values
+table(hw$male); table(hw$white); mean(hw$ed); mean(hw$prst)
 hw_2 <- expand.grid(male = 1, #modal value
@@
                       ed=mean(hw$ed),
-                      age=mean(hw$age),
-                      prst = seq(from = min(hw$prst), to = max(hw$prst), length.out = 2293))
+                      prst=mean(hw$prst),
+                      age = age_rug)
 
@@
 head(cumul_probs)
-plot_data <- data.frame(hw$prst, cumul_probs)
+plot_data <- data.frame(age_rug, cumul_probs)
 
 #plot data
-ggplot(data=plot_data, aes(x=hw.prst))+
-    geom_smooth(data=plot_data, aes(x=hw.prst, y=Strongly.disagree), se=FALSE)+
-    geom_smooth(data=plot_data, aes(x=hw.prst, y=Disagree), se=FALSE)+
-    geom_smooth(data=plot_data, aes(x=hw.prst, y=Agree), se=FALSE)+
-      geom_smooth(data=plot_data, aes(x=hw.prst, y=Strongly.agree), se=FALSE)+
+ggplot(data=plot_data, aes(x=age_rug))+
+    geom_smooth(data=plot_data, aes(x=age_rug, y=Strongly.disagree), se=FALSE)+
+    geom_smooth(data=plot_data, aes(x=age_rug, y=Disagree), se=FALSE)+
+    geom_smooth(data=plot_data, aes(x=age_rug, y=Agree), se=FALSE)+
+      geom_smooth(data=plot_data, aes(x=age_rug, y=Strongly.agree), se=FALSE)+
     geom_ribbon(data=plot_data, aes(ymin=0, ymax=Strongly.disagree, fill="Strongly.disagree"))+
@@
                                     fill="Strongly.agree"))+
-    labs(x="Values of prestige of occupation", y="Cumulative Predicted Probability")+
-    scale_fill_discrete(name="Hard Work")+
+    labs(x="Age of respondent", y="Cumulative Predicted Probability")+
+    scale_fill_discrete(name="Warm Relationship w/ Child")+
     theme_bw()

### 2023-03-29 Lab 10 PoissonNBmodels/Lab10_PoissonNBmodels.Rmd
Add files via upload

@@
 ---
-title: "Lab 10: Poisson and Negative Binomial Models"
+title: 'Lab 10: Poisson and Negative Binomial Models'
 author: "Eric Parajon"
-output: pdf_document
+output:
+  html_document:
+    df_print: paged
 ---
@@
 Load necessary packages:
-```{r setup, include=FALSE}
+```{r setup, include=TRUE}
 knitr::opts_chunk$set(echo = TRUE, messsage = FALSE)
@@
                AER,
-               foreign)
+               foreign,
+               countreg,
+               cowplot)
 ```
@@
 We're going to look at the number of militarized interstate wars based on 'trade network polarization' (lgtrdnpi2) and the number of democratic cliques (clqdemreg) in the system:
+Since we're using count data what should the family argument be?
 
@@
 summary(m1)
+
 ```
@@
 
---- don't look ----
---- don't look ----
---- don't look ----
---- don't look ----
---- don't look ----
---- don't look ----
---- don't look ----
---- don't look ----
---- don't look ----
---- don't look ----
---- don't look ----
---- don't look ----
---- don't look ----
---- don't look ----
+--- don't look\
+--- don't look\
+--- don't look\
+--- don't look\
+--- don't look\
+--- don't look\
+--- don't look\
+--- don't look\
+--- don't look\
+--- don't look\
+--- don't look\
 
@@
 
-m2<- glm.nb(nowars20 ~ lgtrdnpi2 + clqdemreg, 
+m2<- MASS::glm.nb(nowars20 ~ lgtrdnpi2 + clqdemreg, 
             data = count)
@@
 ```{r}
-install.packages("countreg", repos="http://R-Forge.R-project.org")
+#install.packages("countreg", repos="http://R-Forge.R-project.org")
+
 library(countreg)
@@
 ylims <- ylim(-5, 11)  # common scale for comparison
+#Using cowplot to put plots together with labels
 plot_grid(autoplot(root.pois) + ylims, autoplot(root.nb) + ylims, 
@@
 What does this tell us?
+
 The rootogram reflects fitted counts.
@@
 #What if we wanted a single point estimate of our predicted count of wars given our IVs?
-mean(pc_sims)
+mean<-mean(pc_sims)
 
 #And the upper and lower boundaries of a confidence interval based on this simulation?
-quantile(pc_sims, prob= 0.025) 
-quantile(pc_sims, prob= 0.975) 
+lower<-quantile(pc_sims, prob= 0.025) 
+upper<-quantile(pc_sims, prob= 0.975)
+
+#putting it together
+mapply(c,mean,lower,upper)
+
 ```
@@
 ```{r}
+
+#Re-naming count df to temp_data for use in this
 temp_data <- count
@@
 
+#Using map_dfr to directly binds the rows to df we're calling pc_dem_cl
 pc_dem_cl <- purrr::map_dfr(dem_cl, function(x){
@@
 #we can do whatever we want with the distribution of predicted counts
+
 plot_data <- t(apply(pc_dem_cl, 1, quantile, probs = c(.1, .5, .9))) %>% as.data.frame()
 plot_data$dem_cl <- dem_cl
+
+#So this give us the average point estimate (expected count), the low bound of a confidence interval, and the high bound at varying numbers of democratic cliques
+
 ```
+
+# Exercise
 
@@
 
-Below, try using ggplot to plot this information we just generated. Follow these guidelines:
-- X axis: Average number of Democratic Cliques
-- Y axis: Predicted count of Intl. Wars
-- Include a geom_line() for Dem_Cliques
-- Include a geom_ribbon() with CIs (80% confidence interval)
+Below, try using ggplot to plot this information we just generated. Follow these guidelines:\
+- X axis: Average number of Democratic Cliques\
+- Y axis: Predicted count of Intl. Wars\
+- Include a geom_line() for Dem_Cliques\
+- Include a geom_ribbon() with CIs (80% confidence interval)\
 - Include a legend which specifies what the line is and what the CIs are, try putting that legend on the bottom
 
-```{r}
-#Now to plot: 
-colnames(plot_data)<-c("low", "med", "high", "Dem_Cliques")
+```{r, eval=FALSE}
+#Now to plot the above data :
+
+#Renaming columns
+colnames(plot_data)<-c("low", "expected", "high", "Dem_Cliques")
 

### 2023-03-29 Lab 10 PoissonNBmodels/Lab10_PoissonNBmodels_AnswerKey.Rmd
Add files via upload

--- 2023-01-04 Lab 10 PoissonNBmodels/Lab10_PoissonNBmodels_AnswerKey.Rmd
@@
 ---
-title: "Lab 10: Poisson and Negative Binomial Models: Answers"
-author: "TA: Eric Parajon"
-output: pdf_document
+title: 'Lab 10: Poisson and Negative Binomial Models: Answers'
+author: 'TA: Eric Parajon'
+output:
+  html_document:
+    df_print: paged
 ---
@@
                AER,
-               foreign)
+               foreign,
+               cowplot)
 ```
@@
 count <- read_dta("Maoz.dta")
-#count2 <-na.omit(count) # <- don't do this with your data unless you are absolutely
-#sure that you want to drop all nas 
 ```
@@
 
+Since we're using count data what should the family argument be?
+
 ```{r}
-m1 <- glm(data = count, nowars20 ~ lgtrdnpi2 + clqdemreg, family="poisson")
+m1 <- glm(nowars20 ~ lgtrdnpi2 + clqdemreg,
+          data = count,
+          family = "poisson")
 summary(m1)
@@
   theme_bw()
-  #Add legend
 ```
@@
 
-m2<- glm.nb(data = count, nowars20~lgtrdnpi2+clqdemreg)
+m2<- MASS::glm.nb(data = count, nowars20~lgtrdnpi2+clqdemreg)
 summary(m2)
@@
 ylims <- ylim(-5, 11)  # common scale for comparison
+
+#Using cowplot to put plots together with labels
 plot_grid(autoplot(root.pois) + ylims, autoplot(root.nb) + ylims, 
@@
 ```
+
 What does this tell us?
@@
 ```{r}
+#Re-naming count df to temp_data for use in this
 temp_data <- count
@@
 #we can do whatever we want with the distribution of predicted counts
+
 plot_data <- t(apply(pc_dem_cl, 1, quantile, probs = c(.1, .5, .9))) %>% as.data.frame()
 plot_data$dem_cl <- dem_cl
+
+#So this give us the average point estimate (expected count), 
+#the low bound of a confidence interval, and the high bound at varying numbers of democratic cliques
+
 ```
+
+# Exercise
+
+To do on your own:
+
+Below, try using ggplot to plot this information we just generated. Follow these guidelines:\
+- X axis: Average number of Democratic Cliques\
+- Y axis: Predicted count of Intl. Wars\
+- Include a geom_line() for Dem_Cliques\
+- Include a geom_ribbon() with CIs (80% confidence interval)\
+- Include a legend which specifies what the line is and what the CIs are, try putting that legend on the bottom
 
@@
 #Now to plot:
-colnames(plot_data)<-c("low", "med", "high", "Dem_Cliques")
 
-#Collapse information into 3 vectors: for each value of trade in our data, 
-#give us the average point estimate (expected count), the low bound of a confidence interval, and the high bound.
+colnames(plot_data)<-c("low", "expected", "high", "Dem_Cliques")
 
@@
        y = "Predicted Count of Intl. Wars") +
-  geom_line(aes(x = Dem_Cliques, y = med, 
-                color = "Number of Interstate Wars"), size = 1) +
+  geom_line(aes(x = Dem_Cliques, y = expected, 
+                color = "Number of Interstate Wars"), linewidth = 1) +
   geom_ribbon(aes(x = Dem_Cliques, ymin = low, ymax = high, 
@@
 pc_plot
+
 ```

### 2023-04-05 Lab 11 Missing data/Lab11_MissingData.Rmd
Add files via upload

@@
 ```{r}
-#clear global environ
+#clear global environment
 rm(list=ls())
 
-#load in airquality data
-data_clean <- airquality
+#load in iris data
+data_clean <- iris
 summary(data_clean)
 
+#Checking number of NAs
+sum(is.na(data_clean))
+
 #here's our "true model" for future reference
-model_true <- lm(Temp~ Ozone + Solar.R + Wind, data = data_clean)
+model_true <- lm(Sepal.Length~ Sepal.Width+Petal.Length+Petal.Width, data = data_clean)
 
@@
 data_miss <- data_clean
-data_miss[4:9,3] <- rep(NA,6)
-data_miss[1:8,4] <- NA
+data_miss[6:27,2] <- rep(NA,2)
+data_miss[102:119,4] <- rep(NA,2)
+data_miss[35:50,3] <- rep(NA,2)
 summary(data_miss) #check out those *~* NAs ~*~
+
+sum(is.na(data_miss))
 
@@
 #what does this do?
-is.na(data_miss$Ozone)
+is.na(data_miss$Sepal.Width)
 
 #and this?
-which(is.na(data_miss$Ozone))
+which(is.na(data_miss$Sepal.Width))
 
 #and this?
-sum(is.na(data_miss$Ozone))
+sum(is.na(data_miss$Sepal.Width))
 
@@
 #let's create an lm with this model to compare to later iterations
-#model Temp as your y, Ozone, Solar.R, and Wind as your x values
-model_listwise <- lm(Temp ~ Ozone + Solar.R + Wind, data = data_listwise)
+
+model_listwise <- lm(Sepal.Length~ Sepal.Width+Petal.Length+Petal.Width, data = data_listwise)
 
@@
 #uses if else statements to recode NA values as the mean of the rest of that column
-#is.an()
-data_recode$Ozone <- ifelse(is.na(data_recode$Ozone), 
-                            mean(data_recode$Ozone, na.rm = TRUE),
-                            data_recode$Ozone)
+data_recode$Sepal.Width <- ifelse(is.na(data_recode$Sepal.Width),
+                            mean(data_recode$Sepal.Width, na.rm = TRUE),
+                            data_recode$Sepal.Width)
 
-data_recode$Solar.R<- ifelse(is.na(data_recode$Solar.R),
-                            mean(data_recode$Solar.R, na.rm = TRUE),
-                            data_recode$Solar.R)
+data_recode$Petal.Length<- ifelse(is.na(data_recode$Petal.Length),
+                            mean(data_recode$Petal.Length, na.rm = TRUE),
+                            data_recode$Petal.Length)
 
-data_recode$Wind<- ifelse(is.na(data_recode$Wind),
-                            mean(data_recode$Wind, na.rm = TRUE),
-                            data_recode$Wind)
+data_recode$Petal.Width<- ifelse(is.na(data_recode$Petal.Width),
+                            mean(data_recode$Petal.Width, na.rm = TRUE),
+                            data_recode$Petal.Width)
 
-data_recode$Temp <- ifelse(is.na(data_recode$Temp),
-                            mean(data_recode$Temp, na.rm = TRUE),
-                            data_recode$Temp)
+data_recode$Sepal.Length <- ifelse(is.na(data_recode$Sepal.Length),
+                            mean(data_recode$Sepal.Length, na.rm = TRUE),
+                            data_recode$Sepal.Length)
+
 #summarize
@@
 #linear model (same as above, but rename "model_recode")
-model_recode <- lm(Temp~ Ozone + Solar.R + Wind, data = data_recode)
+model_recode <- lm(Sepal.Length~ Sepal.Width+Petal.Length+Petal.Width, data = data_recode)
 
@@
 #pooling
-model_mice <- with(mice_data,lm(Temp~ Ozone + Solar.R + Wind))
+model_mice <- with(mice_data,lm(Sepal.Length~ Sepal.Width+Petal.Length+Petal.Width))
 summary(pool(model_mice))
@@
 
-Great, let's try it out.
-Examine the function by calling `?amelia()` - look at the different arguments.
-Notice that amelia can handle time series data uniquely.
-Let's factor that in by creating a running time variable.
+Great, let's try it out. Examine the function by calling `?amelia()` - look at the different arguments.
 
@@
 
-#create running time variable, lets just assume we're in 2023
-data_miss$time <- make_datetime(month = data_miss$Month,
-                                day = data_miss$Day,
-                                year = 2023)
-
 #impute
 amelia_data <- amelia(x = data_miss,
-                 m = 5,
-                 time = "time")
+                      noms="Species",
+                       m = 5)
 
@@
 for(i in 1:5){
-  models_amelia[[i]] <- lm(Temp ~ Ozone + Solar.R + Wind,
+  models_amelia[[i]] <- lm(Sepal.Length~ Sepal.Width+Petal.Length+Petal.Width,
                         data=amelia_data$imputations[[i]]) #why do we index like this?

### 2023-04-05 Lab 11 Missing data/Lab11_MissingData_AnswerKey.Rmd
Add files via upload

@@
 
-#load in airquality data
-data_clean <- airquality
+#load in iris data
+data_clean <- iris
 summary(data_clean)
 
+#Checking number of NAs
+sum(is.na(data_clean))
+
 #here's our "true model" for future reference
-model_true <- lm(Temp~ Ozone + Solar.R + Wind, data = data_clean)
+model_true <- lm(Sepal.Length~ Sepal.Width+Petal.Length+Petal.Width, data = data_clean)
 
@@
 #let's create an lm with this model to compare to later iterations
-#model Temp as your y, Ozone, Solar.R, and Wind as your x values
-model_listwise <- lm(Temp~ Ozone + Solar.R + Wind, data = data_listwise)
+
+model_listwise <- lm(Sepal.Length~ Sepal.Width+Petal.Length+Petal.Width, data = data_listwise)
 
@@
 #uses if else statements to recode NA values as the mean of the rest of that column
-data_recode$Ozone <- ifelse(is.na(data_recode$Ozone),
-                            mean(data_recode$Ozone, na.rm = TRUE),
-                            data_recode$Ozone)
+data_recode$Sepal.Width <- ifelse(is.na(data_recode$Sepal.Width),
+                            mean(data_recode$Sepal.Width, na.rm = TRUE),
+                            data_recode$Sepal.Width)
 
-data_recode$Solar.R<- ifelse(is.na(data_recode$Solar.R),
-                            mean(data_recode$Solar.R, na.rm = TRUE),
-                            data_recode$Solar.R)
+data_recode$Petal.Length<- ifelse(is.na(data_recode$Petal.Length),
+                            mean(data_recode$Petal.Length, na.rm = TRUE),
+                            data_recode$Petal.Length)
 
-data_recode$Wind<- ifelse(is.na(data_recode$Wind),
-                            mean(data_recode$Wind, na.rm = TRUE),
-                            data_recode$Wind)
+data_recode$Petal.Width<- ifelse(is.na(data_recode$Petal.Width),
+                            mean(data_recode$Petal.Width, na.rm = TRUE),
+                            data_recode$Petal.Width)
 
-data_recode$Temp <- ifelse(is.na(data_recode$Temp),
-                            mean(data_recode$Temp, na.rm = TRUE),
-                            data_recode$Temp)
+data_recode$Sepal.Length <- ifelse(is.na(data_recode$Sepal.Length),
+                            mean(data_recode$Sepal.Length, na.rm = TRUE),
+                            data_recode$Sepal.Length)
 
@@
 #pooling
-model_mice <- with(mice_data,lm(Temp~ Ozone + Solar.R + Wind))
+model_mice <- with(mice_data,lm(Sepal.Length~ Sepal.Width+Petal.Length+Petal.Width))
 
@@
 
-#create running time variable, lets just assume we're in 2023
-data_miss$time <- make_datetime(month = data_miss$Month,
-                                day = data_miss$Day,
-                                year = 2023)
-
-#impute
+#imputing (setting factor to nominal variable)
 amelia_data <- amelia(x = data_miss,
-                 m = 5,
-                 time = "time")
+                      noms="Species",
+                       m = 5)
 
@@
 
-#alternatively!
-all_coefs <- sapply(models_africa, coef) #get coefs
-all_ses <- sapply(models_africa, function(x) sqrt(diag(vcov(x)))) #get ses
-
-mi.meld(q = all_coefs, se = all_ses, byrow = F) 
 ```

### 2023-04-05 Lab 11 Missing data/Lab11_MissingData.Rmd
Add files via upload

@@
 data_miss <- data_clean
+data_miss[52:59,1] <- rep(NA,2)
 data_miss[6:27,2] <- rep(NA,2)

### 2023-04-12 Lab 13  CBPS/Lab12_covariatebalance.Rmd
Add files via upload

--- 2023-01-04 Lab 13  CBPS/Lab13_covariatebalance.Rmd
@@
 ---
-title: "Lab 13: Covariate Balancing and CBPS"
+title: "Lab 12: Covariate Balancing and CBPS"
 author: "Eric Parajon with code adapted from Colin Case"
@@
 
-Randomized controlled trials (RCTs) are often seen as the ``gold standard'' for estimating the effects of treatments, interventions, and exposures on outcomes.
+Randomized controlled trials (RCTs) are often seen as the "gold standard" for estimating the effects of treatments, interventions, and exposures on outcomes.
 However, in the social sciences, it is not always possible to achieve randomization (and even in experimental studies, this doesn't happen either!), and we may have unobserved confounding variables that affect someone's likelihood of receiving treatment.
@@
           se=list(NULL, NULL, corr_se),
-          column.labels = c("No Weights OLS", "PS Weights SWLS", "PS/NCP Weights SWLS"),
+          column.labels = c("No Weights", "PS Weights", "PS/NCP Weights"),
           star.cutoffs = c(.10,.05,.01,.001),
@@
 ```
-
-From the three models reported, we see that there is a small positive effect of being primed the cost of the existing municipal structure.
-Across three models, the effect is significant, p<.10+.
-These primed respondents are likely to increase their rating of the conservative party .1 point more on a 1-7 scale.
-The effect size is modest but it is statistically significant in all three models, regardless of the weighting method used.

### 2023-04-12 Lab 14  RDD/Lab12_covariatebalance_Answers.Rmd
Add files via upload

--- 2023-01-04 Lab 13  CBPS/Lab13_covariatebalance_Answers.Rmd
@@
 ---
-title: 'Lab 13: Covariate Balancing and CBPS: Answers'
+title: 'Lab 12: Covariate Balancing and CBPS: Answers'
 author: "Eric Parajon with code adapted from Colin Case"
 output:
+  html_document: default
   pdf_document: default
-  html_document: default
 header-include:
@@
           se=list(NULL, NULL, corr_se),
-          column.labels = c("No Weights OLS", "PS Weights SWLS", "PS/NCP Weights SWLS"),
+          column.labels = c("No Weights", "PS Weights", "PS/NCP Weights"),
           star.cutoffs = c(.10,.05,.01,.001),
           star.char = c("+","*","**","***"),
-          column.sep.width = "10pt",
+          column.sep.width = "11pt",
           notes = c("p<.10+, p<.05*, p<.01**, p<.001***"),
@@
 
-From the three models reported, we see that there is a small positive effect of being primed the cost of the existing municipal structure.
+**Interpretation**: From the three models reported, we see that there is a small positive effect of being primed the cost of the existing municipal structure.
 Across three models, the effect is significant, p<.10+.

### 2023-04-18 Lab 14  RDD/Lab13_RDD.Rmd
Add files via upload

--- 2023-01-04 Lab 14  RDD/Lab14_RDD.Rmd
@@
 ---
-title: "Lab 14 - Regression Discontinuity (RDD) and Instrumental variables (IV)"
+title: "Lab 13 - Regression Discontinuity (RDD) and Instrumental variables (IV)"
 author: "Eric Parajon, with code adapted from Colin Case and Sean Norton"
 output: html_document
-bibliography: rdd.bib
 ---
@@
 
-#Running iv model using ivreg from the aer package
+#Running basic model
+# models wage as a function of a constant and educ
+reg_ex <- lm(lwage~educ,data=women)
+summary(reg_ex)
+
+#This indicates that every additional year of education increases the wage by almost 11% 
+#The issue with this sort of model is that education is most likely to be correlated with individual characteristics that are important for the person's wage, but not modelled (and hence captured by the error term).
+
+#Enter IV!
+
+#We'll be using parents’ education (fatheduc and motheduc) as instruments for education.
+#Running iv model using ivreg() from the aer package
+
+#Before the vertical line we can see the model that is to be estimated, lwage~educ+age+exper+expersq
+#After the vertical line, we first see the instrumental variables used to instrument educ, fatheduc+motheduc; this is followed by all the explanatory variables that are considered exogenous, (age+exper+expersq), rather than re-writing them we can tell R which variables to remove from the instrument set (education).
+
 reg_iv1 <- ivreg(lwage~ educ+age+exper+expersq| 
@@
 #Running first_stage model
-first_stage <- lm(educ~fatheduc+motheduc,data=women) #regress X on Z 
+first_stage <- lm(educ~age+exper+expersq+fatheduc+motheduc,data=women) #regress X on Z 
 #Running null model
@@
 
-For the exercise we are going to replicate the paper by @lmb2004 : \textit{Do Voters Affect or Elect Policies?
+For the exercise we are going to replicate the paper by David S.
+Lee, Enrico Moretti, Matthew J.
+Butler : \textit{Do Voters Affect or Elect Policies?
 Evidence from the U.S.
@@
 
-```{r}
+```{r, eval=FALSE}
 # Load Data
@@
 # Plot RDD
+lmb.fit <- rdrobust(enricoall4$realada,
+            _________)
+rdplot(____,
+       ______,
+       x.label="Democrat Vote Share at time t",
+       y.label="ADA Score at time t+1",
+       c=0.5)
 
@@
 
-```{r}
+```{r, eval=FALSE}
 # Fit RDD Model specifying clustered standard errors for DW-Nom as DV
 
+lmb.fit2 <- rdrobust(enricoall4$dwnom1,
+                     ________)
+
 # Plot RDD
+rdplot(_____,
+       ______,
+       x.label="Democrat Vote Share at time t",
+       y.label="DW-Nominate Score at time t",
+         c=0.5)
 
@@
 
-Yes, we have successfully replicated their results found in the paper.
-
-## McCrary Test
-
-```{r}
+```{r,eval=FALSE}
 #McCrary Test
 
+DCdensity(______)
 ```
-The McCrary Test does not show sharp jumps at around the threshold (c=0.5), but it does alert us that there are heavy tails in the distribution of the running variable.
-The heavy tails come from Republican strongholds (Democrats have no chance of winning) and Democrat strongholds (Democrats always win).
-You can also see this above with the lack of observations around the cutpoint.
-This does lead to some concerns about the robustness of the RD effect we found, and particularly in the sudden reversal of RD trends at the upper tail of the running variable.
-Substantively this makes sense - if Democrats always win a district, then there's no point to adjust their policy predispositions to election results!
-
-# Bibliography

### 2023-04-18 Lab 14  RDD/Lab13_rdd_Answers.Rmd
Add files via upload

--- 2023-01-04 Lab 14  RDD/Lab14_rdd_Answers.Rmd
@@
 ---
-title: "Lab 14 - Regression Discontinuity (RDD) and Instrumental variables (IV):Answers"
+title: "Lab 13 - Regression Discontinuity (RDD) and Instrumental variables (IV):Answers"
 author: "Eric Parajon, with code adapted from Colin Case and Sean Norton"
 output:
+  html_document: default
   pdf_document: default
-  html_document: default
 ---
@@
 ```{r}
-
 #Reading in the data
@@
 #Based on the output, I find that the instruments are both relevant (confirmed by F-test) and valid (confirmed by Sargan test) to education, and that higher levels of education slightly increases a women's wage (significant at the 0.1 level).
-
 ```
@@
 They used the ADA Score and the DW-Nominate Scores (figures I and VI respectively in the paper), both are quantitative measures of the ideological position of representatives.
+
 They argue that the running variable to the Democratic Vote Share at time $t$ - So if democrats receive more than half the share of votes, they would win in that district and a democratic representative would be elected.
@@
 The heavy tails come from Republican strongholds (Democrats have no chance of winning) and Democrat strongholds (Democrats always win).
+
 You can also see this above with the lack of observations around the cutpoint.
